### 2012-02-11 exam3/1.tex
start on exam 3

% You must \input this file into a file the has previously
% defined the \ifrubric conditional.

\documentclass[12pt]{article}                   % -*- latex -*-

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{./exam}
\usepackage{type1cm} % We are using scalable fonts. Really.
\usepackage{alltt,times,comment,amsmath}
\usepackage{ct} % tight CM font for \tt
\newexercise{exercise}{Exercise}[subsection]

\makeatletter
\newcommand{\ie}{\mbox{\emph{i.e.}}}    % \mbox keeps the last period from
\newcommand{\Ie}{\mbox{\emph{I.e.}}}    % looking like an end-of-sentence.
\newcommand{\etc}{\emph{etc.}}

\newcommand{\itum}[1]{\item{\bf #1}\\*}

\newcommand{\var}[1]{\textrm{\textit{#1}}}

\newenvironment{inset}
 {\bgroup\parskip=1ex plus 1ex\begin{list}{}%
        {\topsep=0pt\rightmargin\leftmargin}%
        \item[]}%
 {\end{list}\leavevmode\egroup\global\@ignoretrue}

\newenvironment{insetverb}%
  {\begin{inset}\begin{verbatim}}%
  {\end{verbatim}\end{end}}%

\def\pts#1{\marginpar{\footnotesize \raggedright  \fbox{#1 {\sc Points}}}}
%\def\pts#1{\relax}

\ifrubric
\newenvironment{solution}{}{}
\else
\excludecomment{solution}
\fi
\newexercise{problem}{Problem}

\newcommand\code[1]{\texttt{#1}}

% Works in math mode; all special chars remain special; cheaper than \cd.
% Will not be correct size in super and subscripts, though.
\newcommand{\ex}[1]{\mbox{\ttt #1}} 
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\vspace*{-1.5cm}
\ifrubric
\centerline{\Large CSU2510H Exam 1 Rubric -- Spring 2012}

\else
\centerline{\Large CSU2510H Exam 1 -- Spring 2012}

\vspace{0.5cm}

\begin{center}
\begin{tabular}{l@{\qquad}l}
Name:                        & \rule{174pt}{1pt} \\[.5cm]
Student Id (last 4 digits):  & \rule{174pt}{1pt} \\[.5cm]
%Section (morning, honors or afternoon):           & \rule{174pt}{1pt} \\[.5cm]
\end{tabular}
\end{center}

\noindent\begin{minipage}{8cm}\sloppy
\begin{itemize}
\item You may use the usual primitives and expression forms of any of
  the \code{class} languages; for everything else, define it.

\item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} for
  {\tt (check-expect {\slshape c e})} and $\lambda$ for
  \code{lambda} to save time writing.

\item To add a method to an existing class definition, you
  may write just the method and indicate the appropriate class name
  rather than re-write the entire class definition.

\item We expect data \emph{and} interface definitions
  (although interfaces may be defined in comments rather than
  with \code{define-interface}).

\item If an interface is given to you, you do not need to repeat the
  contract and purpose statements in your implementations.  Likewise,
  you do not need to repeat any test cases given to you, but you
  should add tests wherever appropriate.

%% DVH: This info is already in the table.
% \item You may obtain a maximum of 55 points: 50 for the first six
% problems; and five extra-credit points for the final problem.

\item Some basic test taking advice: Before you start answering
any problems, read \emph{every} problem, so your brain can be thinking
about the harder problems in background while you knock off the easy ones.
\end{itemize}

\bigskip

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & ?\\ \hline
2 & & ?\\ \hline
3 & & ?\\ \hline
4 & & ?\\ \hline
5 & & ?\\ \hline
Extra & & ? \\ \hline
{\bf Total} & & 83 \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

\vfill\thispagestyle{empty}
\newpage

\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \noindent
% {\bf Home on the Range}

% Come up with funny story here.

% \newpage
\begin{problem}\pts{17} % same as on the previous exam :)

A \emph{range} represents a set of numbers between two endpoints.  To
start with, you only need to consider ranges that \emph{include} the
smaller endpoint and \emph{exclude} the larger endpoint---such ranges
are called \emph{half-open}.  For example, the range $[3,4.7)$
includes all of the numbers between $3$ and $4.7$, including $3$ but
\emph{not} including $4.7$. So $4$ and $3.0000001$ are both in the
range, but $5$ is not.  In the notation used here, the $[$ means
include, and the $)$ means exclude.

\smallskip

\noindent
\begin{enumerate}
\item Design a representation for ranges and implement the
  \texttt{in-range?} method, which determines if a number is in the
  range.  For example, the range $[3,7.2)$ includes the numbers $3$
  and $5.0137$, but not the numbers $-17$ or $7.2$.

\item Extend the data definition and implementation of ranges to
  represent ranges that \emph{exclude} the low end of the range and
  \emph{include} the high end, written $(lo,hi]$.

\item Add a \texttt{union} method to the interface for ranges and
  implement it in all range classes.  This method should consume a
  range and produces a new range that includes all the numbers in this
  range \emph{and} all the numbers in the given range.

  You may extend your data definition for ranges to
  support this method.

  You may apply the abstraction recipe, but you are not required to do
  so; you will not be marked down for duplicated code on this problem.
\end{enumerate}

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

\begin{solution}
\begin{verbatim}
;; Part 1:

;; A Range is one of
;; - (new range% Number Number)                         [1pt]
;; Interp: represents the range between `lo' and `hi'
;;         including `lo', but *not* including `hi'
;; and implements IRange:

;; The IRange interface includes:
;; - in-range? : Number -> Boolean                      [1pt]
;; determine if the given number is in this range
;; - union : Range -> Range                             [1pt]
;; produce the range containing this and the given range

(define-class range%                                    [1pt]
  (fields lo hi)
  (check-expect ((range% 0 1) . in-range? 0) true)      [1pt]
  (check-expect ((range% 0 1) . in-range? 1) false)
  (define/public (in-range? n)                          [2pt]
    (and (>= n (field lo)) (< n (field hi))))

  (define/public (union r)
    (union-range% this r)))

;; Part 2:

;; A Range is also:
;; - (new hi-range% Number Number)                      [1pt]
;; Interp: represents the range between `lo' and `hi'
;;         including `hi', but *not* including `lo'

(define-class hi-range%                                 [1pt]
  (fields lo hi)
  (check-expect ((hi-range% 0 1) . in-range? 0) false)  [1pt]
  (check-expect ((hi-range% 0 1) . in-range? 0.5) true)
  (check-expect ((hi-range% 0 1) . in-range? 1) true)
  (define/public (in-range? n)                          [1pt]
    (and (> n (field lo)) (<= n (field hi))))

  (define/public (union r)
    (union-range% this r)))

;; Part 3:

;; A Range is also:
;; - (new union-range% Range Range)                     [1pt]
;; Interp: including all the numbers in both ranges

(define-class union-range%                              [1pt]
  (fields left right)
  (define/public (in-range? n)                          [2pt]
    (or ((field left) . in-range? n)
        ((field right) . in-range? n)))

  ;; Point is for all the union methods together
  (define/public (union r)                              [1pt]
    (union-range% this r)))

(define r1 (range% 0 1))
(define r2 (hi-range% 0 1))
(define r3 (range% 2 4))

;; union + in-range? test                               [1pt]
(check-expect (r1 . union r2 . union r3 . in-range? 3) 
              true)

;; testing all the union methods                        [1pt]
(check-expect (r1 . union r2 . in-range? 1) true)
(check-expect (r2 . union r1 . in-range? 0) true)
\end{verbatim}
\end{solution}
\end{problem}
%\ifrubric\bigskip\else
\newpage
%\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{18}

Here are data, class, and interface definitions for Shapes:
\begin{verbatim}
;; A Shape is one of:
;; - (rect% Number Number)
;; - (circ% Number)
;; implements:
;; bba : -> Number     (short for "bounding-box-area")
;; Compute the area of the smallest bounding box for this shape.

(define-class rect%
  (fields width height))

(define-class circ%
  (fields radius))
\end{verbatim}
%
Here are some examples of how \code{bounding-box-area} should work:
\begin{verbatim}
(check-expect ((rect% 3 4) . bba) 12)
(check-expect ((circ% 1.5) . bba)  9)
\end{verbatim}

\begin{enumerate}
\item Design the \code{bba} method for the
  \code{rect\%} and \code{circ\%} class.

\item Design a super class of \code{rect\%} and \code{circ\%} and
  lift the \code{bba} method to the super class.  Extend
  the shape interface as needed, but implement any methods you add.

\item Design a new variant of a Shape, Square, which should support
  all of the methods of the interface.
\end{enumerate}

\begin{solution}
\begin{verbatim}
;; Part 1 [6pt]
(define-class rect% 
  (fields width height)                
  (define/public (bba)                 [2pt]
    (* (field width) (field height))))

(define-class circ% 
  (fields radius)                      
  (define/public (bba)                 [2pt]
    (sqr (* 2 (field radius)))))

;; Part 2 [10pt]

;; implements:
;; width : -> Number                         [1pt]
;; height : -> Number
;; Compute the {width,height} of this shape. [1pt]

(define-class shape% 
  (define/public (bounding-box-area)
    (* (this . width) (this . height))))     [3pt]

(define-class rect% 
  (super shape%)                             [1pt]
  (fields width height))                     [1pt]

(define-class circ% 
  (super shape%)                             [1pt]
  (fields radius)
  (define/public (width)                     [1pt]
    (* 2 (field radius)))
  (define/public (height)                    [1pt]
    (width)))

;; Part 3 [4pt]

(define-class square%
  (super shape%)                             [1pt]
  (fields width)                             [1pt]
  (define/public (height)                    [1pt]
    (width)))

(check-expect ((square% 5) . bba) 25)        [1pt]
\end{verbatim}
\end{solution}

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{12}

One perspective that unifies the paradigms of programming with
functions and programming with objects is to view a ``function'' as an
object that understands a method called \code{apply}.  With that in
mind, we can define an interface for functions-as-objects:

\begin{verbatim}
;; A [IFun X Y] implements:
;; apply : X -> Y
;; Apply this function to the given input.
\end{verbatim}

Here we are representing a \code{X -> Y} function as an object that
has an \code{apply} method that consumes an \code{X} and produces a
\code{Y}.

\begin{enumerate}
\item Design a class that wraps a real function with contract \code{(X
  -> Y)} to implement \code{[IFun X Y]}.

\item Using your wrapper class, construct the objects representing the
  ``add 1'' function and ``sub 1'' function.

\item Another useful operation on functions is composition.  Here is
  the interface for a \code{compose} method that composes two functions
  represented as objects:

\begin{verbatim}
;; [IFun X Y] also implements:
;; compose : [IFun Y Z] -> [IFun X Z]
;; Produce a function that applies this function to its input, 
;; then applies the given function to that result.
\end{verbatim}

For example, if \code{addone} and \code{subone} refer to the objects
you constructed in part 2, the following check should succeed:

\begin{verbatim}
(check-expect ((addone . compose subone) . apply 5) 5)
(check-expect ((addone . compose addone) . apply 5) 7)
\end{verbatim}

Implement the \code{compose} method for your wrapper class.
\end{enumerate}

\begin{solution}

\begin{verbatim}
;; Part 1 [5pt]
(define-class fun%
  (fields f)
  (define/public (apply x)
    ((field f) x)))

;; Part 2 [2pt]
(define addone (fun% add1))
(define subone (fun% sub1))

;; Part 3 [5pt]
(define-class fun%
  (fields f)
  (define/public (compose g)
    (fun% (lambda (x)
            (g . apply (apply x))))))
\end{verbatim}

\end{solution}

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{21}

The Northeastern University Registrar is instituting a new course
registration system, in which each student will wait in a ``Virtual
Line'' until every student ahead of them has registered.  A simple way
to represent a line (also known as a \emph{queue}) is by using a list.
But this representation makes it slow to add somebody to the end of
the line (or to take somebody off the front of the line, depending on
whether the front of the list represents the front or rear of the
line).

In order to provide maximal waiting efficiency, you have been tasked
with implementing a representation that uses \emph{two} lists!  The
key idea of this fancy representation is that one list will represent
some portion of the front of the line, while the other will represent
the remainder of the line \emph{in reverse order}.  So if you're the
first element of the first list, you are at the head of the line.  On
the other hand, if you're the first element of the second list, you
are the very last person in line.
%
Here is the interface for queues:
\begin{verbatim}
;; A [IQ X] implements:
;;
;; head : -> X
;; Produce the element at the head of this queue.
;; Assume: this queue is not empty.
;;
;; deq : -> [IQ X]      (Short for "dequeue")
;; Produces a new queue like this queue, but without 
;; this queue's first element.
;; Assume: this queue is not empty.
;;
;; enq : X -> [IQ X]    (Short for "enqueue")
;; Produce new queue with given element added to the
;; END of this queue.
;;
;; emp? : -> Boolean
;; Is this queue empty?
\end{verbatim}

The \code{head} and \code{deq} operations require that the queue be
non-empty when they are used, but this can be assumed and these
operations do not need to check for an empty queue.

Further, the Registrar's office has just learned about
\emph{invariants}, and insists on maintaining the following invariant
about all of their queues:

\begin{quote}
  {\it if the front of the queue is empty, the whole queue must also be empty.}
\end{quote}

\noindent
The Registrar's office has given you three tasks to prepare their
Virtual Line for its launch later this semester:

\begin{enumerate}
\item Design an implementation of the queue data structure to the Registrar's
  specifications.  You must maintain the invariant stated above, and
  you should take advantage of the invariant when implementing the
  operations.
\item Unfortunately, when testing the queue, the Registrar has
  discovered that some queues with the same elements in the same order
  can be represented in multiple ways.  Give an example of two
  different representations of the same queue.  Implement a
  \code{to-list} operation which produces a list of elements going in
  order from the front to the rear of the queue. In your tests, you
  should show how this addresses the problem.
\item The Registrar has a problem with careless data entry.  Design
  and implement a constructor for queues which, given two input lists
  of elements, ensures that the invariant is maintained.  
\end{enumerate}

\begin{solution}
\begin{verbatim}
;; Part 1:
;; invariant: front is empty IFF queue is empty
;; A [Queue X] is (queue% [Listof X] [Listof X])           [1pt]
;; and implements [IQ X]

(define-class queue%                                       [1pt]
  (fields front back)
  ;; Add element to the back
  ;; X -> [Queue X]
  ;; add it to the front IFF the front is empty
  ;; (otherwise invariant is violated)
  (check-expect (eq . enqueue 1) (queue% '(1) '()))        [1pt]
  (check-expect (q . enqueue 5) (queue% '(1 2 3) '(5)))
  (define/public (enqueue i)
    (cond [(empty? (field front))                          [2pt]
           (queue% (list i) empty)]
          [else
           (queue% (field front) (cons i (field back)))])) [2pt]
  ;; get the first element
  ;;  -> X
  ;; assumes this queue is non-empty
  ;; uses the invariant
  (check-expect (q . head) 1)                              [1pt]
  (define/public (head)                                    
    (first (field front)))
  ;; drop the first element
  ;;  -> [Queue X]
  ;; assumes this queue is non-empty
  (check-expect (eq . enqueue 1 . dequeue) eq)             [1pt]
  (check-expect (q . enqueue 5 . dequeue) (queue% '(2 3) '(5)))
  (define/public (dequeue)
    (cond 
      ;; maintain the invariant -- don't let `front' be empty
      [(empty? (rest (field front)))                       [3pt]
       (queue% (reverse (field back)) empty)]
      ;; otherwise, simple
      [else (queue% (rest (field front)) (field back))]))  [2pt]
  ;; is this queue empty?
  ;; emp? : -> Boolean
  ;; uses the invariant
  (check-expect (q . emp?) false)                          [1pt]
  (check-expect (eq . emp?) true)
  (define/public (emp?)
    (empty? (field front)))

;; Part 2:

;; Examples:                                               [1pt]
(queue% (list 1) (list 6)) vs (queue% (list 1 6) empty)
(queue% (list 1 8 12) (list 6)) vs (queue% (list 1 8) (list 12 6))

  ;; convert the queue to a list                           [1pt]
  ;; -> [Listof X] 
  ;; turn examples above into tests:                       [1pt]
  (check-expect ((queue% (list 1) (list 6)) . to-list)
		((queue% (list 1 6) empty) . to-list))
  (check-expect ((queue% (list 1 8 12) (list 6)) . to-list)
		((queue% (list 1 8) (list 12 6)) . to-list))
  (check-expect (q . to-list) '(1 2 3))                    
  (check-expect (q . enqueue 5 . to-list) '(1 2 3 5))   
  (check-expect (q . enqueue 5 . dequeue . to-list) '(2 3 5))
  (define/public (to-list)                                 [1pt]
    (append (field front)
            (reverse (field back)))))

(define q (queue% '(1 2 3) empty))
(define eq (queue% empty empty))

;; Part 3:                                                 [2pt]
;; several good possibilities (others are possible)
(constructor (f b)
  (fields (append f (reverse b)) empty))

(constructor (f b)
  (cond [(empty? f) (fields (reverse b) empty)]
	[else (fields f b)]))

(constructor (f b)
  (cond [(and (empty? f) (empty? b)) (fields f b)]
	[(empty? f) (fields (reverse b) empty)]
	[else (fields f b)]))

\end{verbatim}
\end{solution}
\end{problem}

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}
%
Here is a simple interface and implementation for dictionaries that
associate natural numbers to strings.

\begin{verbatim}
;; A Dict is one of:
;;  - (ld-empty%)
;;  - (ld-cons% Nat String Dict)
;; implements:
;;
;; has-key? : Nat -> Boolean
;; Does this dict. have the given key?
;;
;; lookup : Nat -> String
;; Produce the value associated with the given key in this dict.
;; Assume: the key exists in this dict.
;;
;; set : Nat String -> Dict
;; Associate the given key to the given value in this dict.

(define-class ld-empty%
  (define/public (has-key? k) false)

  (define/public (set k v)
    (ld-cons% k v this)))

(define-class ld-cons%
  (fields key val rest)

  (define/public (has-key? k)
    (cond [(= (field key) k) true]
          [else ((field rest) . has-key? k)]))

  (define/public (lookup k)
    (cond [(= (field key) k) (field val)]
          [else ((field rest) . lookup k)]))

  (define/public (set k v)
    (ld-cons% k v this)))
\end{verbatim}

The dictionary data structure has some important properties, one of
which is that if we associate a value with a key and then look up that
key, we expect to get back the value we associated.  For example, for
any given dictionary, if we associate \code{1} with \code{"Fred"},
then look up \code{1}, we should get \code{"Fred"}.  More generally,
we expect this to be true for \emph{any} key-value pair, not just
\code{1} and \code{"Fred"}.

\begin{enumerate}
\item Codify the general ``set/lookup'' property, described above, as
  a predicate.  Write a successful test in terms of that predicate.
\end{enumerate}

Now consider a generalization of the ``set/lookup'' property which
says that if we set \emph{two} keys to \emph{two} values then look
each of them up, we should get back the original values.  For example, for
any given dictionary, if we associate \code{1} with \code{"Fred"} and
\code{2} with \code{"Wilma"}; then looking up \code{1} will produce
\code{"Fred"} and looking up \code{2} will produce \code{"Wilma\"}.
More generally, we can state this as a property for any two key-value
pairs.

\begin{enumerate}
\setcounter{enumi}{1}
\item Codify the general ``double-set/lookup'' property as a
  predicate.  Write a successful test in terms of that predicate.

\item Does this property hold for all dictionaries and keys-value
  pairs?  If so, why?  If not, write a counter-example, i.e.~write a
  failing test in terms of the predicate you gave in part 2.
\end{enumerate}

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi
\end{problem}

\begin{solution}
\begin{verbatim}
;; Part 1 
;; Dict Nat String -> Boolean            [4pt]
(define (set/lookup d k v)
  (string=? (d . set k v . lookup k) v))

;; [2pt]
(check-expect (set/lookup (ld-empty%) 1 "Fred") true)

;; Part 2
;; Dict Nat String Nat String -> Boolean   [4pt]
(define (double-set/lookup d k1 v1 k2 v2)
  (and (string=? (d . set k1 v1 . set k2 v2 . lookup k1) v1)
       (string=? (d . set k1 v1 . set k2 v2 . lookup k2) v2)))

;; [2pt]
(check-expect (double-set/lookup (ld-empty%) 1 "Fred" 2 "Wilma") true)

;; Part 3  [3pt]
;; Counter-example:
(check-expect (double-set/lookup (ld-empty%) 1 "Fred" 1 "Wilma") true)
\end{verbatim}
\end{solution}

\end{document}

### 2012-02-11 exam3/1.tex
nat and bool problems

@@
 \begin{problem} \pts{15}
-%
-Here is a simple interface and implementation for dictionaries that
-associate natural numbers to strings.
+Last night, Asumu was up late improving the {\tt class} languages.
+Unfortunately, he was too tired to test his work, and his last commit
+broke the numeric system.
+
+This puts us in the dire position of no longer being able to calculate
+the factorial of large numbers, and we turn to you for help.
+
+This is the definition of a class with a {\tt fact} method:
 
 \begin{verbatim}
-;; A Dict is one of:
-;;  - (ld-empty%)
-;;  - (ld-cons% Nat String Dict)
-;; implements:
-;;
-;; has-key? : Nat -> Boolean
-;; Does this dict. have the given key?
-;;
-;; lookup : Nat -> String
-;; Produce the value associated with the given key in this dict.
-;; Assume: the key exists in this dict.
-;;
-;; set : Nat String -> Dict
-;; Associate the given key to the given value in this dict.
-
-(define-class ld-empty%
-  (define/public (has-key? k) false)
-
-  (define/public (set k v)
-    (ld-cons% k v this)))
-
-(define-class ld-cons%
-  (fields key val rest)
-
-  (define/public (has-key? k)
-    (cond [(= (field key) k) true]
-          [else ((field rest) . has-key? k)]))
-
-  (define/public (lookup k)
-    (cond [(= (field key) k) (field val)]
-          [else ((field rest) . lookup k)]))
-
-  (define/public (set k v)
-    (ld-cons% k v this)))
+;; A Fact is (new fact%)
+(define-class fact%
+  (define (fact n)
+    (cond [(zero? n) 1]
+          [else (* n (this . fact (sub1 n)))])))
 \end{verbatim}
 
-The dictionary data structure has some important properties, one of
-which is that if we associate a value with a key and then look up that
-key, we expect to get back the value we associated.  For example, for
-any given dictionary, if we associate \code{1} with \code{"Fred"},
-then look up \code{1}, we should get \code{"Fred"}.  More generally,
-we expect this to be true for \emph{any} key-value pair, not just
-\code{1} and \code{"Fred"}.
+We need to rewrite this method, using the \verb|Natural| interface,
+which has the following methods:
 
-\begin{enumerate}
-\item Codify the general ``set/lookup'' property, described above, as
-  a predicate.  Write a successful test in terms of that predicate.
-\end{enumerate}
+\begin{verbatim}
+;; A Natural implements:
+;; - zero? : -> Boolean
+;; is this Natural the zero value?
+;; - sub1 : -> Natural
+;; produce a Natural one less than this one 
+;; (or zero if this one is zero)
+;; - times : Natural -> Natural
+;; multiply this Natural by the given one
+\end{verbatim}
 
-Now consider a generalization of the ``set/lookup'' property which
-says that if we set \emph{two} keys to \emph{two} values then look
-each of them up, we should get back the original values.  For example, for
-any given dictionary, if we associate \code{1} with \code{"Fred"} and
-\code{2} with \code{"Wilma"}; then looking up \code{1} will produce
-\code{"Fred"} and looking up \code{2} will produce \code{"Wilma\"}.
-More generally, we can state this as a property for any two key-value
-pairs.
+Write a new definition of the \verb|fact%| class using \verb|Natural|,
+and \emph{not} using the built-in numbers of the class system or
+DrRacket.  You can use the constant \verb|one|, which is a \verb|Natural|
+representing $1$.
 
-\begin{enumerate}
-\setcounter{enumi}{1}
-\item Codify the general ``double-set/lookup'' property as a
-  predicate.  Write a successful test in terms of that predicate.
+\ifrubric
+\else
+\fi
 
-\item Does this property hold for all dictionaries and keys-value
-  pairs?  If so, why?  If not, write a counter-example, i.e.~write a
-  failing test in terms of the predicate you gave in part 2.
-\end{enumerate}
+\newpage
 
-\ifrubric\else
-\newpage
-[Here is some more space for the previous problem.]
+Now, design a data defintion that implements \verb|Natural|, and
+supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.  Define the
+constant \verb|one|.
+
+It will be helpful to define the helper methods \verb|add1| and
+\verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
+\cdot m = (n-1) \cdot m + n$.
+
+\ifrubric
+\else
 \newpage
@@
 \begin{verbatim}
-;; Part 1 
-;; Dict Nat String -> Boolean            [4pt]
-(define (set/lookup d k v)
-  (string=? (d . set k v . lookup k) v))
 
-;; [2pt]
-(check-expect (set/lookup (ld-empty%) 1 "Fred") true)
+;; Part 1: N Points
+(define-class fact%
+  (define (fact n)
+    (cond [(n . zero?) one]
+          [else (n . times (this . fact (n . sub1)))])))
 
-;; Part 2
-;; Dict Nat String Nat String -> Boolean   [4pt]
-(define (double-set/lookup d k1 v1 k2 v2)
-  (and (string=? (d . set k1 v1 . set k2 v2 . lookup k1) v1)
-       (string=? (d . set k1 v1 . set k2 v2 . lookup k2) v2)))
+;; Part 2: M points
+;; A Natural is one of
+;; - (new zero%)
+;; - (new succ% Natural)
+(define-class zero%
+  (define (sub1) this)
+  (define (plus n) n)
+  (define (add1) (new succ% this))
+  (define (times n) this)
+  (define (zero?) true))
 
-;; [2pt]
-(check-expect (double-set/lookup (ld-empty%) 1 "Fred" 2 "Wilma") true)
+(define-class succ%
+  (fields pred)
+  (define (zero?) false)
+  (define (add1) (new succ% this))
+  (define (plus n) (this . pred . plus n . add1))
+  (define (times n) (this . pred . times n . plus n))
+  (define (sub1) (this . pred)))
 
-;; Part 3  [3pt]
-;; Counter-example:
-(check-expect (double-set/lookup (ld-empty%) 1 "Fred" 1 "Wilma") true)
+(define one (new succ% (new zero%)))
+(define zero (new zero%))
+
+(check-expect (one . zero?) false)
+(check-expect (one . sub1) zero)
+(check-expect ((new succ% one) . sub1) one)
+(check-expect (zero . zero?) true)
+(check-expect (zero . sub1) zero)
+\end{verbatim}
+\newpage
+\end{solution}
+
+\begin{problem}
+
+\textbf{Extra Credit}
+\\
+\noindent
+After Van Horn discovered the problems with numbers, he attempted to
+fix them.  Unfortunately, not only did he not fix them, he broke
+booleans as well.  Fortunately, we have an idea about how you can
+help, expressed in the following data definitions.
+
+\begin{verbatim}
+;; A Bool implements:
+;; - branch : [Action X] [Action X] -> X
+;; run the first action if this is true, otherwise run the second action
+
+;; An [Action X] implements:
+;; - run : -> X
+;; do this action
+
+\end{verbatim}
+
+Following these interface definitions, design data and class
+defintions for implementations of the \verb|Bool| interface, and
+reimplement the \verb|fact%| class to use them.
+You may modify your natural number implementation if neccessary for
+your solution.
+
+\ifrubric\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+
+\ifrubric\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\end{problem}
+
+\begin{solution}
+\begin{verbatim}
+(define-class fact%
+  (define (fact n)
+    (n . zero? . branch (new const% 0) (new recur% n))))
+(define-class const%
+  (fields run))
+(define-class recur%
+  (fields n)
+  (define (run) ((new fact%) . fact (this . n . sub1) . times n)))
+
+(define-class t%
+  (define (branch t e) (t . run)))
+(define-class f%
+  (define (branch t e) (e . run)))
+
+(define true (new t%))
+(define false (new f%))
 \end{verbatim}
 \end{solution}
-
 \end{document}

### 2012-02-12 exam3/1.tex
Touch ups to nat and bool problem.

@@
 
-\item We expect data \emph{and} interface definitions
-  (although interfaces may be defined in comments rather than
-  with \code{define-interface}).
+\item We expect data \emph{and} interface definitions.
 
@@
 broke the numeric system.
-
+%
 This puts us in the dire position of no longer being able to calculate
@@
 
-This is the definition of a class with a {\tt fact} method:
+This is the definition of a class of functional objects
+with a {\tt fact} method:
 
 \begin{verbatim}
-;; A Fact is (new fact%)
+;; A Fact is (new fact%) and implements 
+;; - fact : Natural -> Natural
+;;   Compute n!.
 (define-class fact%
@@
 
-We need to rewrite this method, using the \verb|Natural| interface,
-which has the following methods:
+Since we can no longer rely on built in numbers to represent
+\verb|Natural|s, we will represent \verb|Natural|s with objects and
+then rewrite the \verb|fact| method to operate over this new kind of
+data.
+
+Here is the \verb|Natural| interface:
 
@@
 ;; - sub1 : -> Natural
-;; produce a Natural one less than this one 
-;; (or zero if this one is zero)
+;;   produce a natural one less than this natural 
+;;   (or zero if this natural is zero)
 ;; - times : Natural -> Natural
-;; multiply this Natural by the given one
+;;   multiply this natural by the given natural
 \end{verbatim}
 
-Write a new definition of the \verb|fact%| class using \verb|Natural|,
+\newpage
+Write a new definition of the \verb|fact| method using \verb|Natural|,
 and \emph{not} using the built-in numbers of the class system or
-DrRacket.  You can use the constant \verb|one|, which is a \verb|Natural|
-representing $1$.
+DrRacket.  You can assume there is a defined constant \verb|one|,
+which is a \verb|Natural| object representing $1$.
 
@@
 Now, design a data defintion that implements \verb|Natural|, and
-supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.  Define the
-constant \verb|one|.
+supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.
+Define the constant \verb|one|.  (\emph{Hint}: remember that the
+natural numbers can be seen as a recursive union; a natural number is
+either zero, or one more than a natural number.)
 
@@
 ;; - branch : [Action X] [Action X] -> X
-;; run the first action if this is true, otherwise run the second action
+;;   run the first action if this bool is true, 
+;;   otherwise run the second action
 
@@
 ;; - run : -> X
-;; do this action
+;; run this action
 
@@
 defintions for implementations of the \verb|Bool| interface, and
-reimplement the \verb|fact%| class to use them.
+reimplement the \verb|fact| method to use them.
 You may modify your natural number implementation if neccessary for
@@
   (define (fact n)
-    (n . zero? . branch (new const% 0) (new recur% n))))
+    (n . zero? . branch (new const% one) (new recur% n))))
 (define-class const%

### 2012-02-12 exam3/1.tex
First crack at problem 1.

@@
 
-A \emph{range} represents a set of numbers between two endpoints.  To
-start with, you only need to consider ranges that \emph{include} the
-smaller endpoint and \emph{exclude} the larger endpoint---such ranges
-are called \emph{half-open}.  For example, the range $[3,4.7)$
-includes all of the numbers between $3$ and $4.7$, including $3$ but
-\emph{not} including $4.7$. So $4$ and $3.0000001$ are both in the
-range, but $5$ is not.  In the notation used here, the $[$ means
-include, and the $)$ means exclude.
+Here is a data definition for a class of objects.  It intentionally
+uses meaningless names to test whether you've sufficiently
+internalized the data-driven development process we've talked about
+since the Fall.
 
-\smallskip
+\begin{verbatim}
+;; A P is one of:          
+;; - A                  ;; An A is a (new a% String)
+;; - B                  ;; A  B is a (new b% String A)
+;; - C                  ;; A  C is a (new c% P)
+
+(define-class a% (fields cookoo))
+(define-class b% (fields gaga bang))
+(define-class c% (fields doodle))
+\end{verbatim}
 
 \noindent
+Design the following methods for \verb|P|s:
+
 \begin{enumerate}
-\item Design a representation for ranges and implement the
-  \texttt{in-range?} method, which determines if a number is in the
-  range.  For example, the range $[3,7.2)$ includes the numbers $3$
-  and $5.0137$, but not the numbers $-17$ or $7.2$.
 
-\item Extend the data definition and implementation of ranges to
-  represent ranges that \emph{exclude} the low end of the range and
-  \emph{include} the high end, written $(lo,hi]$.
+\item \verb|size| - the size of an \verb|A| is $5$; the size of a
+  \verb|B| is $8$, regardless of what's inside; and the size of a
+  \verb|C| contributes $1$ to the size of the \verb|P| inside.
 
-\item Add a \texttt{union} method to the interface for ranges and
-  implement it in all range classes.  This method should consume a
-  range and produces a new range that includes all the numbers in this
-  range \emph{and} all the numbers in the given range.
+\item \verb|replace-gaga| - produces a \verb|P| with the \verb|gaga|
+  field of any \verb|B| within this \verb|P| with a given string.
 
-  You may extend your data definition for ranges to
-  support this method.
+\item \verb|same-gaga-cookoo?| - produces true for any \verb|P| which
+  contains a \verb|B| whose \verb|gaga| is equal to a given string and
+  whose \verb|bang| is an \verb|A| whose \verb|cookoo| is equal to
+  another given string.  Produces false otherwise.
 
-  You may apply the abstraction recipe, but you are not required to do
-  so; you will not be marked down for duplicated code on this problem.
 \end{enumerate}
+
+Be sure to define any interfaces you need so that your program and say
+which classes implement which interfaces so that your program treats
+all values according to their interface.
 
@@
 \begin{verbatim}
-;; Part 1:
 
-;; A Range is one of
-;; - (new range% Number Number)                         [1pt]
-;; Interp: represents the range between `lo' and `hi'
-;;         including `lo', but *not* including `hi'
-;; and implements IRange:
+;; A P implements:
+;; - size : -> Number
+;;   Size of this P.
+;; - replace-gaga : String -> P
+;;   Replace gaga field (if any) of this P with given string.
+;; - same-gaga-cookoo? : String String -> Boolean
+;;   Is there a B within this P with given gaga field and cookoo field?
 
-;; The IRange interface includes:
-;; - in-range? : Number -> Boolean                      [1pt]
-;; determine if the given number is in this range
-;; - union : Range -> Range                             [1pt]
-;; produce the range containing this and the given range
+;; An A implements:
+;; - cookoo : -> String
+;;   Get the cookoo string.
 
-(define-class range%                                    [1pt]
-  (fields lo hi)
-  (check-expect ((range% 0 1) . in-range? 0) true)      [1pt]
-  (check-expect ((range% 0 1) . in-range? 1) false)
-  (define/public (in-range? n)                          [2pt]
-    (and (>= n (field lo)) (< n (field hi))))
+;; In a%:
+(define (size) 5)
+(define (replace-gaga str) this)
+(define (same-gaga-cookoo? g c) false)
 
-  (define/public (union r)
-    (union-range% this r)))
+;; In b%:
+(define (size) 8)
+(define (replace-gaga str)
+  (new b% str (this . bang)))
+(define (same-gaga-cookoo? g c)
+  (and (string=? g (this . gaga))
+       (string=? c (this . bang . cookoo))))
 
-;; Part 2:
+;; In c%:
+(define (size) (add1 (this . doodle . size)))
+(define (replace-gaga str)
+  (new c% (this . doodle . replace-gaga str)))
+(define (same-gaga-cookoo? g c)
+  (this . doodle . same-gaga-cookoo? g c))
 
-;; A Range is also:
-;; - (new hi-range% Number Number)                      [1pt]
-;; Interp: represents the range between `lo' and `hi'
-;;         including `hi', but *not* including `lo'
-
-(define-class hi-range%                                 [1pt]
-  (fields lo hi)
-  (check-expect ((hi-range% 0 1) . in-range? 0) false)  [1pt]
-  (check-expect ((hi-range% 0 1) . in-range? 0.5) true)
-  (check-expect ((hi-range% 0 1) . in-range? 1) true)
-  (define/public (in-range? n)                          [1pt]
-    (and (> n (field lo)) (<= n (field hi))))
-
-  (define/public (union r)
-    (union-range% this r)))
-
-;; Part 3:
-
-;; A Range is also:
-;; - (new union-range% Range Range)                     [1pt]
-;; Interp: including all the numbers in both ranges
-
-(define-class union-range%                              [1pt]
-  (fields left right)
-  (define/public (in-range? n)                          [2pt]
-    (or ((field left) . in-range? n)
-        ((field right) . in-range? n)))
-
-  ;; Point is for all the union methods together
-  (define/public (union r)                              [1pt]
-    (union-range% this r)))
-
-(define r1 (range% 0 1))
-(define r2 (hi-range% 0 1))
-(define r3 (range% 2 4))
-
-;; union + in-range? test                               [1pt]
-(check-expect (r1 . union r2 . union r3 . in-range? 3) 
-              true)
-
-;; testing all the union methods                        [1pt]
-(check-expect (r1 . union r2 . in-range? 1) true)
-(check-expect (r2 . union r1 . in-range? 0) true)
 \end{verbatim}
@@
 \verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
-\cdot m = (n-1) \cdot m + n$.
+\cdot m = (n-1) \cdot m + n$.  Don't overlook the slightly weird
+specification for \verb|sub1|: $0-1 = 0$; this weirdness is common
+amongst mathematicians when working with natural numbers.
 

### 2012-02-12 exam3/1.tex
First draft of problem 2.

@@
 \usepackage{alltt,times,comment,amsmath}
+\usepackage{alltt}
 \usepackage{ct} % tight CM font for \tt
@@
 
-Here are data, class, and interface definitions for Shapes:
+Here are data, class, and interface definitions for a list of numbers:
+\begin{alltt}
+;; An LoN is one of:
+;; - (new mt%)
+;; - (new cons% Number LoN)
+;; and implements:
+;; - sum : -> Number
+;;   Compute the sum of the numbers in this list.
+;; - sqrs : -> LoN
+;;   List of squares of each number in this list.
+
+(define-class mt%
+  (define (sum) 0)
+  (define (sqrs) this))
+
+(define-class cons%
+  (fields first rest)
+  (define (sum)
+    (+ (this . first)
+       (this . rest . sum)))
+  (define (sqrs)
+    (new cons% 
+         (sqr (this . first)) ; (sqr x) = x\(\sp{2}\)
+         (this . rest . sqrs))))
+\end{alltt}
+%
+Here are some examples:
 \begin{verbatim}
-;; A Shape is one of:
-;; - (rect% Number Number)
-;; - (circ% Number)
-;; implements:
-;; bba : -> Number     (short for "bounding-box-area")
-;; Compute the area of the smallest bounding box for this shape.
-
-(define-class rect%
-  (fields width height))
-
-(define-class circ%
-  (fields radius))
-\end{verbatim}
-%
-Here are some examples of how \code{bounding-box-area} should work:
-\begin{verbatim}
-(check-expect ((rect% 3 4) . bba) 12)
-(check-expect ((circ% 1.5) . bba)  9)
+(check-expect ((new mt%) . sum) 0)
+(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum)  9)
+(check-expect ((new mt%) . sqrs) (new mt%))
+(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sqrs)
+              (new cons% 16 (new cons% 25 (new mt%))))
 \end{verbatim}
 
-\begin{enumerate}
-\item Design the \code{bba} method for the
-  \code{rect\%} and \code{circ\%} class.
+\newpage
+\noindent
+Develop the \verb|sum-sqrs| method that computes the sum of the
+squares of a list of numbers.  Here is an example:
+\begin{verbatim}
+(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum-sqrs) 
+              (+ 16 25))
+\end{verbatim}
 
-\item Design a super class of \code{rect\%} and \code{circ\%} and
-  lift the \code{bba} method to the super class.  Extend
-  the shape interface as needed, but implement any methods you add.
-
-\item Design a new variant of a Shape, Square, which should support
-  all of the methods of the interface.
-\end{enumerate}
+If you can, define a super class of \verb|mt%| and \verb|cons%| and
+lift the definition of \verb|sum-sqrs| to this class.  (We will assume
+the appropriate \verb|super| declaration in \verb|mt%| and
+\verb|cons%|.)  If you cannot, define methods in both \verb|mt%| and
+\verb|cons%| for partial credit.
 
@@
 \begin{verbatim}
-;; Part 1 [6pt]
-(define-class rect% 
-  (fields width height)                
-  (define/public (bba)                 [2pt]
-    (* (field width) (field height))))
+(define-class list%
+  (define (sum-sqrs)
+    ((this . sqrs) . sum)))
 
-(define-class circ% 
-  (fields radius)                      
-  (define/public (bba)                 [2pt]
-    (sqr (* 2 (field radius)))))
-
-;; Part 2 [10pt]
-
-;; implements:
-;; width : -> Number                         [1pt]
-;; height : -> Number
-;; Compute the {width,height} of this shape. [1pt]
-
-(define-class shape% 
-  (define/public (bounding-box-area)
-    (* (this . width) (this . height))))     [3pt]
-
-(define-class rect% 
-  (super shape%)                             [1pt]
-  (fields width height))                     [1pt]
-
-(define-class circ% 
-  (super shape%)                             [1pt]
-  (fields radius)
-  (define/public (width)                     [1pt]
-    (* 2 (field radius)))
-  (define/public (height)                    [1pt]
-    (width)))
-
-;; Part 3 [4pt]
-
-(define-class square%
-  (super shape%)                             [1pt]
-  (fields width)                             [1pt]
-  (define/public (height)                    [1pt]
-    (width)))
-
-(check-expect ((square% 5) . bba) 25)        [1pt]
+(check-expect ((new mt%) . sum-sqrs) 0)
 \end{verbatim}
@@
 \ifrubric\else
-[Here is some more space for the previous problem.]
-\newpage
 [Here is some more space for the previous problem.]

### 2012-02-12 exam3/1.tex
add set problem

@@
 
-The Northeastern University Registrar is instituting a new course
-registration system, in which each student will wait in a ``Virtual
-Line'' until every student ahead of them has registered.  A simple way
-to represent a line (also known as a \emph{queue}) is by using a list.
-But this representation makes it slow to add somebody to the end of
-the line (or to take somebody off the front of the line, depending on
-whether the front of the list represents the front or rear of the
-line).
+In order to manage class scheduling, the College decided that they
+needed to create a new implementation of the \verb|Set| data
+structure.  They went with the lowest bidder, which
+turned out to be a bunch of BU students.  Now, having paid for this
+code, the College needs to test it, and they turn to you.
 
-In order to provide maximal waiting efficiency, you have been tasked
-with implementing a representation that uses \emph{two} lists!  The
-key idea of this fancy representation is that one list will represent
-some portion of the front of the line, while the other will represent
-the remainder of the line \emph{in reverse order}.  So if you're the
-first element of the first list, you are at the head of the line.  On
-the other hand, if you're the first element of the second list, you
-are the very last person in line.
-%
-Here is the interface for queues:
+Unfortunately, code the BU students produced is so convoluted that it
+doesn't bear looking at.  All you have to go on is that the \verb|Set|
+data structure obeys the following interface:
+
 \begin{verbatim}
-;; A [IQ X] implements:
-;;
-;; head : -> X
-;; Produce the element at the head of this queue.
-;; Assume: this queue is not empty.
-;;
-;; deq : -> [IQ X]      (Short for "dequeue")
-;; Produces a new queue like this queue, but without 
-;; this queue's first element.
-;; Assume: this queue is not empty.
-;;
-;; enq : X -> [IQ X]    (Short for "enqueue")
-;; Produce new queue with given element added to the
-;; END of this queue.
-;;
-;; emp? : -> Boolean
-;; Is this queue empty?
+;; empty-set is a set
+
+;; A Set implements:
+;; - insert : String -> Set
+;; adds the given string to the set, if it isn't already there
+;; - remove : String -> Set
+;; removes the given string from the set
+;; if the given string is not in the set, produce the original set
+;; - size : -> Number
+;; count the elements of the set
+;; - has : String -> Boolean
+;; is the given String in the set?
 \end{verbatim}
-
-The \code{head} and \code{deq} operations require that the queue be
-non-empty when they are used, but this can be assumed and these
-operations do not need to check for an empty queue.
-
-Further, the Registrar's office has just learned about
-\emph{invariants}, and insists on maintaining the following invariant
-about all of their queues:
+\noindent
+\verb|Set|s should satisfy the following property, called the
+``insert-size'' property, for any element and
+set:
 
 \begin{quote}
-  {\it if the front of the queue is empty, the whole queue must also be empty.}
+If we insert an element into a set, then the resulting set is the
+same size as the original set \emph{if} the element was already in the
+set, and the resulting set is one larger than the original set
+if the element was \emph{not} already in the set.
+\end{quote}
+
+\begin{enumerate}
+\item Codify the ``insert-size'' property as a predicate.  Write a
+  test case that should pass (if the sets work right) using the
+  predicate you have defined.
+\end{enumerate}
+
+\noindent
+The College wants to be really sure that their sets are correct, and
+they've come up with another property that they think is true, called
+the ``insert-remove'' property:
+
+\begin{quote}
+If we insert element $a$ into a set, and then remove element $b$ from
+the resulting set, the result has the same size as if we do the two
+operations in the reverse order.
 \end{quote}
@@
 \noindent
-The Registrar's office has given you three tasks to prepare their
-Virtual Line for its launch later this semester:
+The College is considering adding this to the test suite as well.
 
 \begin{enumerate}
-\item Design an implementation of the queue data structure to the Registrar's
-  specifications.  You must maintain the invariant stated above, and
-  you should take advantage of the invariant when implementing the
-  operations.
-\item Unfortunately, when testing the queue, the Registrar has
-  discovered that some queues with the same elements in the same order
-  can be represented in multiple ways.  Give an example of two
-  different representations of the same queue.  Implement a
-  \code{to-list} operation which produces a list of elements going in
-  order from the front to the rear of the queue. In your tests, you
-  should show how this addresses the problem.
-\item The Registrar has a problem with careless data entry.  Design
-  and implement a constructor for queues which, given two input lists
-  of elements, ensures that the invariant is maintained.  
+\setcounter{enumi}{1}
+
+\item Codify this property as a predicate, and write a test case that
+  should pass using the predicate.
+
+\item Is the ``insert-remove'' property true for all sets?  If so,
+  give an explanation.  If not, write a counter-example (i.e., a
+  failing test case) using your predicate.
+
 \end{enumerate}
@@
 \begin{verbatim}
-;; Part 1:
-;; invariant: front is empty IFF queue is empty
-;; A [Queue X] is (queue% [Listof X] [Listof X])           [1pt]
-;; and implements [IQ X]
+;; pred1 : Set String -> Boolean
+;; check the "insert-size" property
+(define (pred1 s i)
+  (or (and (s . has i) (= (s . insert i . size) (s . size)))
+      (and (not (s . has i))
+           (= (s . insert i . size) (add1 (s . size))))))
 
-(define-class queue%                                       [1pt]
-  (fields front back)
-  ;; Add element to the back
-  ;; X -> [Queue X]
-  ;; add it to the front IFF the front is empty
-  ;; (otherwise invariant is violated)
-  (check-expect (eq . enqueue 1) (queue% '(1) '()))        [1pt]
-  (check-expect (q . enqueue 5) (queue% '(1 2 3) '(5)))
-  (define/public (enqueue i)
-    (cond [(empty? (field front))                          [2pt]
-           (queue% (list i) empty)]
-          [else
-           (queue% (field front) (cons i (field back)))])) [2pt]
-  ;; get the first element
-  ;;  -> X
-  ;; assumes this queue is non-empty
-  ;; uses the invariant
-  (check-expect (q . head) 1)                              [1pt]
-  (define/public (head)                                    
-    (first (field front)))
-  ;; drop the first element
-  ;;  -> [Queue X]
-  ;; assumes this queue is non-empty
-  (check-expect (eq . enqueue 1 . dequeue) eq)             [1pt]
-  (check-expect (q . enqueue 5 . dequeue) (queue% '(2 3) '(5)))
-  (define/public (dequeue)
-    (cond 
-      ;; maintain the invariant -- don't let `front' be empty
-      [(empty? (rest (field front)))                       [3pt]
-       (queue% (reverse (field back)) empty)]
-      ;; otherwise, simple
-      [else (queue% (rest (field front)) (field back))]))  [2pt]
-  ;; is this queue empty?
-  ;; emp? : -> Boolean
-  ;; uses the invariant
-  (check-expect (q . emp?) false)                          [1pt]
-  (check-expect (eq . emp?) true)
-  (define/public (emp?)
-    (empty? (field front)))
+(check-expect (pred1 empty-set "hi") true)
 
-;; Part 2:
+;; pred2 : Set String -> Boolean
+;; check the "remove-insert" property
+(define (pred2 s a b)
+  (= (s . insert a . remove b . size)
+     (s . remove b . insert a . size))
 
-;; Examples:                                               [1pt]
-(queue% (list 1) (list 6)) vs (queue% (list 1 6) empty)
-(queue% (list 1 8 12) (list 6)) vs (queue% (list 1 8) (list 12 6))
+(check-expect (pred2 empty-set "yes" "no"))
 
-  ;; convert the queue to a list                           [1pt]
-  ;; -> [Listof X] 
-  ;; turn examples above into tests:                       [1pt]
-  (check-expect ((queue% (list 1) (list 6)) . to-list)
-		((queue% (list 1 6) empty) . to-list))
-  (check-expect ((queue% (list 1 8 12) (list 6)) . to-list)
-		((queue% (list 1 8) (list 12 6)) . to-list))
-  (check-expect (q . to-list) '(1 2 3))                    
-  (check-expect (q . enqueue 5 . to-list) '(1 2 3 5))   
-  (check-expect (q . enqueue 5 . dequeue . to-list) '(2 3 5))
-  (define/public (to-list)                                 [1pt]
-    (append (field front)
-            (reverse (field back)))))
-
-(define q (queue% '(1 2 3) empty))
-(define eq (queue% empty empty))
-
-;; Part 3:                                                 [2pt]
-;; several good possibilities (others are possible)
-(constructor (f b)
-  (fields (append f (reverse b)) empty))
-
-(constructor (f b)
-  (cond [(empty? f) (fields (reverse b) empty)]
-	[else (fields f b)]))
-
-(constructor (f b)
-  (cond [(and (empty? f) (empty? b)) (fields f b)]
-	[(empty? f) (fields (reverse b) empty)]
-	[else (fields f b)]))
-
+;; counter-example
+(check-expect (pred2 empty-set "yes" "yes"))
 \end{verbatim}
@@
 reimplement the \verb|fact| method to use them.
-You may modify your natural number implementation if neccessary for
-your solution.
+You will need to provide new defintions for \verb|true| and
+\verb|false|, since they were broken by this catastrophe as well.
 
@@
 \fi
-\end{problem}
 
@@
   (fields n)
-  (define (run) ((new fact%) . fact (this . n . sub1) . times n)))
+  (define (run) ((new fact%) . fact (this . n . sub1) . times (this . n))))
 
@@
 \end{solution}
+\end{problem}
 \end{document}

### 2012-02-12 exam3/1.tex
Touch ups to set problem.

@@
 \begin{verbatim}
-;; empty-set is a set
+;; empty-set is a Set value containing no elements
 
@@
 ;; - insert : String -> Set
-;; adds the given string to the set, if it isn't already there
+;;   adds given string to this set, if it isn't already there
+;;   (otherwise does nothing)
 ;; - remove : String -> Set
-;; removes the given string from the set
-;; if the given string is not in the set, produce the original set
+;;   removes given string from this set, if it's already there
+;;   (otherwise does nothing)
 ;; - size : -> Number
-;; count the elements of the set
-;; - has : String -> Boolean
-;; is the given String in the set?
+;;   count the elements of this set
+;; - has? : String -> Boolean
+;;   is the given String in this set?
 \end{verbatim}
@@
 \begin{quote}
-If we insert an element into a set, then the resulting set is the
+If we insert any element into any set, then the resulting set is the
 same size as the original set \emph{if} the element was already in the
@@
 
+\newpage
+\ifrubric\else
+[Here is some more space for the previous problem.]
+
+\newpage
 \noindent
@@
 \begin{quote}
-If we insert element $a$ into a set, and then remove element $b$ from
-the resulting set, the result has the same size as if we do the two
-operations in the reverse order.
+If we insert element any element $a$ into any set, and then remove any
+element $b$ from the resulting set, the result has the same size as if
+we do the two operations in the reverse order, i.e., remove $b$, then
+insert $a$.
 \end{quote}
@@
 
-\newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
 \newpage

### 2012-02-12 exam3/1.tex
improved with points

@@
 
-\item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} for
-  {\tt (check-expect {\slshape c e})} and $\lambda$ for
-  \code{lambda} to save time writing.
+\item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} as
+  shorthand for
+  {\tt (check-expect {\slshape c e})}.
 
@@
 
+\item The extra credit problem is \emph{all or nothing};  no partial
+  credit will be awarded.
+
+\item Unless specifically requested, templates and super classes are
+  \emph{not} required.
+
 \item Some basic test taking advice: Before you start answering
-any problems, read \emph{every} problem, so your brain can be thinking
+any problems, read \emph{every} problem, so your brain can  think
 about the harder problems in background while you knock off the easy ones.
@@
 {\bf Problem} & Points & out of \\ \hline
-1 & & ?\\ \hline
-2 & & ?\\ \hline
-3 & & ?\\ \hline
-4 & & ?\\ \hline
-5 & & ?\\ \hline
-Extra & & ? \\ \hline
-{\bf Total} & & 83 \\ \hline
+1 & & 18\\ \hline
+2 & & 10\\ \hline
+3 & & 15\\ \hline
+4 & & 18\\ \hline
+%5 & & 5\\ \hline
+Extra & & 5 \\ \hline
+{\bf Total} & & 61+5 \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 % \newpage
-\begin{problem}\pts{17} % same as on the previous exam :)
+\begin{problem}\pts{18} % same as on the previous exam :)
 
@@
 \begin{verbatim}
-
+;;> [6 pts -- contract and purpose for each method]
 ;; A P implements:
@@
 
+;;> [0 pts -- not required, but write something]
 ;; An A implements:
@@
 
-;; In a%:
+;;> [2 pts] In a%: 
 (define (size) 5)
@@
 
-;; In b%:
+;;> [3 pts] In b%:
 (define (size) 8)
@@
 
-;; In c%:
+;;> [4 pts] In c%:
 (define (size) (add1 (this . doodle . size)))
@@
   (this . doodle . same-gaga-cookoo? g c))
+
+;; tests
+;;> [3 pts -- 1 per method]
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem}\pts{18}
+\begin{problem}\pts{10}
 
@@
 \begin{verbatim}
+
+;; [2 pts% -- class def]
 (define-class list%
+  ;; [2 pts -- contract and purpose]
+  ;; sum-sqrs : -> Number
+  ;; compute the sum of the squares of the numbers in this list
+  ;; [4 pts -- implementation]
   (define (sum-sqrs)
@@
 
+;; [2 pts -- tests]
 (check-expect ((new mt%) . sum-sqrs) 0)
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem}\pts{12}
-
-One perspective that unifies the paradigms of programming with
-functions and programming with objects is to view a ``function'' as an
-object that understands a method called \code{apply}.  With that in
-mind, we can define an interface for functions-as-objects:
-
-\begin{verbatim}
-;; A [IFun X Y] implements:
-;; apply : X -> Y
-;; Apply this function to the given input.
-\end{verbatim}
-
-Here we are representing a \code{X -> Y} function as an object that
-has an \code{apply} method that consumes an \code{X} and produces a
-\code{Y}.
-
-\begin{enumerate}
-\item Design a class that wraps a real function with contract \code{(X
-  -> Y)} to implement \code{[IFun X Y]}.
-
-\item Using your wrapper class, construct the objects representing the
-  ``add 1'' function and ``sub 1'' function.
-
-\item Another useful operation on functions is composition.  Here is
-  the interface for a \code{compose} method that composes two functions
-  represented as objects:
-
-\begin{verbatim}
-;; [IFun X Y] also implements:
-;; compose : [IFun Y Z] -> [IFun X Z]
-;; Produce a function that applies this function to its input, 
-;; then applies the given function to that result.
-\end{verbatim}
-
-For example, if \code{addone} and \code{subone} refer to the objects
-you constructed in part 2, the following check should succeed:
-
-\begin{verbatim}
-(check-expect ((addone . compose subone) . apply 5) 5)
-(check-expect ((addone . compose addone) . apply 5) 7)
-\end{verbatim}
-
-Implement the \code{compose} method for your wrapper class.
-\end{enumerate}
-
-\begin{solution}
-
-\begin{verbatim}
-;; Part 1 [5pt]
-(define-class fun%
-  (fields f)
-  (define/public (apply x)
-    ((field f) x)))
-
-;; Part 2 [2pt]
-(define addone (fun% add1))
-(define subone (fun% sub1))
-
-;; Part 3 [5pt]
-(define-class fun%
-  (fields f)
-  (define/public (compose g)
-    (fun% (lambda (x)
-            (g . apply (apply x))))))
-\end{verbatim}
-
-\end{solution}
-
-\newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
-\newpage
-[Here is some more space for the previous problem.]
-\newpage
-\fi
-
-\end{problem}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{21}
+\begin{problem} \pts{15}
 
@@
 ;;   adds given string to this set, if it isn't already there
-;;   (otherwise does nothing)
+;;   (otherwise produces this set)
 ;; - remove : String -> Set
 ;;   removes given string from this set, if it's already there
-;;   (otherwise does nothing)
+;;   (otherwise produces this set)
 ;; - size : -> Number
@@
 
+\ifrubric\else
 \newpage
-\ifrubric\else
 [Here is some more space for the previous problem.]
+\fi
 
@@
 \begin{verbatim}
+;; [1 pt -- contract & purpose]
 ;; pred1 : Set String -> Boolean
 ;; check the "insert-size" property
+;; [4 pts -- code]
 (define (pred1 s i)
@@
 
+;; [2 pts -- test case]
 (check-expect (pred1 empty-set "hi") true)
 
+;; [1 pt -- contract & purpose]
 ;; pred2 : Set String -> Boolean
 ;; check the "remove-insert" property
+;; [3 pts -- code]
 (define (pred2 s a b)
@@
 
-(check-expect (pred2 empty-set "yes" "no"))
+;; [1 pt -- successful test]
+(check-expect (pred2 empty-set "yes" "no") true)
 
 ;; counter-example
-(check-expect (pred2 empty-set "yes" "yes"))
+;; [3 pts -- failing test]
+(check-expect (pred2 empty-set "yes" "yes") true) ;; either true or false
 \end{verbatim}
 \end{solution}
-\end{problem}
 
 \newpage
+\ifrubric\else
 [Here is some more space for the previous problem.]
@@
 \fi
+\end{problem}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{15}
+\begin{problem} \pts{18}
 Last night, Asumu was up late improving the {\tt class} languages.
@@
           [else (* n (this . fact (sub1 n)))])))
+
+(check-expect ((new fact%) . fact 0) 1)
+(check-expect ((new fact%) . fact 5) 120)
 \end{verbatim}
@@
 
-\ifrubric
-\else
-\fi
+\begin{verbatim}
+(check-expect ((new fact%) . fact one) one)
+(check-expect ((new fact%) . fact (one . sub1)) one)
+\end{verbatim}
 
@@
 \begin{verbatim}
-
-;; Part 1: N Points
+;; Part 1: [2 pts]
 (define-class fact%
@@
 
-;; Part 2: M points
+;; Part 2:
+;; [2 pts -- data defintion]
 ;; A Natural is one of
@@
 (define-class zero%
+  ;; [4 pts -- methods in zero%]
   (define (sub1) this)
@@
   (fields pred)
+  ;; [7 pts -- methods in succ%]
   (define (zero?) false)
@@
 
+(define zero (new zero%))
+;; [1 pt -- define one]
 (define one (new succ% (new zero%)))
-(define zero (new zero%))
 
+;; [3 pts -- tests]
 (check-expect (one . zero?) false)
@@
 
-\begin{problem}
+\begin{problem}\pts{5}
 
@@
 booleans as well.  Fortunately, we have an idea about how you can
-help, expressed in the following data definitions.
+help, expressed in the following interface definitions.
 
@@
 \begin{verbatim}
+;; [5 pts -- get it right]
 (define-class fact%
@@
 
-(define true (new t%))
-(define false (new f%))
+(define true (new t%)) (define false (new f%))
 \end{verbatim}

### 2012-02-12 exam3/1.tex
Small fixes.

@@
 any problems, read \emph{every} problem, so your brain can  think
-about the harder problems in background while you knock off the easy ones.
+about the harder problems in the background while you knock off the easy ones.
 \end{itemize}
@@
 
-\item \verb|replace-gaga| - produces a \verb|P| with the \verb|gaga|
-  field of any \verb|B| within this \verb|P| with a given string.
+\item \verb|replace-gaga| - produces a \verb|P| like this one, but
+  with any \verb|gaga| field replaced by the given string.
 
@@
 
-Be sure to define any interfaces you need so that your program and say
-which classes implement which interfaces so that your program treats
-all values according to their interface.
+Be sure to define any interfaces you need and say which classes
+implement which interfaces so that your program treats all values
+according to their interface.
 

### 2012-02-13 exam3/1.tex
Small tweaks from taking exam.

@@
 
-(define-class a% (fields cookoo))
-(define-class b% (fields gaga bang))
-(define-class c% (fields doodle))
+(define-class a% (fields w))
+(define-class b% (fields x y))
+(define-class c% (fields z))
 \end{verbatim}
@@
 
-\item \verb|replace-gaga| - produces a \verb|P| like this one, but
-  with any \verb|gaga| field replaced by the given string.
+\item \verb|replace-x| - produces a \verb|P| like this one, but
+  with any \verb|x| field replaced by the given string.
 
-\item \verb|same-gaga-cookoo?| - produces true for any \verb|P| which
-  contains a \verb|B| whose \verb|gaga| is equal to a given string and
-  whose \verb|bang| is an \verb|A| whose \verb|cookoo| is equal to
-  another given string.  Produces false otherwise.
+\item \verb|same-x-and-w?| - given two strings, produce true if this
+  \verb|P| contains a \verb|B| whose \verb|x| is equal to the
+  first given string and whose \verb|y| is an \verb|A| whose
+  \verb|w| is equal to the second given string.  Produces false
+  otherwise.
 
@@
 ;;   Size of this P.
-;; - replace-gaga : String -> P
-;;   Replace gaga field (if any) of this P with given string.
-;; - same-gaga-cookoo? : String String -> Boolean
-;;   Is there a B within this P with given gaga field and cookoo field?
+;; - replace-x : String -> P
+;;   Replace x field (if any) of this P with given string.
+;; - same-x-and-w? : String String -> Boolean
+;;   Is there a B within this P with given x field and w field?
 
@@
 ;; An A implements:
-;; - cookoo : -> String
-;;   Get the cookoo string.
+;; - w : -> String
+;;   Get the w string.
 
@@
 (define (size) 5)
-(define (replace-gaga str) this)
-(define (same-gaga-cookoo? g c) false)
+(define (replace-x str) this)
+(define (same-x-and-w? g c) false)
 
@@
 (define (size) 8)
-(define (replace-gaga str)
-  (new b% str (this . bang)))
-(define (same-gaga-cookoo? g c)
-  (and (string=? g (this . gaga))
-       (string=? c (this . bang . cookoo))))
+(define (replace-x str)
+  (new b% str (this . y)))
+(define (same-x-and-w? g c)
+  (and (string=? g (this . x))
+       (string=? c (this . y . w))))
 
 ;;> [4 pts] In c%:
-(define (size) (add1 (this . doodle . size)))
-(define (replace-gaga str)
-  (new c% (this . doodle . replace-gaga str)))
-(define (same-gaga-cookoo? g c)
-  (this . doodle . same-gaga-cookoo? g c))
+(define (size) (add1 (this . z . size)))
+(define (replace-x str)
+  (new c% (this . z . replace-x str)))
+(define (same-x-and-w? g c)
+  (this . z . same-x-and-w? g c))
 
@@
 \newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
-\newpage
-\fi
 
@@
 
-Unfortunately, code the BU students produced is so convoluted that it
-doesn't bear looking at.  All you have to go on is that the \verb|Set|
-data structure obeys the following interface:
+Unfortunately, the code the BU students produced is so convoluted that
+it doesn't bear looking at.  All you have to go on is that the
+\verb|Set| data structure obeys the following interface:
 
 \begin{verbatim}
-;; empty-set is a Set value containing no elements
-
 ;; A Set implements:
@@
 ;;   is the given String in this set?
+
+;; empty-set is a Set value containing no elements
 \end{verbatim}
@@
 (define (pred1 s i)
-  (or (and (s . has i) (= (s . insert i . size) (s . size)))
-      (and (not (s . has i))
-           (= (s . insert i . size) (add1 (s . size))))))
+  (= (s . insert i . size)
+     (cond [(s . has? i) (add1 (s . size))]
+           [else               (s . size)])))
 
@@
 \newpage
-Write a new definition of the \verb|fact| method using \verb|Natural|,
-and \emph{not} using the built-in numbers of the class system or
-DrRacket.  You can assume there is a defined constant \verb|one|,
-which is a \verb|Natural| object representing $1$.
+Write a new definition of the \verb|fact| method using the
+\verb|Natural| interface definition given above, and \emph{not} using
+the built-in numbers of the class system or DrRacket.  You can assume
+there is a defined constant \verb|one|, which is a \verb|Natural|
+object representing $1$.
 
@@
 
-\ifrubric\else
-\newpage
-[Here is some more space for the previous problem.]
-\fi
-
 \begin{solution}

### 2012-02-13 exam3/1.tex
Add dir problem.

@@
 3 & & 15\\ \hline
-4 & & 18\\ \hline
-%5 & & 5\\ \hline
+4 & & 15\\ \hline
+5 & & 18\\ \hline
 Extra & & 5 \\ \hline
-{\bf Total} & & 61+5 \\ \hline
+{\bf Total} & & 76+5 \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{problem} \pts{15}
+
+Here are data and class definitions for representing the files and
+directories of a computer:
+
+\begin{verbatim}
+;; An Elem is one of
+;; - (new file% String Number)
+;; - (new dir% String LoElem)
+(define-class file% (fields name size))
+(define-class dir% (fields name elems))
+
+;; A LoElem is one of
+;; - (new mt%)
+;; - (new cons% Elem LoElem)
+(define-class mt%)
+(define-class cons% (fields first rest))    
+\end{verbatim}
+
+\noindent
+Here is an example of a directory tree and its representation as an
+\verb|Elem|:
+\begin{verbatim}
+#|
+ User
+   +-- A
+   |   +-- a.txt
+   |   +-- B
+   |       +-- b.rkt
+   +-- C
+       +-- c.c
+|#
+
+(define mt (new mt%))
+(define a (new file% "a.txt" 5))
+(define b (new file% "b.rkt" 10))
+(define c (new file% "c.c" 1000))
+
+(define C/ (new dir% "C" (new cons% c mt)))
+(define B/ (new dir% "B" (new cons% b mt)))
+(define A/ (new dir% "A" (new cons% a (new cons% B/ mt))))
+(define User/ (new dir% "/" (new cons% A/ (new cons% C/ mt))))
+\end{verbatim}
+
+\newpage
+\noindent
+Your job is to implement one of the main tasks of the common Unix
+utility \verb|find|, which gives a list of all the files within an
+element.  To do so, add a \verb|list-files| method of \verb|Elem|
+which produces a list of files.  So for example, we expect:
+\begin{verbatim}
+(check-expect (User/ . list-files) 
+              (new cons% a (new cons% b (new cons% c mt))))
+\end{verbatim}
+\emph{Hint}: it may come in handy to be able to concatenate two
+\verb|LoElem| together.
+
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+
+\begin{solution}
+\begin{verbatim}
+;;> [1pts -- interfaces, contracts, purpose]
+;; An Elem implements
+;; - list-files : -> LoElem
+;;   List all files within this element.
+
+;;> [2pts -- interfaces, contracts, purpose]
+;; A LoElem implements
+;; - list-files : -> LoElem
+;;   List all files within this list of elements.
+;; - append : LoElem -> LoElem
+;;   Append this list of elements to given list.
+
+;;> [1pt -- code]
+;; in dir%
+(define (list-files)
+  (this . elems . list-files))
+
+;;> [1pt -- code]
+;; in file%
+(define (list-files)
+  (new cons% this (new mt%)))
+
+;;> [2pt -- code]
+;; in mt%
+(define (append loe) loe)
+(define (list-files) this)
+
+;;> [4pt -- code]
+;; in cons%
+(define (append loe)
+  (new cons% (this . first) (this . rest . append loe)))
+(define (list-files)
+  (this . first . list-files . append (this . rest . list-files)))
+
+;;> [1pt -- sanity]
+
+;;> [3pts -- tests]
+;; tests
+(check-expect (mt . append (new cons% a mt)) (new cons% a mt))
+(check-expect ((new cons% a mt) . append (new cons% b mt))
+              (new cons% a (new cons% b mt)))
+
+(check-expect (mt . list-files) mt)
+(check-expect ((new cons% a (new cons% b mt)) . list-files)
+              (new cons% a (new cons% b mt)))
+
+(check-expect (a . list-files) (new cons% a mt))
+\end{verbatim}
+\end{solution}
+
+\end{problem}
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{problem} \pts{18}

### 2012-03-25 exam4/2.tex
revise cover page

--- 2012-02-13 exam3/1.tex
@@
 \usepackage{ct} % tight CM font for \tt
+\usepackage[usenames,dvipsnames]{xcolor}
 \newexercise{exercise}{Exercise}[subsection]
@@
 \ifrubric
-\newenvironment{solution}{}{}
+\newenvironment{solution}{\color{Red}}{}
 \else
@@
 \vspace*{-1.5cm}
-\ifrubric
-\centerline{\Large CSU2510H Exam 1 Rubric -- Spring 2012}
-
-\else
-\centerline{\Large CSU2510H Exam 1 -- Spring 2012}
+\centerline{\Large CSU2510H Exam 1 \ifrubric (SOLUTION) \fi-- Spring 2012}
 
@@
 
+\ifrubric\relax\else
 \begin{center}
@@
 \end{center}
+\fi
 
@@
 \item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} as
-  shorthand for
-  {\tt (check-expect {\slshape c e})}.
+  shorthand for writing both {\tt check-expect} and {\tt
+    Tester.checkExpect}.
 
@@
 
-\item The extra credit problem is \emph{all or nothing};  no partial
-  credit will be awarded.
+% \item The extra credit problem is \emph{all or nothing};  no partial
+%   credit will be awarded.
 
@@
 {\bf Problem} & Points & out of \\ \hline
-1 & & 18\\ \hline
-2 & & 10\\ \hline
-3 & & 15\\ \hline
-4 & & 15\\ \hline
-5 & & 18\\ \hline
-Extra & & 5 \\ \hline
-{\bf Total} & & 76+5 \\ \hline
+1 & & ??\\ \hline
+2 & & ??\\ \hline
+3 & & ??\\ \hline
+4 & & ??\\ \hline
+5 & & ??\\ \hline
+%% Extra & & 5 \\ \hline
+{\bf Total} & & ?? \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 \newpage
-
-\fi
 
@@
 \begin{solution}
-\begin{verbatim}
-;;> [6 pts -- contract and purpose for each method]
+\begin{alltt}
+\rubricpts{6 pts -- contract and purpose for each method}
 ;; A P implements:
@@
 
-;;> [0 pts -- not required, but write something]
+\rubricpts{0 pts -- not required, but write something}
 ;; An A implements:
@@
 
-;;> [2 pts] In a%: 
+\rubricpts{2 pts}
+;; In a%: 
 (define (size) 5)
@@
 
-;;> [3 pts] In b%:
+\rubricpts{3 pts}
+;; In b%:
 (define (size) 8)
@@
 
-;;> [4 pts] In c%:
+\rubricpts{4 pts}
+;; In c%:
 (define (size) (add1 (this . z . size)))
@@
 ;; tests
-;;> [3 pts -- 1 per method]
-
-\end{verbatim}
+\rubricpts{3 pts -- 1 per method}
+(define a (new a% "a"))
+(define b (new b% "b" a))
+(define c (new c% b))
+(define d (new c% a))
+(check-expect (a . size) 5)
+(check-expect (b . size) 8)
+(check-expect (c . size) 6)
+(check-expect (d . size) 9)
+(check-expect (a . replace-x "*") a)
+(check-expect (b . replace-x "*") (new b% "*" a))
+(check-expect (c . replace-x "*") (new c% (new b% "*" a)))
+(check-expect (d . replace-x "*") d)
+(check-expect (a . same-x-and-w? "b" "a") false)
+(check-expect (b . same-x-and-w? "b" "a") true)
+(check-expect (c . same-x-and-w? "b" "a") true)
+(check-expect (d . same-x-and-w? "b" "a") false)
+\end{alltt}
 \end{solution}
@@
 \begin{solution}
-\begin{verbatim}
-
-;; [2 pts% -- class def]
+\begin{alltt}
+\rubricpts{2 pts% -- class def}
 (define-class list%
-  ;; [2 pts -- contract and purpose]
+  \rubricpts{2 pts -- contract and purpose}
   ;; sum-sqrs : -> Number
   ;; compute the sum of the squares of the numbers in this list
-  ;; [4 pts -- implementation]
+  \rubricpts{4 pts -- implementation}
   (define (sum-sqrs)
@@
 
-;; [2 pts -- tests]
+\rubricpts{2 pts -- tests}
 (check-expect ((new mt%) . sum-sqrs) 0)
-\end{verbatim}
+\end{alltt}
 \end{solution}
@@
 \begin{solution}
-\begin{verbatim}
-;; [1 pt -- contract & purpose]
-;; pred1 : Set String -> Boolean
-;; check the "insert-size" property
-;; [4 pts -- code]
-(define (pred1 s i)
+\begin{alltt}
+\rubricpts{1 pt -- contract & purpose}
+;; check-insert-size : Set String -> Boolean
+;; check the "insert-size" property on s and i.
+\rubricpts{4 pts -- code}
+(define (check-insert-size s i)
   (= (s . insert i . size)
@@
 
-;; [2 pts -- test case]
-(check-expect (pred1 empty-set "hi") true)
+\rubricpts{2 pts -- test case}
+(check-expect (check-insert-size empty-set "hi") true)
 
-;; [1 pt -- contract & purpose]
-;; pred2 : Set String -> Boolean
-;; check the "remove-insert" property
-;; [3 pts -- code]
-(define (pred2 s a b)
+\rubricpts{1 pt -- contract & purpose}
+;; check-remove-insert : Set String -> Boolean
+;; check the "remove-insert" property on s, a, and b.
+\rubricpts{3 pts -- code}
+(define (check-remove-insert s a b)
   (= (s . insert a . remove b . size)
@@
 
-;; [1 pt -- successful test]
-(check-expect (pred2 empty-set "yes" "no") true)
+\rubricpts{1 pt -- successful test}
+(check-expect (check-remove-insert empty-set "yes" "no") true)
 
-;; counter-example
-;; [3 pts -- failing test]
-(check-expect (pred2 empty-set "yes" "yes") true) ;; either true or false
-\end{verbatim}
+;; Property does not hold.
+;; Counter-example (failing test):
+\rubricpts{3 pts -- failing test}
+(check-expect (check-remove-insert empty-set "yes" "yes") true)
+\end{alltt}
 \end{solution}
@@
 \begin{solution}
-\begin{verbatim}
-;;> [1pts -- interfaces, contracts, purpose]
+\begin{alltt}
+\rubricpts{1pts -- interfaces, contracts, purpose}
 ;; An Elem implements
@@
 
-;;> [2pts -- interfaces, contracts, purpose]
+\rubricpts{2pts -- interfaces, contracts, purpose}
 ;; A LoElem implements
@@
 
-;;> [1pt -- code]
+\rubricpts{1pt -- code}
 ;; in dir%
@@
 
-;;> [1pt -- code]
+\rubricpts{1pt -- code}
 ;; in file%
@@
 
-;;> [2pt -- code]
+\rubricpts{2pt -- code}
 ;; in mt%
@@
 
-;;> [4pt -- code]
+\rubricpts{4pt -- code}
 ;; in cons%
@@
 
-;;> [1pt -- sanity]
+\rubricpts{1pt -- sanity}
 
-;;> [3pts -- tests]
+\rubricpts{3pts -- tests}
 ;; tests
@@
 (check-expect (a . list-files) (new cons% a mt))
-\end{verbatim}
+\end{alltt}
 \end{solution}
@@
 
+\begin{solution}
+\begin{alltt}
+\rubricpts{2 pts}
+(define-class fact%
+  (define (fact n)
+    (cond [(n . zero?) one]
+          [else (n . times (this . fact (n . sub1)))])))
+
+\end{alltt}
+\end{solution}
+
 \newpage
 
-Now, design a data defintion that implements \verb|Natural|, and
+Now, design data and class defintions that implement \verb|Natural|, and
 supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.
@@
 \verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
-\cdot m = (n-1) \cdot m + n$.  Don't overlook the slightly weird
+\cdot m = (n-1) \cdot m + m$.  Don't overlook the slightly weird
 specification for \verb|sub1|: $0-1 = 0$; this weirdness is common
@@
 \begin{solution}
-\begin{verbatim}
-;; Part 1: [2 pts]
-(define-class fact%
-  (define (fact n)
-    (cond [(n . zero?) one]
-          [else (n . times (this . fact (n . sub1)))])))
-
-;; Part 2:
-;; [2 pts -- data defintion]
+\begin{alltt}
+\rubricpts{2 pts -- data defintion}
 ;; A Natural is one of
 ;; - (new zero%)
-;; - (new succ% Natural)
+;; - (new add1% Natural)
 (define-class zero%
-  ;; [4 pts -- methods in zero%]
+  \rubricpts{4 pts -- methods in zero%}
   (define (sub1) this)
+  (define (zero?) true)
+  (define (add1) (new add1% this))
   (define (plus n) n)
-  (define (add1) (new succ% this))
-  (define (times n) this)
-  (define (zero?) true))
+  (define (times n) this))
 
-(define-class succ%
-  (fields pred)
-  ;; [7 pts -- methods in succ%]
+(define-class add1%
+  (fields sub1)
+  \rubricpts{7 pts -- methods in add1%}
   (define (zero?) false)
-  (define (add1) (new succ% this))
-  (define (plus n) (this . pred . plus n . add1))
-  (define (times n) (this . pred . times n . plus n))
-  (define (sub1) (this . pred)))
+  (define (add1) (new add1% this))
+  (define (plus n) (this . sub1 . plus n . add1))
+  (define (times n) (this . sub1 . times n . plus n)))
 
 (define zero (new zero%))
-;; [1 pt -- define one]
-(define one (new succ% (new zero%)))
+\rubricpts{1 pt -- define one}
+(define one (new add1% (new zero%)))
 
-;; [3 pts -- tests]
+\rubricpts{3 pts -- tests}
 (check-expect (one . zero?) false)
 (check-expect (one . sub1) zero)
-(check-expect ((new succ% one) . sub1) one)
+(check-expect ((new add1% one) . sub1) one)
 (check-expect (zero . zero?) true)
 (check-expect (zero . sub1) zero)
-\end{verbatim}
+\end{alltt}
 \newpage
@@
 \begin{solution}
-\begin{verbatim}
-;; [5 pts -- get it right]
+\begin{alltt}
+\rubricpts{5 pts -- get it right}
 (define-class fact%
@@
     (n . zero? . branch (new const% one) (new recur% n))))
+
 (define-class const%
   (fields run))
+
 (define-class recur%
   (fields n)
-  (define (run) ((new fact%) . fact (this . n . sub1) . times (this . n))))
+  (define (run) 
+    ((new fact%) . fact (this . n . sub1) . times (this . n))))
 
@@
 
-(define true (new t%)) (define false (new f%))
-\end{verbatim}
+(define true (new t%)) 
+(define false (new f%))
+
+(check-expect ((new fact%) . fact zero) one)                  ; 0! = 1
+(check-expect ((new fact%) . fact one) one)                   ; 1! = 1
+(check-expect ((new fact%) . fact (one . add1)) (one . add1)) ; 2! = 2
+(check-expect ((new fact%) . fact (one . add1 . add1))        ; 3! = 6
+              (one . add1 . add1 . add1 . add1 . add1))
+\end{alltt}
 \end{solution}

### 2012-03-25 exam4/2.tex
problem 1

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{tikz} 
+\usepackage{qtree} 
 \usepackage{./exam}
@@
 \usepackage{ct} % tight CM font for \tt
-\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{xcolor}
 \newexercise{exercise}{Exercise}[subsection]
@@
 \vspace*{-1.5cm}
-\centerline{\Large CSU2510H Exam 1 \ifrubric (SOLUTION) \fi-- Spring 2012}
+\centerline{\Large CSU2510H Exam 2 \ifrubric (SOLUTION) \fi-- Spring 2012}
 
@@
 
-% \noindent
-% {\bf Home on the Range}
-
-% Come up with funny story here.
-
-% \newpage
 \begin{problem}\pts{18} % same as on the previous exam :)
 
-Here is a data definition for a class of objects.  It intentionally
-uses meaningless names to test whether you've sufficiently
-internalized the data-driven development process we've talked about
-since the Fall.
+  While trying to manage the grade database, Asumu hit upon a new idea
+  for a data structure: the 2/3-tree.  A 2/3-tree is either a leaf
+  (which holds a {\tt String} value), a 2-node, or a 3-node.  A 2-node
+  has two children, and a 3-node has three children.  The twist is
+  that a 2-node can't have any 2-nodes as children, and a 3-node can't
+  have any 3-nodes as children.
 
-\begin{verbatim}
-;; A P is one of:          
-;; - A                  ;; An A is a (new a% String)
-;; - B                  ;; A  B is a (new b% String A)
-;; - C                  ;; A  C is a (new c% P)
+Here's an example of a 2/3-tree:
 
-(define-class a% (fields w))
-(define-class b% (fields x y))
-(define-class c% (fields z))
-\end{verbatim}
+\vspace*{8mm}
+\bigskip
 
-\noindent
-Design the following methods for \verb|P|s:
+\Tree [. {\tt "Cat"} {\tt "Hat"} [. {\tt "Zebra"} {\tt "Rhino"}  ] !{\qbalance} ]
+
+\bigskip
+\bigskip
 
@@
 
-\item \verb|size| - the size of an \verb|A| is $5$; the size of a
-  \verb|B| is $8$, regardless of what's inside; and the size of a
-  \verb|C| contributes $1$ to the size of the \verb|P| inside.
+\item Design data, class, and interface definitions, in Java, to
+  represent 2/3-trees.  Construct the above example using your
+  definitions.
 
-\item \verb|replace-x| - produces a \verb|P| like this one, but
-  with any \verb|x| field replaced by the given string.
-
-\item \verb|same-x-and-w?| - given two strings, produce true if this
-  \verb|P| contains a \verb|B| whose \verb|x| is equal to the
-  first given string and whose \verb|y| is an \verb|A| whose
-  \verb|w| is equal to the second given string.  Produces false
-  otherwise.
+\item Design the {\tt height} method, which computes the height of the
+  tree (i.e., the maximum distance from the root to a leaf).  The
+  example tree has a height of 3.
 
 \end{enumerate}
-
-Be sure to define any interfaces you need and say which classes
-implement which interfaces so that your program treats all values
-according to their interface.
 
@@
 \begin{alltt}
-\rubricpts{6 pts -- contract and purpose for each method}
-;; A P implements:
-;; - size : -> Number
-;;   Size of this P.
-;; - replace-x : String -> P
-;;   Replace x field (if any) of this P with given string.
-;; - same-x-and-w? : String String -> Boolean
-;;   Is there a B within this P with given x field and w field?
-
-\rubricpts{0 pts -- not required, but write something}
-;; An A implements:
-;; - w : -> String
-;;   Get the w string.
-
-\rubricpts{2 pts}
-;; In a%: 
-(define (size) 5)
-(define (replace-x str) this)
-(define (same-x-and-w? g c) false)
-
-\rubricpts{3 pts}
-;; In b%:
-(define (size) 8)
-(define (replace-x str)
-  (new b% str (this . y)))
-(define (same-x-and-w? g c)
-  (and (string=? g (this . x))
-       (string=? c (this . y . w))))
-
-\rubricpts{4 pts}
-;; In c%:
-(define (size) (add1 (this . z . size)))
-(define (replace-x str)
-  (new c% (this . z . replace-x str)))
-(define (same-x-and-w? g c)
-  (this . z . same-x-and-w? g c))
-
-;; tests
-\rubricpts{3 pts -- 1 per method}
-(define a (new a% "a"))
-(define b (new b% "b" a))
-(define c (new c% b))
-(define d (new c% a))
-(check-expect (a . size) 5)
-(check-expect (b . size) 8)
-(check-expect (c . size) 6)
-(check-expect (d . size) 9)
-(check-expect (a . replace-x "*") a)
-(check-expect (b . replace-x "*") (new b% "*" a))
-(check-expect (c . replace-x "*") (new c% (new b% "*" a)))
-(check-expect (d . replace-x "*") d)
-(check-expect (a . same-x-and-w? "b" "a") false)
-(check-expect (b . same-x-and-w? "b" "a") true)
-(check-expect (c . same-x-and-w? "b" "a") true)
-(check-expect (d . same-x-and-w? "b" "a") false)
+\rubricpts{? pts}
 \end{alltt}

### 2012-03-25 exam4/2.tex
problems 1&2 with solns

@@
 \usepackage{tikz} 
+\usepackage[]{xcolor}
 \usepackage{qtree} 
@@
 \usepackage{ct} % tight CM font for \tt
-\usepackage{xcolor}
 \newexercise{exercise}{Exercise}[subsection]
@@
 \ifrubric
-\newenvironment{solution}{\color{Red}}{}
+\newenvironment{solution}{\color{red}}{}
 \else
@@
 {\bf Problem} & Points & out of \\ \hline
-1 & & ??\\ \hline
-2 & & ??\\ \hline
+1 & & 15\\ \hline
+2 & & 22\\ \hline
 3 & & ??\\ \hline
@@
 
-\begin{problem}\pts{18} % same as on the previous exam :)
+\begin{problem}\pts{15} % same as on the previous exam :)
 
@@
 
-\Tree [. {\tt "Cat"} {\tt "Hat"} [. {\tt "Zebra"} {\tt "Rhino"}  ] !{\qbalance} ]
+\Tree [. {\tt "Bread"} {\tt "Hat"} [. [. {\tt "Zebra"} {\tt "Dog"} {\tt
+  "Cat"} ] {\tt "Rhinocerous"}  ] !{\qbalance} ]
 
@@
   tree (i.e., the maximum distance from the root to a leaf).  The
-  example tree has a height of 3.
+  example tree has a height of 4.
 
@@
 
+\ifrubric\else
 \newpage
-\ifrubric\else
 [Here is some more space for the previous problem.]
@@
 \begin{solution}
-\begin{alltt}
-\rubricpts{? pts}
-\end{alltt}
+\begin{verbatim}
+//> 4pts for interfaces
+interface Tree {
+  //> 1pt for purpose
+  // count the height of this tree
+  Integer height();
+}
+
+interface TwoTree extends Tree {}
+interface ThreeTree extends Tree {}
+\end{verbatim}
+\newpage
+\begin{verbatim}
+//> 4pts for class definitions
+//> 3pts for method definitions
+class Leaf implements TwoTree, ThreeTree {
+  String val;
+
+  Integer height() { return 1; }
+}
+
+class TwoNode implements TwoTree {
+  ThreeTree left;
+  ThreeTree right;
+
+  Integer height() { return max(this.left.height(), this.right.height()); }
+}
+
+class ThreeNode implements ThreeTree {
+  TwoTree left;
+  TwoTree mid;
+  TwoTree right;
+
+  Integer height() { return max(this.left.height(),
+                                this.mid.height(),
+                                this.right.height()); }
+}
+
+Tree r = new Leaf("Rhinocerous");
+Tree h = new Leaf("Hat");
+Tree b = new Leaf("Bread");
+Tree t1 = new ThreeNode(new Leaf("Zebra"),new Leaf("Dog"),new Leaf("Cat"));
+Tree t2 = new TwoNode(t1,r);
+Tree t3 = new ThreeNode(b,h,t2);
+
+//> 3pts for tests
+t3.height -> 4
+r.height -> 1
+t2.height -> 3
+\end{verbatim}
 \end{solution}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem}\pts{10}
+\begin{problem}\pts{22}
 
-Here are data, class, and interface definitions for a list of numbers:
-\begin{alltt}
-;; An LoN is one of:
-;; - (new mt%)
-;; - (new cons% Number LoN)
-;; and implements:
-;; - sum : -> Number
-;;   Compute the sum of the numbers in this list.
-;; - sqrs : -> LoN
-;;   List of squares of each number in this list.
+Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
+\verb|visit| method in all of your classes from the previous problem.
+Make sure you design the interfaces and methods so that a visitor can
+return any type of data.
 
-(define-class mt%
-  (define (sum) 0)
-  (define (sqrs) this))
+\begin{solution}
+\begin{verbatim}
+//> 2pts parameterization
+//> 2pts methods
+interface TreeVisitor<T> {
+  T visitLeaf(String s);
+  T visitTwoNode(ThreeTree l, ThreeTree r);
+  T visitThreeNode(TwoTree l, TwoTree r);
+}
 
-(define-class cons%
-  (fields first rest)
-  (define (sum)
-    (+ (this . first)
-       (this . rest . sum)))
-  (define (sqrs)
-    (new cons% 
-         (sqr (this . first)) ; (sqr x) = x\(\sp{2}\)
-         (this . rest . sqrs))))
-\end{alltt}
-%
-Here are some examples:
-\begin{verbatim}
-(check-expect ((new mt%) . sum) 0)
-(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum)  9)
-(check-expect ((new mt%) . sqrs) (new mt%))
-(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sqrs)
-              (new cons% 16 (new cons% 25 (new mt%))))
+//> 2pts overall for getting types right
+//> 3pts overall for getting impls right
+
+// in Leaf:
+<T> T visit(TreeVisitor<T> v) { 
+  return v.visitLeaf(this.val);
+}
+
+// in TwoNode:
+<T> T visit(TreeVisitor<T> v) { 
+  return v.visitTwoNode(this.left, this.right);
+}
+
+// in ThreeNode:
+<T> T visit(TreeVisitor<T> v) { 
+  return v.visitThreeNode(this.left, this.mid, this.right);
+}
+// no tests required!
 \end{verbatim}
+\end{solution}
 
 \newpage
-\noindent
-Develop the \verb|sum-sqrs| method that computes the sum of the
-squares of a list of numbers.  Here is an example:
-\begin{verbatim}
-(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum-sqrs) 
-              (+ 16 25))
-\end{verbatim}
 
-If you can, define a super class of \verb|mt%| and \verb|cons%| and
-lift the definition of \verb|sum-sqrs| to this class.  (We will assume
-the appropriate \verb|super| declaration in \verb|mt%| and
-\verb|cons%|.)  If you cannot, define methods in both \verb|mt%| and
-\verb|cons%| for partial credit.
+\ifrubric\else
+[Here is some more space for the previous problem.]
+\newpage
+\fi
+
+Design the \verb|CountVisitor| class, which counts the number of
+leaves in a 2/3-tree (there are 6 in the given example.)
 
 \begin{solution}
-\begin{alltt}
-\rubricpts{2 pts% -- class def}
-(define-class list%
-  \rubricpts{2 pts -- contract and purpose}
-  ;; sum-sqrs : -> Number
-  ;; compute the sum of the squares of the numbers in this list
-  \rubricpts{4 pts -- implementation}
-  (define (sum-sqrs)
-    ((this . sqrs) . sum)))
+\begin{verbatim}
+//> 2pts class def
+//> 2pts implementations
+class CountVisitor implements TreeVisitor<Integer> {
+  Integer visitLeaf(String s) { return 1; }
+  Integer visitTwoNode(Tree l, Tree r) { return l.visit(this) + r.visit(this); }
+  Integer visitThreeNode(Tree l, Tree m, Tree r) { 
+     return l.visit(this) + r.visit(this) + m.visit(this);
+  }
+}
+//> 2pts tests
+r.visit(new CountVisitor()) -> 1
+t2.visit(new CountVisitor()) -> 4
+t3.visit(new CountVisitor()) -> 6
+\end{verbatim}
+\end{solution}
 
-\rubricpts{2 pts -- tests}
-(check-expect ((new mt%) . sum-sqrs) 0)
-\end{alltt}
+\newpage
+
+Design the \verb|LongestVisitor| class, which produces the longest
+\verb|String| in a 2/3-tree.  The longest string in the example is
+\verb|"Rhinocerous"|.
+
+\begin{solution}
+\begin{verbatim}
+// 2pts class def
+// 3pts methods
+class LongestVisitor implements TreeVisitor<String> {
+  String visitLeaf(String s) { return s; }
+  String visitTwoNode(Tree l, Tree r) { 
+     String s1 = l.visit(this);
+     String s2 = r.visit(this);
+     return (s1.length() > s2.length())?s1:s2;
+  }
+  String visitThreeNode(Tree l, Tree m, Tree r) { 
+     String s1 = l.visit(this);
+     String s2 = r.visit(this);
+     String s3 = m.visit(this);
+     if (s1.length() > s2.length() && s1.length() > s3.length())
+       return s1;
+     else if (s2.length() > s3.length())
+       return s2;
+     else
+       returns s3;
+  }
+}
+//> 2pts tests
+h.visit(new LongestVisitor()) -> "Hat"
+t2.visit(new LongestVisitor()) -> "Rhinocerous"
+t3.visit(new LongestVisitor()) -> "Rhinocerous"
+\end{verbatim}
 \end{solution}

### 2012-03-26 exam4/2.tex
Exam 2h problem 3.

@@
 
-In order to manage class scheduling, the College decided that they
-needed to create a new implementation of the \verb|Set| data
-structure.  They went with the lowest bidder, which
-turned out to be a bunch of BU students.  Now, having paid for this
-code, the College needs to test it, and they turn to you.
-
-Unfortunately, the code the BU students produced is so convoluted that
-it doesn't bear looking at.  All you have to go on is that the
-\verb|Set| data structure obeys the following interface:
+In mathematics, an {\bf unordered} pair is a pair of values in which
+the order of the elements does not matter.  So for example, if
+$\{3,4\}$ is an unordered pair, then it should be considered ``the
+same'' as the unordered pair $\{4,3\}$.  Here is an implementation of
+{\tt UnPair}, which represents unordered pairs.  Notice that both {\tt
+  equals} and {\tt hashCode} have been overridden.
 
 \begin{verbatim}
-;; A Set implements:
-;; - insert : String -> Set
-;;   adds given string to this set, if it isn't already there
-;;   (otherwise produces this set)
-;; - remove : String -> Set
-;;   removes given string from this set, if it's already there
-;;   (otherwise produces this set)
-;; - size : -> Number
-;;   count the elements of this set
-;; - has? : String -> Boolean
-;;   is the given String in this set?
+class UnPair<X> {
+  X left;
+  X right;
+  UnPair(X left, X right) {
+    this.left = left;
+    this.right = right;
+  }
 
-;; empty-set is a Set value containing no elements
+  // Compute the hash of this unordered pair.
+  public int hashCode() {
+    return left.hashCode();
+  }
+
+  // Is this unordered pair the same as the given unordered pair?
+  public boolean same(UnPair<X> p) { ... }
+
+  // Is this unordered pair the same as the given object?
+  public boolean equals(Object that) {
+    return (that instanceof UnPair)
+        && this.same((UnPair<X>)that);
+  }
+}
 \end{verbatim}
-\noindent
-\verb|Set|s should satisfy the following property, called the
-``insert-size'' property, for any element and
-set:
-
-\begin{quote}
-If we insert any element into any set, then the resulting set is the
-same size as the original set \emph{if} the element was already in the
-set, and the resulting set is one larger than the original set
-if the element was \emph{not} already in the set.
-\end{quote}
-
-\begin{enumerate}
-\item Codify the ``insert-size'' property as a predicate.  Write a
-  test case that should pass (if the sets work right) using the
-  predicate you have defined.
-\end{enumerate}
-
-\ifrubric\else
-\newpage
-[Here is some more space for the previous problem.]
-\fi
 
@@
 \noindent
-The College wants to be really sure that their sets are correct, and
-they've come up with another property that they think is true, called
-the ``insert-remove'' property:
-
-\begin{quote}
-If we insert element any element $a$ into any set, and then remove any
-element $b$ from the resulting set, the result has the same size as if
-we do the two operations in the reverse order, i.e., remove $b$, then
-insert $a$.
-\end{quote}
-
-\noindent
-The College is considering adding this to the test suite as well.
-
-\begin{enumerate}
-\setcounter{enumi}{1}
-
-\item Codify this property as a predicate, and write a test case that
-  should pass using the predicate.
-
-\item Is the ``insert-remove'' property true for all sets?  If so,
-  give an explanation.  If not, write a counter-example (i.e., a
-  failing test case) using your predicate.
-
-\end{enumerate}
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{1 pt -- contract & purpose}
-;; check-insert-size : Set String -> Boolean
-;; check the "insert-size" property on s and i.
-\rubricpts{4 pts -- code}
-(define (check-insert-size s i)
-  (= (s . insert i . size)
-     (cond [(s . has? i) (add1 (s . size))]
-           [else               (s . size)])))
-
-\rubricpts{2 pts -- test case}
-(check-expect (check-insert-size empty-set "hi") true)
-
-\rubricpts{1 pt -- contract & purpose}
-;; check-remove-insert : Set String -> Boolean
-;; check the "remove-insert" property on s, a, and b.
-\rubricpts{3 pts -- code}
-(define (check-remove-insert s a b)
-  (= (s . insert a . remove b . size)
-     (s . remove b . insert a . size))
-
-\rubricpts{1 pt -- successful test}
-(check-expect (check-remove-insert empty-set "yes" "no") true)
-
-;; Property does not hold.
-;; Counter-example (failing test):
-\rubricpts{3 pts -- failing test}
-(check-expect (check-remove-insert empty-set "yes" "yes") true)
-\end{alltt}
-\end{solution}
+Part (A): Implement the omitted {\tt same} method that compares this
+unordered pair to a given unordered pair.  This method should work
+regardless of the order of the pair, so for example {\tt new
+  UnPair<Integer>(3,4)} is the same (according to {\tt same}) as {\tt
+  new UnPair<Integer>(4,3)}.
 
 \newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
+\noindent
+Part (B): Assuming {\tt same} and {\tt equals} work as expected, is
+the given {\tt hashCode} method valid?  That is, does it satisfy or
+violate the law of {\tt hashCode}?  If it is valid, give an argument
+for \emph{why}.  If it is not, give a counter-example to the law of
+{\tt hashCode}.
+\end{problem}
 \newpage
-\fi
-\end{problem}
 

### 2012-03-26 exam4/2.tex
Exam 2H, problem 4.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{15}
+\begin{problem} \pts{??}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{15}
+\begin{problem} \pts{??}
 
-Here are data and class definitions for representing the files and
-directories of a computer:
+Lists are nice, but sometimes it would be better to have a list
+without an end.  For example, we might want to model the days of the
+week as a list that wraps around at the end so that Monday follows
+Sunday.  To do so, we need lists that are \emph{circular}, and in
+order to make circular lists, we need to use mutation.  Here's an idea
+for a list method that makes is possible to create cyclic lists:
 
 \begin{verbatim}
-;; An Elem is one of
-;; - (new file% String Number)
-;; - (new dir% String LoElem)
-(define-class file% (fields name size))
-(define-class dir% (fields name elems))
-
-;; A LoElem is one of
-;; - (new mt%)
-;; - (new cons% Elem LoElem)
-(define-class mt%)
-(define-class cons% (fields first rest))    
+interface List<X> {
+   // EFFECT: set the rest of the last cons in this list
+   // to the given list.
+   void setTail(List<X> ls);
+}
 \end{verbatim}
@@
 \noindent
-Here is an example of a directory tree and its representation as an
-\verb|Elem|:
+It works as follows:
+
 \begin{verbatim}
-#|
- User
-   +-- A
-   |   +-- a.txt
-   |   +-- B
-   |       +-- b.rkt
-   +-- C
-       +-- c.c
-|#
+List<Integer> waltz =
+  new Cons<Integer>(1,
+    new Cons<Integer>(2,
+      new Cons<Integer>(3,
+        new MT<Integer>())));
 
-(define mt (new mt%))
-(define a (new file% "a.txt" 5))
-(define b (new file% "b.rkt" 10))
-(define c (new file% "c.c" 1000))
+\end{verbatim}
+At this point, {\tt waltz} is just a list of three elements {\tt 1, 2,
+  3}.  To make the list circular we can call {\tt setTail} giving {\tt
+  waltz} as the new tail:
 
-(define C/ (new dir% "C" (new cons% c mt)))
-(define B/ (new dir% "B" (new cons% b mt)))
-(define A/ (new dir% "A" (new cons% a (new cons% B/ mt))))
-(define User/ (new dir% "/" (new cons% A/ (new cons% C/ mt))))
+\begin{verbatim}
+waltz.setTail(waltz);
 \end{verbatim}
 
+\noindent
+Now {\tt waltz} is a list of elements: {\tt 1, 2, 3, 1, 2, 3, 1, 2, 3,
+  \dots}, and so on, \emph{ad infinitum}.
+
+Of course, it's also possible to use {\tt setTail} to update a list in
+a non-circular way.  Imagine instead we had just done:
+
+\begin{verbatim}
+waltz.setTail(new Cons<Integer>(4, new MT<Integer>()));
+\end{verbatim}
+
+\noindent
+In this case, {\tt waltz} is now just {\tt 1, 2, 3, 4}.
+
 \newpage
-\noindent
-Your job is to implement one of the main tasks of the common Unix
-utility \verb|find|, which gives a list of all the files within an
-element.  To do so, add a \verb|list-files| method of \verb|Elem|
-which produces a list of files.  So for example, we expect:
-\begin{verbatim}
-(check-expect (User/ . list-files) 
-              (new cons% a (new cons% b (new cons% c mt))))
-\end{verbatim}
-\emph{Hint}: it may come in handy to be able to concatenate two
-\verb|LoElem| together.
-
-\ifrubric
-\else
-\newpage
-[Here is some more space for the previous problem.]
-\fi
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{1pts -- interfaces, contracts, purpose}
-;; An Elem implements
-;; - list-files : -> LoElem
-;;   List all files within this element.
-
-\rubricpts{2pts -- interfaces, contracts, purpose}
-;; A LoElem implements
-;; - list-files : -> LoElem
-;;   List all files within this list of elements.
-;; - append : LoElem -> LoElem
-;;   Append this list of elements to given list.
-
-\rubricpts{1pt -- code}
-;; in dir%
-(define (list-files)
-  (this . elems . list-files))
-
-\rubricpts{1pt -- code}
-;; in file%
-(define (list-files)
-  (new cons% this (new mt%)))
-
-\rubricpts{2pt -- code}
-;; in mt%
-(define (append loe) loe)
-(define (list-files) this)
-
-\rubricpts{4pt -- code}
-;; in cons%
-(define (append loe)
-  (new cons% (this . first) (this . rest . append loe)))
-(define (list-files)
-  (this . first . list-files . append (this . rest . list-files)))
-
-\rubricpts{1pt -- sanity}
-
-\rubricpts{3pts -- tests}
-;; tests
-(check-expect (mt . append (new cons% a mt)) (new cons% a mt))
-(check-expect ((new cons% a mt) . append (new cons% b mt))
-              (new cons% a (new cons% b mt)))
-
-(check-expect (mt . list-files) mt)
-(check-expect ((new cons% a (new cons% b mt)) . list-files)
-              (new cons% a (new cons% b mt)))
-
-(check-expect (a . list-files) (new cons% a mt))
-\end{alltt}
-\end{solution}
+Design an implementation of {\tt setTail} that works as described.
+Revise the {\tt List<X>} interface with any methods you need to add to
+make {\tt setTail} work.  (You'll notice that the effect statement for
+{\tt setTail} implies that this list cannot be empty; if you try to
+{\tt setTail} on an empty list, you may raise an exception.)
 
 \end{problem}
-\newpage
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{18}
-Last night, Asumu was up late improving the {\tt class} languages.
-Unfortunately, he was too tired to test his work, and his last commit
-broke the numeric system.
-%
-This puts us in the dire position of no longer being able to calculate
-the factorial of large numbers, and we turn to you for help.
-
-This is the definition of a class of functional objects
-with a {\tt fact} method:
-
-\begin{verbatim}
-;; A Fact is (new fact%) and implements 
-;; - fact : Natural -> Natural
-;;   Compute n!.
-(define-class fact%
-  (define (fact n)
-    (cond [(zero? n) 1]
-          [else (* n (this . fact (sub1 n)))])))
-
-(check-expect ((new fact%) . fact 0) 1)
-(check-expect ((new fact%) . fact 5) 120)
-\end{verbatim}
-
-Since we can no longer rely on built in numbers to represent
-\verb|Natural|s, we will represent \verb|Natural|s with objects and
-then rewrite the \verb|fact| method to operate over this new kind of
-data.
-
-Here is the \verb|Natural| interface:
-
-\begin{verbatim}
-;; A Natural implements:
-;; - zero? : -> Boolean
-;;   is this Natural the zero value?
-;; - sub1 : -> Natural
-;;   produce a natural one less than this natural 
-;;   (or zero if this natural is zero)
-;; - times : Natural -> Natural
-;;   multiply this natural by the given natural
-\end{verbatim}
-
-\newpage
-Write a new definition of the \verb|fact| method using the
-\verb|Natural| interface definition given above, and \emph{not} using
-the built-in numbers of the class system or DrRacket.  You can assume
-there is a defined constant \verb|one|, which is a \verb|Natural|
-object representing $1$.
-
-\begin{verbatim}
-(check-expect ((new fact%) . fact one) one)
-(check-expect ((new fact%) . fact (one . sub1)) one)
-\end{verbatim}
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{2 pts}
-(define-class fact%
-  (define (fact n)
-    (cond [(n . zero?) one]
-          [else (n . times (this . fact (n . sub1)))])))
-
-\end{alltt}
-\end{solution}
-
-\newpage
-
-Now, design data and class defintions that implement \verb|Natural|, and
-supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.
-Define the constant \verb|one|.  (\emph{Hint}: remember that the
-natural numbers can be seen as a recursive union; a natural number is
-either zero, or one more than a natural number.)
-
-It will be helpful to define the helper methods \verb|add1| and
-\verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
-\cdot m = (n-1) \cdot m + m$.  Don't overlook the slightly weird
-specification for \verb|sub1|: $0-1 = 0$; this weirdness is common
-amongst mathematicians when working with natural numbers.
-
-\ifrubric
-\else
-\newpage
-[Here is some more space for the previous problem.]
-\newpage
-\fi
-\end{problem}
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{2 pts -- data defintion}
-;; A Natural is one of
-;; - (new zero%)
-;; - (new add1% Natural)
-(define-class zero%
-  \rubricpts{4 pts -- methods in zero%}
-  (define (sub1) this)
-  (define (zero?) true)
-  (define (add1) (new add1% this))
-  (define (plus n) n)
-  (define (times n) this))
-
-(define-class add1%
-  (fields sub1)
-  \rubricpts{7 pts -- methods in add1%}
-  (define (zero?) false)
-  (define (add1) (new add1% this))
-  (define (plus n) (this . sub1 . plus n . add1))
-  (define (times n) (this . sub1 . times n . plus n)))
-
-(define zero (new zero%))
-\rubricpts{1 pt -- define one}
-(define one (new add1% (new zero%)))
-
-\rubricpts{3 pts -- tests}
-(check-expect (one . zero?) false)
-(check-expect (one . sub1) zero)
-(check-expect ((new add1% one) . sub1) one)
-(check-expect (zero . zero?) true)
-(check-expect (zero . sub1) zero)
-\end{alltt}
-\newpage
-\end{solution}
-
-\begin{problem}\pts{5}
-
-\textbf{Extra Credit}
-\\
-\noindent
-After Van Horn discovered the problems with numbers, he attempted to
-fix them.  Unfortunately, not only did he not fix them, he broke
-booleans as well.  Fortunately, we have an idea about how you can
-help, expressed in the following interface definitions.
-
-\begin{verbatim}
-;; A Bool implements:
-;; - branch : [Action X] [Action X] -> X
-;;   run the first action if this bool is true, 
-;;   otherwise run the second action
-
-;; An [Action X] implements:
-;; - run : -> X
-;; run this action
-
-\end{verbatim}
-
-Following these interface definitions, design data and class
-defintions for implementations of the \verb|Bool| interface, and
-reimplement the \verb|fact| method to use them.
-You will need to provide new defintions for \verb|true| and
-\verb|false|, since they were broken by this catastrophe as well.
-
-\ifrubric\else
-\newpage
-[Here is some more space for the previous problem.]
-\fi
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{5 pts -- get it right}
-(define-class fact%
-  (define (fact n)
-    (n . zero? . branch (new const% one) (new recur% n))))
-
-(define-class const%
-  (fields run))
-
-(define-class recur%
-  (fields n)
-  (define (run) 
-    ((new fact%) . fact (this . n . sub1) . times (this . n))))
-
-(define-class t%
-  (define (branch t e) (t . run)))
-(define-class f%
-  (define (branch t e) (e . run)))
-
-(define true (new t%)) 
-(define false (new f%))
-
-(check-expect ((new fact%) . fact zero) one)                  ; 0! = 1
-(check-expect ((new fact%) . fact one) one)                   ; 1! = 1
-(check-expect ((new fact%) . fact (one . add1)) (one . add1)) ; 2! = 2
-(check-expect ((new fact%) . fact (one . add1 . add1))        ; 3! = 6
-              (one . add1 . add1 . add1 . add1 . add1))
-\end{alltt}
-\end{solution}
-\end{problem}
 \end{document}

### 2012-03-27 exam4/UnPair.java
unpair

@@ -1,0 +1,37 @@
+import tester.*;
+
+class UnPair<X> {
+  X left;
+  X right;
+  UnPair(X left, X right) {
+    this.left = left;
+    this.right = right;
+  }
+
+  public int hashCode() {
+    return left.hashCode();
+  }
+
+  public boolean same(UnPair<X> p) {
+     return (this.left.equals(p.left) &&
+             this.right.equals(p.right))
+         || (this.left.equals(p.right) &&
+             this.right.equals(p.left));
+  }
+
+  public boolean equals(Object that) {
+    return (that instanceof UnPair)
+        && this.same((UnPair<X>)that);
+  }
+}
+
+class Examples {
+    void testUnPair(Tester t) {
+	t.checkExpect(new UnPair<Integer>(5,7).equals(new UnPair<Integer>(5,7)),
+		      true);
+	t.checkExpect(new UnPair<Integer>(5,7).equals(new UnPair<Integer>(7,5)),
+		      true);
+	t.checkExpect(new UnPair<Integer>(5,7).equals(new UnPair<Integer>(7,7)),
+		      false);
+    }
+}

### 2012-03-27 exam4/2.tex
problem 5

@@
 
+\newpage
+
+\begin{problem}
+\pts{???}
+
+Iterators are an extremely useful concept.  In particular, they aren't
+limited to iterating over the contents of data structures such as {\tt
+ArrayList}s.  We can also construct iterators directly, for sequences
+we are interested in.
+
+\bigskip
+
+Design the class {\tt RangeIterator}, which takes two {\tt Integer}s
+as constructor arguments, and iterates over all of the {\tt Integer}s
+between them, including \emph{both} end points.  {\tt RangeIterator}
+must implement the {\tt Iterator} interface, given below.
+
+\begin{verbatim}
+interface Iterator<T> {
+  Boolean hasNext();
+  T next();
+  // The Java definition of Iterator includes the remove method.
+  // You do *not* need to implement remove.
+}
+\end{verbatim}
+
+\begin{solution}
+\begin{verbatim}
+class RangeIterator implements Iterator<Integer> {
+  Integer lo;
+  Integer hi;
+
+  RangeIterator (Integer lo, Integer hi) { 
+    this.lo = lo; this.hi = hi; 
+  }
+
+  Boolean hasNext() { return this.lo <= this.hi; }
+
+  Integer next() { this.lo = this.lo + 1; return this.lo - 1; }
+}
+
+// Tests
+RangeIterator ri = new RangeIterator(0,2);
+ri.hasNext() -> true
+ri.next() -> 0
+ri.next() -> 1
+ri.hasNext() -> true
+ri.next() -> 2
+ri.hasNext() -> false
+\end{verbatim}
+\end{solution}
+\newpage
+
+\noindent
+Recall our definition of function objects:
+\begin{verbatim}
+interface Function<T,U> {
+  U call(T x);
+}
+\end{verbatim}
+\noindent
+Define the {\tt Square} class, which is a \verb|Function| that squares
+\verb|Integer|s.
+
+\ifrubric
+\begin{solution}
+\begin{verbatim}
+class Square implements Function<Integer,Integer> {
+  Integer call(Integer x) { return x * x; }
+} 
+Square s = new Square();
+s.call(3) -> 9
+s.call(0) -> 0
+\end{verbatim}
+\end{solution}
+\else
+\newpage
+\fi
+
+Using this interface, design and implement the {\tt MapIterator}
+class, which takes a \verb|Function| and an \verb|Iterator| as
+arguments, and iterates over the values produced by applying the
+function to each result produced by the iterator.
+
+\ifrubric
+\begin{solution}
+\begin{verbatim}
+class MapIterator<T,U> implements Iterator<U> {
+  Function<T,U> f;
+  Iterator<T> i;
+  MapIterator<T,U>(Function<T,U> f, Iterator<T> i) {
+   this.f = f; this.i = i;
+  }
+
+  boolean hasNext() { return i.hasNext(); }
+  U next() { return f.call(i.next()); }
+}
+
+//tests
+Iterator<Integer> i = new MapIterator<Integer,Integer>(new Square(), 
+                           new RangeIterator(1,5));
+i.next() -> 1
+i.next() -> 4
+i.next() -> 9
+i.hasNext() -> true
+
+\end{verbatim}
+\end{solution}
+\else
+\fi
+\newpage
+
+\noindent
+Now design the method \verb|sumSquares|, which takes two
+\verb|Integer|s, and uses the classes you've just defined, along with
+a \verb|for| or \verb|while| loop, to compute the sum of the squares
+of all the numbers between the given \verb|Integer|s, inclusive.
+
+\ifrubric
+\begin{solution}
+\begin{verbatim}
+class Test {
+  Integer sumSquares(Integer a, Integer b) {
+    Iterator<Integer> i = new MapIterator<Integer,Integer>(new Square(), 
+                             new RangeIterator(a,b));
+    Integer sum = 0;
+    while (i.hasNext()) { sum += i.next(); }
+    return sum;
+  }
+}
+
+sumSquares(1,2) -> 5
+sumSquares(0,0) -> 0
+sumSquares(10,10) -> 100
+\end{verbatim}
+\end{solution}
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+
+\end{problem}
+
 \end{document}

### 2012-03-27 exam4/SetTail.java
setTail.

@@ -1,0 +1,50 @@
+import tester.*;
+
+interface List<X> {
+    void setTail(List<X> ls);
+    void setTailAcc(List<X> ls, Cons<X> last);
+}
+
+class MT<X> implements List<X> {
+    public void setTail(List<X> ls) {
+	throw new RuntimeException("setTail on MT");
+    }
+    public void setTailAcc(List<X> ls, Cons<X> last) {
+	last.rest = ls;
+    }
+}
+
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+    Cons(X first, List<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+    public void setTail(List<X> ls) {
+	this.rest.setTailAcc(ls, this);
+    }
+    public void setTailAcc(List<X> ls, Cons<X> last) {
+	this.rest.setTailAcc(ls, this);
+    }
+}
+
+class Examples {
+    void testSetTail(Tester t) {
+	List<Integer> one = new Cons<Integer>(1, new MT<Integer>());
+	List<Integer> two = new Cons<Integer>(2, new MT<Integer>());
+	one.setTail(two);
+	t.checkExpect(one, new Cons<Integer>(1,
+					     new Cons<Integer>(2,
+							       new MT<Integer>())));
+	t.checkExpect(((Cons)one).rest == two, true);
+    }
+    void testSetTailCycle(Tester t) {
+	List<Integer> ones = new Cons<Integer>(1, new MT<Integer>());
+	ones.setTail(ones);
+	t.checkExpect(((Cons)ones).first, 1);
+	t.checkExpect(((Cons)((Cons)ones).rest).first, 1);
+	t.checkExpect(((Cons)((Cons)((Cons)ones).rest).rest).first, 1);
+	t.checkExpect(((Cons)ones).rest == ones, true);
+    }
+}

### 2012-03-27 exam4/2.tex
done with soln

@@
 4 & & ??\\ \hline
-5 & & ??\\ \hline
+5 & & 20\\ \hline
 %% Extra & & 5 \\ \hline
@@
 //> 3pts for tests
-t3.height -> 4
-r.height -> 1
-t2.height -> 3
+t3.height() -> 4
+r.height() -> 1
+t2.height() -> 3
 \end{verbatim}
@@
 //> 3pts overall for getting impls right
+
+// in Tree:
+  <T> T visit(TreeVisitor<T> v);
 
@@
 \begin{problem}
-\pts{???}
+\pts{20}
 
@@
 \begin{verbatim}
+//> 4pts
 class RangeIterator implements Iterator<Integer> {
@@
 
-  Boolean hasNext() { return this.lo <= this.hi; }
+  public boolean hasNext() { return this.lo <= this.hi; }
 
-  Integer next() { this.lo = this.lo + 1; return this.lo - 1; }
+  public Integer next() { this.lo = this.lo + 1; return this.lo - 1; }
 }
 
-// Tests
+//> Tests 2pts
 RangeIterator ri = new RangeIterator(0,2);
@@
 \begin{verbatim}
+//> 3pts
 class Square implements Function<Integer,Integer> {
-  Integer call(Integer x) { return x * x; }
+  public Integer call(Integer x) { return x * x; }
 } 
@@
 \begin{verbatim}
+//> 4pts
 class MapIterator<T,U> implements Iterator<U> {
@@
 
-  boolean hasNext() { return i.hasNext(); }
-  U next() { return f.call(i.next()); }
+  public boolean hasNext() { return i.hasNext(); }
+  public U next() { return f.call(i.next()); }
 }
 
-//tests
+//tests 2pts
 Iterator<Integer> i = new MapIterator<Integer,Integer>(new Square(), 
@@
 \begin{verbatim}
+//> 4pts
 class Test {
@@
 
+//> Tests 1pt
 sumSquares(1,2) -> 5

### 2012-03-27 exam4/2-3.java

@@ -1,0 +1,109 @@
+import tester.*;
+
+interface Tree {
+    //> 1pt for purpose
+    // count the height of this tree
+    Integer height();
+    <T> T visit(TreeVisitor<T> v);    
+}
+interface TwoTree extends Tree {}
+interface ThreeTree extends Tree {}
+interface TreeVisitor<T> {
+  T visitLeaf(String s);
+  T visitTwoNode(ThreeTree l, ThreeTree r);
+  T visitThreeNode(TwoTree l, TwoTree m, TwoTree r);
+}
+
+//> 4pts for class definitions
+//> 3pts for method definitions
+class Leaf implements TwoTree, ThreeTree {
+    String val;
+    Leaf(String v) { val = v; }
+
+    public <T> T visit(TreeVisitor<T> v) { 
+	return v.visitLeaf(this.val);
+    }
+
+    public Integer height() { return 1; }
+}
+
+class TwoNode implements TwoTree {
+    ThreeTree left;
+    ThreeTree right;
+
+    TwoNode(ThreeTree l, ThreeTree r) { left = l; right = r; }
+
+    public <T> T visit(TreeVisitor<T> v) { 
+	return v.visitTwoNode(this.left, this.right);
+    }
+
+    public Integer height() { return Math.max(this.left.height(), this.right.height()); }
+}
+
+class ThreeNode implements ThreeTree {
+    TwoTree left;
+    TwoTree mid;
+    TwoTree right;
+
+    ThreeNode(TwoTree l, TwoTree m, TwoTree r) { left = l; right = r; mid = m; }
+
+    public <T> T visit(TreeVisitor<T> v) { 
+	return v.visitThreeNode(this.left, this.mid, this.right);
+    }
+    public Integer height() { return Math.max(this.left.height(),
+					      Math.max(this.mid.height(),
+						       this.right.height())); }
+}
+
+class LongestVisitor implements TreeVisitor<String> {
+    public String visitLeaf(String s) { return s; }
+    public String visitTwoNode(ThreeTree l, ThreeTree r) { 
+	String s1 = l.visit(this);
+	String s2 = r.visit(this);
+	return (s1.length() > s2.length())?s1:s2;
+    }
+    public String visitThreeNode(TwoTree l, TwoTree m, TwoTree r) { 
[…]

### 2012-03-27 exam4/iters.java

@@ -1,0 +1,72 @@
+import tester.*;
+import java.util.*;
+
+class MapIterator<T,U> implements Iterator<U> {
+  Function<T,U> f;
+  Iterator<T> i;
+  MapIterator(Function<T,U> f, Iterator<T> i) {
+   this.f = f; this.i = i;
+  }
+
+  public boolean hasNext() { return i.hasNext(); }
+  public U next() { return f.call(i.next()); }
+  public void remove() { throw new RuntimeException("remove"); }
+}
+
+interface Function<T,U> {
+  U call(T x);
+}
+
+class RangeIterator implements Iterator<Integer> {
+  Integer lo;
+  Integer hi;
+  
+  RangeIterator (Integer lo, Integer hi) { 
+    this.lo = lo; this.hi = hi; 
+  }
+
+  public boolean hasNext() { return this.lo <= this.hi; }
+
+  public Integer next() { this.lo = this.lo + 1; return this.lo - 1; }
+
+  public void remove() { throw new RuntimeException("remove"); }
+
+}
+
+class Square implements Function<Integer,Integer> {
+  public Integer call(Integer x) { return x * x; }
+} 
+
+
+class Test {
+    Integer sumSquares(Integer a, Integer b) {
+	Iterator<Integer> i = new MapIterator<Integer,Integer>(new Square(), 
+							       new RangeIterator(a,b));
+	Integer sum = 0;
+	while (i.hasNext()) { sum += i.next(); }
+	return sum;
+    }
+    
+    void test(Tester t) {
+	//> Tests 2pts
+	RangeIterator ri = new RangeIterator(0,2);
+	t.checkExpect(ri.hasNext(), true);
+	t.checkExpect(ri.next(), 0);
+	t.checkExpect(ri.next(), 1);
+	t.checkExpect(ri.hasNext(), true);
+	t.checkExpect(ri.next(), 2);
+	t.checkExpect(ri.hasNext(), false);
+	Iterator<Integer> i = new MapIterator<Integer,Integer>(new Square(), 
+							       new RangeIterator(1,5));
+	t.checkExpect(i.next(), 1);
+	t.checkExpect(i.next(), 4);
+	t.checkExpect(i.next(), 9);
+	t.checkExpect(i.hasNext(), true);
+	//> Tests 1pt
+	t.checkExpect(sumSquares(1,2), 5);
+	t.checkExpect(sumSquares(0,0), 0);
+	t.checkExpect(sumSquares(10,10), 100);
+    }
+	
+}
+

### 2012-03-27 exam4/2.tex
fix compilation problems in soln

@@
   tree (i.e., the maximum distance from the root to a leaf).  The
-  example tree has a height of 4.
+  example tree has a height of 4. You may find the function
+  \verb|Math.max| useful; it takes two numbers and produces the
+  largest.
 
@@
   String val;
+  Leaf(String v) { val = v; }
 
@@
   ThreeTree right;
+  TwoNode(ThreeTree l, ThreeTree r) { left = l; right = r; }
 
-  Integer height() { return max(this.left.height(), this.right.height()); }
+  Integer height() { return Math.max(this.left.height(), this.right.height()); }
 }
@@
 
-  Integer height() { return max(this.left.height(),
-                                this.mid.height(),
-                                this.right.height()); }
+  TwoNode(ThreeTree l, ThreeTree r) { left = l; right = r; }
+
+  Integer height() { return Math.max(this.left.height(),
+                                     Math.max(this.mid.height(),
+                                              this.right.height())); }
 }
-
-Tree r = new Leaf("Rhinocerous");
-Tree h = new Leaf("Hat");
-Tree b = new Leaf("Bread");
-Tree t1 = new ThreeNode(new Leaf("Zebra"),new Leaf("Dog"),new Leaf("Cat"));
-Tree t2 = new TwoNode(t1,r);
+ThreeTree r = new Leaf("Rhinocerous");
+TwoTree h = new Leaf("Hat");
+TwoTree b = new Leaf("Bread");
+ThreeTree t1 = new ThreeNode(new Leaf("Zebra"),new Leaf("Dog"),new Leaf("Cat"));
+TwoTree t2 = new TwoNode(t1,r);
 Tree t3 = new ThreeNode(b,h,t2);
-
 //> 3pts for tests
@@
   T visitTwoNode(ThreeTree l, ThreeTree r);
-  T visitThreeNode(TwoTree l, TwoTree r);
+  T visitThreeNode(TwoTree l, TwoTree m, TwoTree r);
 }
@@
   Integer visitLeaf(String s) { return 1; }
-  Integer visitTwoNode(Tree l, Tree r) { return l.visit(this) + r.visit(this); }
-  Integer visitThreeNode(Tree l, Tree m, Tree r) { 
+  Integer visitTwoNode(ThreeTree l, ThreeTree r) { return l.visit(this) + r.visit(this); }
+  Integer visitThreeNode(TwoTree l, TwoTree m, TwoTree r) { 
      return l.visit(this) + r.visit(this) + m.visit(this);
@@
   String visitLeaf(String s) { return s; }
-  String visitTwoNode(Tree l, Tree r) { 
+  String visitTwoNode(ThreeTree l, ThreeTree r) { 
      String s1 = l.visit(this);
@@
   }
-  String visitThreeNode(Tree l, Tree m, Tree r) { 
+  String visitThreeNode(TwoTree l, TwoTree m, TwoTree r) { 
      String s1 = l.visit(this);
@@
      else
-       returns s3;
+       return s3;
   }

### 2012-03-27 exam4/2.tex
Final touches.

@@
 2 & & 22\\ \hline
-3 & & ??\\ \hline
-4 & & ??\\ \hline
+3 & & 15\\ \hline
+4 & & 15\\ \hline
 5 & & 20\\ \hline
 %% Extra & & 5 \\ \hline
-{\bf Total} & & ?? \\ \hline
+{\bf Total} & & 87 \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 \begin{problem}\pts{22}
-
-Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
+\
+\begin{enumerate}
+\item Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
 \verb|visit| method in all of your classes from the previous problem.
@@
 
-Design the \verb|CountVisitor| class, which counts the number of
+\item Design the \verb|CountVisitor| class, which counts the number of
 leaves in a 2/3-tree (there are 6 in the given example.)
@@
 
-Design the \verb|LongestVisitor| class, which produces the longest
+\item Design the \verb|LongestVisitor| class, which produces the longest
 \verb|String| in a 2/3-tree.  The longest string in the example is
 \verb|"Rhinocerous"|.  
+\end{enumerate}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{??}
+\begin{problem} \pts{15}
 
@@
 \newpage
-\noindent
-Part (A): Implement the omitted {\tt same} method that compares this
+\begin{enumerate}
+\item Implement the omitted {\tt same} method that compares this
   unordered pair to a given unordered pair.  This method should work
   regardless of the order of the pair, so for example {\tt new
-  UnPair<Integer>(3,4)} is the same (according to {\tt same}) as {\tt
-  new UnPair<Integer>(4,3)}.
+    UnPair<Integer>(3,4)} is the same (according to {\tt same}) as
+  {\tt new UnPair<Integer>(4,3)}.
 
 \newpage
-\noindent
-Part (B): Assuming {\tt same} and {\tt equals} work as expected, is
+\item Assuming {\tt same} and {\tt equals} work as expected, is
 the given {\tt hashCode} method valid?  That is, does it satisfy or
@@
 {\tt hashCode}.
+\end{enumerate}
 \end{problem}
@@
 
-\begin{problem} \pts{??}
+\begin{problem} \pts{15}
 
@@
 
+\newpage
+[Here is some more space for the previous problem.]
+
 \end{problem}
@@
 limited to iterating over the contents of data structures such as {\tt
-ArrayList}s.  We can also construct iterators directly, for sequences
+  ArrayList}s.  We can also construct iterators directly for sequences
 we are interested in.
@@
 
-Design the class {\tt RangeIterator}, which takes two {\tt Integer}s
+\begin{enumerate}
+\item Design the class {\tt RangeIterator}, which takes two {\tt Integer}s
 as constructor arguments, and iterates over all of the {\tt Integer}s
@@
   T next();
-  // The Java definition of Iterator includes the remove method.
-  // You do *not* need to implement remove.
+  // Java requires an Iterator to define remove,
+  // but you do *not* need to implement remove.
 }
 \end{verbatim}
+
+\newpage
+[Here is some more space for the previous problem.]
 
@@
 
-\noindent
+\item
 Recall our definition of function objects:
@@
 
-Using this interface, design and implement the {\tt MapIterator}
+\item Using this interface, design and implement the {\tt MapIterator}
 class, which takes a \verb|Function| and an \verb|Iterator| as
@@
 \noindent
-Now design the method \verb|sumSquares|, which takes two
+\item Now design the method \verb|sumSquares|, which takes two
 \verb|Integer|s, and uses the classes you've just defined, along with
@@
 \fi
-
+\end{enumerate}
 \end{problem}

### 2012-03-27 exam4/2.tex
all java

@@
 \begin{itemize}
-\item You may use the usual primitives and expression forms of any of
-  the \code{class} languages; for everything else, define it.
+\item All problems on this exam are in Java. You may use the the
+  primitives and expression forms of Java, and the libraries we have
+  used in class and in lab; anything else must be defined.
 
 \item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} as
-  shorthand for writing both {\tt check-expect} and {\tt
-    Tester.checkExpect}.
+  shorthand for writing {\tt
+    Tester.checkExpect}; the {\tt Examples} class and {\tt test}
+  method around the tests are not required.
 

### 2012-03-27 exam4/2.tex
Space.

@@
 \begin{itemize}
-\item All problems on this exam are in Java. You may use the the
-  primitives and expression forms of Java, and the libraries we have
+\item All problems must be done in Java. You may use any Java we have
   used in class and in lab; anything else must be defined.

### 2012-05-06 web/12.su.reg/exams/exam2a.tex
Copy for upcoming class.

\documentclass[11pt]{article}

%% ------------------------------------------------------------------
%% SOLUTIONS:
\def\thel{\noindent\rule{2.5cm}{.5pt}}
\long\def\begsol#1 #2\endsol{#1}
\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
%%uncomment the above line to get the solutions printed

\newcommand\code[1]{\texttt{#1}}
\newcommand\bcode[1]{\texttt{\textbf{#1}}}

%% PROBLEMS:
\def\pts#1{\marginpar{\footnotesize \raggedright  \fbox{#1 {\sc Points}}}}
\newcounter{Pctr}
\newenvironment{problem}{\stepcounter{Pctr}%
\begin{description}
\item[\noindent{\bf Problem} \arabic{Pctr}] 
\end{description}}{\relax}
%% ------------------------------------------------------------------

%% SUBPARTS:
\newcounter{parts}
\renewcommand{\theparts}{\Alph}
%% ------------------------------------------------------------------

\begin{document}

\renewcommand{\theenumi}{\Alph{enumi}}
\setcounter{Pctr}{0}

%% ------------------------------------------------------------------

\vfill
\centerline{\Large CS 2510 Exam 2 -- Spring 2012}

~\\[2cm]

\begin{center}
\begin{tabular}{l@{\qquad}l}
Name:                        & \rule{200pt}{.1pt} \\[.5cm]
Student Id (last 4 digits):  & \rule{200pt}{.1pt} \\[.5cm]
\end{tabular}
\end{center}

\noindent\begin{minipage}{7.5cm} $\bullet$ Write down the answers in the
space provided.

$\bullet$ You may use all syntax that you know from \textit{FunJava}
(that is, the parts of Java we have studied in class), although there
are several features you will \emph{not} need.

$\bullet$ When defining methods, you do not need to give a complete
class definition---just indicate in which class your method definition
should be placed.

$\bullet$ For tests you only need to provide the expression that
computes the actual value, connecting it with an arrow to the expected
value. For example \code{s.method() -> true} is sufficient.

$\bullet$ Remember that the phrase ``design a class'' or ``design a
method'' means more than just providing a definition. It means to
design them according to the \textbf{design recipe}.  You are
\textit{not} required to provide a method template unless the problem
specifically asks for one.  However, be prepared to struggle if you
choose to skip the template step.

$\bullet$ We will not answer \textit{any} questions during the exam.

\bigskip

\textit{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{4.5cm}
\begin{tabular}{|c|l|@/r|}
\hline
\textbf{Problem} & Points & \\ \hline
A & & 5\\ \hline
B & & 5\\ \hline
C & & 10\\ \hline
D & & 10\\ \hline
\textbf{Total} &  & 30 \\ \hline
\end{tabular}
\end{minipage}

\vfill\thispagestyle{empty}
\newpage

%% -----------------------------------------------------------------------------
%% 
\pts{34}
\begin{problem}
Amateur radio (aka ham radio) broadcasts work as follow: there are
three kinds of transmitter nodes: those that simply transmit a signal
to another node, those that split a signal and transmit it to two
other nodes, and finally those that are the end points of
communication that don't broadcast any further.  When a broadcast goes
out, the network is set up in such a way that no broadcast goes to the
same node twice (there are no cycles in the network).

The following class diagram represents a radio broadcast network:

\begin{verbatim}
                     +--------------+<-------------------------+
                     |    Node<T>   |<-----------------------+ |
                     +--------------+<---------------------+ | |
                     | T data       |                      | | |
                     +--------------+                      | | |
                     / \   / \   / \                       | | |
                     ---   ---   ---                       | | |
                      |     |     |                        | | |
        +-------------+     |     +-----------+            | | |
        |                   |                 |            | | |
        |                   |                 |            | | |
 +-------------+      +--------------+  +---------------+  | | |
 | Terminal<T> |      | Repeat<T>    |  | Split<T>      |  | | |
 +-------------+      +--------------+  +---------------+  | | |
                   +--| Node<T> next |  | Node<T> left  |--+ | |
                   |  +--------------+  | Node<T> right |----+ |
                   |                    +---------------+      |
                   +-------------------------------------------+
\end{verbatim}

\begin{enumerate}

\newpage
%-----------------------------------------------
\item {\em{(5 points)}}

Here is a graphical view of a local ham radio network:

\begin{verbatim}
           "NEU"
          /     \
         /       \
   "CCIS"         "CE"
      |           / |
   "Clubs"       /  |
    /   \       /   |
"NUACM" "CISters"  "Robotics"
               \    /
                \  /
            "Robocisters"
\end{verbatim}

When a node appears above another with a line between them, it means
the upper node broadcasts to the lower one.  The strings given show
the data each node carries.

Translate the above information into examples of data that represent
the network.

\begsol{\vspace{0.5in}}
\begin{verbatim}
Node<String> rc = new Terminal<String>("Robocisters");
Node<String> rs = new Repeat<String>("Robotics", rc);
Node<String> cs = new Repeat<String>("CISters", rs);
Node<String> ce = new Split<String>("CE", cs, rs);
Node<String> nu = new Terminal<String>("NUACM");
Node<String> cb = new Split<String>("Clubs", nu, cs);
Node<String> is = new Repeat<String>("CCIS", cb);
Node<String> ne = new Split<String>("NEU", is, ce);
\end{verbatim}

\endsol

\newpage
\item  {\em{(5 points)}}

Here is the data representation of another network (this one carries
integers rather than strings):

\begin{verbatim}
Node<Integer> h = new Terminal<Integer>(800);
Node<Integer> b = new Terminal<Integer>(12);
Node<Integer> a = new Split<Integer>(1, h, b);
Node<Integer> z = new Terminal<Integer>(7);
Node<Integer> c = new Split<Integer>(9, z, a);
Node<Integer> j = new Repeat<Integer>(11, a);
Node<Integer> d = new Repeat<Integer>(44, j);
Node<Integer> g = new Split<Integer>(19, c, d);
Node<Integer> t = new Terminal<Integer>(4);
Node<Integer> s = new Split<Integer>(13, t, c);
Node<Integer> k = new Split<Integer>(14, s, g);
\end{verbatim}

Translate the above into a graphical representation of the network
similar to that given in part A.

\begsol{\vspace{0.5in}}
\begin{verbatim}
     14
    /   \
   /     19
  /    /   \
 13   |    44     
 |  \ /    |     
 4   9     11  
     / \  /
    7   1 
       /  \
     800   12
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(10 points)}}

One measure of the networks robustness is the number of paths from 
a node to any terminal.

Design the method {\tt countAllPaths} that computes the number of
paths from this node to a terminal node.

\begsol{\vspace{0.5in}}
\begin{verbatim}
// in Node<T>
// count all possible paths from this node to any terminal
abstract public int countAllPaths();

// in Terminal<T>
public int countAllPaths() {
   return 1;
}

// in Repeat<T>
public int countAllPaths() {
   return this.next.countAllPaths();
}

// in Split<T>
public int countAllPaths() {
   return this.left.countAllPaths() + this.right.countAllPaths();
}

rc.countAllPaths() -> 1
rs.countAllPaths() -> 1
ce.countAllPaths() -> 2
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(10 points)}}

The quality of radio transmissions deteriorate as they are broadcast
across more and more nodes.  We would like to measure the degree of
distortion in a network.

Design the method {\tt distanceTo} that finds the length of the
shortest network route from this node to the node containing the given
data.

\begsol{\vspace{0.5in}}
\begin{verbatim}
// In Node<T>
// compute the length of the path to the node with the given data 
abstract public int distanceTo(T data);

// compute the length of the path to the node with the given data
public int distanceTo(T data){
  if (this.data.equals(data))
    return 1;
  else
    return 1000; // nonsense value - we will never have 1000 nodes
}

// compute the length of the path to the node with the given data
public int distanceTo(T data){
  if (this.data.equals(data))
    return 1;
  else
    return 1 + this.next.distanceTo(data);
}

// compute the length of the path to the node with the given data
public int distanceTo(T data){
  if (this.data.equals(data))
    return 1;
  else
    return 1 + Math.min(this.left.distanceTo(data), 
                        this.right.distanceTo(data));
}

// PLUS TESTS
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\end{enumerate}
\end{problem}
\end{document}

### 2012-05-06 web/12.su.reg/exams/exam2b.tex
Copy for upcoming class.

\documentclass[11pt]{article}

%% ------------------------------------------------------------------
%% SOLUTIONS:
\def\thel{\noindent\rule{2.5cm}{.5pt}}
\long\def\begsol#1 #2\endsol{#1}
\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
%%uncomment the above line to get the solutions printed

\newcommand\code[1]{\texttt{#1}}
\newcommand\bcode[1]{\texttt{\textbf{#1}}}

%% PROBLEMS:
\def\pts#1{\marginpar{\footnotesize \raggedright  \fbox{#1 {\sc Points}}}}
\newcounter{Pctr}
\newenvironment{problem}{\stepcounter{Pctr}%
\begin{description}
\item[\noindent{\bf Problem} \arabic{Pctr}] 
\end{description}}{\relax}
%% ------------------------------------------------------------------

%% SUBPARTS:
\newcounter{parts}
\renewcommand{\theparts}{\Alph}
%% ------------------------------------------------------------------

\begin{document}

\renewcommand{\theenumi}{\Alph{enumi}}
\setcounter{Pctr}{0}

%% ------------------------------------------------------------------

\vfill
\centerline{\Large CS 2510 Exam 2 -- Spring 2012}

~\\[2cm]

\begin{center}
\begin{tabular}{l@{\qquad}l}
Name:                        & \rule{200pt}{.1pt} \\[.5cm]
Student Id (last 4 digits):  & \rule{200pt}{.1pt} \\[.5cm]
\end{tabular}
\end{center}

\noindent\begin{minipage}{7.5cm} $\bullet$ Write down the answers in the
space provided.

$\bullet$ You may use all syntax that you know from \textit{FunJava}
(that is, the parts of Java we have studied in class), although there
are several features you will \emph{not} need.

$\bullet$ When defining methods, you do not need to give a complete
class definition---just indicate in which class your method definition
should be placed.

$\bullet$ For tests you only need to provide the expression that
computes the actual value, connecting it with an arrow to the expected
value. For example \code{s.method() -> true} is sufficient.

$\bullet$ Remember that the phrase ``design a class'' or ``design a
method'' means more than just providing a definition. It means to
design them according to the \textbf{design recipe}.  You are
\textit{not} required to provide a method template unless the problem
specifically asks for one.  However, be prepared to struggle if you
choose to skip the template step.

$\bullet$ We will not answer \textit{any} questions during the exam.

\bigskip

\textit{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{4.5cm}
\begin{tabular}{|c|l|@/r|}
\hline
\textbf{Problem} & Points & \\ \hline
A & & 5\\ \hline
B & & 5\\ \hline
C & & 10\\ \hline
D & & 10\\ \hline
\textbf{Total} &  & 30 \\ \hline
\end{tabular}
\end{minipage}

\vfill\thispagestyle{empty}
\newpage

%% -----------------------------------------------------------------------------
%% 
\pts{34}
\begin{problem}

The Boston power grid consists of a network of connected electric
transformers (like the one that burst into flames last night).  There
are three basic kinds of transformers: those that simply transmit
power to another transformer, those that split electricity to two
other transformers, and those that are the end points of the power
network that don't connect to any further transformers.  When a power
grid is set up, it's done in such a way that no power is transmitted
from any transformer back to itself (there are no cycles in the grid);
cycles cause fires.  Transformers also carry some data to help
identify each point in the grid.

The following class diagram represent power grids:

\begin{verbatim}
                     +--------------+<-------------------------+
                     |    Trans<T>  |<-----------------------+ |
                     +--------------+<---------------------+ | |
                     | T data       |                      | | |
                     +--------------+                      | | |
                     / \   / \   / \                       | | |
                     ---   ---   ---                       | | |
                      |     |     |                        | | |
        +-------------+     |     +-----------+            | | |
        |                   |                 |            | | |
        |                   |                 |            | | |
 +-------------+      +---------------+  +--------------+  | | |
 |   End<T>    |      | Link<T>       |  | Couple<T>    |  | | |
 +-------------+      +---------------+  +--------------+  | | |
                   +--| Trans<T> conn |  | Trans<T> one |--+ | |
                   |  +---------------+  | Trans<T> two |----+ |
                   |                    +---------------+      |
                   +-------------------------------------------+
\end{verbatim}

\begin{enumerate}

\newpage
%-----------------------------------------------
\item {\em{(5 points)}}

Here is a graphical view of a local power grid:

\begin{verbatim}
             "BOS"
            /     \
           /       \
       "Logan"     "BPD"
       /    \       |
 "Seaport" "ICA"   "MFA"
       \    /  \    | 
      "Ferry"   "NEU"
                 |
              "Ruggles"
\end{verbatim}

When a transformer appears above another with a line between them, it
means the upper transformer transmits power to the lower one.  The
strings given show the data each transformer carries.

Translate the above information into examples of data that represent
the grid.

\begsol{\vspace{0.5in}}
\begin{verbatim}
Trans<String> r = new End<String>("Ruggles");
Trans<String> n = new Link<String>("NEU", r);
Trans<String> m = new Link<String>("MFA", n);
Trans<String> f = new End<String>("Ferry");
Trans<String> i = new Couple<String>("ICA", f, n);
Trans<String> s = new Link<String>("Seaport", f);
Trans<String> l = new Couple<String>("Logan", s, i);
Trans<String> p = new Link<String>("BPD", m);
Trans<String> b = new Couple<String>("BOS", l, p);
\end{verbatim}
\endsol

\newpage
\item  {\em{(5 points)}}

Here is the data representation of another grid:

\begin{verbatim}
Node<String> t1 = new Terminal<String>("h");
Node<String> t2 = new Terminal<String>("b");
Node<String> t3 = new Split<String>("a", t1, t2);
Node<String> t4 = new Terminal<String>("z");
Node<String> t5 = new Split<String>("c", t4, t3);
Node<String> t6 = new Repeat<String>("j", t3);
Node<String> t7 = new Repeat<String>("d", t6);
Node<String> t8 = new Split<String>("g", t5, t7);
Node<String> t9 = new Terminal<String>("t");
Node<String> t10 = new Split<String>("s", t9, t5);
Node<String> t11 = new Split<String>("k", t10, t8);
\end{verbatim}

Translate the above into a graphical representation of the grid
similar to that given in part A.

\begsol{\vspace{0.5in}}
\begin{verbatim}
     "k"
    /   \
   /    "g"
  /    /   \
 "s"   |   "d"     
 |  \  /    |     
"t" "c"   "j"  
     / \  /
  "z"   "a" 
       /  \
     "h"  "b"
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(10 points)}}

One measure of the power grids robustness is the number of paths 
between two transformers.

Design the method {\tt countPaths} that computes the number of paths
from this transformer to a tranformer containing the given data.

\begsol{\vspace{0.5in}}

\begin{verbatim}
// in Trans<T>:
// Count the number of paths from this transformer to one containing 
// the given data
abstract public int countPaths(T data);

// in End<T>:
public int countPaths(T data) {
  if (this.data.equals(data)
    return 1;
  else 
    return 0;
}

// in Link<T>
public int countPaths(T data) {
  if (this.data.equals(data)
    return 1;
  else 
    return this.conn.countPaths(data);
}

// in Couple<T>
public int countPaths(T data) {
  if (this.data.equals(data))
    return 1;
  else
    return this.one.countPaths(data) + this.two.countPaths(data)
}
r.countPaths("Ruggles") -> 1
r.countPaths("NEU") -> 0
n.countPaths("Ruggles") -> 1
b.countPaths("Ruggles") -> 2
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(10 points)}}

The efficiency of a power grid can be determined by the longest path
that power is transfered on.

Design the method {\tt maxLength} that finds the length of the longest
connection from this node to any end transformer.

\begsol{\vspace{0.5in}}
\begin{verbatim}
// in Trans<T>:
// Compute the length of the longest path power can travel on.
abstract int countPaths(T data);

// in End<T>:
public int maxLength() {
  return 1;  // or 0.
}

// in Link<T>
public int maxLength(T data) {
  return 1 + this.conn.maxLength();
}

// in Couple<T>
public int countPaths(T data) {
  return 1 + Math.max(this.one.maxLength(), this.two.maxLength());
}

r.maxLength() -> 1
b.maxLength() -> 4
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\end{enumerate}
\end{problem}
\end{document}

### 2012-05-06 web/12.su.reg/assignments/Capitals.java

@@ -1,0 +1,159 @@
+import java.util.*;
+
+interface Capitals{
+	ArrayList<City> capitals = new ArrayList<City>(Arrays.asList{
+		new City("36105", "Montgomery", "AL", 86.310449, 32.32573),
+		new City("85009", "Phoenix", "AZ", 112.128368, 33.456373),
+		new City("72201", "Little Rock", "AR", 92.281939, 34.748342),
+		new City("95819", "Sacramento", "CA", 121.436634, 38.568293),
+		new City("80204", "Denver", "CO", 105.025854, 39.734022),
+		new City("06120", "Hartford", "CT", 72.675807, 41.78596),
+		new City("19901", "Dover", "DE", 75.535983, 39.156639),
+		new City("32304", "Tallahassee", "FL", 84.321132, 30.447752),
+		new City("30306", "Atlanta", "GA", 84.351418, 33.786027),
+		new City("83705", "Boise", "ID", 116.219104, 43.585077),
+		new City("62701", "Springfield", "IL", 89.649531, 39.80004),
+		new City("46219", "Indianapolis", "IN", 86.049533, 39.782092),
+		new City("50314", "Des Moines", "IA", 93.632993, 41.603003),
+		new City("66610", "Topeka", "KS", 95.746061, 38.982213),
+		new City("40212", "Frankfort", "KY", 85.804479, 38.265116),
+		new City("70815", "Baton Rouge", "LA", 91.059558, 30.455809),
+		new City("04330", "Augusta", "ME", 69.766548, 44.323228),
+		new City("21403", "Annapolis", "MD", 76.49103, 38.952394),
+		new City("02115", "Boston", "MA", 71.092215, 42.342706),
+		new City("48823", "Lansing", "MI", 84.476409, 42.738805),
+		new City("55103", "Saint Paul", "MN", 93.121594, 44.960798),
+		new City("39209", "Jackson", "MS", 90.244626, 32.318422),
+		new City("65101", "Jefferson City", "MO", 92.152462, 38.546212),
+		new City("59635", "Helena", "MT", 111.905089, 46.597324),
+		new City("68503", "Lincoln", "NE", 96.676623, 40.823339),
[…]

### 2012-05-06 web/12.su.reg/assignments/FilesDirectories.java

@@ -1,0 +1,133 @@
+// Assignment 4
+// CS 2510 Spring 2012
+// FilesDirectories.java
+
+/*
+             +-----------------------+
+             |  +-----------------+  |
+             |  |                 |  |
+             v  v                 |  |
+         +-------+                |  |
+         | ILoFD |                |  |
+         +-------+                |  |
+            / \                   |  | 
+            ---                   |  |
+             |                    |  |
+      ---------------             |  |
+      |             |             |  |
+  +--------+  +------------+      |  |
+  | MtLoFD |  | ConsLoFD   |      |  |
+  +--------+  +------------+      |  |
+            +-| IFD first  |      |  |
+            | | ILoFD rest |------+  |
+            | +------------+         |
+            |                        |
+            v                        |
+           +-------+                 |
+           | IFD   |                 |
+           +-------+                 |
+               / \                   |
+               ---                   |
+                |                    |
+        -------------------          |
+        |                 |          |
+  +-------------+  +--------------+  |
+  | File        |  | Directory    |  |
+  +-------------+  +--------------+  |
+  | String name |  | String name  |  |
+  | int size    |  | ILoFD fdlist |--+
+  | String kind |  +--------------+  
+  +-------------+ 
+*/
+
+
+// to represent a list of files and directories in a computer file system
+interface ILoFD{
+}
+
+// to represent an empty list of files and directories in a computer file system
+class MtLoFD implements ILoFD{
+  MtLoFD(){}  
+}
+
+// to represent a nonempty list of files and directories in a computer file system
+class ConsLoFD implements ILoFD{
+  IFD first;
+  ILoFD rest;
+  
+  ConsLoFD(IFD first, ILoFD rest){
+    this.first = first;
+    this.rest = rest;
+  }  
+
+}
+
+// to represent files and directories in a computer file system
+interface IFD{
+}
+
+
+// to represent files in a computer file system
+class File implements IFD{
+  String name;
+  int size;
+  String kind;
+  
+  File(String name, int size, String kind){
+    this.name = name;
+    this.size = size;
+    this.kind = kind;
+  }
+}
+
+// to represent a directory in a computer file system
+class Directory implements IFD{
+  String name;
+  ILoFD fdlist;
+  
+  Directory(String name, ILoFD fdlist){
[…]

### 2012-05-06 web/12.su.reg/assignments/Interactions.java

@@ -1,0 +1,57 @@
+/**
+ * Interactions driver for the Eliza game.
+ * 
+ * @since 19 March 2012 
+ */
+
+import java.io.*;
+
+public class Interactions{
+  
+  /**
+   * Run the program - starting with the <code>eliza</code>
+   * method.
+   * 
+   * @param args unused
+   */
+  public static void main(String[] args) {   
+    
+    Interactions i = new Interactions();
+    i.eliza();
+  }
+  
+  /**
+   * Run the Eliza game
+   */ 
+  public void eliza(){
+  
+    BufferedReader input =
+      new BufferedReader(new InputStreamReader(System.in));
+    
+    // define data you need to play the game ... REPLACE THIS WITH YOUR CODE !!!
+    // ...
+    
+    System.out.println("Ask me a question. \n:>");
+    try{
+      String s = input.readLine();
+      while (!(s == null) && s.length() > 0){
+
+        
+        // mock code: echo every line
+        System.out.println(s);   // REPLACE THIS WITH YOUR CODE !!!
+                                 // that finds out the reply to the question s
+                                 // and prints the reply
+        
+        System.out.println(":>");
+        s = input.readLine();
+        if (s == null || s.length() == 0){
+          System.out.println("Goodbye");
+        }
+      }
+    }
+    catch(IOException e){
+      System.out.println("Goodbye");
+    }  
+  }
+}
+

### 2012-05-06 web/12.su.reg/blog/EqualDoubleDispatch.java

@@ -1,0 +1,177 @@
+/***********************************************
+ *  CS2510 Spring 2012 
+ *  Structural equality via double-dispatch
+ ***********************************************/
+
+import tester.*;
+
+// Represents a Location
+class Loc {
+    Integer x;
+    Integer y;
+    
+    Loc(Integer x, Integer y){
+        this.x = x;
+        this.y = y;
+    }
+    
+    // Is this Loc the same as the given one
+    Boolean sameLoc(Loc that){
+        return (that.x == this.x &&
+                that.y == this.y);
+    }
+}
+
+
+// Represents a Shape
+interface IShape {
+    // Is that shape the same as this Shape
+    public Boolean sameShape(IShape that);
+
+    // Is that Circle the same as this Shape
+    public Boolean sameCirc(Circ that);
+    // Is that Rect the same as this Shape
+    public Boolean sameRect(Rect that);
+    // Is that Combo the same as this Shape
+    public Boolean sameCombo(Combo that);
+}
+
+// Represents a Circle
+class Circ implements IShape {
+    Loc loc;
+    Integer rad;
+
+    Circ(Loc loc, Integer rad) {
+        this.loc = loc; this.rad = rad;
+    }
+    
+    /** Template:
+     *    Fields:
+     *      ... this.loc ... -- Loc
+     *      ... this.rad ... -- int
+     *      
+     *    Methods:
+     *      ... this.sameShape(IShape) ... -- Boolean
+     *      ... this.sameCirc(Circ) ... -- Boolean
+     *      ... this.sameRect(Rect) ... -- Boolean
+     *      ... this.sameCombo(Combo) ... -- Boolean
+     *      
+     *    Methods of Fields:
+     *      ... this.loc.sameLoc() ... -- Boolean
+     */
+
+    // Is that Shape the same as this Circ
+    public Boolean sameShape(IShape that) {
+        return that.sameCirc(this);
+    }
+    // Is that Circle the same as this Circ
+    public Boolean sameCirc(Circ that) {
+        return (this.loc.sameLoc(that.loc) &&
+                this.rad == that.rad);
+    }
+    // Is that Rect the same as this Circ
+    public Boolean sameRect(Rect that) {
+        return false;
+    }
+    // Is that Combo the same as this Circ
+    public Boolean sameCombo(Combo that) {
+        return false;
+    }
+}
+
+// Represents a Rectangle
[…]

### 2012-05-06 web/12.su.reg/blog/Marathon.java

@@ -1,0 +1,289 @@
+import tester.*;
+
+// Interface for two kinds of functional objects:
+// - predicates on runners
+// - comparison between runners
+
+// Represents a predicate on Runners.
+interface RunnerQuestion {
+    // Ask this question of the runner.
+    Boolean ask(Runner r);
+}
+
+// Represent a comparison between two Runners.
+interface RunnerComparison {
+    // Is r1 "better" than r2?
+    Boolean compare(Runner r1, Runner r2);
+}
+
+// Represents the comparison: is r1 faster than r2?
+class IsFaster implements RunnerComparison {
+    // Is r1 faster than r2?
+    public Boolean compare(Runner r1, Runner r2) {
+	return r1.time < r2.time;
+    }
+}
+
+class HasLongerName implements RunnerComparison {
+    // Does r1 have a longer name than r2?
+    public Boolean compare(Runner r1, Runner r2) {
+	return r1.name.length() > r2.name.length();
+    }
+}
+
+class HasLowerBibNumber implements RunnerComparison {
+    // Does r1 have a lower bib number than r2?
+    public Boolean compare(Runner r1, Runner r2) {
+	return r1.bib < r2.bib;
+    }
+}
+
+// Represents the predicate: is given runner older than 50?
+class IsOld implements RunnerQuestion {
+    Integer age;
+
+    // By default, old is > 50 years.
+    IsOld() {
+	this(50);
+    }
+
+    IsOld(Integer age) {
+	this.age = age;
+    }
+
+    // Is the runner's age greater than this's age?
+    public Boolean ask(Runner r) {
+	return r.age > this.age;
+    }
+}
+
+// Represents the predicate: is given runner's time 130 or better?
+class IsFast implements RunnerQuestion {
+    public Boolean ask(Runner r) {
+	return r.time <= 130;
+    }
+}
+
+// Represents the predicate: is given runner named "Shorter"?
+class IsNamedShorter implements RunnerQuestion {
+    public Boolean ask(Runner r) {
+	return r.name.equals("Shorter");
+    }
+}
+
+// Represents the predicate: if given runner is old, are they fast?
+class IsOldFast implements RunnerQuestion {
+    public Boolean ask(Runner r) {
+	return r.age <= 50 || r.time <= 200;
+    }
+}
+
+// Represents a list of runners
+interface ILoR {    
[…]

### 2012-05-06 web/12.su.reg/labs/ExamplesArrayList.java

@@ -1,0 +1,117 @@
+import tester.*;
+import java.util.*;
+
+/**
+ * A class designed to explore and validate the methods defined in the 
+ * <code>ArrayList</code> class.
+ * 
+ * @since 13 March 2012
+ */
+class ExamplesArrayList{
+	ExamplesArrayList(){}
+	
+	/** A sample <code>ArrayList</code> */
+	ArrayList<Person> arlist = new ArrayList<Person>();
+	
+	Person pete = new Person("Pete", 23);
+	Person jill = new Person("Jill", 19);
+	Person kim = new Person("Kim", 20);
+	Person pat = new Person("Pat", 21);
+	
+	/**
+	 * Test the method newAge for the class Person
+	 *
+	 * @param t The tester that runs the tests and reports the results
+	 */
+	public void testNewAge(Tester t){
+		t.checkExpect(this.jill.newAge(25), new Person("Jill", 25));
+	}
+	
+	/**
+	 * Test the method birthday for the class Person
+	 *
+	 * @param t The tester that runs the tests and reports the results
+	 */
+	public void testBirthday(Tester t){
+		// invoke the method
+		this.jill.birthday();
+		
+		// verify the effects
+		t.checkExpect(this.jill, new Person("Jill", 20));
+	}
+	
+	/**
+	 * EFFECT:
+	 * Initialize the array list with four data items
+	 */
+	public void initList(){
+		this.arlist.clear();
+		this.arlist.add(this.pete);
+		this.arlist.add(this.jill);
+		this.arlist.add(this.kim);
+		this.arlist.add(this.pat);
+	}
+	
+	/**
+	 * EFFECT:
+	 * Run the tests to validate the design of the <code>ArrayList</code> class.
+	 * 
+	 * @param t The tester that runs the tests and reports the results
+	 */
+	public void testArrayList(Tester t){
+		this.initList();
+		t.checkExpect(this.arlist.size(), 4);		
+	}
+	
+	/**
+	 * Run the tests defined in the <code>ExamplesArrayList</code> class.
+	 * 
+	 * @param argv unused
+	 */
+	public static void main(String[] argv){
+		ExamplesArrayList e = new ExamplesArrayList();
+		// run tests and report the results: print all test results, print all data
+		Tester.runReport(e, true, true);
+	}
+}
+
+/**
+ * A class of objects to include in the sample ArrayList
+ */
+class Person{
+	
+	/** the name of this person */
+	String name;
+	
+	/** the age of this person */
+	int age;
+	
+	/**
[…]

### 2012-05-06 web/12.su.reg/labs/Files.java

@@ -1,0 +1,130 @@
+// CS 2510 Spring 2012
+// Lab 4
+
+import tester.*;
+
+// to represent different files in a computer
+interface IFile{
+  
+  // compute the size of this file
+  int size();
+  
+  // compute the time (in seconds) to download this file
+  // at the given download rate
+  int downloadTime(int rate);
+  
+  // is the owner of this file the same 
+  // as the owner of the given file?
+  boolean sameOwner(IFile that);
+}
+
+// to represent a text file
+class TextFile implements IFile{
+  String name;
+  String owner;
+  int length;   // in bytes
+  
+  TextFile(String name, String owner, int length){
+    this.name = name;
+    this.owner = owner;
+    this.length = length;
+  }
+  
+  // compute the size of this file
+  int size(){
+    return this.length;
+  }  
+  
+  // compute the time (in seconds) to download this file
+  // at the given download rate
+  int downloadTime(int rate){
+    return 0;
+  }
+  
+  // is the owner of this file the same 
+  // as the owner of the given file?
+  boolean sameOwner(IFile that){
+    return true;
+  }
+}
+
+//to represent an image file
+class ImageFile implements IFile{
+  String name;
+  String owner;
+  int width;   // in pixels
+  int height;  // in pixels
+  
+  ImageFile(String name, String owner, int width, int height){
+    this.name = name;
+    this.owner = owner;
+    this.width = width;
+    this.height = height;
+  }
+  
+  // compute the size of this file
+  int size(){
+    return this.width * this.height;
+  }  
+  
+  // compute the time (in seconds) to download this file
+  // at the given download rate
+  int downloadTime(int rate){
+    return 0;
+  }
+  
+  // is the owner of this file the same 
+  // as the owner of the given file?
+  boolean sameOwner(IFile that){
+    return true;
+  }
+}
+
+
+//to represent an audio file
+class AudioFile implements IFile{
+  String name;
+  String owner;
+  int speed;   // in bytes per second
+  int length;  // in seconds of recording time
+    
+  AudioFile(String name, String owner, int speed, int length){
+    this.name = name;
+    this.owner = owner;
+    this.speed = speed;
+    this.length = length;
+  }
+  
+  // compute the size of this file
+  int size(){
+    return this.speed * this.length;
+  }  
+  
+  // compute the time (in seconds) to download this file
[…]

### 2012-05-06 web/12.su.reg/labs/Shapes.java

@@ -1,0 +1,55 @@
+import tester.*;
+
+// to represent a geometric shape
+interface IShape{}
+
+// to represent a circle
+class Circle implements IShape{
+	int x;
+	int y;
+	int rad;
+	
+	Circle(int x, int y, int rad){
+		this.x = x;
+		this.y = y;
+		this.rad = rad;
+	}
+}
+
+// to represent a rectangle
+class Rect implements IShape{
+	int x;
+	int y;
+	int w;
+	int h;
+	
+	Rect(int x, int y, int w, int h){
+		this.x = x;
+		this.y = y;
+		this.w = w;
+		this.h = h;
+	}
+}
+
+// to represent a combined shape
+class Combo implements IShape{
+	IShape top;
+	IShape bot;
+	
+	Combo(IShape top, IShape bot){
+		this.top = top;
+		this.bot = bot;
+	}
+}
+
+class ExamplesShapes{
+	ExamplesShapes(){}
+	
+	IShape circle = new Circle(50, 50, 50);
+	IShape rleft = new Rect(20, 20, 20, 20);
+	IShape rBot = new Rect(20, 60, 60, 20);
+	
+	IShape addMouth = new Combo(this.rBot, this.circle);
+	IShape addLeftEye = new Combo(this.rleft, this.addMouth);
+	IShape face = new Combo(new Rect(60, 20, 20, 20), this.addLeftEye);	
+}

### 2012-05-06 web/12.su.reg/labs/Strings.java

@@ -1,0 +1,69 @@
+// CS 2510 Spring 2012
+// Lab 4
+
+import tester.*;
+
+// to represent a list of Strings
+interface ILoS{
+  // combine all Strings in this list into one
+  String combine();
+}
+
+// to represent an empty list of Strings
+class MtLoS implements ILoS{
+  MtLoS(){}
+  
+  // combine all Strings in this list into one
+  String combine(){
+    return "";
+  }  
+}
+
+// to represent a nonempty list of Strings
+class ConsLoS implements ILoS{
+  String first;
+  ILoS rest;
+  
+  ConsLoS(String first, ILoS rest){
+    this.first = first;
+    this.rest = rest;  
+  }
+
+ /*
+  TEMPLATE
+  FIELDS:
+  ... this.first ...         -- String
+  ... this.rest ...          -- ILoS
+  
+  METHODS
+  ... this.combine() ...      -- String
+  
+  METHODS FOR FIELDS
+  ... this.first.concat(String) ...       -- String
+  ... this.first.compareTo(String) ...    -- int
+  ... this.rest.combine() ...             -- String
+  
+  */
+  
+  // combine all Strings in this list into one
+  String combine(){
+    return this.first.concat(this.rest.combine());
+  }  
+
+}
+
+// to represent examples for lists of strings
+class ExamplesStrings{
+  ExamplesStrings(){}
+  
+  ILoS mary = new ConsLoS("Mary ",
+               new ConsLoS("had ",
+                new ConsLoS("a ",
+                 new ConsLoS("little ",
+                  new ConsLoS("lamb.", new MtLoS())))));
+                          
+  boolean testCombine(Tester t){
+    return 
+    t.checkExpect(this.mary.combine(), "Mary had a little lamb.");
+  }
+}

### 2012-05-06 web/12.su.reg/lect/Calendar.java

@@ -1,0 +1,254 @@
+import tester.*;
+
+// to represent a time of day in hours and minutes
+class ClockTime {
+    Integer minutes;  // must be in the range [0, 24)
+    Integer hours;    // must be in the range [0, 60)
+	
+    ClockTime(Integer hours, Integer minutes) {
+	this.hours = hours;
+	this.minutes = minutes;
+    }
+    
+    /* TEMPLATE
+     * FIELDS:
+     * ... this.hours ...         -- Integer
+     * ... this.minutes ...       -- Integer
+     * METHODS:
+     * ... this.toMinutes() ...   -- Integer
+     * ... this.isBefore() ...    -- Boolean
+     * ...
+     */
+    
+    // convert the time on this clock to total minutes since midnight
+    public Integer toMinutes() {
+	return this.minutes + (60 * this.hours);
+    }
+    
+    // Is this time before that time?
+    public Boolean isBefore(ClockTime that) {
+	return this.toMinutes() <= that.toMinutes();
+    }
+}
+
+// to represent an event in a calendar
+class Event {
+    String name;
+    ClockTime startTime;
+    ClockTime endTime;
+    
+    Event(String name, ClockTime startTime, ClockTime endTime){
+	this.startTime = startTime;
+	this.endTime = endTime;
+    }
+    
+    /* TEMPLATE
+     * FIELDS:
+     * ... this.name ...                    -- String
+     * ... this.startTime ...               -- ClockTime
+     * ... this.endTime ...                 -- ClockTime
+     * METHODS:
+     * ... this.duration() ...              -- Integer
+     * ... this.endsBefore(Event) ...       -- Boolean
+     * ...
+     * METHODS FOR FIELDS:
+     * ... this.startTime.toMinutes() ...   -- Integer
+     * ... this.endTime.toMinutes() ...     -- Integer
+     * ... this.startTime.isBefore(ClockTime) ...    -- Boolean
+     * ... this.endTime.isBefore(ClockTime) ...    -- Boolean
+     */	
+    
+    // compute the duration of this event in minutes
+    public Integer duration() {
+	return this.endTime.toMinutes() - this.startTime.toMinutes();
+    }
+
+    // Does this event end before that event starts?
+    public Boolean endsBefore(Event that) {
+	return (this.startTime.toMinutes() + this.duration())
+	    <= that.startTime.toMinutes();
+    }
+
[…]

### 2012-05-06 web/12.su.reg/lect/EqualGoodCoffee.java

@@ -1,0 +1,79 @@
+/***********************************************
+ *  CS2510 Spring 2012 
+ *  Coffee, done right
+ ***********************************************/
+
+import tester.*;
+
+interface Drink {
+    // Is this drink the same as that drink?
+    Boolean same(Drink that);
+
+    // Is this drink the same as that tea?
+    Boolean sameTea(Tea that);
+    // Is this drink the same as that coffee?
+    Boolean sameCoffee(Coffee that);
+    // Is this drink the same as that latte?
+    Boolean sameLatte(Latte latte);
+}
+
+abstract class ADrink {
+    public Boolean sameTea(Tea that) { return false; }
+    public Boolean sameCoffee(Coffee that) { return false; }
+    public Boolean sameLatte(Latte that) { return false; }
+}
+
+class Tea extends ADrink implements Drink {
+    // Is this tea the same as that drink?
+    public Boolean same(Drink that) {
+	return that.sameTea(this);
+    }
+
+    public Boolean sameTea(Tea that) {
+	return true;
+    }
+}
+
+class Coffee extends ADrink implements Drink {
+    // Is this coffee the same as that drink?
+    public Boolean same(Drink that) {
+	return that.sameCoffee(this);
+    }
+
+    public Boolean sameCoffee(Coffee that) { 
+	return true;
+    }
+}
+
+class Latte extends Coffee {
+    public Boolean same(Drink that) {
+	return that.sameLatte(this);
+    }
+    
+    public Boolean sameLatte(Latte that) {
+	return true;
+    }
+
+    public Boolean sameCoffee(Coffee that) {
+	return false;
+    }
+}
+
+class Examples {
+    Drink t = new Tea();
+    Drink c = new Coffee();
+    Drink l = new Latte();
+    
+    public Boolean testSameDrink(Tester t) {
+	return t.checkExpect(this.t.same(this.t), true)
+	    && t.checkExpect(this.c.same(this.c), true)
+	    && t.checkExpect(this.t.same(this.c), false)
+	    && t.checkExpect(this.c.same(this.t), false);
+    }
+    public Boolean testSameCoffee(Tester t) {
+	return t.checkExpect(this.l.same(this.c), false)  // works!
+	    && t.checkExpect(this.c.same(this.l), false); // still works!
+    }
+
+}
+

### 2012-05-06 web/12.su.reg/lect/Pizza.java

@@ -1,0 +1,181 @@
+import tester.*;
+
+/*
+Make examples of data - including a pizza with at least three toppings 
+   
+Design the method that counts the number of toppings on a pizza
+
+Design the method that determines whether the pizza
+ has the given crust and cheese?
+
+Design the method that will produce a new pizza, same as before, 
+ but with the given crust instead
+*/
+
+interface IPizza{
+  
+  // count the toppings on this pizza
+  public int countToppings();
+  
+  // is this a pizza with the given crust and cheese?
+  public boolean sameCrustCheese(String crust, String cheese);   
+  
+  // make a new pizza just like this one but with the given crust
+  public IPizza newCrust(String crust);
+}
+
+class Crust{
+  String name;
+  
+  Crust(String name){
+    this.name = name;
+  }
+  
+  // is this a pizza with the given crust?
+  public boolean sameCrust(String crust){
+    return this.name.equals(crust);
+  }
+}
+
+
+class Cheese implements IPizza{
+  String kind;
+  Crust crust;
+  
+  Cheese(String kind, Crust crust){
+    this.kind = kind;
+    this.crust = crust;
+  }
+
+  // count the toppings on this pizza
+  public int countToppings(){
+    return 0;
+  }
+  
+  // is this a pizza with the given crust and cheese?
+  public boolean sameCrustCheese(String crust, String cheese){
+    if (this.kind.equals(cheese))
+      return this.crust.sameCrust(crust);
+    else
+      return false;
+  }  
+  
+  // make a new pizza just like this one but with the given crust
+  public IPizza newCrust(String crust){
+    return new Cheese(this.kind, new Crust(crust));
+  }
+}
+
+class Topping implements IPizza{
+  IPizza base;
+  
+  Topping(IPizza base){
+    this.base = base;
+  }
+
+  // count the toppings on this pizza
+  public int countToppings(){
+    return 1 + this.base.countToppings();
+  }
+  
+  // is this a pizza with the given crust and cheese?
+  public boolean sameCrustCheese(String crust, String cheese){
+    return
+    this.base.sameCrustCheese(crust, cheese);
+  }  
+  
+  // make a new pizza just like this one but with the given crust
+  public IPizza newCrust(String crust){
+    return new Topping(this.base.newCrust(crust));
+  }
+}
+
+// examples and tests for the IPizza class hierarchy
[…]

### 2012-05-06 web/12.su.reg/lect/bst/ABT.java

@@ -1,0 +1,8 @@
+// Common code to implementations of IBT<X>.
+public abstract class ABT<X> implements IBT<X> {
+
+    // Number of values in this tree.
+    public Integer size() {
+	return this.accept(new Size<X>());
+    }
+}

### 2012-05-06 web/12.su.reg/lect/bst/Alpha.java

@@ -1,0 +1,8 @@
+import java.util.*;
+
+// Represents lexicographic order on strings
+public class Alpha implements Comparator<String> {
+    public int compare(String s1, String s2) {
+	return s1.compareTo(s2);
+    }
+}

### 2012-05-06 web/12.su.reg/lect/bst/BSTExamples.java

@@ -1,0 +1,32 @@
+import tester.*;
+
+class BSTExamples {
+    IBT<String> mt = new Leaf<String>();
+    IBT<String> n1 = new Node<String>("Alice", mt, mt);
+    IBT<String> n2 = new Node<String>("Carol", mt, mt);
+    IBT<String> n3 = new Node<String>("Bobby", n1, n2);
+    IBT<String> n4 = new Node<String>("Bobby", mt, n2);
+    
+    IBST<String> smt = new BST<String>(new Alpha(), mt);
+    IBST<String> s1 = new BST<String>(new Alpha(), n1);
+    IBST<String> s2 = new BST<String>(new Alpha(), n3);
+    IBST<String> s3 = new BST<String>(new Alpha(), n4);
+
+    void testSize(Tester t) {
+	t.checkExpect(s1.size(), 1);
+	t.checkExpect(s2.size(), 3);
+    }
+
+    void testMax(Tester t) {
+	t.checkExpect(s1.max(), "Alice");
+	t.checkExpect(s1.min(), "Alice");
+	t.checkExpect(s2.max(), "Carol");
+	t.checkExpect(s2.min(), "Alice");
+    }
+
+    void testInsert(Tester t) {
+	t.checkExpect(smt.insert("Alice"), s1);
+	t.checkExpect(s3.insert("Alice"), s2);
+    }
+}
+	    

### 2012-05-06 web/12.su.reg/lect/bst/Examples.java

@@ -1,0 +1,22 @@
+import tester.*;
+
+class Examples {
+    IBT<Integer> mt = new Leaf<Integer>();
+    IBT<Integer> n1 = new Node<Integer>(7, mt, mt);
+    IBT<Integer> n2 = new Node<Integer>(8, mt, mt);
+    IBT<Integer> n3 = new Node<Integer>(2, n1, n2);
+    IBT<Integer> n4 = new Node<Integer>(1, mt, mt);
+    IBT<Integer> n5 = new Node<Integer>(4, n3, n1);
+
+    void testSize(Tester t) {
+	t.checkExpect(n1.size(), 1);
+	t.checkExpect(n3.size(), 3);
+	t.checkExpect(n5.size(), 5);
+    }
+
+    void testSizeVisitor(Tester t) {
+	t.checkExpect(n1.accept(new Size<Integer>()), 1);
+        t.checkExpect(n3.accept(new Size<Integer>()), 3);
+	t.checkExpect(n5.accept(new Size<Integer>()), 5);
+    }
+}

### 2012-05-06 web/12.su.reg/lect/bst/IBT.java

@@ -1,0 +1,8 @@
+// Represents a binary tree.
+public interface IBT<X> {
+    // Accept the given visitor and visit the data in this tree.
+    public <T> T accept(IVisitorBT<X,T> v);
+
+    // Number of values in this tree.
+    Integer size();
+}

### 2012-05-06 web/12.su.reg/lect/bst/IVisitorBT.java

@@ -1,0 +1,5 @@
+// Represents a T computation over a binary tree of Xs.
+interface IVisitorBT<X,T> {
+    T visitNode(X val, IBT<X> left, IBT<X> right);
+    T visitLeaf();
+}

### 2012-05-06 web/12.su.reg/lect/bst/IsSorted.java

@@ -1,0 +1,19 @@
+import java.util.*;
+
+public class IsSorted<X> implements IVisitorBT<X,Boolean> {
+    Comparator<X> c;
+    IsSorted(Comparator<X> c) {
+	this.c = c;
+    }
+
+    public Boolean visitLeaf() {
+	return true;
+    }
+
+    public Boolean visitNode(X val, IBT<X> left, IBT<X> right) {
+	return left.accept(this)
+	    && right.accept(this)
+	    && c.compare(left.accept(new RightMostAcc<X>(val)), val) <= 0
+	    && c.compare(val, right.accept(new LeftMostAcc<X>(val))) <= 0;
+    }
+}

### 2012-05-06 web/12.su.reg/lect/bst/Node.java

@@ -1,0 +1,16 @@
+// Represents a node in a binary tree of X.
+public class Node<X> extends ABT<X> {
+    X val;
+    IBT<X> left;
+    IBT<X> right;
+
+    Node(X val, IBT<X> left, IBT<X> right) {
+	this.val = val;
+	this.left = left;
+	this.right = right;
+    }
+
+    public <T> T accept(IVisitorBT<X,T> v) {
+	return v.visitNode(this.val, this.left, this.right);
+    }
+}

### 2012-05-17 web/12.su.reg/exams/exam2.tex
Exam 2.

--- 2012-05-06 web/12.su.reg/exams/exam2b.tex
@@
 \long\def\begsol#1 #2\endsol{#1}
-\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
+%\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
 %%uncomment the above line to get the solutions printed
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Spring 2012}
+\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
 
@@
 
-$\bullet$ You may use all syntax that you know from \textit{FunJava}
-(that is, the parts of Java we have studied in class), although there
-are several features you will \emph{not} need.
-
-$\bullet$ When defining methods, you do not need to give a complete
-class definition---just indicate in which class your method definition
-should be placed.
+$\bullet$ You may use all syntax of Java that we have studied in
+class.
 
@@
 
-$\bullet$ We will not answer \textit{any} questions during the exam.
-
 \bigskip
@@
 \end{minipage}\hfil\begin{minipage}[t]{4.5cm}
-\begin{tabular}{|c|l|@/r|}
+\begin{tabular}{|c|l@{\qquad\qquad}|r|}
 \hline
-\textbf{Problem} & Points & \\ \hline
-A & & 5\\ \hline
-B & & 5\\ \hline
-C & & 10\\ \hline
-D & & 10\\ \hline
-\textbf{Total} &  & 30 \\ \hline
+\textbf{Score} &  & 30 \\ \hline
 \end{tabular}
@@
 %% 
-\pts{34}
+\pts{30}
 \begin{problem}
 
-The Boston power grid consists of a network of connected electric
-transformers (like the one that burst into flames last night).  There
-are three basic kinds of transformers: those that simply transmit
-power to another transformer, those that split electricity to two
-other transformers, and those that are the end points of the power
-network that don't connect to any further transformers.  When a power
-grid is set up, it's done in such a way that no power is transmitted
-from any transformer back to itself (there are no cycles in the grid);
-cycles cause fires.  Transformers also carry some data to help
-identify each point in the grid.
-
-The following class diagram represent power grids:
+Binary trees are one of the most widely used data structures in
+computer science.  Here are three examples of binary trees of numbers:
 
 \begin{verbatim}
-                     +--------------+<-------------------------+
-                     |    Trans<T>  |<-----------------------+ |
-                     +--------------+<---------------------+ | |
-                     | T data       |                      | | |
-                     +--------------+                      | | |
-                     / \   / \   / \                       | | |
-                     ---   ---   ---                       | | |
-                      |     |     |                        | | |
-        +-------------+     |     +-----------+            | | |
-        |                   |                 |            | | |
-        |                   |                 |            | | |
- +-------------+      +---------------+  +--------------+  | | |
- |   End<T>    |      | Link<T>       |  | Couple<T>    |  | | |
- +-------------+      +---------------+  +--------------+  | | |
-                   +--| Trans<T> conn |  | Trans<T> one |--+ | |
-                   |  +---------------+  | Trans<T> two |----+ |
-                   |                    +---------------+      |
-                   +-------------------------------------------+
+         5        1           3
+                   \         / \
+                    2       4  18
+                              /  \
+                             9    12
+                            /    /
+                           7    3
 \end{verbatim}
 
+Nodes in a binary tree carry data, which in this case are numbers, and
+a node has a left and right subtree.  Binary trees, like lists, may
+contain \emph{any} number of elements, including 5, 8, 19, and 0.
+
 \begin{enumerate}
+\item Design a data definition for binary trees.  It must adequately
+  represent any possible binary tree.
 
-\newpage
-%-----------------------------------------------
-\item {\em{(5 points)}}
+\item Give class diagrams corresponding to the data definition from
+  part A.
 
-Here is a graphical view of a local power grid:
+\item Design a {\tt mirror} method that computes the mirror image of a
+  binary tree.  For example, the mirror image of the above trees would
+  be:
 
 \begin{verbatim}
-             "BOS"
+     5        1           3
+             /           / \
+            2           18  4
                        /  \
-           /       \
-       "Logan"     "BPD"
-       /    \       |
- "Seaport" "ICA"   "MFA"
-       \    /  \    | 
-      "Ferry"   "NEU"
-                 |
-              "Ruggles"
+                     12    9
+                       \    \
+                        3    7
 \end{verbatim}
+\end{enumerate}
 
-When a transformer appears above another with a line between them, it
-means the upper transformer transmits power to the lower one.  The
-strings given show the data each transformer carries.
-
-Translate the above information into examples of data that represent
-the grid.
-
-\begsol{\vspace{0.5in}}
-\begin{verbatim}
-Trans<String> r = new End<String>("Ruggles");
-Trans<String> n = new Link<String>("NEU", r);
-Trans<String> m = new Link<String>("MFA", n);
-Trans<String> f = new End<String>("Ferry");
-Trans<String> i = new Couple<String>("ICA", f, n);
-Trans<String> s = new Link<String>("Seaport", f);
-Trans<String> l = new Couple<String>("Logan", s, i);
-Trans<String> p = new Link<String>("BPD", m);
-Trans<String> b = new Couple<String>("BOS", l, p);
-\end{verbatim}
-\endsol
-
+\end{problem}
 \newpage
-\item  {\em{(5 points)}}
-
-Here is the data representation of another grid:
-
-\begin{verbatim}
-Node<String> t1 = new Terminal<String>("h");
-Node<String> t2 = new Terminal<String>("b");
-Node<String> t3 = new Split<String>("a", t1, t2);
-Node<String> t4 = new Terminal<String>("z");
-Node<String> t5 = new Split<String>("c", t4, t3);
-Node<String> t6 = new Repeat<String>("j", t3);
-Node<String> t7 = new Repeat<String>("d", t6);
-Node<String> t8 = new Split<String>("g", t5, t7);
-Node<String> t9 = new Terminal<String>("t");
-Node<String> t10 = new Split<String>("s", t9, t5);
-Node<String> t11 = new Split<String>("k", t10, t8);
-\end{verbatim}
-
-Translate the above into a graphical representation of the grid
-similar to that given in part A.
-
-\begsol{\vspace{0.5in}}
-\begin{verbatim}
-     "k"
-    /   \
-   /    "g"
-  /    /   \
- "s"   |   "d"     
- |  \  /    |     
-"t" "c"   "j"  
-     / \  /
-  "z"   "a" 
-       /  \
-     "h"  "b"
-\end{verbatim}
-\endsol
-
-\newpage
-%-----------------------------------------------
-\item {\em{(10 points)}}
-
-One measure of the power grids robustness is the number of paths 
-between two transformers.
-
-Design the method {\tt countPaths} that computes the number of paths
-from this transformer to a tranformer containing the given data.
-
-\begsol{\vspace{0.5in}}
-
-\begin{verbatim}
-// in Trans<T>:
-// Count the number of paths from this transformer to one containing 
-// the given data
-abstract public int countPaths(T data);
-
-// in End<T>:
-public int countPaths(T data) {
-  if (this.data.equals(data)
-    return 1;
-  else 
-    return 0;
-}
-
-// in Link<T>
-public int countPaths(T data) {
-  if (this.data.equals(data)
-    return 1;
-  else 
-    return this.conn.countPaths(data);
-}
-
-// in Couple<T>
-public int countPaths(T data) {
-  if (this.data.equals(data))
-    return 1;
-  else
-    return this.one.countPaths(data) + this.two.countPaths(data)
-}
-r.countPaths("Ruggles") -> 1
-r.countPaths("NEU") -> 0
-n.countPaths("Ruggles") -> 1
-b.countPaths("Ruggles") -> 2
-\end{verbatim}
-\endsol
-
-\newpage
-%-----------------------------------------------
-\item {\em{(10 points)}}
-
-The efficiency of a power grid can be determined by the longest path
-that power is transfered on.
-
-Design the method {\tt maxLength} that finds the length of the longest
-connection from this node to any end transformer.
-
-\begsol{\vspace{0.5in}}
-\begin{verbatim}
-// in Trans<T>:
-// Compute the length of the longest path power can travel on.
-abstract int countPaths(T data);
-
-// in End<T>:
-public int maxLength() {
-  return 1;  // or 0.
-}
-
-// in Link<T>
-public int maxLength(T data) {
-  return 1 + this.conn.maxLength();
-}
-
-// in Couple<T>
-public int countPaths(T data) {
-  return 1 + Math.max(this.one.maxLength(), this.two.maxLength());
-}
-
-r.maxLength() -> 1
-b.maxLength() -> 4
-\end{verbatim}
-\endsol
-
-\newpage
-%-----------------------------------------------
-\end{enumerate}
-\end{problem}
+\ 
 \end{document}

### 2012-05-24 web/12.su.reg/exams/exam3.su.12.tex
Rename to avoid clash with PDF file.

--- 2012-05-17 web/12.su.reg/exams/exam2.tex
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 3 -- Summer 2012}
 
@@
 \hline
-\textbf{Score} &  & 30 \\ \hline
+\textbf{Score} &  & 45 \\ \hline
 \end{tabular}
@@
 %% 
-\pts{30}
+\pts{15}
 \begin{problem}
-
-Binary trees are one of the most widely used data structures in
-computer science.  Here are three examples of binary trees of numbers:
+We've seen multiple instances of objects representing functions, such
+as predicates and comparison functions.  The general idea for
+representing a function with contract {\tt A -> B} is to use an object
+with an {\tt apply} method that consumes an {\tt A} and produces a
+{\tt B}:
 
 \begin{verbatim}
-         5        1           3
-                   \         / \
-                    2       4  18
-                              /  \
-                             9    12
-                            /    /
-                           7    3
+// Represents a function from A to B.
+interface Fun<A,B> {
+  // Apply this function to given argument.
+  B apply(A a);
+}
 \end{verbatim}
 
-Nodes in a binary tree carry data, which in this case are numbers, and
-a node has a left and right subtree.  Binary trees, like lists, may
-contain \emph{any} number of elements, including 5, 8, 19, and 0.
-
-\begin{enumerate}
-\item Design a data definition for binary trees.  It must adequately
-  represent any possible binary tree.
-
-\item Give class diagrams corresponding to the data definition from
-  part A.
-
-\item Design a {\tt mirror} method that computes the mirror image of a
-  binary tree.  For example, the mirror image of the above trees would
-  be:
+Design an implementation of {\tt Fun<Posn,Double>} that calculates the
+distance from the origin to the posn.  You may assume the following
+data definition for {\tt Posn} (with the usual constructor):
 
 \begin{verbatim}
-     5        1           3
-             /           / \
-            2           18  4
-                       /  \
-                     12    9
-                       \    \
-                        3    7
+// Represents a point on a Cartesian plane
+class Posn {
+  Integer x;
+  Integer y;
+}
 \end{verbatim}
-\end{enumerate}
 
+Hint: the distance between two points is given by
+$\sqrt{(x_2-x_1)^2+(y_2-y_1)^2}$ and {\tt Math.sqrt} and {\tt
+  Math.sqr} calculate what their names suggest.
+
+\end{problem}
+\newpage
+\pts{15}
+\begin{problem}
+Assume the following definition for {\tt ListVisitor<X,R>}:
+
+\begin{verbatim}
+// Represents a computation over a list of X producing an R.
+interface ListVisitor<X,R> {
+  R visitEmpty();
+  R visitCons(X first, List<X> rest);
+}
+\end{verbatim}
+
+Design an implementation of {\tt ListVisitor<X,List<Y>>} that maps a
+given function (of type {\tt Fun<X,Y>}) over a list of {\tt X} to
+produce a list of {\tt Y}.
+
+\end{problem}
+\newpage
+
+\pts{15}
+\begin{problem}
+Design an implementation of {\tt ListVisitor<X,X>} that produces the
+first element of a non-empty list that minimizes the result of a given
+{\tt Fun<X,Double>} function.  It should signal an error if the list
+is empty.
+
+For example, if given the distance function from problem 1 and a list
+of {\tt Posn}s representing (4,5), (2,3), and (1,2), it should produce
+the posn representing (1,2) since it closest to the origin (it
+minimizes the distance function).
+
+Hint: you should use an accumulator-based design and therefore may
+need to develop two visitors.
 \end{problem}

### 2012-05-25 admin/charter/CS2500.tex
Initial version of course charters.

\documentclass[11pt]{article}

\usepackage{fullpage}

\title{CS~2500: Course Charter \& Syllabus}

\author{Jay Aslam, Will Clinger, David Van Horn \\
        College of Computer and Information Science \\
        Northeastern University}

\date{February 21, 2012}

\begin{document}

\maketitle

\section{CS~2500 Course Charter}

\paragraph{Course Title:}
%
CS~2500 \emph{Fundamentals of Computer Science 1} (4 SH)

\paragraph{Course Description:}
%
Introduces the fundamentals of program design via study of increasingly
complex data definitions: structures, containment, unions, self-referential,
mutually referential. In addition, the course studies design via generative
recursion, accumulator-style functions, and mutation of variables and
structures. The course also introduces models of computations, which explain
what programs do when they are applied to inputs.

\paragraph{Prerequisites:}
%
The course requires no previous experience in writing computer programs, and
only assumes mathematics at the level of high-school Algebra. Students are
expected to have an interest in learning about Computer Science in depth.

\paragraph{Textbooks:}
%
\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, 
MIT Press, 2000.

\paragraph{Topics Covered:}
%
Computers and the nature of computation. (Formal) models of computation.
Functions; function definitions. Conditional computation structures. Structure
definitions. Modeling information with data. The structure of data:
composite, containment, union, self-referential, mutually referential,
cyclically defined. Program design and design recipes: design based on the
structure of data. Extensional equality of structured data. (Automated) testing.
Accumulators. Designing abstractions. Generative recursion.

\paragraph{Course Outcomes:}
%
Upon completion of this course, a student should be able to \emph{design programs},
and, consequently, have some understanding of the fundamentals of computation.
In particular, the student should be able to: Complete data analysis and
definition for problems that require the use of composition, containment,
union, and self-reference. Design programs to solve problems that consume and
produce data structures based on composition, containment, union, and
self-reference. Design test suites for these programs and evaluate the
outcomes. Design simple abstractions and test the design. Reason about the
performance characteristics of programs in a basic way.

\paragraph{Measurement of Course Outcomes:}
%
Outcomes are measured and verified through: Weekly programming homeworks.
Daily short paper-and-pencil quizzes. Two to four major exams. Inclusion of
student's work in an electronic portfolio.

\pagebreak

\section{CS~2500 Sample Syllabus}

The course sequences through the following topics:
\begin{itemize}
\item overview, intro to simple arithmetic calculation
\item atomic \& compound data: 
  \begin{itemize}
    \item integers
    \item booleans \& conditional computation
    \item records
  \end{itemize}
\item representation: information vs. data
\item data-driven programming
\item testing
\item unions
\item self-referential unions 1---recursion
\item self-referential unions 2---list processing
\item designing larger programs: top-down design; iterative refinement
\item recursion in all forms: trees, s-expressions, forests
\item abstraction: interfaces and implementations
\item abstraction: parametric data definitions and contracts
\item notions of equality
\item local name scope; simple asymptotic complexity
\item abstracting control with higher-order functions
\item constructing higher-order functions with lambda
\item infinite data structures and computations
\item generative recursion: root finding, sorting
\item generative recursion: graphs \& search; accumulators
\item review and wrap up
\end{itemize}

\end{document}

### 2012-05-25 admin/charter/CS2510.tex
Initial version of course charters.

\documentclass[11pt]{article}

\usepackage{fullpage}

\title{CS~2510: Course Charter \& Syllabus}

\author{Jay Aslam, Will Clinger, David Van Horn \\
        College of Computer and Information Science \\
        Northeastern University}

\date{February 21, 2012}

\begin{document}

\maketitle

\section{CS~2510 Course Charter}

\paragraph{Course Title:}
%
CS~2510 \emph{Fundamentals of Computer Science 2}	(4 SH)

\paragraph{Course Description:}
%
Introduces the fundamentals of the design of class hierarchies in object-oriented programming with emphasis on the use of abstractions to design reusable code. The course also covers basic reasoning techniques concerning the algorithmic behavior of functions and methods, especially for linear structures (e.g., lists, stacks, queues) and value-dependent structures (e.g., binary search trees, hash tables).

\paragraph{Prerequisites:}
%
CS~2110. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2500 or concurrently.

\paragraph{Textbooks:}
%
\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000. \emph{How to Design Class Hierarchies}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.

\paragraph{Topics Covered:}
%
Classes, class containment, class hierarchies: data definitions and methods.
%
Designing abstractions and reusable code: interfaces, abstract classes, subclasses, function objects, abstracting over traversals, polymorphism, abstract data type, algorithms parametrized over the data type and functional behavior. 
%
Equality and unit test design.
%
Introduction to algorithm complexity illustrated on linear structures and their algorithms: lists, queues, stacks; direct access structures and their algorithms: sorting and searching, value-dependent (parametric) data structures and related algorithms: binary search trees, heaps, hash tables.
%
Composite structures and algorithms: graphs, graph traversals and searches, backtracking.
%
Program re-use: use of existing libraries.
%
Designing user interactions: the Model-View-Control architecture.
%
Exposure to design patterns: Chain of Responsibility, Template and Hook, Visitor, Iterator, Command, State, Adapter

\paragraph{Course Outcomes:}
%
Upon completion of this course, a student should be able to: 
(1) Design, implement, and test class hierarchies with appropriate methods. 
(2) Read the documentation of class libraries and write programs that use them appropriately. 
(3) Understand the algorithmic behavior of individual methods.

\paragraph{Measurement of Course Outcomes:}
%
Weekly programming homeworks. Final project. Weekly short paper and pencil quizzes. Two to four major exams. Inclusion of student's work in an electronic portfolio.

\section{CS~2510 Sample Syllabus}

The course studies the class-based program design and the design of abstractions that support the design of reusable software and libraries. It covers the principles of object oriented program design, the basic rules of program evaluation, the use of abstractions in support of the design of reusable code, and examines the relationship between algorithms and data structures, as well as basic techniques for analyzing algorithm complexity.

The course is suitable for both CS majors and non-majors. It assumes that student has been introduced to the basic principles of program design and computation.

\paragraph{Course Objectives:}
%
The goal is to help the student understand the principles of class based program design using an object-oriented programming language(s)---not just Java. Java-like syntax is used early, with the transition to standard Java later in the course, to illustrate how the program design principles are used in practical applications. Using a concrete language also presents an opportunity to discuss the strengths and weaknesses of languages and language paradigms.

\paragraph{MODULE 1:}
%
Review of Fundies 1 in the context of a class-based statically typed language (2 weeks)
%
\begin{itemize}
\item Review of the program design and abstractions in functional style;
  design recipe, abstractions, loops with accumulators
\item Defining classes and class hierarchies to represent complex information:
  simple classes, classes with fields that are instances of another class, 
  unions of classes, self-referential and mutually referential class   
  hierarchies: lists, trees, graphs
\item Designing methods for classes: static type system, dynamic dispatch,
  delegation (chain of responsibility)
\item Design Recipe for data definitions, for method definitions
\end{itemize}

\paragraph{MODULE 2:}
%
Designing Abstractions (3 weeks)
%
\begin{itemize}
\item Review of the Design Recipe for Abstractions
\item Abstracting over the data: abstract classes; subclasses
\item Abstracting over the functional behavior: function objects, functional iterator
\item Understanding equality of complex data
\end{itemize}

\paragraph{MODULE 3:}
%
Stateful Programming (3 weeks)
%
\begin{itemize}
\item Designing methods with effects; Designing tests for effects
\item Understanding equality in the presence of mutation
\item Direct access data structures: (e.g. Java ArrayList)
\item Loops
\end{itemize}

\paragraph{MODULE 4:}
%
Abstracting over the Data Type; Designing and using libraries (6 weeks)
%
\begin{itemize}
\item Polymorphism: parametrized classes, methods
\item Abstract Data Type --- stack, queue, list, set
\item Abstract Data Type --- map, hashmap
\item HashCode method and overriding equals method
\item Algorithm complexity --- estimates; Stress tests
\item Java Collection Framework
\end{itemize}

\paragraph{Notes:}
%
\begin{itemize}

\item A number of algorithms (filter, andmap, ormap, linear search, binary search, binary search trees, insertion sort, selection sort, quicksort, heap-based priority queue, heapsort, table lookup, use of hashmap, and some graph traversal algorithms (BFS, DFS, and possibly shortest path) are used to illustrate the program design, the need for appropriate data representation, and the need for understanding the complexity of algorithms.

\item Stress tests are used to reinforce the need for understanding the algorithm complexity.

\item Data structures (stacks, queues, lists, trees, arraylist, hashmap, lookup tables) are presented in the context of problems that motivate their use.

\item The goal is to make sure the students begin to understand the design of class based libraries, know how to look up the needed information, and can begin using libraries in their programs.

\item A comprehensive final project challenges the students to apply the design principles and the knowledge of the libraries to a creative larger program with up to 1000 LOC.

\end{itemize}

\end{document}

### 2012-05-25 admin/charter/CS2510.tex
Revised charter for 2510.

@@
 %
-Introduces the fundamentals of the design of class hierarchies in object-oriented programming with emphasis on the use of abstractions to design reusable code.
-The course also covers basic reasoning techniques concerning the algorithmic behavior of functions and methods, especially for linear structures (e.g., lists, stacks, queues) and value-dependent structures (e.g., binary search trees, hash tables).
+
+Introduces the fundamentals of the design of class hierarchies in
+object-oriented programming with emphasis on representation-dependent
+program design with objects. The course also covers basic reasoning
+techniques concerning the correctness and algorithmic behavior of
+computations.
 
@@
 %
-CS~2110. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2500 or concurrently.
+CS~2500. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2510 or concurrently.
 
@@
 %
-\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000.
-\emph{How to Design Class Hierarchies}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.
+\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000. \emph{How to Design Classes}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.
 
@@
 %
-Classes, class containment, class hierarchies: data definitions and methods.
+Objects, classes, class hierarchies.  Data definitions and methods.
 %
-Designing abstractions and reusable code: interfaces, abstract classes, subclasses, function objects, abstracting over traversals, polymorphism, abstract data type, algorithms parametrized over the data type and functional behavior.
+Class diagrams.
 %
-Equality and unit test design.
+Interfaces, abstract classes, subclasses.  Types.  Polymorphism.
 %
-Introduction to algorithm complexity illustrated on linear structures and their algorithms: lists, queues, stacks; direct access structures and their algorithms: sorting and searching, value-dependent (parametric) data structures and related algorithms: binary search trees, heaps, hash tables.
+Accumulators, iterators, loops.
 %
-Composite structures and algorithms: graphs, graph traversals and searches, backtracking.
+Function objects, comparators, visitors.  Sorting: quick sort, merge
+sort, heap sort.  Notions of equality.
 %
-Program re-use: use of existing libraries.
+Java collections library.
 %
-Designing user interactions: the Model-View-Control architecture.
+Overriding {\tt equals} and {\tt hashCode}.
 %
-Exposure to design patterns: Chain of Responsibility, Template and Hook, Visitor, Iterator, Command, State, Adapter
+Testing: unit testing, stress testing, and property-based random testing.
+%
+Binary trees, binary search, heaps, hash tables.
+%
+Invariants.
+%
+Graphs, graph traversal, generative recursion.  Depth-first,
+breadth-first, and priority-queue based search.
+%
+Design patterns.
+%
+Mutation and stateful programming.
+%
+Indexed data structures.
 
@@
 %
-The goal is to help the student understand the principles of class based program design using an object-oriented programming language(s)---not just Java.
-Java-like syntax is used early, with the transition to standard Java later in the course, to illustrate how the program design principles are used in practical applications.
-Using a concrete language also presents an opportunity to discuss the strengths and weaknesses of languages and language paradigms.
+The goal is to help the student understand the principles of class
+based program design using an object-oriented programming
+language. Java is used to illustrate how the program design principles
+are used in practical applications. Using a concrete language also
+presents an opportunity to discuss the strengths and weaknesses of
+languages and language paradigms.
 

### 2012-05-31 web/12.su.reg/exams/exam4.su.12.tex
Exam 4

--- 2012-05-17 web/12.su.reg/exams/exam2.tex
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 4 -- Summer 2012}
 
@@
 %% 
-\pts{30}
+\pts{45}
 \begin{problem}
+An association list is a way of representing an association between
+values.  For example, you might you use an association list to
+associate students with their grades, or battleships with their
+position, or bank account numbers with their account balances.
 
-Binary trees are one of the most widely used data structures in
-computer science.  Here are three examples of binary trees of numbers:
+The representation of an association list is much like a typical list,
+except that every ``cons'' has \emph{two} elements and the rest of the
+list.  One element is a ``key'', the other is its association.
+
+Here is a data definition for association lists parameterized over the
+types of keys and their values:
 
 \begin{verbatim}
-         5        1           3
-                   \         / \
-                    2       4  18
-                              /  \
-                             9    12
-                            /    /
-                           7    3
+interface Assoc<K,V> {}
+
+class EmptyAssoc<K,V> implements Assoc<K,V> {}
+
+class ConsAssoc<K,V> implements Assoc<K,V> {
+  K key;
+  V val;
+  Assoc<K,V> rest;
+  ConsAssoc(K key, V val, Assoc<K,V> rest) {
+    this.key = key;
+    this.val = val;
+    this.rest = rest;
+  }
+}
 \end{verbatim}
 
-Nodes in a binary tree carry data, which in this case are numbers, and
-a node has a left and right subtree.  Binary trees, like lists, may
-contain \emph{any} number of elements, including 5, 8, 19, and 0.
-
 \begin{enumerate}
-\item Design a data definition for binary trees.  It must adequately
-  represent any possible binary tree.
-
-\item Give class diagrams corresponding to the data definition from
-  part A.
-
-\item Design a {\tt mirror} method that computes the mirror image of a
-  binary tree.  For example, the mirror image of the above trees would
-  be:
-
-\begin{verbatim}
-     5        1           3
-             /           / \
-            2           18  4
-                       /  \
-                     12    9
-                       \    \
-                        3    7
-\end{verbatim}
+\item Develop the visitor pattern for association lists.
+\item Design an implementation of an association list visitor that
+  looks up the (first) value associated to a given key.  If the key is
+  not in the list, the visitor shoud throw an exception.
+\item Using the double dispatch approach, design a {\tt same} method
+  that determines if this association list is structurally equal to a
+  given association list.
 \end{enumerate}
-
 \end{problem}
@@
 \ 
+\newpage
+\ 
 \end{document}

### 2012-06-14 web/12.su.reg/exams/exam5.su.12.alt.tex
Exam 5 alt, Exam 6, exam solutions, and assignment 6.

--- 2012-05-17 web/12.su.reg/exams/exam2.tex
@@
 \documentclass[11pt]{article}
-
+\usepackage{amsmath}
 %% ------------------------------------------------------------------
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 5 (Alternate) -- Summer 2012}
 
@@
 %% 
-\pts{30}
+\pts{45}
 \begin{problem}
+Develop an implementation of {\tt Comparator<Square>} that orders
+squares by their area in such a way that the smaller the area, the
+\emph{greater} the square is considered.  So for example, there is no
+square greater than the one with width $= 0$ and height $= 0$.
 
-Binary trees are one of the most widely used data structures in
-computer science.  Here are three examples of binary trees of numbers:
+You may rely on the following definition of {\tt Square}:
 
 \begin{verbatim}
-         5        1           3
-                   \         / \
-                    2       4  18
-                              /  \
-                             9    12
-                            /    /
-                           7    3
+// Represents a square with given height and width.
+class Square {
+  Integer width;
+  Integer height;
+  Square(Integer width, Integer height) {
+    this.width = width;
+    this.height = height;
+  }
+}
 \end{verbatim}
+\end{problem}
 
-Nodes in a binary tree carry data, which in this case are numbers, and
-a node has a left and right subtree.  Binary trees, like lists, may
-contain \emph{any} number of elements, including 5, 8, 19, and 0.
+\newpage
+\begin{problem}
+Design a method:
+\begin{center}
+{\tt <T> Boolean noSameNeighbors(ArrayList<T> ls, Comparator<T> c)}
+\end{center}
+that determines if the given array list has no adjacent elements
+(i.e.~neighboring elements) that are considered of equal size
+according to the given comparator.
+\end{problem}
+\newpage
+\begin{problem}
+Design a method:
+\begin{verbatim}
+<A> Comparator<ArrayList<A>> dict(Comparator<A> ca)
+\end{verbatim}
+that consumes a comparator for {\tt A}s and and produces a comparator
+for array lists of {\tt A}s that is the \emph{dictionary order} (aka
+\emph{alphabetic order}).
 
-\begin{enumerate}
-\item Design a data definition for binary trees.  It must adequately
-  represent any possible binary tree.
-
-\item Give class diagrams corresponding to the data definition from
-  part A.
-
-\item Design a {\tt mirror} method that computes the mirror image of a
-  binary tree.  For example, the mirror image of the above trees would
-  be:
-
-\begin{verbatim}
-     5        1           3
-             /           / \
-            2           18  4
-                       /  \
-                     12    9
-                       \    \
-                        3    7
-\end{verbatim}
-\end{enumerate}
-
+In other words, given an ordering on individual elements, produce an
+ordering on \emph{lists of} elements that corresponds to the ordering
+used in dictionaries.  A dictionary uses an ordering on individual
+letters (single elements) to make an ordering on strings of letters
+(lists of elements) so that ``abc'' $<$ ``abcd'', ``abc'' $<$ ``adc'',
+``abc'' $>$ ``ab'', and ``abc'' $=$ ``abc'', etc.
 \end{problem}
+\ 
+\newpage
+\ 
 \newpage

### 2012-06-14 web/12.su.reg/blog/Exam5a.java

@@ -1,0 +1,135 @@
+import tester.*;
+import java.util.Comparator;
+import java.util.ArrayList;
+
+class Square {
+    Integer width;
+    Integer height;
+    Square(Integer width, Integer height) {
+	this.width = width;
+	this.height = height;
+    }
+    // Produce area of this square
+    Integer area() {
+	return this.width * this.height;
+    }
+}
+
+// Compare two squares so that smaller area means "larger".
+class SquareCompare implements Comparator<Square> {
+    public int compare(Square a, Square b) {
+	Integer ad = a.area();
+	Integer bd = b.area();
+	return ad < bd ? 1 : ad.equals(bd) ? 0 : -1;
+    }
+}
+
+class NoSameNeighbors {
+    // Is given array list in ascending sorted order according to c?
+    <T> Boolean noSameNeighbors(ArrayList<T> ls, Comparator<T> c) {
+	if (ls.size() < 2) {
+	    return true;
+	} else {
+	    return noSameAcc(ls, c, 1);
+	}
+    }
+
+    <T> Boolean noSameAcc(ArrayList<T> ls, Comparator<T> c, Integer i) {
+	if (i.equals(ls.size())) {
+	    return true;
+	} else {
+	    return !(c.compare(ls.get(i-1), ls.get(i)) == 0)
+		&& this.noSameAcc(ls, c, i+1);
+	}
+    }
+}
+
+class Dict<A> implements Comparator<ArrayList<A>> {
+    Comparator<A> ca;
+    Dict(Comparator<A> ca) {
+	this.ca = ca;
+    }
+
+    // Compare x and y using dictionary ordering.
+    public int compare(ArrayList<A> x, ArrayList<A> y) {
+        return compareAcc(x, y, 0);
+    }
+
+    int compareAcc(ArrayList<A> x, ArrayList<A> y, Integer i) {
+	if (i.equals(x.size())) {
+	    return x.size() - y.size();
+	} else if (i.equals(y.size())) {
+	    return 1;
+	} else {
+	    int first = ca.compare(x.get(i), y.get(i));
+	    if (first == 0) {
+		return compareAcc(x, y, i+1);
+	    } else {
+		return first;
+	    }
+	}
+    }
+}
+
+class Alg {
+    <A> Comparator<ArrayList<A>> dict(Comparator<A> ca) {
+	return new Dict<A>(ca);
[…]

### 2012-06-14 web/12.su.reg/blog/Exam6.java

@@ -1,0 +1,131 @@
+import java.util.*;
+import tester.*;
+
+class Alg {
+    Double avg(Iterable<Double> nums) {
+	Double sum = 0.0;
+	Double cnt = 0.0;
+	for (Double n : nums) {
+	    sum = sum + n;
+	    cnt = cnt + 1;
+	}
+	if (cnt.equals(0.0)) {
+	    throw new RuntimeException("No nums");
+	} else {
+	    return sum / cnt;
+	}
+    }
+}
+
+class SumSquares implements Iterable<Integer>, Iterator<Integer> {
+    Iterator<Integer> xi;
+    Integer sum;
+    SumSquares(Iterable<Integer> x) {
+	this.xi = x.iterator();
+	this.sum = 0;
+    }
+    public Iterator<Integer> iterator() {
+	return this;
+    }
+    public boolean hasNext() {
+	return this.xi.hasNext();
+    }
+    public Integer next() {
+	Integer n = this.xi.next();
+	this.sum = this.sum + (n * n);
+	return this.sum;
+    }
+    public void remove() {}
+}
+
+class Pair<X,Y> {
+    X left;
+    Y right;
+    Pair(X left, Y right) {
+	this.left = left;
+	this.right = right;
+    }
+}
+
+class Zip<X,Y> implements Iterable<Pair<X,Y>>, Iterator<Pair<X,Y>> {
+    Iterator<X> xi;
+    Iterator<Y> yi;
+    Zip(Iterable<X> x, Iterable<Y> y) {
+	this.xi = x.iterator();
+	this.yi = y.iterator();
+    }
+    public Iterator<Pair<X,Y>> iterator() {
+	return this;
+    }
+    public boolean hasNext() {
+	return this.xi.hasNext()
+	    && this.yi.hasNext();
+    }
+    public Pair<X,Y> next() {
+	return new Pair<X,Y>(this.xi.next(), this.yi.next());
+    }
+    public void remove() {}
+}
+
+// A magic utility
+class Iter<X> implements Iterable<X> {
+    X[] vals;
+    Iter(X... vals) {
+	this.vals = vals;
+    }
+    public Iterator<X> iterator() {
+	return Arrays.asList(this.vals).iterator();
+    }
+}
+
+class Nat implements Iterable<Integer>, Iterator<Integer> {
+    Integer n;
[…]

### 2012-06-14 web/12.su.reg/blog/Exam6.java
Slight tweak to exam 6 solution.

@@ -4,15 +4,15 @@ import tester.*;
 class Alg {
     Double avg(Iterable<Double> nums) {
 	Double sum = 0.0;
-	Double cnt = 0.0;
+	Integer ct = 0;
 	for (Double n : nums) {
 	    sum = sum + n;
-	    cnt = cnt + 1;
+	    ct = ct + 1;
 	}
-	if (cnt.equals(0.0)) {
+	if (ct.equals(0)) {
 	    throw new RuntimeException("No nums");
 	} else {
-	    return sum / cnt;
+	    return sum / ct;
 	}
     }
 }

### 2013-02-12 exam5/1.tex
start on new exam

--- 2012-02-13 exam3/1.tex
@@
 \vspace*{-1.5cm}
-\ifrubric
-\centerline{\Large CSU2510H Exam 1 Rubric -- Spring 2012}
-
-\else
-\centerline{\Large CSU2510H Exam 1 -- Spring 2012}
+\centerline{\Large CSU2510H Exam 1 \ifrubric (SOLUTION) \fi-- Spring 2013}
 
@@
 2 & & 10\\ \hline
-3 & & 15\\ \hline
-4 & & 15\\ \hline
-5 & & 18\\ \hline
 Extra & & 5 \\ \hline
-{\bf Total} & & 76+5 \\ \hline
+{\bf Total} & & 28+5 \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 
-\fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \noindent
-% {\bf Home on the Range}
-
-% Come up with funny story here.
-
-% \newpage
-\begin{problem}\pts{18} % same as on the previous exam :)
-
-Here is a data definition for a class of objects.  It intentionally
-uses meaningless names to test whether you've sufficiently
-internalized the data-driven development process we've talked about
-since the Fall.
-
-\begin{verbatim}
-;; A P is one of:          
-;; - A                  ;; An A is a (new a% String)
-;; - B                  ;; A  B is a (new b% String A)
-;; - C                  ;; A  C is a (new c% P)
-
-(define-class a% (fields w))
-(define-class b% (fields x y))
-(define-class c% (fields z))
-\end{verbatim}
-
-\noindent
-Design the following methods for \verb|P|s:
-
-\begin{enumerate}
-
-\item \verb|size| - the size of an \verb|A| is $5$; the size of a
-  \verb|B| is $8$, regardless of what's inside; and the size of a
-  \verb|C| contributes $1$ to the size of the \verb|P| inside.
-
-\item \verb|replace-x| - produces a \verb|P| like this one, but
-  with any \verb|x| field replaced by the given string.
-
-\item \verb|same-x-and-w?| - given two strings, produce true if this
-  \verb|P| contains a \verb|B| whose \verb|x| is equal to the
-  first given string and whose \verb|y| is an \verb|A| whose
-  \verb|w| is equal to the second given string.  Produces false
-  otherwise.
-
-\end{enumerate}
-
-Be sure to define any interfaces you need and say which classes
-implement which interfaces so that your program treats all values
-according to their interface.
-
-\newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
-\newpage
-[Here is some more space for the previous problem.]
-\newpage
-\fi
-
-\begin{solution}
-\begin{verbatim}
-;;> [6 pts -- contract and purpose for each method]
-;; A P implements:
-;; - size : -> Number
-;;   Size of this P.
-;; - replace-x : String -> P
-;;   Replace x field (if any) of this P with given string.
-;; - same-x-and-w? : String String -> Boolean
-;;   Is there a B within this P with given x field and w field?
-
-;;> [0 pts -- not required, but write something]
-;; An A implements:
-;; - w : -> String
-;;   Get the w string.
-
-;;> [2 pts] In a%: 
-(define (size) 5)
-(define (replace-x str) this)
-(define (same-x-and-w? g c) false)
-
-;;> [3 pts] In b%:
-(define (size) 8)
-(define (replace-x str)
-  (new b% str (this . y)))
-(define (same-x-and-w? g c)
-  (and (string=? g (this . x))
-       (string=? c (this . y . w))))
-
-;;> [4 pts] In c%:
-(define (size) (add1 (this . z . size)))
-(define (replace-x str)
-  (new c% (this . z . replace-x str)))
-(define (same-x-and-w? g c)
-  (this . z . same-x-and-w? g c))
-
-;; tests
-;;> [3 pts -- 1 per method]
-
-\end{verbatim}
-\end{solution}
-\end{problem}
-%\ifrubric\bigskip\else
-\newpage
-%\fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem}\pts{10}
-
-Here are data, class, and interface definitions for a list of numbers:
-\begin{alltt}
-;; An LoN is one of:
-;; - (new mt%)
-;; - (new cons% Number LoN)
-;; and implements:
-;; - sum : -> Number
-;;   Compute the sum of the numbers in this list.
-;; - sqrs : -> LoN
-;;   List of squares of each number in this list.
-
-(define-class mt%
-  (define (sum) 0)
-  (define (sqrs) this))
-
-(define-class cons%
-  (fields first rest)
-  (define (sum)
-    (+ (this . first)
-       (this . rest . sum)))
-  (define (sqrs)
-    (new cons% 
-         (sqr (this . first)) ; (sqr x) = x\(\sp{2}\)
-         (this . rest . sqrs))))
-\end{alltt}
-%
-Here are some examples:
-\begin{verbatim}
-(check-expect ((new mt%) . sum) 0)
-(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum)  9)
-(check-expect ((new mt%) . sqrs) (new mt%))
-(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sqrs)
-              (new cons% 16 (new cons% 25 (new mt%))))
-\end{verbatim}
-
-\newpage
-\noindent
-Develop the \verb|sum-sqrs| method that computes the sum of the
-squares of a list of numbers.  Here is an example:
-\begin{verbatim}
-(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum-sqrs) 
-              (+ 16 25))
-\end{verbatim}
-
-If you can, define a super class of \verb|mt%| and \verb|cons%| and
-lift the definition of \verb|sum-sqrs| to this class.  (We will assume
-the appropriate \verb|super| declaration in \verb|mt%| and
-\verb|cons%|.)  If you cannot, define methods in both \verb|mt%| and
-\verb|cons%| for partial credit.
-
-\begin{solution}
-\begin{verbatim}
-
-;; [2 pts% -- class def]
-(define-class list%
-  ;; [2 pts -- contract and purpose]
-  ;; sum-sqrs : -> Number
-  ;; compute the sum of the squares of the numbers in this list
-  ;; [4 pts -- implementation]
-  (define (sum-sqrs)
-    ((this . sqrs) . sum)))
-
-;; [2 pts -- tests]
-(check-expect ((new mt%) . sum-sqrs) 0)
-\end{verbatim}
-\end{solution}
-
-\newpage
-
-\end{problem}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In order to manage class scheduling, the College decided that they
+% Set/Property problem
+
+In order to manage the distribution of ID , the College decided that they
 needed to create a new implementation of the \verb|Set| data
@@
 \begin{problem} \pts{15}
+
+% JSON problem
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{18}
-Last night, Asumu was up late improving the {\tt class} languages.
-Unfortunately, he was too tired to test his work, and his last commit
-broke the numeric system.
-%
-This puts us in the dire position of no longer being able to calculate
-the factorial of large numbers, and we turn to you for help.
-
-This is the definition of a class of functional objects
-with a {\tt fact} method:
-
-\begin{verbatim}
-;; A Fact is (new fact%) and implements 
-;; - fact : Natural -> Natural
-;;   Compute n!.
-(define-class fact%
-  (define (fact n)
-    (cond [(zero? n) 1]
-          [else (* n (this . fact (sub1 n)))])))
-
-(check-expect ((new fact%) . fact 0) 1)
-(check-expect ((new fact%) . fact 5) 120)
-\end{verbatim}
-
-Since we can no longer rely on built in numbers to represent
-\verb|Natural|s, we will represent \verb|Natural|s with objects and
-then rewrite the \verb|fact| method to operate over this new kind of
-data.
-
-Here is the \verb|Natural| interface:
-
-\begin{verbatim}
-;; A Natural implements:
-;; - zero? : -> Boolean
-;;   is this Natural the zero value?
-;; - sub1 : -> Natural
-;;   produce a natural one less than this natural 
-;;   (or zero if this natural is zero)
-;; - times : Natural -> Natural
-;;   multiply this natural by the given natural
-\end{verbatim}
-
-\newpage
-Write a new definition of the \verb|fact| method using the
-\verb|Natural| interface definition given above, and \emph{not} using
-the built-in numbers of the class system or DrRacket.  You can assume
-there is a defined constant \verb|one|, which is a \verb|Natural|
-object representing $1$.
-
-\begin{verbatim}
-(check-expect ((new fact%) . fact one) one)
-(check-expect ((new fact%) . fact (one . sub1)) one)
-\end{verbatim}
-
-\newpage
-
-Now, design a data defintion that implements \verb|Natural|, and
-supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.
-Define the constant \verb|one|.  (\emph{Hint}: remember that the
-natural numbers can be seen as a recursive union; a natural number is
-either zero, or one more than a natural number.)
-
-It will be helpful to define the helper methods \verb|add1| and
-\verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
-\cdot m = (n-1) \cdot m + n$.  Don't overlook the slightly weird
-specification for \verb|sub1|: $0-1 = 0$; this weirdness is common
-amongst mathematicians when working with natural numbers.
-
-\ifrubric
-\else
-\newpage
-[Here is some more space for the previous problem.]
-\newpage
-\fi
-\end{problem}
-
-\begin{solution}
-\begin{verbatim}
-;; Part 1: [2 pts]
-(define-class fact%
-  (define (fact n)
-    (cond [(n . zero?) one]
-          [else (n . times (this . fact (n . sub1)))])))
-
-;; Part 2:
-;; [2 pts -- data defintion]
-;; A Natural is one of
-;; - (new zero%)
-;; - (new succ% Natural)
-(define-class zero%
-  ;; [4 pts -- methods in zero%]
-  (define (sub1) this)
-  (define (plus n) n)
-  (define (add1) (new succ% this))
-  (define (times n) this)
-  (define (zero?) true))
-
-(define-class succ%
-  (fields pred)
-  ;; [7 pts -- methods in succ%]
-  (define (zero?) false)
-  (define (add1) (new succ% this))
-  (define (plus n) (this . pred . plus n . add1))
-  (define (times n) (this . pred . times n . plus n))
-  (define (sub1) (this . pred)))
-
-(define zero (new zero%))
-;; [1 pt -- define one]
-(define one (new succ% (new zero%)))
-
-;; [3 pts -- tests]
-(check-expect (one . zero?) false)
-(check-expect (one . sub1) zero)
-(check-expect ((new succ% one) . sub1) one)
-(check-expect (zero . zero?) true)
-(check-expect (zero . sub1) zero)
-\end{verbatim}
-\newpage
-\end{solution}
 
 \begin{problem}\pts{5}
+
+% Visitor on Binary Trees
 

### 2013-02-12 exam5/1.tex
problem 1

@@
 {\bf Problem} & Points & out of \\ \hline
-1 & & 18\\ \hline
+1 & & 19\\ \hline
 2 & & 10\\ \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{15}
+\begin{problem} \pts{19}
 
@@
 
-In order to manage the distribution of ID , the College decided that they
-needed to create a new implementation of the \verb|Set| data
-structure.  They went with the lowest bidder, which
-turned out to be a bunch of BU students.  Now, having paid for this
-code, the College needs to test it, and they turn to you.
-
-Unfortunately, the code the BU students produced is so convoluted that
-it doesn't bear looking at.  All you have to go on is that the
-\verb|Set| data structure obeys the following interface:
+In order to manage the distribution of ID cards, the University needs
+a data structure for managing sets of numbers.  Van Horn started on
+the implementation, resulting in what you see below.
 
 \begin{verbatim}
-;; A Set implements:
-;; - insert : String -> Set
-;;   adds given string to this set, if it isn't already there
-;;   (otherwise produces this set)
-;; - remove : String -> Set
-;;   removes given string from this set, if it's already there
-;;   (otherwise produces this set)
-;; - size : -> Number
-;;   count the elements of this set
-;; - has? : String -> Boolean
-;;   is the given String in this set?
+;; A Set is one of:
+;; - (new empty%)
+;; - (new singelton% Number)
+;; - (new union% Set Set)
+;; Interpretation: a set that contains all the *unique* numbers in the data
+;; and implements
+;; - sum : -> Number
+;;   Produce the sum of all the unique elements of this set
+;; - sum/acc : Set -> Number
+;;   Produce the sum of all the unique elements of this set that are
+;;   *not* in the given set
+;; - contains? : Number -> Boolean
+;;   Does this set contain the given number?
+;; - union : Set -> Set
+;;   Combine this set with the given set
 
-;; empty-set is a Set value containing no elements
+(define-class empty%
+  (define (sum) 0)
+  (define (sum/acc s) 0))
+
+(define-class singleton%
+  (fields n)
+  (define (sum) n)
+  (define (sum/acc s) (cond [(s . contains? (this . n)) 0]
+                            [else (this . n)])))
+
+(define-class union%
+  (fields l r)
+  (define (sum) (this . sum/acc (new empty%)))
+  (define (sum/acc s) (+ (this . l . sum/acc s)
+                         (this . r . sum/acc (s . union (this . l))))))
 \end{verbatim}
-\noindent
+
+\newpage
+
+Unfortunately, as you can see, Van Horn fell asleep before finishing,
+and the deadline is today.  Help him out by implementing the
+\verb|union| and \verb|contains?| methods for all \verb|Set|s.
+
+\ifrubric
+\begin{alltt}
+\rubricpts{1 pt -- union method}
+\rubricpts{3 pts -- contains method}
+;; in empty%
+(define (union s) s)
+(define (contains n) #f)
+;; also ok:
+(define (union s) (new union% this s)
+;; in singleton%
+(define (union s) (new union% this s)
+(define (contains n) (= n (this . n)))
+;; in union%
+(define (union s) (new union% this s)
+(define (contains n) (or (this . l . contains n)
+                         (this . r . contains n)))
+(define e (new empty%))
+\rubricpts{2 pts -- tests}
+(check-expect (e . contains 5) #f)
+(check-expect ((new singleton% 5) . contains 5) #t)
+(check-expect ((new singleton% 6) . contains 5) #f)
+(check-expect ((new singleton% 6) . union  e . contains 5) #f)
+(check-expect (e . union (new singleton% 5) . contains 5) #t)
+(check-expect ((new singleton% 5) . union (new singleton% 5) . contains 5) #t)
+\end{alltt}
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
 \verb|Set|s should satisfy the following property, called the
-``insert-size'' property, for any element and
-set:
+``union-contains'' property, for any number $n$ and
+any two sets $s_1$ and $s_2$:
 
 \begin{quote}
-If we insert any element into any set, then the resulting set is the
-same size as the original set \emph{if} the element was already in the
-set, and the resulting set is one larger than the original set
-if the element was \emph{not} already in the set.
+Given two sets $s_1$ and $s_2$, if $n$ is contained in the union of $s_1$ and $s_2$, then $n$ is either
+contained in $s_1$ or it is contained in $s_2$.
 \end{quote}
@@
 \begin{enumerate}
-\item Codify the ``insert-size'' property as a predicate.  Write a
-  test case that should pass (if the sets work right) using the
+\item Codify the ``union-contains'' property as a predicate.  Write a
+  test case that should pass (if your implementations work right) using the
   predicate you have defined.
@@
 
-\ifrubric\else
+\ifrubric
+\begin{alltt}
+\rubricpts{1 pt -- contract & purpose}
+;; union-contains : Set Set Number -> Boolean
+\rubricpts{3 pts -- code}
+(define (union-contains s1 s2 n)
+  (check-expect (s1 . union s2 . contains n)
+                (or (s1 . contains n) (s2 . contains n))))
+\rubricpts{1 pt -- successful test}
+(union-contains e (new singleton% 17) 17)
+\end{alltt}
+\else
 \newpage
@@
 \noindent
-The College wants to be really sure that their sets are correct, and
-they've come up with another property that they think is true, called
-the ``insert-remove'' property:
+The University doesn't really trust Van Horn, and they want to make
+sure that \verb|Set|s satisfy another property, called the
+``union-sum'' property:
 
 \begin{quote}
-If we insert element any element $a$ into any set, and then remove any
-element $b$ from the resulting set, the result has the same size as if
-we do the two operations in the reverse order, i.e., remove $b$, then
-insert $a$.
+Given two sets $s_1$ and $s_2$, computing the sum of $s_1$ and the sum
+of $s_2$ and then adding the two results is equal to the sum of the
+union of  $s_1$ and $s_2$.
 \end{quote}
@@
 \noindent
-The College is considering adding this to the test suite as well.
+The University is considering adding this to the test suite as well.
 
@@
 
-\item Is the ``insert-remove'' property true for all sets?  If so,
+\item Is the ``union-sum'' property true for all sets?  If so,
   give an explanation.  If not, write a counter-example (i.e., a
@@
 \rubricpts{1 pt -- contract & purpose}
-;; check-insert-size : Set String -> Boolean
-;; check the "insert-size" property on s and i.
-\rubricpts{4 pts -- code}
-(define (check-insert-size s i)
-  (= (s . insert i . size)
-     (cond [(s . has? i) (add1 (s . size))]
-           [else               (s . size)])))
-
-\rubricpts{2 pts -- test case}
-(check-expect (check-insert-size empty-set "hi") true)
-
-\rubricpts{1 pt -- contract & purpose}
-;; check-remove-insert : Set String -> Boolean
-;; check the "remove-insert" property on s, a, and b.
+;; union-sum : Set Set -> Boolean
 \rubricpts{3 pts -- code}
-(define (check-remove-insert s a b)
-  (= (s . insert a . remove b . size)
-     (s . remove b . insert a . size))
+(define (union-contains s1 s2 n)
+  (check-expect (s1 . union s2 . sum)
+                (+ (s1 . sum) (s2 . sum))))w
 
 \rubricpts{1 pt -- successful test}
-(check-expect (check-remove-insert empty-set "yes" "no") true)
-
-;; Property does not hold.
-;; Counter-example (failing test):
+;; successful test
+(union-sum (new singleton% 17) e)
 \rubricpts{3 pts -- failing test}
-(check-expect (check-remove-insert empty-set "yes" "yes") true)
+;; It's false, counterexample:
+(union-sum (new singleton% 17) (new singleton% 17))
 \end{alltt}

### 2013-02-13 exam5/1.tex
fix a bunch of bugs in solutions

@@
 ;;   *not* in the given set
-;; - contains? : Number -> Boolean
+;; - contains : Number -> Boolean
 ;;   Does this set contain the given number?
@@
   (fields n)
-  (define (sum) n)
-  (define (sum/acc s) (cond [(s . contains? (this . n)) 0]
+  (define (sum) (this . n))
+  (define (sum/acc s) (cond [(s . contains (this . n)) 0]
                             [else (this . n)])))
@@
 and the deadline is today.  Help him out by implementing the
-\verb|union| and \verb|contains?| methods for all \verb|Set|s.
+\verb|union| and \verb|contains| methods for all \verb|Set|s.
 
@@
 ;; also ok:
-(define (union s) (new union% this s)
+(define (union s) (new union% this s))
 ;; in singleton%
-(define (union s) (new union% this s)
+(define (union s) (new union% this s))
 (define (contains n) (= n (this . n)))
 ;; in union%
-(define (union s) (new union% this s)
+(define (union s) (new union% this s))
 (define (contains n) (or (this . l . contains n)
@@
 \rubricpts{3 pts -- code}
-(define (union-contains s1 s2 n)
+(define (union-sum s1 s2 n)
   (check-expect (s1 . union s2 . sum)
-                (+ (s1 . sum) (s2 . sum))))w
+                (+ (s1 . sum) (s2 . sum))))
 

### 2013-02-13 exam5/1.tex
First draft of JSON problem.

@@
 
+\begin{problem} \pts{??}
+
+JSON is a lightweight data-interchange format.  It is widely used on
+the web and plays exactly the role of S-Expressions in universe
+programs, but for JavaScript programs.  It's used by folks like Google
+and there's tons of data on the web that's available in the JSON
+format.  So as to not be left out of the fun, your co-op employer has
+asked you to design an object representation of JSON to incorporate
+into their 2 million lines of {\tt class/2} code.  To start things off
+simply, here are some examples of what we'll call ``JSON, Jr.'':
+
+\begin{verbatim}
+"this is JSON"
+{ "this is JSON", "Jr." }
+{ {}, { "So", "is" }, "this" }
+\end{verbatim}
+The first example is just a string.  The second is an ``array'' of two
+JSON elements, which are both strings.  The third is another array,
+but of three elements: the first is an array of zero elements, the
+second is an array of two strings, and the third is a string.
+
+\begin{enumerate}
+\item Design an object-based data representation for JSON, Jr.
+\newpage
+\item Design a method for counting the number of strings in a JSON,
+  Jr.~object.
+\newpage
+\item Among the 2M lines of code your employer maintains are 30,000
+  lines of code devoted to implementations of the {\tt
+    StringPred} interface, which is:
+\begin{verbatim}
+;; A StringPred implements
+;; - apply : String -> Boolean
+\end{verbatim}
+These objects represent predicates on strings.  Now that your JSON,
+Jr. library is in place, they've decided to put their predicates to
+use in searching a large corpus of JSON data.  To accomodate this,
+design the following method for JSON, Jr.~objects:
+\begin{verbatim}
+;; Find the first string in this JSON Jr object satisfying
+;; given predicate, or #f if there's no such string.
+;; find : StringPred -> String or #f
+\end{verbatim}
+\end{enumerate}
+
+\end{problem}
+
+\newpage
+\ifrubric\else
+[Here is some more space for the previous problem.]
+\newpage
+\fi
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-02-13 exam5/1.tex
{} to [] and remove dir problem.

@@
 "this is JSON"
-{ "this is JSON", "Jr." }
-{ {}, { "So", "is" }, "this" }
+[ "this is JSON", "Jr." ]
+[ [], [ "So", "is" ], "this" ]
 \end{verbatim}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{15}
 
-% JSON problem
-
-Here are data and class definitions for representing the files and
-directories of a computer:
-
-\begin{verbatim}
-;; An Elem is one of
-;; - (new file% String Number)
-;; - (new dir% String LoElem)
-(define-class file% (fields name size))
-(define-class dir% (fields name elems))
-
-;; A LoElem is one of
-;; - (new mt%)
-;; - (new cons% Elem LoElem)
-(define-class mt%)
-(define-class cons% (fields first rest))    
-\end{verbatim}
-
-\noindent
-Here is an example of a directory tree and its representation as an
-\verb|Elem|:
-\begin{verbatim}
-#|
- User
-   +-- A
-   |   +-- a.txt
-   |   +-- B
-   |       +-- b.rkt
-   +-- C
-       +-- c.c
-|#
-
-(define mt (new mt%))
-(define a (new file% "a.txt" 5))
-(define b (new file% "b.rkt" 10))
-(define c (new file% "c.c" 1000))
-
-(define C/ (new dir% "C" (new cons% c mt)))
-(define B/ (new dir% "B" (new cons% b mt)))
-(define A/ (new dir% "A" (new cons% a (new cons% B/ mt))))
-(define User/ (new dir% "/" (new cons% A/ (new cons% C/ mt))))
-\end{verbatim}
-
-\newpage
-\noindent
-Your job is to implement one of the main tasks of the common Unix
-utility \verb|find|, which gives a list of all the files within an
-element.  To do so, add a \verb|list-files| method of \verb|Elem|
-which produces a list of files.  So for example, we expect:
-\begin{verbatim}
-(check-expect (User/ . list-files) 
-              (new cons% a (new cons% b (new cons% c mt))))
-\end{verbatim}
-\emph{Hint}: it may come in handy to be able to concatenate two
-\verb|LoElem| together.
-
-\ifrubric
-\else
-\newpage
-[Here is some more space for the previous problem.]
-\fi
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{1pts -- interfaces, contracts, purpose}
-;; An Elem implements
-;; - list-files : -> LoElem
-;;   List all files within this element.
-
-\rubricpts{2pts -- interfaces, contracts, purpose}
-;; A LoElem implements
-;; - list-files : -> LoElem
-;;   List all files within this list of elements.
-;; - append : LoElem -> LoElem
-;;   Append this list of elements to given list.
-
-\rubricpts{1pt -- code}
-;; in dir%
-(define (list-files)
-  (this . elems . list-files))
-
-\rubricpts{1pt -- code}
-;; in file%
-(define (list-files)
-  (new cons% this (new mt%)))
-
-\rubricpts{2pt -- code}
-;; in mt%
-(define (append loe) loe)
-(define (list-files) this)
-
-\rubricpts{4pt -- code}
-;; in cons%
-(define (append loe)
-  (new cons% (this . first) (this . rest . append loe)))
-(define (list-files)
-  (this . first . list-files . append (this . rest . list-files)))
-
-\rubricpts{1pt -- sanity}
-
-\rubricpts{3pts -- tests}
-;; tests
-(check-expect (mt . append (new cons% a mt)) (new cons% a mt))
-(check-expect ((new cons% a mt) . append (new cons% b mt))
-              (new cons% a (new cons% b mt)))
-
-(check-expect (mt . list-files) mt)
-(check-expect ((new cons% a (new cons% b mt)) . list-files)
-              (new cons% a (new cons% b mt)))
-
-(check-expect (a . list-files) (new cons% a mt))
-\end{alltt}
-\end{solution}
-
-\end{problem}
 \newpage

### 2013-02-13 exam5/1.tex
extra credit

@@
 \noindent
-After Van Horn discovered the problems with numbers, he attempted to
-fix them.  Unfortunately, not only did he not fix them, he broke
-booleans as well.  Fortunately, we have an idea about how you can
-help, expressed in the following interface definitions.
+Labich was unhappy to discover that the class languages don't have his
+favorite programming language feature---pattern matching.  After
+thinking hard about this problem, he came up with a solution, which he
+calls \emph{pattern objects}.
 
+Consider the following data defintion:
 \begin{verbatim}
-;; A Bool implements:
-;; - branch : [Action X] [Action X] -> X
-;;   run the first action if this bool is true, 
-;;   otherwise run the second action
+;; A BTree is one of:
+;; - (new leaf% Number)
+;; - (new node% BTree Number BTree)
+;; and implements
+;; - match : [Pattern X] -> X
+;;   applies the given pattern object to this tree
 
-;; An [Action X] implements:
-;; - run : -> X
-;; run this action
+;; A [Pattern X] implements:
+;; - leaf : Number -> X
+;; - node : BTree Number BTree -> X
 
+(define-class leaf%
+  (fields n)
+  ;; apply the appropriate method from the given pattern object
+  (define (match p) (p . leaf (this . n))))
+
+(define-class node%
+  (fields left n right)
+  ;; apply the appropriate method from the given pattern object
+  (define (match p) 
+    (p . node (this . left) (this . n) (this . right))))
 \end{verbatim}
 
-Following these interface definitions, design data and class
-defintions for implementations of the \verb|Bool| interface, and
-reimplement the \verb|fact| method to use them.
-You will need to provide new defintions for \verb|true| and
-\verb|false|, since they were broken by this catastrophe as well.
+A pattern object is like the implementation of two different functions
+at once: one that implements a function for leaves from a binary tree,
+and one function that works on nodes.
+
+Write a pattern object that computes the product of all of the numbers
+in a binary tree.
 
@@
 \rubricpts{5 pts -- get it right}
-(define-class fact%
-  (define (fact n)
-    (n . zero? . branch (new const% one) (new recur% n))))
+;; product% implements [Pattern Number]
+(define-class product%
+  ;; the product of a leaf
+  (define (leaf n) n)
+  ;; the product of a node
+  (define (node l n r) (* n (l . match this) (r . match this))))
 
-(define-class const%
-  (fields run))
-
-(define-class recur%
-  (fields n)
-  (define (run) 
-    ((new fact%) . fact (this . n . sub1) . times (this . n))))
-
-(define-class t%
-  (define (branch t e) (t . run)))
-(define-class f%
-  (define (branch t e) (e . run)))
-
-(define true (new t%)) 
-(define false (new f%))
-
-(check-expect ((new fact%) . fact zero) one)                  ; 0! = 1
-(check-expect ((new fact%) . fact one) one)                   ; 1! = 1
-(check-expect ((new fact%) . fact (one . add1)) (one . add1)) ; 2! = 2
-(check-expect ((new fact%) . fact (one . add1 . add1))        ; 3! = 6
-              (one . add1 . add1 . add1 . add1 . add1))
+(define p (new product%))
+(check-expect ((new leaf% 5) . match p) 5)
+(check-expect ((new node%  (new leaf% 2) 5 (new leaf% 7)) . match p) 70)
 \end{alltt}

### 2013-02-25 web/12.su.reg/lect/Graph.java
Some old summer stuff.

@@ -1,0 +1,144 @@
+import java.util.*;
+import tester.*;
+
+class Graph {
+    Hashtable<String,ArrayList<Edge>> g;
+    Graph() {
+	this.g = new Hashtable<String,ArrayList<Edge>>();
+    }
+
+    // EFFECT: Add edge between node1 and node2 in this graph.
+    void addEdge(String node1, String node2) {
+	if (!this.g.containsKey(node1)) {
+	    this.g.put(node1, new ArrayList<Edge>());
+	}
+	if (!this.g.containsKey(node2)) {
+	    this.g.put(node2, new ArrayList<Edge>());
+	}
+
+	this.g.get(node1).add(new Edge(node1, node2));
+	this.g.get(node2).add(new Edge(node2, node1));
+    }
+
+    // Get the neighbors of the given node in this graph.
+    ArrayList<Edge> getNeighbors(String node) {
+	return this.g.get(node);
+    }
+
+    // Does this graph contain the given node?
+    Boolean containsNode(String node) {
+	return this.g.containsKey(node);
+    }
+
+    // Is there an edge between node1 and node2?
+    Boolean edgeBetween(String node1, String node2) {
+	if (!this.containsNode(node1)) { return false; }
+	if (!this.containsNode(node2)) { return false; }
+	for (Edge e : this.getNeighbors(node1))
+	    if (e.to.equals(node2))
+		return true;
+	return false;
+    }
+}
+
+class Edge {
+    String from;
+    String to;
+    Edge(String from, String to) {
+	this.from = from;
+	this.to = to;
+    }
+}
+
+// Represents all the nodes we've already visited.
+class Seen {
+    ArrayList<String> seen;
+    Seen() {
+	this.seen = new ArrayList<String>();
+    }
+
+    // EFFECT: add given node to this seen.
+    void add(String node) {
+	this.seen.add(node);
+    }
+
+    // Has the given node been seen already?
+    Boolean hasSeen(String node) {
+	for (String n : seen) {
+	    if (n.equals(node)) {
+		return true;
+	    }
+	}
+	return false;
+    }
+}
+
+class Algorithms {
+
+    // Is there a path from start to end in g?
[…]

### 2013-02-25 web/12.su.reg/lect/Lab9/Balloon.java

@@ -1,0 +1,20 @@
+
+/** Represents a round Balloon */
+public class Balloon {
+    int x;
+    int y;
+    int radius;
+    String color;
+  
+    public Balloon (int x, int y, int radius, String color){
+        this.x = x;
+        this.y = y;
+        this.radius = radius;
+        this.color = color;
+    }
+    
+    /** Compute the distance of this balloon from the top of window */
+    int distanceFromTop(){
+        return this.y - this.radius;
+    } 
+}

### 2013-02-25 web/12.su.reg/lect/Lab9/IList.java

@@ -1,0 +1,16 @@
+import tester.Traversal;
+
+/** Represents a list of T */
+public interface IList<T> extends Traversal<T>{
+    
+    /** Is this list empty? */
+    public boolean isEmpty();  
+    
+    /** Produce the first item in this list */
+    public T getFirst();
+
+    /** Produce a {@link Traversal} for the rest of this list */
+    public Traversal<T> getRest();
+}
+
+

### 2013-02-25 web/12.su.reg/lect/Lab9/ISelect.java

@@ -1,0 +1,7 @@
+
+/** Represents a generic predicate */
+public interface ISelect<T>{
+    
+    /** Should the given thing be selected? */
+    public boolean select(T t);
+}

### 2013-02-25 web/12.su.reg/lect/Lab9/MtList.java

@@ -1,0 +1,20 @@
+import tester.*;
+
+/** Represents an empty list of T */
+public class MtList<T> implements IList<T>{
+
+    public MtList(){}
+
+    /** Is this list empty? */
+    public boolean isEmpty(){ return true; }  
+
+    /** Produce the first item in this list */
+    public T getFirst(){
+        throw new IllegalUseOfTraversalException("No first element");
+    }
+  
+    /** Produce a {@link Traversal} for the rest of this list */
+    public Traversal<T> getRest(){
+        throw new IllegalUseOfTraversalException("No more elements");
+    }
+}       

### 2013-02-25 web/12.su.reg/lect/Lab9/SmallBalloon.java

@@ -1,0 +1,17 @@
+
+/** Represents a predicate for selecting Balloons smaller 
+ *   than a given radius */
+public class SmallBalloon implements ISelect<Balloon>{
+
+    /** The (upper) limit for selected Balloons */
+    int limit;
+
+    SmallBalloon(int limit){
+        this.limit = limit;
+    }
+    
+    /** Is the given Balloon smaller than the limit? */
+    public boolean select(Balloon balloon){
+        return balloon.radius < this.limit;
+    }
+}

### 2013-02-25 web/12.su.reg/lect/MySort.java

@@ -1,0 +1,101 @@
+import tester.*;
+import java.util.*;
+
+interface Sort {
+    // Sort given list in ascending order.
+    ArrayList<Integer> sort(ArrayList<Integer> list);
+}
+
+class Selection implements Sort {
+
+    // Sort given list in ascending order.
+    // EFFECT: given list is sorted in-place.
+    public ArrayList<Integer> sort(ArrayList<Integer> list) {
+	Integer i = 0;
+	while (!i.equals(list.size())) {
+	    this.swap(list, i, this.findMin(list, i));
+	    i++;
+	}
+	return list;
+    }
+
+    // Starting at idx, find index of minimum element in list.
+    Integer findMin(ArrayList<Integer> list, Integer idx) {
+	Integer i = idx;
+	Integer min = list.get(i);
+	while (!i.equals(list.size())) {
+	    if (list.get(i) < min) {
+		min = list.get(i);
+	    }
+	    i++;
+	}
+	return min;
+    }
+
+    // EFFECT: Swap elements at index i and j in list.
+    void swap(ArrayList<Integer> list, Integer i, Integer j) {
+	list.set(j, list.set(i, list.get(j)));
+    }
+}
+
+class Insertion implements Sort {
+
+    // Sort given list in ascending order.
+    // EFFECT: given list is sorted in-place.
+    public ArrayList<Integer> sort(ArrayList<Integer> list) {
+	Integer i = 0;
+	while (!i.equals(list.size())) {
+	    this.insert(i, list);
+	    i++;
+	}
+	return list;
+    }
+
+    // EFFECT: Insert element of list at idx into
+    // sorted prefix of list.
+    void insert(Integer idx, ArrayList<Integer> list) {
+        Integer i = 0;
+        while (i < idx && list.get(idx) > list.get(i))
+            i++;
+        list.add(i, list.remove(idx.intValue()));
+    }
+}
+
+
+class Examples {
+    Sort isort = new Insertion();
+    Sort ssort = new Selection();
+
+    ArrayList<Sort> sorts = new ArrayList<Sort>();{
+	sorts.add(isort);
+	sorts.add(ssort);
+    }
+
+    ArrayList<Integer> a = new ArrayList<Integer>();
+
+    void init() {
+	a.clear();
+	a.add(7);
+	a.add(1);
+	a.add(9);
+	a.add(4);
+    }
+
[…]

### 2013-02-25 web/12.su.reg/lect/Sort.java

@@ -1,0 +1,83 @@
+import tester.*;
+import java.util.*;
+
+interface Sort {
+    // Sort given list in ascending order.
+    ArrayList<Integer> sort(ArrayList<Integer> list);
+}
+
+class Selection implements Sort {
+
+    // Sort given list in ascending order.
+    // EFFECT: sorts the given list in-place.
+    public ArrayList<Integer> sort(ArrayList<Integer> list) {
+	Integer i = 0;
+	while(!i.equals(list.size())) {
+	    this.swap(list, i, this.findMin(list, i));
+	    i++;
+	}
+	return list;
+    }
+
+    // EFFECT: swaps the elements at i and j in the list.
+    void swap(ArrayList<Integer> list, Integer i, Integer j) {
+	list.set(j, list.set(i, list.get(j)));
+    }
+
+    // Find the index of the first minimum element of the given list,
+    // starting at the given index.
+    Integer findMin(ArrayList<Integer> list, Integer idx) {
+	Integer i = idx;
+	Integer iMin = idx;
+	while(!i.equals(list.size())) {
+	    if (list.get(i) < list.get(iMin)) {
+		iMin = i;
+	    }
+	    i++;
+	}
+	return iMin;
+    }
+}
+
+class Insertion {
+    
+}
+
+class Examples {
+    Sort ssort = new Selection();
+    Sort isort = new Insertion();
+
+    ArrayList<Sort> sorts = new ArrayList<Sort>();{
+	sorts.add(isort);
+	sorts.add(ssort);
+    }
+
+    ArrayList<Integer> a = new ArrayList<Integer>();
+
+    void init() {
+	this.a.clear();
+	this.a.add(7);
+	this.a.add(1);
+	this.a.add(9);
+	this.a.add(4);
+    }
+
+    void testSorts(Tester t) {
+	for(Sort s : this.sorts) {
+	    this.init();
+	    ArrayList<Integer> b = s.sort(this.a);
+	    t.checkExpect(b.size(), 4);
+	    t.checkExpect(b.get(0), 1);
+	    t.checkExpect(b.get(1), 4);
+	    t.checkExpect(b.get(2), 7);
+	    t.checkExpect(b.get(3), 9);
+	}
+    }
+
+    void testSelectionInPlace(Tester t) {
+	this.init();
+	ArrayList<Integer> b = this.ssort.sort(this.a);
+	t.checkExpect(b == a, true);
+    }
+}
+

### 2013-02-25 web/12.su.reg/lect/StringAppend.java

@@ -1,0 +1,6 @@
+public class StringAppend implements IVisitorBT<String,String> {
+    public String visitLeaf() { return ""; }
+    public String visitNode(String val, IBT<String> left, IBT<String> right) {
+	return val + left.accept(this) + right.accept(this);
+    }
+}

### 2013-02-25 web/12.su.reg/lect/bt/Biggest.java

@@ -1,0 +1,42 @@
+import java.util.Comparator;
+
+class LongerString implements Comparator<String> {
+    Integer compare(String s1, String s2) {
+	...
+    }
+}
+
+class Biggest<X> implements IVisitorBT<X,X> {
+
+    Comparator<X> comp;
+    Biggest(Comparator<X> comp) {
+	this.comp = comp;
+    }
+
+    public Integer visitLeaf() {
+	throw new RuntimeException("No biggest number.");
+    }
+
+    public Integer visitNode(Integer val, 
+		      IBT<Integer> left, 
+		      IBT<Integer> right) {
+	return Math.max(left.accept(new BiggestSoFar(val)),
+			right.accept(new BiggestSoFar(val)));
+    }
+}
+
+class BiggestSoFar implements IVisitorBT<Integer,Integer> {
+    Integer seen;
+    BiggestSoFar(Integer seen) {
+	this.seen = seen;
+    }
+
+    public Integer visitLeaf() { return this.seen; }
+    public Integer visitNode(Integer val, 
+		      IBT<Integer> left, 
+		      IBT<Integer> right) {
+	return Math.max(val, Math.max(left.accept(this),
+				      right.accept(this)));
+    }
+}
+			      

### 2013-02-25 web/12.su.reg/lect/bt/Leaf.java

@@ -1,0 +1,8 @@
+// Represents a leaf in a binary tree of X.
+public class Leaf<X> extends ABT<X> {
+    Leaf() {}
+
+    public <T> T accept(IVisitorBT<X,T> v) {
+	return v.visitLeaf();
+    }
+}

### 2013-02-25 web/12.su.reg/lect/bt/Size.java

@@ -1,0 +1,7 @@
+// BT visitor that computes the size of the visited bt.
+public class Size<X> implements IVisitorBT<X,Integer> {
+    public Integer visitLeaf() { return 0; }
+    public Integer visitNode(X val, IBT<X> left, IBT<X> right) {
+	return 1 + left.accept(this) + right.accept(this);
+    }
+}

### 2013-02-25 web/12.su.reg/world/OceanWorld.java

@@ -1,0 +1,186 @@
+import tester.*;
+import funworld.*;
+import geometry.*;
+import colors.*;
+
+import java.awt.*;
+
+// CS 2510 Lab 5: Designing World Games
+// 7 February 2012
+
+// A Shark is a new Shark(Integer, Integer)
+// Interp: new Shark(y) represents shark at height y in graphics coordinates
+class Shark {
+  Integer y;
+  Integer health;
+  WorldImage sharkImage;
+
+  Shark(Integer y, Integer health) {
+    this.y = y;
+    this.health = health;
+    this.sharkImage = new FromFileImage(new CartPt(30, y), "shark.png");
+  }
+
+  // This shark chomps the fish
+  public Shark chomp(Fish f) {
+    // this.y this.health f.p f.health
+    return new Shark(this.y, this.health + f.health);
+  } 
+  
+  //move this shark up or down depending of the key
+  public Shark onKey(String ke){
+    // move this shark up or down depending of the key
+    return this;
+  }
+
+}
+
+// A Fish swims in the ocean and provides some food value for the shark
+class Fish {
+  CartPt p;          // the current osition of this fish
+  Integer health;    // the food value of this fish for the shark
+  
+  // make a new fish
+  Fish(CartPt p, Integer health) {
+    this.p = p;
+    this.health = health;
+  }
+
+  // Move this fish left.
+  public Fish moveLeft() {
+    return new Fish(this.p.moveLeft(3), this.health);
+  }
+
+  // Is this fish close to the given shark?
+  /*
+    public Boolean closeByHuh(Shark that) {
+    ...
+    }
+   */
+  
+  // produce an image of this fish at this fish's location
+  WorldImage makeFishImage(){
+    return new FromFileImage(this.p, "pink-fish.png");
+  }
+
+}
+
+// a CartPt extends the Posn class that represents the (x,y) 
+// in graphics coordinates and allows for adding new methods
+class CartPt extends Posn {
+  
+  // the super constructor does the job
+  CartPt(Integer x, Integer y) {
+    super(x, y);
+  }
+
+  // Move this posn left dx units.
+  public CartPt moveLeft(Integer dx) {
+    return new CartPt(this.x-dx, this.y);
+  }
+}
+
+// A School is one of:
+// - new EmptySchool()
+// - new ConsSchool(Fish, School)
+
+// to represent a school of fish
+interface School {
+
+  // produce an image of this fishies at this their locations
[…]

### 2013-02-25 web/12.su.reg/blog/CheckedException.java
Some old blogged Java.

@@ -1,0 +1,36 @@
+import tester.*;
+
+// An example of how to test for exceptional behavior
+
+// Represents a function A -> B
+interface Function<A,B> {
+    // Apply this function to given argument.
+    B apply(A a);
+}
+
+// Represents inversion operation
+class Divider implements Function<Double,Double> {
+    Divider() {}
+
+    // Produce 1/x.
+    public Double apply(Double x) {
+	if (x.equals(0.0)) {
+	    throw new RuntimeException("Divide by zero.");
+	} else {
+	    return 1/x;
+	}
+    }
+}
+
+class Examples {
+    void testInv(Tester t) {
+	t.checkException(// The expected exception
+			 new RuntimeException("Divide by zero."),
+			 // The object that throws
+			 new Divider(),
+			 // The method (as a string!) that throws
+			 "apply",
+			 // The arguments to the method
+			 0.0);
+    }
+}

### 2013-02-25 web/12.su.reg/blog/Lab.java

@@ -1,0 +1,228 @@
+import tester.*;
+
+// Represents a association between keys and values.
+interface Assoc<K,V> {
+    // Accept visitor and compute result.
+    <R> R accept(AssocVisitor<K,V,R> v);
+
+    // Is this assoc interpreted the same as that assoc?
+    Boolean same(Assoc<K,V> that);
+
+    // Is this a subset of that?
+    Boolean subset(Assoc<K,V> that);
+    // Does this assoc contain given key, value association?
+    Boolean contains(K key, V val);
+    // Remove this key value pair from this assoc
+    Assoc<K,V> remove(K key);
+
+    // EFFECT: modify value to given value for given key
+    // Produce old value
+    V update(K key, V val);
+}
+
+interface AssocVisitor<K,V,R> {
+    R visitEmptyAssoc();
+    R visitConsAssoc(K key, V val, Assoc<K,V> rest);
+}
+
+abstract class AAssoc<K,V> implements Assoc<K,V> {
+    AAssoc() {}
+
+    // Is this assoc interpreted the same as that assoc?
+    public Boolean same(Assoc<K,V> that) {
+	return this.subset(that)
+	    && that.subset(this);
+    }
+
+    public boolean equals(Object that) {
+	return (that instanceof Assoc)
+	    && this.same((Assoc<K,V>)that);
+    }
+}
+
+class EmptyAssoc<K,V> extends AAssoc<K,V> {
+    EmptyAssoc() {}
+    public <R> R accept(AssocVisitor<K,V,R> v) {
+	return v.visitEmptyAssoc();
+    }
+    public Boolean subset(Assoc<K,V> that) { return true; }
+    public Boolean contains(K key, V val) { return false; }
+    public Assoc<K,V> remove(K key) { return this; }
+    public int hashCode() { return 42; }
+    public V update(K key, V val) {
+	throw new RuntimeException("No such key");
+    }
+}
+
+class ConsAssoc<K,V> extends AAssoc<K,V> {
+    K key;
+    V val;
+    Assoc<K,V> rest;
+    ConsAssoc(K key, V val, Assoc<K,V> rest) {
+	this.key = key;
+	this.val = val;
+	this.rest = rest;
+    }
+    public <R> R accept(AssocVisitor<K,V,R> v) {
+	return v.visitConsAssoc(this.key, this.val, this.rest);
+    }
[…]

### 2013-02-25 web/12.su.reg/today/BTIter.java
Old summer today folder.

@@ -1,0 +1,131 @@
+interface BT<X> {
+    <R> R accept(BTVisitor<X,R> v);
+}
+
+interface BTVisitor<X,R> {
+    R visitLeaf();
+    R visitNode(X val, BT<X> left, BT<X> right);
+}
+
+class MaybeCons<X> implements BTVisitor<X,List<Node<X>>> {
+    List<Node<X>> ls;
+    MaybeCons(List<Node<X>> ls) {
+	this.ls = ls;
+    }
+    public List<Node<X>> visitLeaf() { return this.ls; }
+    public List<Node<X>> visitNode(X val, BT<X> left, BT<X> right) {
+	return new Cons<Node<X>>(new Node<X>(val, left, right));
+    }
+}
+
+class Leaf<X> implements BT<X> {
+    public <R> R accept(BTVisitor<X,R> v) {
+	return v.visitLeaf();
+    }
+}
+
+class Node<X> implements BT<X> {
+    X val;
+    BT<X> left;
+    BT<X> right;
+    Node(X val, BT<X> left, BT<X> right) {
+	this.val = val;
+	this.left = left;
+	this.right = right;
+    }
+    public <R> R accept(BTVisitor<X,R> v) {
+	return v.visitNode(this.val, this.left, this.right);
+    }
+}
+
+class BTIterator<X> implements Iterator<X> {
+    List<Node<X>> ls;
+    BTIterator(Node<X> bt) {
+	this.ls = new Cons<Node<X>>(bt, new Empty<Node<X>>());
+    }
+    public X next() {
+	X first = ls.accept(new First<X>());
+	BTVisitor<X,List<Node<X>>> v1 = new MaybeCons(ls.accept(new Rest<X>()));
+	BTVisitor<X,List<Node<X>>> v2 = new MaybeCons(v1);
+	ls = ls.accept(v2);
+	return first;
+    }
+    public boolean hasNext() {
+	for (Node<X> n : ls) return true;
+	return false;
+    }
+    public void remove() {}
+}
+
+interface ListVisitor<X,R> {
+    R visitEmpty();
+    R visitCons(X first, List<X> rest);
+}
+
+class First<X> implements ListVisitor<X,X> {
+    public X visitEmpty() { throw new RuntimeException("Undefined"); }
+    public X visitCons(X first, List<X> rest) { return first; }
+}
+
+class Rest<X> implements ListVisitor<X,List<X>> {
[…]

### 2013-02-25 web/12.su.reg/today/Books.java

@@ -1,0 +1,88 @@
+import tester.*;
+
+class Author {
+    String name;
+    LoB bs;
+    Author(String name) {
+	this.name = name;
+	this.bs = new MTB();
+    }
+
+    // EFFECT: add given book to this author's list of books.
+    void addBook(Book b) {
+	this.bs = new ConsB(b, this.bs);
+    }
+}
+
+class Book {
+    String title;
+    LoA as;
+    Book(String title, LoA as) {
+	this.title = title;
+	this.as = as;
+	as.addBook(this);
+    }
+}
+
+interface LoB {}
+
+class MTB implements LoB {
+    MTB(){}
+}
+
+class ConsB implements LoB {
+    Book b;
+    LoB rest;
+    ConsB(Book b, LoB rest) {
+	this.b = b;
+	this.rest = rest;
+    }
+}
+
+interface LoA {
+    // EFFECT: add given book to these authors' books.
+    void addBook(Book b);
+}
+
+class MTA implements LoA {
+    MTA(){}
+
+    // EFFECT: add given book to these authors' books.
+    public void addBook(Book b) {}
+}
+
+class ConsA implements LoA {
+    Author a;
+    LoA rest;
+    ConsA(Author a, LoA rest) {
+	this.a = a;
+	this.rest = rest;
+    }
+
+    // EFFECT: add given book to these authors' books.
+    public void addBook(Book b) {
+	this.a.addBook(b);
+	this.rest.addBook(b);
+    }
+}
+
+
+class Examples {
+
+    Author mf = new Author("Matthias");
+    Author vkp = new Author("Viera");
+
+    Book htdc = new Book("How to Design Classes",
+			 new ConsA(mf, new ConsA(vkp, new MTA())));
+
+    Book htdp = new Book("How to Design Programs",
+			 new ConsA(mf, new MTA()));
+
+    void testBook(Tester t) {
+	t.checkExpect(mf.bs, new ConsB(htdp,
+				       new ConsB(htdc,
+						 new MTB())));
+	t.checkExpect(htdc.as, new ConsA(mf, new ConsA(vkp, new MTA())));
+	t.checkExpect(vkp.bs, new ConsB(htdc, new MTB()));
+    }
+}

### 2013-02-25 web/12.su.reg/today/Date.java

@@ -1,0 +1,49 @@
+import tester.*;
+
+// Represents a date
+// Month has to be in [1,12], Day in [1,31].
+class Date {
+    Integer year;
+    Integer month;
+    Integer day;
+    Date(Integer year, Integer month, Integer day) {
+	this.year = year;
+	if (month > 12) {
+	    throw new RuntimeException("There are only 12 months on this planet.");
+	}
+	this.month = month;
+	if (day > 31) {
+	    throw new RuntimeException("There are never more than 31 days in a month.");
+	}
+	this.day = day;
+    }
+
+    private Date() {
+	this.year = 1970;
+	this.month = 1;
+	this.day = 1;
+    }
+
+    Date blah() {
+	return new Date();
+    }
+
+    Date(String s) {
+	this();
+	if (s.equals("My birthday")) {
+	    this.year = 2002;
+	    this.month = 2;
+	    this.day = 2;
+	} else {
+
+	}
+    }
+}
+
+class Examples {
+    Date d1 = new Date(1992, 11, 1);
+    // Date d2 = new Date(-45, 34, 783);  // Reject
+    Date d3 = new Date(d1);
+    Date d4 = new Date("My birthday");
+    Date d5 = new Date("Not my birthday");
+}

### 2013-02-25 web/12.su.reg/today/DoubleDispatch.java

@@ -1,0 +1,67 @@
+import tester.*;
+
+// Double dispatch approach to structural equality.
+
+// Represents a geometric shape
+interface Shape {
+    // Is this shape the same as that shape?
+    Boolean same(Shape that);
+
+    // Is this shape the same as that square?
+    Boolean sameSq(Sq that);
+    // Is this shape the same as that circle?
+    Boolean sameCirc(Circ that);
+}
+
+abstract class AShape implements Shape {
+    public Boolean sameSq(Sq that) { return false; }
+    public Boolean sameCirc(Circ that) { return false; }
+}
+
+// Represents a square
+class Sq extends AShape {
+    Integer size;
+    Sq(Integer size) {
+	this.size = size;
+    }
+
+    // Is this square the same as that shape?
+    public Boolean same(Shape that) {
+	return that.sameSq(this);
+    }
+
+    // Is this square the same as that square?
+    public Boolean sameSq(Sq that) {
+	return this.size.equals(that.size);
+    }
+}
+
+// Represents a circle
+class Circ extends AShape {
+    Integer radius;
+    Circ(Integer radius) {
+	this.radius = radius;
+    }
+
+    // Is this circle the same as that shape?
+    public Boolean same(Shape that) {
+	return that.sameCirc(this);
+    }
+
+    // Is this circle the same as that circle?
+    public Boolean sameCirc(Circ that) {
+	return this.radius.equals(that.radius);
+    }
+}
+
+class Examples {
+    Examples() {}
+
+    void testSame(Tester t) {
+	t.checkExpect(new Sq(5).same(new Sq(5)), true);
+	t.checkExpect(new Sq(5).same(new Sq(7)), false);
+	t.checkExpect(new Circ(5).same(new Circ(5)), true);
+	t.checkExpect(new Circ(5).same(new Circ(7)), false);
+	t.checkExpect(new Circ(5).same(new Sq(5)), false);
+    }
+}

### 2013-02-25 web/12.su.reg/today/Drink.java

@@ -1,0 +1,59 @@
+import tester.*;
+
+// An Example showing the evil of instanceof and casts.
+// To fix this problem: don't use instanceof and casts,
+// use either
+// - safe casting
+// - double dispatch
+// - visitors
+
+// Represents a drink
+interface Drink {
+    // Is this drink the same as that drink?
+    Boolean same(Drink that);
+}
+
+// Represents a cup of coffee.
+class Coffee implements Drink {
+    Integer sugars;
+    Coffee(Integer sugars) {
+	this.sugars = sugars;
+    }
+
+    // Is this coffee the same as that drink?
+    public Boolean same(Drink that) {
+	return (that instanceof Coffee)
+	    && ((Coffee)that).sugars.equals(this.sugars);
+    }
+}
+
+// Represents a cup of tea.
+class Tea implements Drink {
+    Tea() {}
+
+    // Is this tea the same as that drink?
+    public Boolean same(Drink that) {
+	return (that instanceof Tea);
+    }
+}
+
+// Represents a cup of coffee with Bailey's in it.
+class Baileys extends Coffee {
+    Baileys(Integer sugar) { super(sugar); }
+
+    // Is this coffee w/ Bailey's the same as that drink?
+    public Boolean same(Drink that) {
+	return (that instanceof Baileys)
+	    && ((Baileys)that).sugars.equals(this.sugars);
+    }
+}
+
+class Examples {
+    Examples() {}
+
+    void testSame(Tester t) {
+	t.checkExpect(new Baileys(0).same(new Coffee(0)), false);
+	// This test fails because our program is BROKEN.
+	t.checkExpect(new Coffee(0).same(new Baileys(0)), false);
+    }
+}

### 2013-02-25 web/12.su.reg/today/EqualVisitor.java

@@ -1,0 +1,141 @@
+import tester.*;
+
+// Represents a computation over a list of integers producing a Result
+interface ListVisitor<Result> {
+    Result visitEmpty();
+    Result visitCons(Integer first, ListInteger rest);
+}
+
+class Equal implements ListVisitor<Boolean> {
+    ListInteger other;
+    Equal(ListInteger other) {
+	this.other = other;
+    }
+    public Boolean visitEmpty() {
+	return other.accept(new IsEmpty());
+    }
+    public Boolean visitCons(Integer first, ListInteger rest) {
+	return other.accept(new SameCons(first, rest));
+    }
+}
+
+class IsEmpty implements ListVisitor<Boolean> {
+    IsEmpty() {}
+    public Boolean visitEmpty() { return true; }
+    public Boolean visitCons(Integer first, ListInteger rest) { return false; }
+}
+
+class SameCons implements ListVisitor<Boolean> {
+    Integer first;
+    ListInteger rest;
+    SameCons(Integer first, ListInteger rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+    public Boolean visitEmpty() { return false; }
+    public Boolean visitCons(Integer first, ListInteger rest) {
+	return this.first == first
+	    && this.rest.accept(new Equal(rest));
+    }
+}
+
+
+// Represents the sum computation over lists of integers.
+class Sum implements ListVisitor<Integer> {
+    Sum() {}
+    public Integer visitEmpty() {
+	return 0;
+    }
+    public Integer visitCons(Integer first, ListInteger rest) {
+	return first + rest.accept(this);
+    }
+}
+
+interface ListInteger {
+    // Compute the largest element in a non-empty list.
+    Integer max();
+
+    // Compute the largest number in this list and acc.
+    // ACC: the largest number seen so far.
+    Integer maxAcc(Integer acc);
+
+    // Accept the given visitor and compute a result.
+    <R> R accept(ListVisitor<R> v);
+}
+
+class Empty implements ListInteger {
+    Empty() {}
+
+    // Compute the largest element in a non-empty list.
+    public Integer max() {
+	throw new RuntimeException("Can't take the max of an empty list, mom.");
+    }
+
+    // Compute the largest element in this list and acc, or acc if no elements.
[…]

### 2013-02-25 web/12.su.reg/today/Exam3.java

@@ -1,0 +1,159 @@
+/*********************************************************************/
+/* Given */
+
+import tester.*;
+
+// Represents a function from A to B.
+interface Fun<A,B> {
+    // Apply this function to the given argument.
+    B apply(A a);
+}
+
+// Represents a point on the Cartesian plane.
+class Posn {
+    Integer x;
+    Integer y;
+    Posn(Integer x, Integer y) {
+	this.x = x;
+	this.y = y;
+    }
+}
+
+// Represents a computation over a list of X producing an R.
+interface ListVisitor<X,R> {
+    R visitEmpty();
+    R visitCons(X first, List<X> rest);
+}
+
+// Represents a list of X.
+interface List<X> {
+    <R> R accept(ListVisitor<X,R> v);
+}
+
+// Represents an empty list of X.
+class Empty<X> implements List<X> {
+    Empty() {}
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitEmpty();
+    }
+}
+
+// Represents a non-empty list of X.
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+    Cons(X first, List<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitCons(this.first, this.rest);
+    }
+}
+
+/*********************************************************************/
+/* Problem 1 */
+
+class Dist implements Fun<Posn,Double> {
+    Dist() {}
+    public Double apply(Posn p) {
+	return Math.sqrt((p.x * p.x) + (p.y * p.y));
+    }
+}
+
+/*********************************************************************/
+/* Problem 2 */
+
+// Map given function over a list and produce a list of results.
+class Map<X,Y> implements ListVisitor<X,List<Y>> {
+    Fun<X,Y> f;
+    Map(Fun<X,Y> f) {
+	this.f = f;
+    }
+    public List<Y> visitEmpty() {
+	return new Empty<Y>();
+    }
+    public List<Y> visitCons(X first, List<X> rest) {
+	return new Cons<Y>(this.f.apply(first), rest.accept(this));
+    }
+}
+
+/*********************************************************************/
+/* Problem 3 */
+
[…]

### 2013-02-25 web/12.su.reg/today/FactIter.java

@@ -1,0 +1,32 @@
+import tester.*;
+import java.util.*;
+
+class HalfOpen implements Iterable<Integer>, Iterator<Integer> {
+    Integer lo;
+    Integer hi;
+    HalfOpen(Integer lo, Integer hi) {
+	this.lo = lo;
+	this.hi = hi;
+    }
+
+    public Iterator<Integer> iterator() { return this; }
+    public Integer next() {
+	lo++;
+	return lo - 1;
+    }
+    public boolean hasNext() {
+	return this.lo < this.hi;
+    }
+    public void remove() {}
+}
+
+class Examples {
+    public static void main(String[] in) {
+	// void testFact(Tester t) {
+	Integer acc = 1;
+	for (Integer i : new HalfOpen(1,6)) {
+	    acc = acc * i;
+	}
+	System.out.println(acc);
+    }
+}

### 2013-02-25 web/12.su.reg/today/Flinstones.java

@@ -1,0 +1,51 @@
+import java.util.ArrayList;
+import tester.*;
+
+class Posn {
+    Integer x;
+    Integer y;
+    Posn(Integer x, Integer y) {
+	this.x = x;
+	this.y = y;
+    }
+    /*
+    public boolean equals(Object o) {
+	return true;
+    }
+    */
+}
+
+class Examples {
+    void testItOut(Tester t) {
+	ArrayList<String> a = new ArrayList<String>() {{
+		this.add("Fred");
+		this.add("Wilma");
+	    }};
+
+	// a.add("Fred");
+	// a.add("Wilma");
+	a.add(2, "Barney");
+	t.checkExpect(a.size(), 3);
+	t.checkExpect(a.contains("Barney"), true);
+	t.checkExpect(a.contains("Homer"), false);
+	t.checkExpect(a.contains(7), false);
+    }
+
+    void testPosn(Tester t) {
+	final Posn p = new Posn(3,4);
+	ArrayList<Posn> ps = new ArrayList<Posn>() {{
+		this.add(p);
+	    }};
+	t.checkExpect(ps.contains(p), true);
+	t.checkExpect(ps.indexOf(p), 0);
+	t.checkExpect(ps.indexOf(7), -1);
+	t.checkExpect(ps.isEmpty(), false);
+	t.checkExpect(ps.remove(0), p);
+	t.checkExpect(ps.isEmpty(), true);
+    }
+
+    // EFFECT: swap elements at index i and j
+    <E> void swap(int i, int j, ArrayList<E> as) {
+	as.set(j, as.set(i, as.get(j)));
+    }
+}

### 2013-02-25 web/12.su.reg/today/HashTableTest.java

@@ -1,0 +1,259 @@
+import java.util.*;
+import tester.*;
+
+interface ListVisitor<X,R> {
+    R visitEmpty();
+    R visitCons(X first, List<X> rest);
+}
+
+interface List<X> extends Iterable<X> {
+    // add more list stuff...
+    Boolean isEmpty();
+    X getFirst();
+    List<X> getRest();
+    <R> R accept(ListVisitor<X,R> v);
+    Integer size();
+}
+
+abstract class AList<X> implements List<X> {
+    public Iterator<X> iterator() {
+	return new ListIterator<X>(this);
+    }
+}
+
+class Empty<X> extends AList<X> {
+    Empty() {}
+    public Boolean isEmpty() { return true; }
+    public X getFirst() { throw new RuntimeException("Empty list"); }
+    public List<X> getRest() { throw new RuntimeException("Empty list"); }
+    public <R> R accept(ListVisitor<X,R> v) { return v.visitEmpty(); }
+    public Integer size() { return 0; }
+}
+
+class Cons<X> extends AList<X> {
+    X first;
+    List<X> rest;
+    Cons(X first, List<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+    public Boolean isEmpty() { return false; }
+    public X getFirst() { return this.first; }
+    public List<X> getRest() { return this.rest; }
+
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitCons(this.first, this.rest);
+    }
+
+    public Integer size() {
+	return 1 + this.rest.size();
+    }
+}
+
+// Represents an iteration through a list.
+class ListIterator<X> implements Iterator<X> {
+    List<X> list;
+    ListIterator(List<X> list) {
+	this.list = list;
+    }
+
+    public boolean hasNext() {
+	return !(this.list.isEmpty());
+    }
+
+    public X next() {
+	X res = this.list.getFirst();
+	this.list = this.list.getRest();
+	return res;
+    }
+
+    public void remove() {}
+}
+
+class Pair<A,B> {
+    A left;
+    B right;
+    Pair(A left, B right) {
+	this.left = left;
+	this.right = right;
+    }
+}
+
+class HashTable<K,V> {
[…]

### 2013-02-25 web/12.su.reg/today/Iterators.java

@@ -1,0 +1,81 @@
+import tester.*;
+import java.util.*;
+
+class Posn implements Iterable<Integer> {
+    Integer x;
+    Integer y;
+    Posn(Integer x, Integer y) {
+	this.x = x;
+	this.y = y;
+    }
+
+    public Iterator<Integer> iterator() {
+	return new PosnIterator(this);
+    }
+}
+
+class PosnIterator implements Iterator<Integer> {
+    Posn p;
+    Integer i; // i is in [0,2]
+    PosnIterator(Posn p) {
+	this.p = p;
+	this.i = 0;
+    }
+
+    public void remove() {
+	// This space is intentionally left blank.
+    }
+
+    public boolean hasNext() {
+	return i < 2;
+    }
+
+    public Integer next() {
+	this.i = this.i+1;
+	if (this.i.equals(1)) {
+	    return this.p.x;
+	} else if (this.i.equals(2)) {
+	    return this.p.y;
+	} else {
+	    throw new NoSuchElementException();
+	}
+    }
+}
+
+class Alg {
+    Integer prod(Iterable<Integer> xs) {
+	Integer acc = 1;
+	for (Integer x : xs) {
+	    acc = acc * x;
+	}
+	return acc;
+    }
+}
+
+
+
+class Examples {
+    void testPosnIterator(Tester t) {
+	Posn p = new Posn(3,4);
+	Iterator<Integer> ip = p.iterator();
+	Iterator<Integer> jp = p.iterator();
+	t.checkExpect(ip.hasNext(), true);
+	t.checkExpect(ip.next(), 3);
+	t.checkExpect(ip.hasNext(), true);
+	t.checkExpect(ip.next(), 4);
+	t.checkExpect(ip.hasNext(), false);
+	t.checkExpect(jp.hasNext(), true);
+    }
+
+    void testProd(Tester t) {
+	ArrayList<Integer> as = new ArrayList<Integer>();
+	as.add(2);
+	as.add(3);
+	as.add(4);
+	Posn p = new Posn(3, 4);
+	Alg a = new Alg();
+	t.checkExpect(a.prod(as), 24);
+	t.checkExpect(a.prod(p), 12);
+    }
+}
+

### 2013-02-25 web/12.su.reg/today/Lecture_17.java

@@ -1,0 +1,151 @@
+/***********************************************
+ *  CS2510 Spring 2011 
+ *  Lecture #17
+ *  ArrayLists and Binary Search
+ ***********************************************/
+
+import tester.*;
+
+// Import the ArrayList class
+import java.util.ArrayList;
+
+// Represents a Dinosaur
+class Dino{
+    String eyecolor;
+    boolean male;
+    int teeth;
+    
+    Dino(String eyecolor, boolean male, int teeth){
+        this.eyecolor = eyecolor;
+        this.male = male;
+        this.teeth = teeth;
+    }
+}
+
+// Examples and Tests of mutation and ArrayList methods
+class LectureExamples{
+    // The initial List
+    ArrayList<Dino> alist = new ArrayList<Dino>();
+    
+    // Some Dinos for testing
+    Dino d1 = new Dino("blue", true, 2);
+    Dino d2 = new Dino("yellow", true, 70);
+    Dino d3 = new Dino("black", false, 6);
+    Dino d4 = new Dino("chartreuse", false, 50);
+    Dino d5 = new Dino("red", false, 8);
+    Dino d6 = new Dino("magenta", true, 40);
+    Dino d7 = new Dino("indigo", true, 33);
+    
+    // Reset our data... add the first three Dinos
+    void reset(){
+        this.alist.clear();
+        this.alist.add(this.d1);
+        this.alist.add(this.d2);
+        this.alist.add(this.d3);
+    }
+    
+    // Examples/Tests of all the ArrayList methods
+    void testAll(Tester t){
+        /** Useful methods for ArrayList<E>
+         *     void clear()
+         *     boolean isEmpty()
+         *     int size()
+         *     boolean add(E element)
+         *     void add(int index, E element)
+         *     E get(int index)
+         *     E set(int index, E element)
+         *     E remove(int index)
+         */
+
+        this.reset();
+        t.checkExpect(this.alist.get(1).eyecolor, "yellow");
+        t.checkExpect(this.alist.get(2).eyecolor, "black");
+
+        // Throws an Exception when:  size() <= index < 0
+        t.checkException(new IndexOutOfBoundsException("Index: 6, Size: 3"),
+                this.alist, "add", 6, this.d4);
+        
+        this.alist.remove(2);
+        t.checkExpect(this.alist.size(), 2);
+        
+        this.alist.add(2, this.d3);
+        t.checkExpect(this.alist.size(), 3);
[…]

### 2013-02-25 web/12.su.reg/today/Lecture_29.java

@@ -1,0 +1,107 @@
+/***********************************************
+ *  CS2510 Spring 2011 
+ *  Lecture #28-ish
+ *  Sorting Demonstration(s)
+ ***********************************************/
+
+import tester.*;
+
+/** Represents an Expression in our "<i>language</i>" */
+interface Expr{
+    
+}
+
+/** Represents a binary expression
+ *  (e.g., addition/subtraction/multiplication) */
+abstract class Bin implements Expr{
+    Expr left;
+    Expr right;
+    
+    Bin(Expr left, Expr right){
+        this.left = left;
+        this.right = right;
+    }
+}
+
+/** Represents a multiplication expression,
+ *    e.g.: (20*y) */
+class Mult extends Bin{
+    
+    Mult(Expr left, Expr right){
+        super(left, right);
+    }
+}
+
+/** Represents an addition expression,
+ *    e.g.: (x+10) */
+class Sum extends Bin{
+    
+    Sum(Expr left, Expr right){
+        super(left, right);
+    }
+}
+
+/** Represents a variable expression,
+ *    e.g.: x */
+class Var implements Expr{
+    String name;
+    
+    Var(String name){
+        this.name = name;
+    }
+}
+
+/** Represents an integer value/expression,
+ *    e.g.: 45 */
+class Val implements Expr{
+    int i;
+    
+    Val(int i){
+        this.i = i;
+    }
+}
+
+/** Represents an equality/definition/assignment,
+ *    e.g.: y = (2*(x*x)) */
+class Equation{
+    Var v;
+    Expr e;
+    
+    Equation(Var v, Expr e){
+        this.v = v;
+        this.e = e;
+    }
+}
+
+/** Represents an arbitrary equation of expressions,
+ *    e.g.: (x*3) = (y+45)*/
+class JoshEquation{
+    Expr el;
+    Expr er;
+    
+    JoshEquation(Expr el, Expr er){
+        this.el = el;
+        this.er = er;
+    }
+}
+
+
+/** Examples/Tests */
+class LectureExamples{
+    LectureExamples(){}
+
+    Expr ten = new Val(10);
+    Expr twenty = new Val(20);
+    
+    Var x = new Var("x");
+    Var y = new Var("y");
+    Var z = new Var("z");
+    
+    Expr e1 = new Sum(new Mult(this.x, this.y),
+                      new Val(4));
+    
+    Equation eq1 = new Equation(this.x, this.ten);
[…]

### 2013-02-25 web/12.su.reg/today/ListVisitor.java

@@ -1,0 +1,116 @@
+
+
+
+
+
+
+
+
+
+// Represents a computation over a list of X producing a Result
+interface ListVisitor<X,Result> {
+    Result visitEmpty();
+    Result visitCons(X first, List<X> rest);
+}
+
+interface List<X> {
+    <R> R accept(ListVisitor<X,R> v);
+}
+
+class Empty<X> implements List<X> {
+    Empty() {}
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitEmpty();
+    }
+}
+
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+    Cons(X first, List<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitCons(this.first, this.rest);
+    }
+}
+
+class Contains implements ListVisitor<String,Boolean> {
+    String element;
+    Contains(String element) {
+	this.element = element;
+    }
+    public Boolean visitEmpty() {
+	return false;
+    }
+
+    public Boolean visitCons(String first, List<String> rest) {
+	return first.equals(this.element)
+	    || rest.accept(this);
+    }
+}
+
+/*
+ ;; [List Integer] -> Integer
+ (define (max ls)
+    (cond [(empty? ls) (error "boo")]
+          [(cons? ls) (max-acc (rest ls) (first ls))]))
+
+ ;; [List Integer] Integer -> Integer
+ (define (max-acc ls a)
+   (cond [(empty? ls) a]
+         [(cons? ls)
+          (cond [(> a (first ls)) (max-acc (rest ls) a)]
+                [else             (max-acc (rest ls) (first ls))])]))
+ */
+
+class Max implements ListVisitor<Integer,Integer> {
+    public Integer visitEmpty() {
+	throw new RuntimeException("boo");
+    }
+    public Integer visitCons(Integer first, List<Integer> rest) {
+	return rest.accept(new MaxAcc(first));
+    }
+}
+
+class MaxAcc implements ListVisitor<Integer,Integer> {
+    Integer a;
+    MaxAcc(Integer a) {
+	this.a = a;
+    }
+    public Integer visitEmpty() {
+	return this.a;
+    }
+    public Integer visitCons(Integer first, List<Integer> rest) {
+	if (this.a > first) {
+	    return rest.accept(this);
[…]

### 2013-02-25 web/12.su.reg/today/QList.java

@@ -1,0 +1,230 @@
+import java.util.Iterator;
+
+interface List<X> extends Iterable<X> {
+   // Cons given element on to this list.
+   List<X> cons(X x);
+
+   // Get the first element of this list (only defined on non-empty lists).
+   X first();
+
+   // Get the rest of this list (only defined on non-empty lists).
+   List<X> rest();
+
+   // Get the ith element of this list
+   // (only defined for lists of i+1 or more elements).
+   X get(Integer i);
+
+   // Compute the number of elements in this list.
+   Integer size();
+
+   // Is this list empty?
+   Boolean isEmpty();
+
+   // Accept given visitor
+   <R> R accept(ListVisitor<X,R> v);
+}
+
+interface ListVisitor<X,R> {
+    R visitEmpty();
+    R visitCons(X first, List<X> rest);
+}
+
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+    Cons(X first, List<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+
+    public List<X> cons(X x) {
+	return new Cons<X>(x, this);
+    }
+    public List<X> rest() {
+	return this.rest;
+    }
+    public X first() {
+	return this.first;
+    }
+    public Integer size() {
+	return 1 + this.rest.size();
+    }
+    public X get(Integer i) {
+	return i.equals(0) ? this.first : this.rest.get(i-1);
+    }
+    public Iterator<X> iterator() {
+	return new ListIterator<X>(this);
+    }
+    public Boolean isEmpty() { return false; }
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitCons(this.first, this.rest);
+    }
+}
+
+class ListIterator<X> implements Iterator<X> {
+    List<X> ls;
+    ListIterator(List<X> ls) {
+	this.ls = ls;
+    }
+    public boolean hasNext() {
+	return ls.isEmpty();
+    }
+    public X next() {
+	X r = this.ls.first();
+	this.ls = this.ls.rest();
+	return r;
+    }
+    public void remove() {}
+}
+
+interface QList<X> extends List<X> {
+    // Cons given element on to this list.
+    QList<X> cons(X x);
+
+    // Get the rest of this list (only defined on non-empty lists).
+    QList<X> rest();
+}
+
[…]

### 2013-02-25 web/12.su.reg/today/QListTests.java

@@ -1,0 +1,33 @@
+
+interface Fun<X,Y> {
+    Y apply(X x);
+}
+
+class EmptyConsSizeOne implements Fun<Integer, Boolean> {
+    Boolean apply(Integer i) {
+	return new QEmpty<Integer>().cons(i).size() == 1;
+    }
+}
+
+
+void testQList(Tester t) {
+
+    Fun<Integer,Boolean> p = new EmptyConsSizeOne();
+
+    QList<String> mt = new QEmpty<String>();
+    t.checkExpect(mt.size(), 0);
+
+    Random r = new Random();
+    for (Integer j : new HalfOpen(0, 1000)) {
+	t.checkExpect(p.apply(r.nextInt()), true);
+    }
+
+    t.checkExpect(p.apply(45), true);
+    t.checkExpect(p.apply(36), true);
+
+
+    // t.checkExpect(mt.cons("Hi").size(), 1);
+    // t.checkExpect(mt.cons("Bye").size(), 1);
+    t.checkExpect(mt.cons("Hi").cons("there").size(), 2);
+    //...
+}

### 2013-02-25 web/12.su.reg/today/Round.java

@@ -1,0 +1,12 @@
+import tester.*;
+
+class Examples {
+    void testRound(Tester t) {
+	Integer x = 3 / 4;
+	Integer y = 1 / 4;
+	Integer z = 5 / 4;
+	t.checkExpect(x, 0);
+	t.checkExpect(y, 0);
+	t.checkExpect(z, 1);
+    }
+}

### 2013-02-25 web/12.su.reg/today/SafeCast.java

@@ -1,0 +1,79 @@
+import tester.*;
+
+// Safe casting approach to structural equality.
+
+// Represents a geometric shape
+interface Shape {
+    // Is this shape the same as that shape?
+    Boolean same(Shape that);
+
+    // Is this shape a square?
+    Boolean isSq();
+    // Is this shape a circle?
+    Boolean isCirc();
+    // Convert this shape to a square
+    Sq toSq();
+    // Convert this shape to a circle
+    Circ toCirc();
+}
+
+abstract class AShape implements Shape {
+    public Boolean isSq() { return false; }
+    public Boolean isCirc() { return false; }
+    public Sq toSq() {
+	throw new RuntimeException("BARF");
+    }
+    public Circ toCirc() {
+	throw new RuntimeException("BARF");
+    }
+}
+
+// Represent a square
+class Sq extends AShape {
+    Integer size;
+    Sq(Integer size) {
+	this.size = size;
+    }
+
+    // Is this square the same as that shape?
+    public Boolean same(Shape that) {
+	return that.isSq()
+	    && this.size.equals(that.toSq().size);
+    }
+
+    // Is this square a square?
+    public Boolean isSq() { return true; }
+    // Convert this square to a square
+    public Sq toSq() { return this; }
+}
+
+// Represents a circle
+class Circ extends AShape {
+    Integer radius;
+    Circ(Integer radius) {
+	this.radius = radius;
+    }
+
+    // Is this circle a circle?
+    public Boolean isCirc() { return true; }
+    // Convert this circle to a circle
+    public Circ toCirc() { return this; }
+
+    // Is this circle the same as that shape?
+    public Boolean same(Shape that) {
+	return that.isCirc()
+	    && this.radius.equals(that.toCirc().radius);
+    }
+}
+
+class Examples {
+    Examples() {}
+
+    void testSame(Tester t) {
+	t.checkExpect(new Sq(5).same(new Sq(5)), true);
+	t.checkExpect(new Sq(5).same(new Sq(7)), false);
+	t.checkExpect(new Circ(5).same(new Circ(5)), true);
+	t.checkExpect(new Circ(5).same(new Circ(7)), false);
+	t.checkExpect(new Circ(5).same(new Sq(5)), false);
+    }
+}

### 2013-02-25 web/12.su.reg/today/Sort.java

@@ -1,0 +1,106 @@
+import tester.*;
+
+// Represents a list of integers
+interface LoI {
+    // Sort this list of integers in ascending order.
+    LoI sort();
+
+    // Insert the given integer into this sorted list of integers,
+    // to produce a sorted list in ascending order.
+    LoI insert(Integer i);
+
+    // Reverse the elements of this list.
+    LoI reverse();
+
+    // Put the given element at the end of this list.
+    LoI putAtEnd(Integer i);
+}
+
+// Represents an empty list of integers
+class Empty implements LoI {
+    Empty() {}
+
+    // Sort this empty list of integers in ascending order.
+    public LoI sort() {
+	return this;
+    }
+
+    public LoI insert(Integer i) {
+	return new Cons(i, this);
+    }
+
+    // Reverse the elements of this empty list.
+    public LoI reverse() {
+	return this;
+    }
+
+    // Put the given element at the end of this empty list.
+    public LoI putAtEnd(Integer i) {
+	return new Cons(i, this);
+    }
+}
+
+// Represents an non-empty list of integers
+class Cons implements LoI {
+    Integer first;
+    LoI rest;
+    Cons(Integer first, LoI rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+
+    // Sort this non-empty list of integers in ascending order.
+    public LoI sort() {
+	return this.rest.sort().insert(this.first);
+    }
+
+    public LoI insert(Integer i) {
+	if (i > this.first) {
+	    return new Cons(this.first,
+			    this.rest.insert(i));
+	} else {
+	    return new Cons(i, this);
+	}
+    }
+
+    // Reverse the elements of this non-empty list.
+    public LoI reverse() {
+	return this.rest.reverse().putAtEnd(this.first);
+    }
+
+    // Put the given element at the end of this non-empty list.
+    public LoI putAtEnd(Integer i) {
+	return new Cons(this.first,
+			this.rest.putAtEnd(i));
+    }
+}
+
+class Examples {
+    Examples() {}
+
+    LoI mt = new Empty();
+    LoI one = new Cons(1, mt);
+    LoI l1 = new Cons(4, new Cons(3, new Cons(1, mt)));
+    LoI l2 = new Cons(2, new Cons(3, new Cons(1, mt)));
[…]

### 2013-02-25 web/12.su.reg/today/Sort2.java

@@ -1,0 +1,201 @@
+import tester.*;
+
+// foldr : [List X] [X Y -> Y] Y -> Y
+// (foldr (list x1 x2 ... xn) f b)
+// = (f x1 (f x2 ... (f xn b)))
+/*
+(define (foldr xs f b)
+  (cond [(empty? xs) b]
+        [(cons? xs)
+         (f (first xs)
+            (foldr (rest xs) f b))]))
+*/
+
+// Represents a list of X
+    interface List<X> {
+    // Sort this list of integers in ascending order.
+    List<X> sort(Comparison<X> c);
+
+    // Insert the given integer into this sorted list of integers,
+    // to produce a sorted list in ascending order.
+    List<X> insert(X i, Comparison<X> c);
+
+    // Reverse the elements of this list.
+    List<X> reverse();
+
+    // Put the given element at the end of this list.
+    List<X> putAtEnd(X i);
+
+    // Fold given combiner function over the elements of this list
+    <Y> Y foldr(Combiner<X,Y> combine, Y base);
+}
+
+// Represents a function that combines a given X and Y into a Y.
+interface Combiner<X,Y> {
+    Y apply(X x, Y y);
+}
+
+// Represents an empty list of integers
+class Empty<X> implements List<X> {
+    Empty() {}
+
+    // Sort this empty list of integers in ascending order.
+    public List<X> sort(Comparison<X> c) {
+	return this;
+    }
+
+    public List<X> insert(X i, Comparison<X> c) {
+	return new Cons<X>(i, this);
+    }
+
+    // Reverse the elements of this empty list.
+    public List<X> reverse() {
+	return this;
+    }
+
+    // Put the given element at the end of this empty list.
+    public List<X> putAtEnd(X i) {
+	return new Cons<X>(i, this);
+    }
+
+    public <Y> Y foldr(Combiner<X,Y> combine, Y base) {
+	return base;
+    }
+}
+
+// Represents an ordering on elements of type X.
+interface Comparison<X> {
+    // Is a greater than b?
+    Boolean greaterThan(X a, X b);
+}
+
+class IntegerGreaterThan implements Comparison<Integer> {
+    IntegerGreaterThan() {}
+    public Boolean greaterThan(Integer a, Integer b) {
+	return a > b;
+    }
+}
+
+class Runner {
[…]

### 2013-02-25 web/12.su.reg/lect/AL.java
More summer lecture stuff.

@@ -1,0 +1,112 @@
+import tester.*;
+
+import java.util.ArrayList;
+
+// Represents a dinosaur.
+class Dino {
+    String eyecolor;
+    Integer teeth;
+    
+    Dino(String eyecolor, Integer teeth) {
+	this.eyecolor = eyecolor;
+	this.teeth = teeth;
+    }
+}
+
+class LectureExamples {
+
+    ArrayList<Dino> alist = new ArrayList<Dino>();
+
+    Dino d1 = new Dino("blue", 2);
+    Dino d2 = new Dino("yellow", 70);
+    Dino d3 = new Dino("black", 6);
+    Dino d4 = new Dino("chartreuse", 50);
+    Dino d5 = new Dino("red", 8);
+    
+    // Methods of ArrayList
+    // void clear()
+    // boolean isEmpty()
+    // int size()
+    // boolean add(E element)
+    // void add(int index, E element)
+    // E get(int index)
+    // E set(int index, E element)
+    // E remove(int index)
+
+    void testAll(Tester t) {
+	t.checkExpect(this.alist.isEmpty(), true);
+	t.checkExpect(this.alist.size(), 0);
+	this.alist.add(this.d1);
+	t.checkExpect(this.alist.isEmpty(), false);
+	t.checkExpect(this.alist.size(), 1);
+	t.checkExpect(this.alist.get(0), this.d1);
+	this.alist.add(this.d2);
+	t.checkExpect(this.alist.get(0), this.d1);
+	t.checkExpect(this.alist.get(1), this.d2);
+	this.alist.clear();
+	t.checkExpect(this.alist.isEmpty(), true);
+	this.alist.add(this.d1);
+	this.alist.add(this.d2);
+	this.alist.add(0, this.d3);
+	t.checkExpect(this.alist.get(0), this.d3);
+	
+	this.alist.add(this.d4);
+	// this.alist.add(5);
+	t.checkExpect(this.alist.get(0).eyecolor, "black");
+
+	// d3, d1, d2, d4
+	t.checkExpect(this.alist.set(0, this.d5), this.d3);
+	// d5, d1, d2, d4
+	t.checkExpect(this.alist.get(0), this.d5);
+	this.swap(this.alist, 0, 3);
+	t.checkExpect(this.alist.get(0), this.d4);
+	t.checkExpect(this.alist.get(3), this.d5);
+	// d4, d1, d2, d5
+	this.alist.remove(1);
[…]

### 2013-02-25 web/12.su.reg/lect/BT.java

@@ -1,0 +1,155 @@
+import tester.*;
+
+// Interp: An IBT represents a binary tree of
+// integers with data only at the leaves.
+interface IBT {
+    // Count the number of leaves in this tree.
+    Integer countLeaves();
+
+    // Is this tree in non-decreasing order from left to right?    
+    Boolean isOrdered();
+
+    // Get the value of the leftmost leaf of this tree.
+    Integer leftMost();
+    
+    // Get the value of the rightmost leaf of this tree.
+    Integer rightMost();
+
+    // Insert the given number into this ordered tree.
+    IBT insert(Integer n);
+}
+    
+class Leaf implements IBT {
+    Integer data;
+    Leaf(Integer data) {
+	this.data = data;
+    }
+
+    // TEMPLATE
+    // FIELDS:
+    // this.data     -- Integer
+    // METHODS:
+    // this.countLeaves()    -- Integer
+    // this.isOrdered()      -- Boolean
+    // this.leftMost()       -- Integer
+    // this.rightMost()      -- Integer
+
+    // Count the number of leaves in this leaf.
+    public Integer countLeaves() {
+	return 1;
+    }
+    
+    // Is this leaf in non-decreasing order from left to right?
+    public Boolean isOrdered() {
+	return true;
+    }
+
+    // Get the value of the leftmost leaf of this leaf.
+    public Integer leftMost() {
+	return this.data;
+    }
+
+    // Get the value of the rightmost leaf of this leaf.
+    public Integer rightMost() {
+	return this.data;
+    }
+
+    // Insert the given number into this ordered leaf.
+    public IBT insert(Integer n) {
+	if (this.data <= n) {
+	    return new Node(this, new Leaf(n));
+	} else {
+	    return new Node(new Leaf(n), this);
+	}
+    }
+}
+
+class Node implements IBT {
+    IBT left;
+    IBT right;
+    Node(IBT left, IBT right) {
+	this.left = left;
+	this.right = right;
+    }
+
+    // TEMPLATE
+    // FIELDS:
+    // this.left     -- IBT
+    // this.right    -- IBT
+    // METHODS:
+    // this.countLeaves()        -- Integer
+    // this.isOrdered()          -- Boolean
+    // this.leftMost()           -- Integer
+    // this.rightMost()          -- Integer    
+    // METHODS FOR FIELDS:
[…]

### 2013-02-25 web/12.su.reg/lect/BlobWorldFun.java

@@ -1,0 +1,395 @@
+import java.awt.Color;
+import java.util.Random;
+
+import tester.*;
+
+import funworld.*;
+import geometry.*;
+import colors.*;
+
+/**
+ * Copyright 2012 Viera K. Proulx
+ * This program is distributed under the terms of the 
+ * GNU Lesser General Public License (LGPL)
+ */
+
+/** Class that represents a colored disk that moves around the Canvas */
+class Blob{
+	
+	Posn center;
+	int radius;
+	IColor col;
+	//ImageMaker image = new ImageMaker("shark.png");
+	
+	/** The constructor */
+	Blob(Posn center, int radius, IColor col) {
+		this.center = center;
+		this.radius = radius;
+		this.col = col;
+	}
+	
+	/** draw this blob in the given World */
+	boolean draw(Canvas c) {
+		//return c.drawImage(image, this.center);
+		return c.drawDisk(this.center, this.radius, this.col);
+	} 
+	
+	/** produce the image of this blob at its current location and color */
+	WorldImage blobImage(){
+		//return new DiskImage(this.center, this.radius, this.col);
+		return new FromFileImage(this.center, "shark.png");
+	}
+	
+	/** move this blob 20 pixels in the direction given by the ke
+	 or change its color to Green, Red or Yellow */
+	public Blob moveBlob(String ke){
+		if (ke.equals("right")){
+			return new Blob(new Posn(this.center.x + 5, this.center.y),
+							this.radius, this.col);
+		}
+		else if (ke.equals("left")){
+			return new Blob(new Posn(this.center.x - 5, this.center.y),
+							this.radius, this.col);
+		}
+		else if (ke.equals("up")){
+			return new Blob(new Posn(this.center.x, this.center.y - 5),
+							this.radius, this.col);
+		}
+		else if (ke.equals("down")){
+			return new Blob(new Posn(this.center.x, this.center.y + 5),
+							this.radius, this.col);
+		}
+		// change the color to Y, G, R
+		else if (ke.equals("Y")){
+			return new Blob(this.center, this.radius, new Yellow());
+		}    
+		else if (ke.equals("G")){
+			return new Blob(this.center, this.radius, new Green());
+		}    
+		else if (ke.equals("R")){
[…]

### 2013-02-25 web/12.su.reg/lect/Book.java

@@ -1,0 +1,1 @@
+import tester.*;

### 2013-02-25 web/12.su.reg/lect/CartPt.java

@@ -1,0 +1,125 @@
+import tester.*;
+
+// CS 2510 - Spring 2012
+// Assignment 4 Problem 2
+// 8 February 2012
+
+// to represent a point on canvas
+class Posn{
+  int x;
+  int y;
+  
+  Posn(int x, int y){
+    this.x = x; 
+    this.y = y;
+  }
+}
+
+// to represent a location in the game program
+public class CartPt extends Posn{
+  double doublex;
+  double doubley;
+  
+  static CartPt ptN = new CartPt(0.0, -1.0);
+  static CartPt ptS = new CartPt(0.0, 1.0);
+  static CartPt ptE = new CartPt(1.0, 0.0);
+  static CartPt ptW = new CartPt(-1.0, 0.0);
+  
+  static double sin45 = Math.sin(Math.PI/4);
+  
+  static CartPt ptNE = new CartPt(CartPt.sin45, -CartPt.sin45);
+  static CartPt ptNW = new CartPt(-CartPt.sin45, -CartPt.sin45);
+  static CartPt ptSW = new CartPt(-CartPt.sin45, CartPt.sin45);
+  static CartPt ptSE = new CartPt(CartPt.sin45, CartPt.sin45);
+  
+  CartPt(Double doublex, Double doubley){
+    super(doublex.intValue(), doubley.intValue());
+    this.doublex = doublex;
+    this.doubley = doubley;
+  }
+  
+  CartPt(int x, int y){
+    super(x, y);
+    this.doublex = 1.0 * x;
+    this.doubley = 1.0 * y;
+  }
+  
+  // compute the distance from this point to the given point
+  double distTo(CartPt that){
+    return 
+    Math.sqrt((this.doublex - that.doublex) * (this.doublex - that.doublex) +
+              (this.doubley - that.doubley) * (this.doubley - that.doubley));
+  }
+  
+  // produce the point on the unit circle in the same direction as this point
+  CartPt unitDirectionTo(CartPt that){
+    return new CartPt(that.doublex / this.distTo(that),
+                      that.doubley / this.distTo(that));
+  }
+  
+  // produce a String that describes the direction for traveling 
+  // from this point to the given one
+  String directionTo(CartPt that){
+    
+    if (CartPt.ptN.distTo(this.unitDirectionTo(that)) < 0.3)
[…]

### 2013-02-25 web/12.su.reg/lect/Dir.java

@@ -1,0 +1,114 @@
+// Non-empty list of X
+interface ConsListof<X> {
+    // Finds the (first) element of the list that minimizes the output
+    // of the function.
+    X argmin(IFun<X,Double> f);
+}
+
+// List of X
+interface Listof<X> {
+    // Finds the (first) element of the list or minX, that minimizes
+    // the output of the function.
+    X argminAcc(IFun<X,Double> f, X minX);
+}
+
+// [X -> Y]
+interface IFun<X,Y> {
+    Y apply(X x);
+}
+
+class Empty<X> implements Listof<X> {
+    Empty() {}
+
+    public X argminAcc(IFun<X,Double> f, X minX) {
+	return minX;
+    }
+}
+
+class Cons<X> implements Listof<X>, ConsListof<X> {
+    X first;
+    Listof<X> rest;
+
+    Cons(X first, Listof<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+    
+    // Finds the (first) element of the list that minimizes the output
+    // of the function.
+    public X argmin(IFun<X,Double> f) {
+	return this.rest.argminAcc(f, this.first);
+    }
+
+    public X argminAcc(IFun<X,Double> f, X minX) {
+	Double fMinX = f.apply(minX);
+	Double fThisX = f.apply(this.first);
+	if (fThisX < fMinX) {
+	    return this.rest.argminAcc(f, this.first);
+	} else {
+	    return this.rest.argminAcc(f, minX);
+	}
+    }		  
+}
+
+class Posn {
+    Double x;
+    Double y;
+
+    Posn(Double x, Double y) {
+	this.x = x;
+	this.y = y;
+    }
+
+    // Euclidean distance between this and that posn.
+    public Double dist(Posn that) {
+	return Math.sqrt(Math.pow(this.x - that.x, 2) +
+			 Math.pow(this.y - that.y, 2));
+    }
+
+    // Move this posn by given vector.
+    public Posn moveBy(Posn v) {
+	return new Posn(this.x + v.x, this.y + v.y);
+    }
+
+    // Choose best direction that moves this toward that posn.
+    public Posn chooseDir(Posn that) {
+	return Dir.directions.argmin(new DistMoveBy(this, that));
+    }
+}
+
+// Represents: (λ (d) (dist (move-by origin d) dest))
[…]

### 2013-02-25 web/12.su.reg/lect/Employees.java

@@ -1,0 +1,156 @@
+// CS U2510 Fall 2011
+// Lab 4: Employees.java
+
+/*
+ *                      +-----------+
+ *                      |    Emp    |<-----------------------+
+ *                      +-----+-----+                        |
+ *                           / \                             |
+ *                          +---+                            |
+ *                            |                              |
+ *            +---------------+-----------------+            |
+ *            |                                 |            |
+ *    +-------+--------+                 +------+---------+  |
+ *    |    Worker      |                 |    Boss        |  |
+ *    +----------------+                 +----------------+  |
+ *    | String name    |                 | String name    |  |
+ *    | Number tasks  |                 | String unit    |  |
+ *    +----------------+                 | Number tasks  |  |
+ *                              +--------| ILoE   peons   |  |
+ *                              |        +----------------+  |
+ *                              V                            |
+ *                        +--------------+                   |
+ *                        |     ILoE     |<------------------+--+
+ *                        +------+-------+                   |  |
+ *                              / \                          |  |
+ *                             +---+                         |  |
+ *                               |                           |  |
+ *                +--------------+--------------+            |  |
+ *                |                             |            |  |
+ *        +-------+------+              +-------+-------+    |  |
+ *        |     MtLoE    |              |    ConsLoE    |    |  |
+ *        +--------------+              +---------------+    |  |
+ *        +--------------+              | Emp  first    |----+  |
+ *                                      | ILoE rest     |-------+
+ *                                      +---------------+
+ */
+
+// Represents an Employee of a company
+interface Emp{
+}
+
+// Represents an employee without subordinates
+class Worker implements Emp{
+    String name;
+    Number tasks;
+
+    Worker(String name, Number tasks){
+        this.name = name;
+        this.tasks = tasks;
+    }
+
+    /* Template:
+     *   Fields:
+     *     ... this.name ...   -- String
[…]

### 2013-02-25 web/12.su.reg/lect/ExamplesWorld.java

@@ -1,0 +1,96 @@
+
+
+import java.awt.Color;
+
+import funworld.*;
+import geometry.*;
+import colors.*;
+
+
+//-----------------------------------------------------------------------------
+// A very simple example of a world
+public class ExamplesWorld extends World{
+  Posn pos;
+  boolean worldEnd = false;
+
+  ExamplesWorld(Posn pos, boolean worldEnd){
+    this.pos = pos;
+    this.worldEnd = worldEnd;
+  }
+  
+  ExamplesWorld(Posn pos){
+    this(pos, false);
+  }
+
+  // just a red ball moving across the screen
+  public WorldImage makeImage(){
+    return 
+    new OverlayImages(new RectangleImage(this.pos, 60, 20, new Red()),
+      new OverlayImages(new TextImage(this.pos, "hello", 12, 0, new Blue()),
+        new LineImage(this.pos, new Posn(this.pos.x + 5, this.pos.y - 5), 
+            new Black())));
+  }
+
+  // test all kinds of actions using key events:
+  // left and right keys move the ball in the given direction
+  // up key returns the ball to the left edge
+  // "E" ends the world
+  // "s" invokes endOfTime method
+  // "q" throws a runtime exception and quits
+  public World onKeyEvent(String ke){
+    System.out.println("Key event " + ke);
+    if (ke.equals("right"))
+      return new ExamplesWorld(new Posn(this.pos.x + 20, this.pos.y));
+    if (ke.equals("left"))
+      return new ExamplesWorld(new Posn(this.pos.x - 20, this.pos.y));
+    if (ke.equals("up"))
+      return new ExamplesWorld(new Posn(0, this.pos.y));
+    if (ke.equals("E"))
+      return new ExamplesWorld(this.pos, true);
+    if (ke.equals("q"))
+      throw new RuntimeException("We quit!!!");
+    return this;
+  }
+
+  /**
+   * Check whether the World ended.
+   */
+  public WorldEnd worldEnds (){
+    // if the blob is outside the canvas, stop
+    if (this.worldEnd)
+      return 
+      new WorldEnd(true, 
+          this.makeLastImage("End of the World!!", 
+                             new Posn(100, 40), Color.red));
+    else
+       return new WorldEnd(false, this.makeImage());
+   }
+  
+  
+  // at each tick print the current position of the ball
+  public World onTick(){
[…]

### 2013-02-25 web/12.su.reg/lect/Gift.java

@@ -1,0 +1,178 @@
+import tester.*;
+
+/*
+Make examples of data - including a gift with at least three wraps 
+   
+Design the method that counts the wraps for a gift
+
+Design the method that determines whether a gift contains
+ the given jewel in the given box?
+
+Design the method that will produce a new gift, same as before, 
+ but with the given jewel inside
+*/
+
+// to represent a gift of a jewel in a box and wrapped several times
+interface IGift{
+
+  // count the wraps for this gift?
+  public int countWraps();
+  
+  //is this a gift with the given jewel and box?
+  public boolean sameJewelBox(String jewel, String box); 
+
+  // make a new gift just like this one but with the given jewel inside
+  public IGift newJewel(String jewel);    
+}
+
+// to represent the jewel inside the gift box
+class Jewel{
+  String name;
+  
+  Jewel(String name){
+    this.name = name;
+  }
+  
+  // does this jewel have the given name?
+  public boolean sameJewel(String name){
+    return this.name.equals(name);
+  }
+}
+
+// to represent the gift box with a jewel inside
+class Box implements IGift{
+  String kind;
+  Jewel jewel;
+  
+  Box(String kind, Jewel jewel){
+    this.kind = kind;
+    this.jewel = jewel;
+  }
+
+  //count the wraps for this gift?
+  public int countWraps(){
+    return 0;
+  }
+  
+  //is this a a gift with the given jewel and box?
+  public boolean sameJewelBox(String jewel, String box){
+    if (this.kind.equals(box))
+      return this.jewel.sameJewel(jewel);
+    else
+      return false;
+  } 
+
+  // make a new gift just like this one but with the given jewel inside
+  public IGift newJewel(String jewel){
+    return new Box(this.kind, new Jewel(jewel));
+  }
+}
+
+// to represent a gift of a jewel in a box, wrapped several times
+class Wrap implements IGift{
+  IGift gift;
+  
+  Wrap(IGift gift){
+    this.gift = gift;
+  }
+
+  //count the wraps for this gift?
+  public int countWraps(){
+    return 1 + this.gift.countWraps();
+  }
+  
+  //is this a a gift with the given jewel and box?
+  public boolean sameJewelBox(String jewel, String box){
+    return
+    this.gift.sameJewelBox(jewel, box);
+  }
+
[…]

### 2013-02-25 web/12.su.reg/lect/IBT.java

@@ -1,0 +1,34 @@
+import tester.*;
+
+// to represent a binary tree with no information in the nodes
+// and integer data in the leaves
+interface IBT {}
+
+// to represent a leaf of a binary tree
+class Leaf implements IBT{
+    Integer data;
+    
+    Leaf(Integer data) {
+	this.data = data;
+    }
+}
+
+// to represent a node in a binary tree
+class Node implements IBT {
+    IBT left;
+    IBT right;
+    
+    Node(IBT left, IBT right){
+	this.left = left;
+	this.right = right;
+    }
+}
+
+// Examples for binary trees
+class ExamplesBTs{
+    ExamplesBTs(){}
+    
+    IBT l3 = new Leaf(3);
+    IBT l5 = new Leaf(5);
+    IBT tree = new Node(this.l3, new Node(this.l5, this.l3));
+}

### 2013-02-25 web/12.su.reg/lect/IBTmethods.java

@@ -1,0 +1,204 @@
+import tester.*;
+
+// Interp: represents a binary tree with no information in the nodes
+// and integer data in the leaves
+interface IBTmethods {    
+    // count the leaves in this binary tree
+    public Integer countLeaves();
+    
+    // are the leaf values ordered from left to right?
+    public Boolean orderedHuh();
+    
+    // find the value of the rightmost leaf in this tree
+    public Integer rightMost();
+    
+    // find the value of the leftmost leaf in this tree
+    public Integer leftMost();    
+
+    // compute maximal element of this tree.
+    public Integer max();
+
+    // compute minimal element of this tree.
+    public Integer min();    
+}
+
+// to represent a leaf of a binary tree
+class Leaf implements IBTmethods{
+    Integer data;
+    
+    Leaf(Integer data){
+	this.data = data;
+    }
+    
+    /* TEMPLATE
+     * FIELDS:
+     * ... this.data ...                    -- Integer
+     * METHODS:
+     * ... this.countLeaves() ...           -- Integer
+     * ... this.orderedHuh() ...            -- Boolean
+     * ... this.rightMost() ...             -- Integer
+     * ... this.leftMost() ...              -- Integer
+     */
+
+    // count the leaves in this leaf
+    public Integer countLeaves() {
+	return 1;
+    }  
+    
+    // are the leaf values ordered from left to right?
+    public Boolean orderedHuh() {
+	return true;
+    }
+    
+    // find the value of the rightmost leaf in this leaf
+    public Integer rightMost() {
+	return this.data;
+    }
+    
+    // find the value of the leftmost leaf in this leaf
+    public Integer leftMost(){
+	return this.data;
+    }
+
+    // compute maximal element of this leaf.
+    public Integer max() {
+	return this.data;
+    }
+
+    // compute minimal element of this leaf.
+    public Integer min() {
+	return this.data;
+    }    
+}
+
+// to represent a node in a binary tree
+class Node implements IBTmethods {
+    IBTmethods left;
+    IBTmethods right;
+    
+    Node(IBTmethods left, IBTmethods right){
+	this.left = left;
+	this.right = right;
+    }
+    
+    /* TEMPLATE
+     * FIELDS:
+     * ... this.left ...                    -- IBT
[…]

### 2013-02-25 web/12.su.reg/lect/Lecture_09.java

@@ -1,0 +1,161 @@
+/***********************************************
+ *  CS2510 Spring 2011
+ *  Lecture #9
+ *  Constructor Overloading, Real Java,
+ *    and Visibility
+ ***********************************************/
+import tester.*;
+
+// Represents the Time
+class Time{
+    int hour;
+    int min;
+    int sec;
+
+    // Complete constructor: verifies inputs
+    Time(int hour, int min, int sec){
+        if (0 <= hour && hour < 24)
+            this.hour = hour;
+        else
+            throw new RuntimeException("Incorrect hours data");
+        if (0 <= min && min < 60)
+            this.min = min;
+        else
+            throw new RuntimeException("Incorrect minutes data");
+        if (0 <= sec && sec < 60)
+            this.sec = sec;
+        else
+            throw new RuntimeException("Incorrect seconds data");
+    }
+
+    // Initialize hours/min, with default of zero seconds
+    Time(int hour, int min){
+        this(hour, min, 0);
+    }
+
+    // Initialize hours/min/sec using AM/PM variant
+    Time(int hour, int min, int sec, boolean am){
+        this(hour, min, sec);
+
+        // Adjust for the PM hours, but verify the hours again
+        if (hour < 12 && !am)
+            this.hour = hour + 12;
+        else
+            throw new RuntimeException("Incorrect hours data");
+    }
+
+    // Produce the time after the given hours and minutes have elapsed
+    // Throws an exception if teh resulting time is the next day
+    Time addTime(int hours, int minutes){
+        if ((this.hour + hours + ((min + minutes) / 60)) > 23)
+            throw new RuntimeException("Tomorrow");
+        else
+            return new Time(this.hour + hours + ((min + minutes) / 60),
+                            (this.min + minutes) % 60,
+                            this.sec);
+    }
+}
+
+// Represents the Time, Part II
+class Time2{
+    // Seconds since midnight
+    int sec;
+
+    // We make this private, so *we* have the power
+    private Time2(int sec){
+        this.sec = sec;
+    }
+
+    // Time2 is now a direct replacement for Time1
+    Time2(int hours, int min, int sec){
+        this(hours * 3600 + min * 60 + sec);
+    }
+    // And, we can overload constructors as before (with defaults)
+    Time2(int hours, int min){
+        this(hours, min, 0);
+    }
+
+    /** Template
+     *    Fields:
+     *     ... this.sec ... -- int
+     *
+     *    Methods:
[…]

### 2013-02-25 web/12.su.reg/lect/Shapes.java

@@ -1,0 +1,96 @@
+import tester.*;
+
+// to represent a geometric shape
+interface IShape {
+    public <Fred> Fred accept(IShapeVisitor<Fred> v);
+}
+
+// Represents a Double computation over shapes
+interface IShapeVisitor<T> {
+    T visitCircle(Integer x, Integer y, Integer rad);
+    T visitRect(Integer x, Integer y, Integer w, Integer h);
+    T visitCombo(IShape top, IShape bot);
+}
+
+class ShapeArea implements IShapeVisitor<Double> {
+    public Double visitCircle(Integer x, Integer y, Integer rad) {
+	return Math.pow(rad, 2) * Math.PI;
+    }
+
+    public Double visitRect(Integer x, Integer y, Integer w, Integer h) {
+	return w * h * 1.0;
+    }
+
+    public Double visitCombo(IShape top, IShape bot) {
+	return top.accept(this) + bot.accept(this);
+    }
+}
+
+// to represent a circle
+class Circle implements IShape {
+    Integer x;
+    Integer y;
+    Integer rad;
+	
+    Circle(Integer x, Integer y, Integer rad) {
+	this.x = x;
+	this.y = y;
+	this.rad = rad;
+    }
+
+    public <Fred> Fred accept(IShapeVisitor<Fred> v) {
+	return v.visitCircle(this.x, this.y, this.rad);
+    }
+}
+
+// to represent a rectangle
+class Rect implements IShape {
+    Integer x;
+    Integer y;
+    Integer w;
+    Integer h;
+	
+    Rect(Integer x, Integer y, Integer w, Integer h) {
+	this.x = x;
+	this.y = y;
+	this.w = w;
+	this.h = h;
+    }
+
+    public <Fred> Fred accept(IShapeVisitor<Fred> v) {
+	return v.visitRect(this.x, this.y, this.w, this.h);
+    }
+}
+
+// to represent a combined shape
+class Combo implements IShape {
+    IShape top;
+    IShape bot;
+	
+    Combo(IShape top, IShape bot) {
+	this.top = top;
+	this.bot = bot;
+    }
+
+    public <Fred> Fred accept(IShapeVisitor<Fred> v) {
+	return v.visitCombo(this.top, this.bot);
+    }
+}
+
+class ExamplesShapes{
+    ExamplesShapes() {}
+	
+    IShape circle = new Circle(50, 50, 50);
[…]

### 2013-02-25 web/12.su.reg/lect/Shark.java

@@ -1,0 +1,407 @@
+import tester.*;
+  
+/* --- THE OCEAN WORLD -----------------------------------------------
+Fish swim across the screen from right to left. There may be one fish,
+or a whole school of fish, or none at a time. A hungry shark swims in
+place at the left side, moving only up and down, controlled by the
+\scheme{"up"} and \scheme{"down"} keys. It tries to catch and eat the
+fish. It gets bigger with each fish it eats, it keeps getting hungrier
+and hungrier as the time goes on between the catches. It dies of
+starvation, if not fed in time.
+--------------------------------------------------------------------*/
+
+import colors.*;
+import draw.*;
+import geometry.*;
+
+import image.*;
+
+import java.util.Random;
+import java.awt.Color;
+
+/*
++----------+
+| Shark    |
++----------+
+| int y    |
+| int life |
++----------+
+
+*/
+
+// to represent a shark in an ocean world
+class Shark {
+  int y;
+  int life;
+
+  //Image sharkPic = new FromFile("shark.png");
+  
+  
+  Shark(int y, int life) {
+    this.y = y;
+    this.life = life;
+  }
+  
+  // produce a new shark from this shark moved in response to the key press
+  Shark onKeyEvent(String ke){
+    if (ke.equals("up"))
+      return new Shark(this.y - 3, this.life);
+    else if (ke.equals("down"))
+      return new Shark(this.y + 3, this.life);
+    else
+      return this;
+  }
+  
+  // draw this shark on the given Canvas
+  boolean draw(Canvas c){
+    return c.drawDisk(new Posn(15, this.y), this.life / 10 + 2, new Black()) &&
+           c.drawLine(new Posn(15, this.y), 
+                      new Posn(40, this.y), 
+                      new Black());
+  }
+  
+  // the image of this shark, its size depends on its life health
+  Image sharkImage(){
+    return 
+    //this.sharkPic.overlay(
+        new Circle(this.life / 10 + 2, "solid", new Black());//);
+  }
+  
+//the image of this shark, its size depends on its life health
+  Scene drawShark(Scene scene){
+    //return scene.placeImage(this.sharkImage(), new Posn(15, this.y));
+    return scene.placeImage(this.sharkImage(), 
[…]

### 2013-02-25 web/12.su.reg/lect/SimpleWorld.java

@@ -1,0 +1,75 @@
+import tester.*;
+import funworld.*;
+import geometry.*;
+import colors.*;
+import java.awt.Color;
+
+class NotClicked extends World {
+    NotClicked() {}
+    public World onMouseClicked(Posn mouse) {
+	//....
+    }
+
+    public WorldImage makeImage() {...}
+}
+
+class Clicked extends World {
+    Posn mouse;
+    Clicked(Posn mouse) {
+	this.mouse = mouse;
+    }
+    
+    public World onMouseClicked(Posn mouse) {
+	//....
+    }
+
+    public WorldImage makeImage() {...}
+}
+
+    
+
+class SimpleWorld extends World {
+
+    String key;
+    Posn mouse;
+
+    SimpleWorld() {
+	this("");
+    }
+
+    SimpleWorld(String key, Posn mouse) {
+	this.key = key;
+	this.mouse = mouse;
+    }
+
+    public World onTick() { return this; }
+
+    public World onMouseClicked(Posn mouse) {
+	// ...
+    }
+
+    public WorldImage makeImage() {
+	return 
+	    new RectangleImage(new Posn(300, 150), 
+			       600, 
+			       300, 
+			       Color.blue)
+	    .overlayImages(new TextImage(new Posn(300, 150), 
+					 this.key, 
+					 100, 
+					 Color.white));
+    }
+
+    public World onKeyEvent(String key) {
+	if (key.equals("q")) {
+	    this.stopWorld();
+	    return this;
+	} else {
+	    return new SimpleWorld(key);
+	}
+    }
+
+    public static void main(String[] args) {
+	new SimpleWorld().bigBang(600, 300, 0.1);
+    }
+}

### 2013-02-25 web/12.su.reg/lect/TickyTack.java

@@ -1,0 +1,231 @@
+import java.awt.Color;
+import java.util.Random;
+
+import tester.*;
+
+import funworld.*;
+import geometry.*;
+import colors.*;;
+
+/**
+ * Copyright 2011 Viera K. Proulx
+ * This program is distributed under the terms of the 
+ * GNU Lesser General Public License (LGPL)
+ */
+
+class House{
+  Posn loc;    // of the SW corner of the house base
+  int width;   // the width of the house
+  int height;  // the height of the house
+  Color color; // the color of the house
+  
+  House(Posn loc, int width, int height, Color color){
+    this.loc = loc;
+    this.width = width;
+    this.height = height;
+    this.color = color;
+  }
+  
+  // make the image of this house
+  WorldImage houseImage(){
+    return 
+    new RectangleImage(
+        // the pinhole in the center
+        new Posn(this.loc.x + this.width / 2, this.loc.y - this.height / 2), 
+        this.width, this.height, this.color).
+    overlayImages(
+        // the roof - the height is half the height of the house
+        new TriangleImage(new Posn(this.loc.x, this.loc.y - this.height), 
+                          new Posn(this.loc.x + this.width, this.loc.y - this.height),
+                          new Posn(this.loc.x + this.width / 2, 
+                                   this.loc.y - 3 * this.height / 2),
+                          Color.red),
+        // the door - black in the middle
+        new RectangleImage(new Posn(this.loc.x + this.width / 2,
+                                    this.loc.y - this.height / 4),
+                           this.width / 2, this.height / 2, Color.gray));
+  }
+}
+
+class Tree{
+  Posn loc;         // the left base of the trunk
+  int trunkHeight;  // the height of the trunk
+  int width;        // the width of the crown of the tree
+  int height;       // the height of the crown of the tree
+  
+  Tree(Posn loc, int trunkHeight, int width, int height){
+    this.loc = loc;
+    this.trunkHeight = trunkHeight;
+    this.width = width;
+    this.height = height;
+  }
+  
+  WorldImage treeImage(){
+    return
+    // the trunk
[…]

### 2013-02-25 web/12.su.reg/lect/TimeLect.java

@@ -1,0 +1,164 @@
+/***********************************************
+ *  CS2510 Spring 2011 
+ *  Lecture #9
+ *  Constructor Overloading, Real Java,
+ *    and Visibility
+ ***********************************************/
+import tester.*;
+
+class Time{
+	int hour;
+	int min;
+	int sec;
+
+	// complete constructor: verifies the input data
+	Time(int hour, int min, int sec){
+		if (0 <= hour && hour < 24)
+			this.hour = hour;
+		else
+			throw new RuntimeException("Incorrect hours data");
+		if (0 <= min && min < 60)
+			this.min = min;
+		else
+			throw new RuntimeException("Incorrect minutes data");
+		if (0 <= sec && sec < 60)
+			this.sec = sec;
+		else
+			throw new RuntimeException("Incorrect seconds data");
+	}
+
+	// initialize hours and minutes with default zero for seconds
+	Time(int hour, int min){
+		this(hour, min, 0);
+	}
+
+	// initialize hours using AM/PM variant
+	Time(int hour, int min, int sec, boolean am){
+		this(hour, min, sec);
+
+		// adjust for the PM hours, but be careful to verify the hours again
+		if (hour < 12 && !am)
+			this.hour = hour + 12;
+		else 
+			throw new RuntimeException("Incorrect hours data");			
+	}
+
+	// produce the time after the given hours and minutes elapsed
+	// throw an exception if this brings us to the next day
+	Time addTime(int hours, int minutes){
+		if (this.hour + hours + (min + minutes) / 60 > 23)
+			throw new RuntimeException("Tomorrow");
+		else
+			return new Time(this.hour + hours + (min + minutes) / 60, 
+					(this.min + minutes) % 60, this.sec);
+	}	
+}
+
+// Represents the Time, Part II
+class Time2{
+    // Seconds since midnight
+    int sec;
+
+    // We make this private, so *we* have the power 
+    private Time2(int sec){
+        this.sec = sec;
+    }
+
+    // Time2 is now a direct replacement for Time1
+    Time2(int hours, int min, int sec){
+        this(hours * 3600 + min * 60 + sec);
+    }
+    // And, we can overload constructors as before (with defaults)
+    Time2(int hours, int min){
+        this(hours, min, 0);
+    }
+
+    /** Template
+     *    Fields:
+     *     ... this.sec ... -- int
+     *    
+     *    Methods:
+     *     ... this.tick() ...     -- Time2
[…]

### 2013-02-25 web/12.su.reg/lect/WormWorldImage.java

@@ -1,0 +1,425 @@
+import java.awt.Color;
+
+import draw.*;
+import geometry.*;
+import colors.*;
+
+import image.*;
+
+//------------------------------------------------------------------------------
+//represent the world of a Worm Game: Worm, Food, and the Bounding Box
+public class WormWorldImage extends World {
+
+  protected WormImage w;
+  protected FoodImage f;
+  protected BoxImage  b;
+
+  public WormWorldImage(WormImage w, FoodImage f, BoxImage b) {
+    this.w = w;
+    this.f = f;
+    this.b = b;
+  }
+
+  // what happens when the clock ticks
+  public World onTick() {
+    //System.out.println("Tick: (" + this.w.head.x + ", " + this.w.head.y + ")");
+    return this.move(this.w.move(this.w.whereTo()));
+  }
+
+  // what happens when the player presses a key
+  public World onKeyEvent(String ke) {
+    //System.out.println("Key event " + ke);
+    //System.out.println("Worm at (" + this.w.head.x + ", " + this.w.head.y + ")");
+    return this.move(this.w.move(ke));
+  }
+
+  // what happens when the player presses a key
+  public World onMouseClicked(Posn loc) {
+    //System.out.println("Worm at (" + this.w.head.x + ", " + this.w.head.y + ")");
+    //System.out.println("Mouse click at (" + loc.x + ", " + loc.y + ")");
+    return this.moveFood(loc);
+  }
+  
+  // keep the worm, but move the food to the new location
+  protected World moveFood(Posn loc){
+    return new WormWorldImage(
+        this.w,
+        new FoodImage(loc.x, loc.y),
+        this.b);
+  }
+  
+  public WorldEnd worldEnds(){
+ // worm ate itself
+    if (this.w.ateItself()){
+      return new WorldEnd(true, this.lastScene("Your worm ate itself."));
+    }
+
+    // worm ran into the wall   
+    else if (this.w.ranIntoWall(this.b)){
+      return new WorldEnd(true, this.lastScene("Your worm ran into a wall."));
+    }
+    else
+      return new WorldEnd(false, this.toDraw());
+  }
+
+  // see where the worm moved to - and proceed with the game 
+  protected World move(WormImage newWorm) {
+ 
+    // worm found food
+    //else 
+      if (newWorm.canEat(this.f)){
+      return new WormWorldImage(
+          this.w = newWorm.eat(this.f),
+          this.f.create(this.b),
+          this.b);
+    }
+
[…]

### 2013-02-25 web/12.su.reg/exams/exam1.tex
Summer exam stuff.

--- 2012-05-17 web/12.su.reg/exams/exam2.tex
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 1 -- Summer 2012}
 
@@
 
-Binary trees are one of the most widely used data structures in
-computer science.  Here are three examples of binary trees of numbers:
+A writing utensil may either be a pen or pencil.  A pen has a color
+and an amount of ink, measured in microlitres (\(\mu\)l).  A pencil
+has a graphite lead, measured in millimetres (mm).
 
-\begin{verbatim}
-         5        1           3
-                   \         / \
-                    2       4  18
-                              /  \
-                             9    12
-                            /    /
-                           7    3
-\end{verbatim}
-
-Nodes in a binary tree carry data, which in this case are numbers, and
-a node has a left and right subtree.  Binary trees, like lists, may
-contain \emph{any} number of elements, including 5, 8, 19, and 0.
+Five microlitres of ink will write a line seven metres long, while a
+graphite lead of 3 millimetres will write a line 22 metres long.
 
 \begin{enumerate}
-\item Design a data definition for binary trees.  It must adequately
-  represent any possible binary tree.
+\item Design a data representation of writing utensils.
 
-\item Give class diagrams corresponding to the data definition from
-  part A.
+\item Give three examples of writing utensils.
 
-\item Design a {\tt mirror} method that computes the mirror image of a
-  binary tree.  For example, the mirror image of the above trees would
-  be:
-
-\begin{verbatim}
-     5        1           3
-             /           / \
-            2           18  4
-                       /  \
-                     12    9
-                       \    \
-                        3    7
-\end{verbatim}
+\item Design a method for writing utensils that determines how long
+  of a line it can write.
 \end{enumerate}
-
 \end{problem}

### 2013-02-25 web/12.su.reg/labs/ExamplesArrayList2.java
Summer lab stuff.

@@ -1,0 +1,192 @@
+import tester.*;
+import java.util.*;
+
+/**
+ * A class designed to explore and validate the methods defined in the 
+ * <code>ArrayList</code> class.
+ * 
+ * @since 13 March 2012
+ */
+class ExamplesArrayList{
+  ExamplesArrayList(){}
+  
+  /** A sample <code>ArrayList</code> */
+  ArrayList<Person> arlist = new ArrayList<Person>();
+  
+  ArrayList<Person> arlistSame = arlist;
+  ArrayList<Person> arlistOther;
+  
+  Person pete = new Person("Pete", 23);
+  Person jill = new Person("Jill", 19);
+  Person kim = new Person("Kim", 20);
+  Person pat = new Person("Pat", 21);
+  
+  /**
+   * Test the method newAge for the class Person
+   *
+   * @param t The tester that runs the tests and reports the results
+   */
+  public void testNewAge(Tester t){
+    t.checkExpect(this.jill.newAge(25), new Person("Jill", 25));
+  }
+  
+  /**
+   * Test the method birthday for the class Person
+   *
+   * @param t The tester that runs the tests and reports the results
+   */
+  public void testBirthday(Tester t){
+    // invoke the method
+    this.jill.birthday();
+    
+    // verify the effects
+    t.checkExpect(this.jill, new Person("Jill", 20));
+    
+    // reset
+    this.jill.age = 19;
+  }
+  
+  /**
+   * EFFECT:
+   * Initialize the array list with four data items
+   */
+  public void initList(){
+    this.arlist.clear();
+    this.arlist.add(this.pete);
+    this.arlist.add(this.jill);
+    this.arlist.add(this.kim);
+    this.arlist.add(this.pat);
+    this.arlistOther = new ArrayList<Person>(this.arlist); 
+  }
+  
+  /**
+   * EFFECT:
+   * Run the tests to validate the design of the <code>ArrayList</code> class.
+   * 
+   * @param t The tester that runs the tests and reports the results
+   */
+  public void testArrayList(Tester t){
+    // check the initial values
+    t.checkExpect(this.arlist.size(), 0);
+    t.checkExpect(this.arlistSame.size(), 0);
+    t.checkExpect(this.arlistOther, null);
+    
+    // initialize the lists and check the effects
+    this.initList();
+    t.checkExpect(this.arlist.size(), 4);
+    t.checkExpect(this.arlistSame.size(), 4);
[…]

### 2013-02-25 web/12.sp.reg/exams/exam3.tex
Spring 2012 reg exam stuff.

\documentclass[11pt]{article}

%% ------------------------------------------------------------------
%% SOLUTIONS:
\def\thel{\noindent\rule{2.5cm}{.5pt}}
\long\def\begsol#1 #2\endsol{#1}
\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
%%uncomment the above line to get the solutions printed

\newcommand\code[1]{\texttt{#1}}
\newcommand\bcode[1]{\texttt{\textbf{#1}}}

%% PROBLEMS:
\def\pts#1{\marginpar{\footnotesize \raggedright  \fbox{#1 {\sc Points}}}}
\newcounter{Pctr}
\newenvironment{problem}{\stepcounter{Pctr}%
\begin{description}
\item[\noindent{\bf Problem} \arabic{Pctr}]
\end{description}}{\relax}
%% ------------------------------------------------------------------

%% SUBPARTS:
\newcounter{parts}
\renewcommand{\theparts}{\Alph}
%% ------------------------------------------------------------------

\begin{document}

\renewcommand{\theenumi}{\Alph{enumi}}
\setcounter{Pctr}{0}

%% ------------------------------------------------------------------

\vfill
\centerline{\Large CS 2510 Exam 3 -- Spring 2012}

~\\[2cm]

\begin{center}
\begin{tabular}{l@{\qquad}l}
Name:                        & \rule{200pt}{.1pt} \\[.5cm]
Student Id (last 4 digits):  & \rule{200pt}{.1pt} \\[.5cm]
\end{tabular}
\end{center}

\noindent\begin{minipage}{7.5cm} $\bullet$ Write down answers in the
space provided.

$\bullet$ You may use all syntax you know from class (that is, the
parts of Java we have studied in class), and you may use any Java
Collection Framework classes you know how to use.

$\bullet$ When defining methods, you do not need to give a complete
class definition---just indicate in which class your method definition
should be placed.

$\bullet$ For tests you only need to provide the expression that
computes the actual value, connecting it with an arrow to the expected
value. For example \code{s.method() -> true} is sufficient.

$\bullet$ Remember that the phrase ``design a class'' or ``design a
method'' means more than just providing a definition. It means to
design them according to the \textbf{design recipe}.  You are
\textit{not} required to provide a method template unless the problem
specifically asks for one.  However, be prepared to struggle if you
choose to skip the template step.

$\bullet$ A word of advice: Decompose complex problems into small
steps.

$\bullet$ We will not answer \textit{any} questions during the exam.

\bigskip

\textit{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{4.5cm}
\begin{tabular}{|c|l|@/r|}
\hline
\textbf{Problem} & Points & \\ \hline
A & & 10\\ \hline
B & & 15\\ \hline
C & & 10\\ \hline
\textbf{Total} &  & 30 \\ \hline
\end{tabular}
\end{minipage}

\vfill\thispagestyle{empty}
\newpage

%% -----------------------------------------------------------------------------
%% 
\pts{10}
\begin{problem}
Design the \texttt{merge} method of the the \texttt{MergeAlgorithm}
class, whose purpose is to take two sorted lists (both are assumed to
be ordered by a given comparator) and to produce a single merged list
that contains all the elements of the two lists in sorted order.

\begin{verbatim}
import java.util.*;
import tester.*;

class MergeAlgorithm {

  <T> ArrayList<T> merge(ArrayList<T> list1,
                         ArrayList<T> list2,
                         Comparator<T> comp) {
         ...
  }
}
\end{verbatim}

\noindent
The \texttt{merge} method should not mutate any of its arguments.

\begsol{\vspace{0.5in}}
\begin{verbatim}
class Merge {

    // Merge two sorted lists into a single sorted list.
    // ASSUME: list1, list2 are sorted according to comp.
    <T> ArrayList<T> merge(ArrayList<T> list1,
			   ArrayList<T> list2,
			   Comparator<T> comp) {

	ArrayList<T> result = new ArrayList<T>();
	Integer i1 = 0;
	Integer i2 = 0;

	while (!i1.equals(list1.size()) &&
	       !i2.equals(list2.size())) {

	    if (comp.compare(list1.get(i1), list2.get(i2)) < 0) {
		result.add(list1.get(i1));
		i1 = i1 + 1;
	    } else {
		result.add(list2.get(i2));
		i2 = i2 + 1;
	    }
	}
	while (!i1.equals(list1.size())) {
	    result.add(list1.get(i1));
	    i1 = i1 + 1;
	}
	while (!i2.equals(list2.size())) {
	    result.add(list2.get(i2));
	    i2 = i2 + 1;
	}

	return result;
    }
}

class CharComp implements Comparator<Character>{
  public int compare(Character c1, Character c2){
    return c1.compareTo(c2);
  }
}

class ExamplesMerge{
  ExamplesMerge(){}

  ArrayList<Character> alphabet1 = 
      new ArrayList<Character>(Arrays.asList(
          'a', 'c', 'e', 'g', 'i', 
          'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's'));

  ArrayList<Character> alphabet2 = 
      new ArrayList<Character>(Arrays.asList(
          'b', 'd', 'f', 'h', 'j', 
          't', 'u', 'v', 'w', 'x', 'y', 'z'));

  ArrayList<Character> alphabet = 
      new ArrayList<Character>(Arrays.asList(
          'a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 
          'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 
          't', 'u', 'v', 'w', 'x', 'y', 'z'));

  ArrayList<Character> result = 
      new ArrayList<Character>();

  CharComp comp = new CharComp();
  Merge ma = new Merge();

  void testMerge(Tester t){
    t.checkExpect(ma.merge(alphabet1, alphabet2, comp, result), alphabet);
  }
}
\end{verbatim}
\endsol

\end{problem}

\newpage
[Here is some more space for the previous problem.]

\newpage
%% -----------------------------------------------------------------------------
%% 
\pts{15}
\begin{problem}
Class \texttt{CircularList} is similar to the \texttt{Dequeue} you
have designed in your homeworks, but has only one head node and the
last node in the list is connected back to the head - making the list
circular.

\begin{verbatim}
// to represent one node in a circular list
class Node<T>{
  T data;
  Node<T> prev;
  Node<T> next;

  Node(T data, Node<T> prev, Node<T> next){
    this.data = data;
    this.prev = prev;
    this.next = next;
  }

}

// to represent a circular list
class Circular<T>{
  Node<T> ht;

  Circular(){
    this.ht = new Node(null, null, null);
    this.ht.prev = this.ht;
    this.ht.next = this.ht;
  }
...
}
\end{verbatim}

Design the following methods for this circular list:

\begin{enumerate}
\newpage
\item Design the method \texttt{isEmpty()} for the
  \texttt{Circularlist}.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  // is this list empty?
  public boolean isEmpty(){
    return(this.ht.next == this.ht);
  }

\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{void addAtFront(T t)} for the
  \texttt{Circularlist} that adds the node with given item as its
  value at the front of this list.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  // add the given item to the front of this list
  public void addAtFront(T t){
    Node<T> oldLast = this.ht.next;
    Node<T> n = new Node<T>(t, this.ht, oldLast);
    this.ht.next = n;
    oldLast.prev = n;
  }
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{void addAtTail(T t)} for the
  \texttt{Circularlist} that adds the node with given item as its
  value at the tail of this list.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  //add the given item to the tail of this list
  public void addAtTail(T t){
    Node<T> oldLast = this.ht.prev;
    Node<T> n = new Node<T>(t, oldLast, this.ht);
    this.ht.prev = n;
    oldLast.next = n;
  }
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{ T removeFromFront()} for the
  \texttt{Circularlist} that removes the node at the front and returns
  its value.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  //remove the given item from the front of this list
  public T removeFromFront(){
    if (this.isEmpty())
      throw new RuntimeException("Cannor remove from an empty list");
    else{
      Node<T> removed = ht.next;
      ht.next = removed.next;
      removed.next.prev = ht;
      return removed.data;
    }
  }
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{ T removeFromTail()} for the
  \texttt{Circularlist} that removes the node at the tail and returns
  its value.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  //remove the given item from the tail of this list
  public T removeFromTail(){
    if (this.isEmpty())
      throw new RuntimeException("Cannot remove from an empty list");
    else{
      Node<T> removed = ht.prev;
      ht.prev = removed.prev;
      removed.prev.next = ht;
      return removed.data;
    }
  }
\end{verbatim}
\endsol

\newpage
[Here is some more space for the previous problem.]

\begsol{\vspace{0.5in}}
\begin{verbatim}
class ExamplesCircular{
  ExamplesCircular(){}

  Circular<String> cir = new Circular<String>();

  void testCircular(Tester t){
    t.checkExpect(this.cir.isEmpty(), true);
    //cir.addAtFront("hello");
    //cir.addAtFront("ciao");
    //cir.addAtFront("bye");

    cir.addAtTail("hello");
    cir.addAtTail("ciao");
    cir.addAtTail("bye");

    System.out.println("Adding at the tail:");

    System.out.println(this.cir.ht.next.data);
    System.out.println(this.cir.ht.next.next.data);
    System.out.println(this.cir.ht.next.next.next.data);

    System.out.println(this.cir.ht.prev.data);
    System.out.println(this.cir.ht.prev.prev.data);
    System.out.println(this.cir.ht.prev.prev.prev.data);

    t.checkExpect(this.cir.isEmpty(), false);
    t.checkExpect(this.cir.removeFromTail(), "bye");
    t.checkExpect(this.cir.removeFromTail(), "ciao");
    t.checkExpect(this.cir.removeFromFront(), "hello");

    t.checkExpect(this.cir.isEmpty(), true);

    cir.addAtFront("hello");
    cir.addAtFront("ciao");
    cir.addAtFront("bye");

    System.out.println("Adding at the front:");

    System.out.println(this.cir.ht.next.data);
    System.out.println(this.cir.ht.next.next.data);
    System.out.println(this.cir.ht.next.next.next.data);

    System.out.println(this.cir.ht.prev.data);
    System.out.println(this.cir.ht.prev.prev.data);
    System.out.println(this.cir.ht.prev.prev.prev.data);

    t.checkExpect(this.cir.isEmpty(), false);
    t.checkExpect(this.cir.removeFromFront(), "bye");
    t.checkExpect(this.cir.removeFromFront(), "ciao");
    t.checkExpect(this.cir.removeFromTail(), "hello");

    t.checkExpect(this.cir.isEmpty(), true);
  }
}
\end{verbatim}
\endsol

\end{enumerate}
\end{problem}

\newpage
%% -----------------------------------------------------------------------------
%% 
\pts{10}
\begin{problem}
You are now masters of encoding and decoding algorithms. Here is a new
one for you to implement: When encoding the letters, leave the first
letter the same as the original. Replace the second letter of the
message with a letter that is one space later in the alphabet. (So, if
the second letter was `e', you would replace it with `f'.) Replace the
third letter with one that is two letters later in the alphabet. So,
if the third letter was `l', you would replace it with `n'.) Proceed
like this till the end. Of course, circle around to the beginning of
the alphabet, if you get to the end. (So, if `w' appears as the 7th
letter, it will be replaced by `b' --- we included a 27th character,
the space.)

For example, ``hello world'' would be encoded as ``hfnosebvzun''.

We are giving you the \texttt{ArrayList<Character>} that contains all
letters of the alphabet and the space character at the end:

\begin{verbatim}
interface DTree {
  /** The original list of characters used in encoding and decoding */
  ArrayList<Character> alphabet = 
        new ArrayList<Character>(Arrays.asList(
            'a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 
            'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 
            't', 'u', 'v', 'w', 'x', 'y', 'z', ' '));
}
\end{verbatim}

\newpage
\begin{enumerate}
\item Design the method \texttt{encode} that consumes a message
  \texttt{String} and produces the encoded \texttt{String}

\begsol{\vspace{0.5in}}
\begin{verbatim}
class Encoder implements DTree{
  Encoder(){}

  String encode(String source, String acc){
    if (source.length() > 0)
    return encode(source.substring(1),
        acc + alphabet.get((alphabet.indexOf(source.charAt(0)) + acc.length()) % 27));
    else 
      return acc;
  }

  String decode(String source, String acc){
    if (source.length() > 0)
    return decode(source.substring(1),
        acc + alphabet.get((alphabet.indexOf(source.charAt(0)) - acc.length()+ 27) % 27));
    else 
      return acc;
  }

}
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{decode} that consumes an encoded message
  \texttt{String} and produces the original message \texttt{String}

\begsol{\vspace{0.5in}}
\begin{verbatim}
class ExamplesExam3{
  ExamplesExam3(){}

  Encoder enc = new Encoder();

  void testEncoder(Tester t){
    t.checkExpect(enc.encode("hello world", ""), "hfnosebvzun");
    t.checkExpect(enc.encode("how are you", ""), "hpycewkgfxd");

    t.checkExpect(enc.decode("hfnosebvzun", ""), "hello world");
    t.checkExpect(enc.decode("hpycewkgfxd", ""), "how are you");
  }

  public static void main(String[] argv){
    ExamplesExam3 e = new ExamplesExam3();
    Encoder enc = new Encoder();

    System.out.println("hello world" + " encoded as " + 
        enc.encode("hello world", ""));
    System.out.println("how are you" + " encoded as " + 
        enc.encode("how are you", ""));

    System.out.println("hello world" + " decoded as " + 
        enc.decode("hfnosebvzun", ""));
    System.out.println("how are you" + " decoded as " + 
        enc.decode("hpycewkgfxd", ""));

    Tester.runFullReport(e);
  }
}
\end{verbatim}
\endsol

\end{enumerate}
\end{problem}

\end{document}

### 2013-02-25 web/12.sp.reg/exams/Try.java

@@ -1,0 +1,33 @@
+import java.util.*;
+
+interface DTree{
+  /** The original list of characters used in encoding and decoding */
+  ArrayList<Character> alphabet = 
+        new ArrayList<Character>(Arrays.asList(
+            'a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 
+            'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 
+            't', 'u', 'v', 'w', 'x', 'y', 'z', ' '));
+}
+
+class Try implements DTree{
+    public static void main(String[] args) {
+	for (String s : args)
+	    System.out.println(encode(s, ""));
+    }
+
+    static String encode(String source, String acc){
+	if (source.length() > 0)
+	    return encode(source.substring(1),
+			  acc + alphabet.get((alphabet.indexOf(source.charAt(0)) + acc.length()) % 27));
+	else
+	    return acc;
+    }
+
+    static String decode(String source, String acc){
+	if (source.length() > 0)
+	    return decode(source.substring(1),
+			  acc + alphabet.get((alphabet.indexOf(source.charAt(0)) - acc.length()+ 27) % 27));
+	else
+	    return acc;
+    }
+}

### 2013-02-25 web/12.sp.reg/exams/exam3.java

@@ -1,0 +1,45 @@
+import java.util.*;
+import tester.*;
+
+class Merge {
+
+    // Merge two sorted lists into a single sorted list.
+    // ASSUME: list1, list2 are sorted according to comp.
+    <T> ArrayList<T> merge(ArrayList<T> list1,
+			   ArrayList<T> list2,
+			   Comparator<T> comp) {
+
+	ArrayList<T> result = new ArrayList<T>();
+	Integer i1 = 0;
+	Integer i2 = 0;
+
+	while (!i1.equals(list1.size()) &&
+	       !i2.equals(list2.size())) {
+
+	    if (comp.compare(list1.get(i1), list2.get(i2)) < 0) {
+		result.add(list1.get(i1));
+		i1 = i1 + 1;
+	    } else {
+		result.add(list2.get(i2));
+		i2 = i2 + 1;
+	    }
+	}
+	while (!i1.equals(list1.size())) {
+	    result.add(list1.get(i1));
+	    i1 = i1 + 1;
+	}
+	while (!i2.equals(list2.size())) {
+	    result.add(list2.get(i2));
+	    i2 = i2 + 1;
+	}
+
+	return result;
+    }
+}
+
+class LessThan implements Comparator<Integer> {
+    public int compare(Integer i, Integer j) {
+	return i.compareTo(j);
+    }
+}
+

### 2013-02-25 web/12.sp.reg/lect/Biggest.java
Old lecture stuff.

@@ -1,0 +1,21 @@
+import java.util.Comparator;
+
+// Find the largest element in the visited tree.
+public class Biggest<X> implements IVisitorBT<X,X> {
+    Comparator<X> comp;
+    Biggest(Comparator<X> comp) {
+	this.comp = comp;
+    }
+
+    public X visitLeaf() { throw new RuntimeException("No elements in tree."); }
+    public X visitNode(X val, IBT<X> left, IBT<X> right) {
+	X biggestLeft = left.accept(new BiggestAcc<X>(val));
+	X biggestRight = right.accept(new BiggestAcc<X>(val));
+	
+	if (comp.compare(biggestLeft, biggestRight) < 0) {
+	    return biggestRight;
+	} else {
+	    return biggestLeft;
+	}
+    }
+}

### 2013-02-25 web/12.sp.reg/lect/BiggestAcc.java

@@ -1,0 +1,14 @@
+// Find the largest element in the visited tree or val.
+public class BiggestAcc implements IVisitorBT<Integer,Integer> {
+    Integer val;
+    BiggestAcc(Integer val) {
+	this.val = val;
+    }
+
+    public Integer visitLeaf() { return this.val; }
+    public Integer visitNode(Integer val, IBT<Integer> left, IBT<Integer> right) {	
+	return Math.max(val,
+			Math.max(left.accept(this),
+				 right.accept(this)));
+    }
+}

### 2013-02-25 web/12.sp.reg/lect/Hanoi.java

@@ -1,0 +1,26 @@
+class Hanoi {
+    Peg a;
+    Peg b;
+    Peg c;
+    Hanoi(Integer n) {
+	this.a = ;
+	this.b = ;
+	this.c = ;
+    }
+}
+
+interface Peg {
+    Peg makePeg(Integer n);
+}
+
+class MTPeg {}
+
+class ConsPeg {
+    Integer first;
+    Peg rest;
+    ConsPegt(Integer first, Peg rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+}
+

### 2013-02-25 web/12.sp.reg/lect/Hash.java

@@ -1,0 +1,126 @@
+import tester.*;
+import java.util.*;
+
+interface HT<K,V> {
+    // Does this hash table have given key?
+    Boolean hasKey(K key);
+
+    // Lookup given key in this hash table.
+    // Throws exception if key is not in table.
+    V lookup(K key);
+
+    // Lookup given key in this hash table.
+    // Produce "nothing" if key is not in table.
+    V or Nothing lookupMaybe(K key);
+
+    // Add given key, value to this hash table.
+    // Overwrites any existing map.
+    void add(K key, V val);
+
+    // Clear all data from this hash table.
+    void clear();
+}
+
+interface KeyVals<K,V> {
+    // Does this list contain the given key?
+    Boolean contains(K key);
+
+    // Get the value associated with the given key.
+    V get(K key);
+}
+
+class MT<K,V> implements KeyVals<K,V> {
+    public Boolean contains(K key) {
+	return false;
+    }
+
+    public V get(K key) {
+	throw new RuntimeException("No such key");
+    }
+}
+
+class Cons<K,V> implements KeyVals<K,V> {
+    K firstKey;
+    V firstVal;
+    KeyVals<K,V> rest;
+    Cons(K firstKey, V firstVal, KeyVals<K,V> rest) {
+	this.firstKey = firstKey;
+	this.firstVal = firstVal;
+	this.rest = rest;
+    }
+
+    public Boolean contains(K key) {
+	if (this.firstKey.equals(key)) {
+	    return true;
+	} else {
+	    return this.rest.contains(key);
+	}
+    }
+
+    public V get(K key) {
+	if (this.firstKey.equals(key)) {
+	    return this.firstVal;
+	} else {
+	    return this.rest.get(key);
+	}
+    }
+}
+
+class Hash<K,V> implements HT<K,V> {
+    ArrayList<KeyVals<K,V>> list;
+    Integer capacity = 500;
+
+    Hash() {
+	this.list = new ArrayList<KeyVals<K,V>>(this.capacity);
+	this.clear();
+    }
+
+    // Does this hash table have given key?
+    public Boolean hasKey(K key) {
+	return this.list.get(key.hashCode() % this.capacity).contains(key);
+    }
+
+    // Lookup given key in this hash table.
+    // Throws exception if key is not in table.
+    public V lookup(K key) {
[…]

### 2013-02-25 web/12.sp.reg/lect/Lecture_17.java

@@ -1,0 +1,143 @@
+import tester.*;
+
+// Import the ArrayList class
+import java.util.ArrayList;
+
+// Represents a Dinosaur
+class Dino{
+    String eyecolor;
+    Integer teeth;
+    
+    Dino(String eyecolor, Integer teeth){
+        this.eyecolor = eyecolor;
+        this.teeth = teeth;
+    }
+}
+
+// Examples and Tests of mutation and ArrayList methods
+class LectureExamples{
+    // The initial List
+    ArrayList<Dino> alist = new ArrayList<Dino>();
+    
+    // Some Dinos for testing
+    Dino d1 = new Dino("blue", 2);
+    Dino d2 = new Dino("yellow", 70);
+    Dino d3 = new Dino("black", 6);
+    Dino d4 = new Dino("chartreuse", 50);
+    Dino d5 = new Dino("red", 8);
+    Dino d6 = new Dino("magenta", 40);
+    Dino d7 = new Dino("indigo", 33);
+    
+    // Reset our data... add the first three Dinos
+    void reset(){
+        this.alist.clear();
+        this.alist.add(this.d1);
+        this.alist.add(this.d2);
+        this.alist.add(this.d3);
+    }
+    
+    // Examples/Tests of all the ArrayList methods
+    void testAll(Tester t){
+        /** Useful methods for ArrayList<E>
+         *     void clear()
+         *     boolean isEmpty()
+         *     int size()
+         *     boolean add(E element)
+         *     void add(int index, E element)
+         *     E get(int index)
+         *     E set(int index, E element)
+         *     E remove(int index)
+         */
+
+        this.reset();
+        t.checkExpect(this.alist.get(1).eyecolor, "yellow");
+        t.checkExpect(this.alist.get(2).eyecolor, "black");
+
+        // Throws an Exception when:  size() <= index < 0
+        t.checkException(new IndexOutOfBoundsException("Index: 6, Size: 3"),
+                this.alist, "add", 6, this.d4);
+        
+        this.alist.remove(2);
+        t.checkExpect(this.alist.size(), 2);
+        
+        this.alist.add(2, this.d3);
+        t.checkExpect(this.alist.size(), 3);
+        t.checkExpect(this.alist.get(2).teeth, 6);
+        
+        this.alist.clear();
+        t.checkExpect(this.alist.isEmpty(), true);
+        t.checkExpect(this.alist.size(), 0);
+        
+        this.reset();
[…]

### 2013-02-25 web/12.sp.reg/lect/Lecture_20.java

@@ -1,0 +1,222 @@
+/***********************************************
+ *  CS2510 Spring 2011 
+ *  Lecture #20
+ *  Loops in all forms (Recursive, While, and For)
+ ***********************************************/
+
+import tester.*;
+import java.util.ArrayList;
+
+
+/** Represents a Predicate */
+interface ISelect<T>{
+    /** Selected? */
+    boolean select(T t);
+}
+
+/** Class that implements different Algorithms */
+class Algos{
+    /** Count the number of Selected Elements from the given Traversal */
+    <T> int countSuch(Traversal<T> tr, ISelect<T> comp){
+        // Call ACC method with the BASE-VALUE
+        return this.countAcc(tr, comp, 0);
+    }
+    
+    /** Count using an Accumulator
+     *    ACC: The number selected so far */
+    <T> int countAcc(Traversal<T> tr, ISelect<T> comp, int acc){
+        // BASE-CASE (TERMINATION-PREDICATE 2000
+        if(tr.isEmpty()){
+            return acc;
+        }else{
+            // THE RECURSIVE CALL
+            return this.countAcc(
+                    // ADVANCE
+                    tr.getRest(), comp,
+                    // UPDATE NEW ACC
+                    this.updateAcc(tr.getFirst(), comp, acc));
+        }
+    }
+    /** Update the Accumulator for the countAcc Method */
+    <T> int updateAcc(T first, ISelect<T> comp, int acc){
+        if(comp.select(first))
+            return acc+1;
+        else{
+            return acc;
+        }
+    }
+    
+    /** Count the Selected elements using a While Loop */
+    <T> int countWhile(Traversal<T> tr, ISelect<T> comp){
+        // INIT
+        int acc = 0;
+        // CONTINUATION-CONDITION
+        while(!tr.isEmpty()){
+            // UPDATE
+            acc = this.updateAcc(tr.getFirst(), comp, acc);
+            // ADVANCE
+            tr = tr.getRest();
+        }
+        // POST STUFF
+        return acc;
+    }
+    
+    /** Count the Selected elements using a For Loop */
+    <T> int countFor(Traversal<T> tr, ISelect<T> comp){
+        // INIT
+        int acc = 0;
+        for(;
+            // CONTINUATION-CONDITION
+            !tr.isEmpty();
+            // ADVANCE
+            tr = tr.getRest())
+        {
+           // UPDATE 
+           acc = this.updateAcc(tr.getFirst(), comp, acc);
[…]

### 2013-02-25 web/12.sp.reg/lect/Lecture_21.java

@@ -1,0 +1,114 @@
+/***********************************************
+ *  CS2510 Spring 2011 
+ *  Lecture #21
+ *  More Loops, with a little World Fun
+ ***********************************************/
+
+import tester.*;
+import image.*;
+import world.*;
+import java.util.*;
+
+/** Crazy Sparkling Animation */
+class EyecolorWorld extends VoidWorld{
+    int WIDTH = 500;
+    int HEIGHT = 500;
+
+
+    // A list of all the sparkles
+    ArrayList<Posn> l = new ArrayList<Posn>();
+    // Random number generator
+    Random randy = new Random();
+    
+    // List of possible Colors
+    ArrayList<String> colors;
+
+    /** Construct and EyeColorWorld. Create the list of possible Colors */
+    EyecolorWorld(){
+        this.colors = new ArrayList<String>();
+        this.colors.add("indigo");
+        this.colors.add("chartreuse");
+        this.colors.add("red");
+        this.colors.add("goldenrod");
+        this.colors.add("blue");
+        this.colors.add("turquoise");
+        this.colors.add("green");
+        this.colors.add("purple");
+    }
+    
+    /** Draw random stars at the Posn locations */
+    public Scene onDraw(){
+        return this.drawAll();
+    }
+    
+    /** Move all the sparkles and remove any that drifted offscreen */
+    public void onTick(){
+        this.l = this.moveAll(this.l);
+        this.l = this.rmOff();
+    }
+    
+    /** Get the color for the given index */
+    String color(int i){
+        return this.colors.get(i);
+    }
+    
+    /** Produce a new list with all the Posns randomly moved */
+    ArrayList<Posn> moveAll(ArrayList<Posn> l){
+        ArrayList<Posn> acc = new ArrayList<Posn>();
+        Iterator<Posn> it = l.iterator();
+        
+        while(it.hasNext()){
+            Posn p = it.next();
+            acc.add(new Posn(p.x + this.randy.nextInt(21)-10,
+                             // Modified from the original to slide them
+                             //   down off screen so we don't get stuck
+                             p.y + this.randy.nextInt(21)));
+        }
+        return acc;
+    }
+
+    /** Draw all the Posns */
+    Scene drawAll(){
+        Scene acc = new EmptyScene(this.WIDTH, this.HEIGHT);
+        for(int i = 0; i < this.l.size(); i = i+1){
+            acc = acc.placeImage(new Star(this.randy.nextInt(10)+15,
[…]

### 2013-02-25 web/12.sp.reg/lect/Lecture_23.java

@@ -1,0 +1,137 @@
+/***********************************************
+ *  CS2510 Spring 2011 
+ *  Lecture #23
+ *  HasMap, HashCode, Equals (Object)
+ ***********************************************/
+
+import java.util.ArrayList;
+import tester.*;
+
+/** Our own home-brew HashTable */
+class Hasher<V>{
+    // A List of Lists with direct access
+    ArrayList<ArrayList<Pair<V>>> alist =
+        new ArrayList<ArrayList<Pair<V>>>();
+    int capacity;
+    
+    /** Create a Hasher with the given capacity */
+    Hasher(int capacity){
+        this.capacity = capacity;
+        // Mark all places as Empty
+        for(int i = 0; i < capacity; i++)
+            alist.add(new ArrayList<Pair<V>>());
+    }
+    
+    /** Return the index of the given integer in our List */
+    int hash(int i){
+        return i % this.capacity;
+    }
+    
+    /** Does this Hasher contain the given Key? */
+    boolean contains(int key){
+        for(Pair p : alist.get(hash(key))){
+            if(p.key == key)
+                return true;
+        }
+        return false;
+    }
+    
+    /** Put a mapping for [key -> val] into this Hasher */
+    void put(int key, V val){
+        // Duplicates are ok for simplicity
+        alist.get(hash(key)).add(new Pair<V>(key, val));
+    }
+    
+    /** Get the Value associated with the given Key */
+    V get(int key){
+        for(Pair<V> p : alist.get(hash(key))){
+            if(p.key == key)
+                return p.val;
+        }
+        throw new RuntimeException("No Mapping for: "+key);
+    }
+}
+
+/** Represents a pair of a Key/Value */
+class Pair<V>{
+    // Just int for simplicity
+    int key;
+    V val;
+    
+    Pair(int i, V v){
+        this.key = i;
+        this.val = v;
+    }
+    
+    /** Compute this Pair's hasCode */
+    public int hashCode(){
+        return this.key % 123457;
+    }
+    /** Is this Pair equal to the given Object? */
+    public boolean equals(Object o){
+        // Cast to Pair... possible ClassCastException at
+        //   runtime.
+        Pair<V> apear = (Pair<V>)o;
+        
+        return apear.key == this.key &&
+               apear.val.equals(this.val);
+    }
+}
+
+/** Represents a Person. Just an example of a computed HashCode */
[…]

### 2013-02-25 web/12.sp.reg/lect/Lecture_24.java

@@ -1,0 +1,183 @@
+import java.util.*;
+import tester.*;
+
+/** Represents a Graph */
+class Graph {
+    HashMap<String, ArrayList<Neighbor>> graph
+	= new HashMap<String, ArrayList<Neighbor>>();
+
+    /** Get the Neighbors of the given Node */
+    ArrayList<Neighbor> getNeighbors(String node){
+        if (!this.graph.containsKey(node)) {
+            throw new RuntimeException("Not there!!");
+	} else {
+	    return this.graph.get(node);
+	}
+    }
+
+    /** Add an Edge */
+    void addEdge(String from, String to, Integer dist){
+        if(!this.graph.containsKey(from))
+            this.graph.put(from, new ArrayList<Neighbor>());
+        if(!this.graph.containsKey(to))
+            this.graph.put(to, new ArrayList<Neighbor>());
+
+        this.graph.get(from).add(new Neighbor(to, dist));
+        this.graph.get(to).add(new Neighbor(from, dist));
+    }
+}
+
+/** Represents a Neighbor with a Distance */
+class Neighbor {
+    String name;
+    Integer distance;
+
+    Neighbor(String name, double Distance){
+        this.name = name;
+        this.distance = distance;
+    }
+}
+
+/** Represent an Edge in a Path */
+class Edge {
+    String src;
+    String targ;
+    Integer distance;
+
+    Edge(String src, String targ, Integer distance) {
+        this.src = src;
+        this.targ = targ;
+        this.distance = distance;
+    }
+}
+
+/** Represents a Path in a Graph */
+class Path {
+    ArrayList<Edge> path = new ArrayList<Edge>();
+
+    /** EFFECT: Add a new Edge to this Path */
+    void add(Edge e){
+        this.path.add(e);
+    }
+
+    /** Does this Path connect to the Target */
+    Boolean containsTarget(String targ){
+        for(Edge e : path) {
+            if(e.targ.equals(targ)) {
+                return true;
+            }
+        }
+        return false;
+    }
+
+    /** Get Edge that matches Targ */
+    Edge get(String targ){
+        for(Edge e : path){
+            if(e.targ.equals(targ)) {
+                return e;
+            }
+        }
+        throw new RuntimeException("No There!!");
+    }
+}
+
+/** Represents a ToDo Visit List */
+interface ToDo {
+
+    /** Add an Edge to this ToDo list */
+    void add(Edge s);
+
[…]

### 2013-02-25 web/12.sp.reg/lect/Lecture_28.java

@@ -1,0 +1,312 @@
+/***********************************************
+ *  Sorting Demonstration(s)
+ ***********************************************/
+
+import java.util.*;
+import tester.*;
+import image.*;
+import world.*;
+
+/** Demonstration of Insertion Sort */
+class Insertion extends ShowSort {
+    /** The partition of sorted/unsorted elements */
+    Integer part;
+
+    /** Take one step of this sorting algorithm */
+    void sortStep() {
+        insert(this.part);
+        this.part++;
+        this.current = this.part;
+        if (part.equals(this.alist.size()))
+            this.done = true;
+    }
+
+    /** Insert the element at the given index into place
+     *    at the front of the list */
+    void insert(Integer idx){
+        Integer i = 0;
+        while (i < idx && alist.get(idx) < alist.get(i))
+            i++;
+        alist.add(i, alist.remove(idx));
+    }
+    /** The Name of this Sorting Algorithm */
+    String name() { return "Insertion Sort"; }
+    /** Start with the partition at 0 */
+    void init(Integer size){
+        super.init(size);
+        this.part = 0;
+    }
+}
+
+/** Demonstration of Selection Sort */
+class Selection extends Insertion {
+    /** Swap the next maximum with the first of the
+     *    unsorted partition */
+    void sortStep() {
+        swap(findMaxIdx(this.alist), this.part);
+        this.part++;
+        this.current = findMaxIdx(this.alist);
+        if (part == this.alist.size())
+            this.done = true;
+    }
+    /** Compute the index of the maximum element */
+    int findMaxIdx(ArrayList<Integer> alist){
+        int maxI = this.part;
+        for(int i = maxI; i < alist.size(); i++)
+            if(alist.get(i) > alist.get(maxI))
+                maxI = i;
+        return maxI;
+    }
+    /** The Name of this Sorting Algorithm */
+    String name(){ return "Selection Sort"; }
+    /** First current is the maximum */
+    void init(int size){
+        super.init(size);
+        this.current = findMaxIdx(this.alist);
+    }
+}
+
+/** Demonstration of Sort */
+class Bubble extends Insertion{
+    /** Take one step of this sorting algorithm */
+    void sortStep(){
+        this.bubbleOne(this.alist);
+        this.current = this.part;
+        this.part++;
+        if(part <= 0)
+            this.done = true;
+    }
[…]

### 2013-02-25 web/12.sp.reg/lect/Today.java

@@ -1,0 +1,310 @@
+import tester.*;
+import java.util.*;
+
+class WrapAL<T> implements Traversal<T> {
+    ArrayList<T> list;
+    Integer position;
+
+    WrapAL(ArrayList<T> list) {
+	this.list = list;
+	this.position = 0;
+    }
+
+    private WrapAL(ArrayList<T> list, Integer position) {
+	this.list = list;
+	this.position = position;
+    }
+
+    public boolean isEmpty() {
+	return this.position.equals(this.list.size());
+    }
+
+    public T getFirst() {
+	return this.list.get(this.position);
+    }
+
+    public Traversal<T> getRest() {
+	return new WrapAL<T>(this.list, this.position + 1);
+    }
+}
+
+
+interface List<T> extends Traversal<T>, Iterable<T> {
+}
+
+class ListIterator<T> implements Iterator<T> {
+    Traversal<T> elems;
+    ListIterator(Traversal<T> elems) {
+	this.elems = elems;
+    }
+
+    public boolean hasNext() {
+	return !this.elems.isEmpty();
+    }
+
+    public T next() {
+	T first = this.elems.getFirst();
+	this.elems = this.elems.getRest();
+	return first;
+    }
+
+    public void remove() {
+	throw new RuntimeException("We didn't implement remove.");
+    }
+}
+
+class MT<T> implements List<T> {
+    MT() {}
+
+    public Iterator<T> iterator() {
+	return new ListIterator<T>(this);
+    }
+
+    public boolean isEmpty() {
+	return true;
+    }
+
+    public T getFirst() {
+	throw new RuntimeException("getFirst of the empty list");
+    }
+
+    public Traversal<T> getRest() {
+	throw new RuntimeException("getRest of the empty list");
+    }
+}
+
+class Cons<T> implements List<T> {
+    T first;
+    List<T> rest;
+    Cons(T first, List<T> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+
+    public Iterator<T> iterator() {
+	return new ListIterator<T>(this);
+    }
+
+    public boolean isEmpty() {
+	return false;
+    }
+
+    public T getFirst() {
+	return this.first;
+    }
+
[…]

### 2013-02-25 web/12.sp.reg/lect/bt/Examples.java

@@ -1,0 +1,39 @@
+import tester.*;
+
+class Examples {
+    IBT<Integer> mt = new Leaf<Integer>();
+    IBT<Integer> n1 = new Node<Integer>(7, mt, mt);
+    IBT<Integer> n2 = new Node<Integer>(8, mt, mt);
+    IBT<Integer> n3 = new Node<Integer>(2, n1, n2);
+    IBT<Integer> n4 = new Node<Integer>(1, mt, mt);
+    IBT<Integer> n5 = new Node<Integer>(4, n3, n1);
+
+    IBT<String> smt = new Leaf<String>();
+    IBT<String> s1 = new Node<String>("D", smt, smt);
+    IBT<String> s2 = new Node<String>("C", s1, smt);
+    IBT<String> s3 = new Node<String>("B", smt, smt);
+    IBT<String> s4 = new Node<String>("A", s3, s2);
+
+    void testStringAppend(Tester t) {
+	t.checkExpect(smt.accept(new StringAppend()), "");
+	t.checkExpect(s4.accept(new StringAppend()), "BADC");
+    }
+
+
+    void testBiggest(Tester t) {
+	t.checkExpect(n1.accept(new Biggest()), 7);
+	t.checkExpect(n5.accept(new Biggest()), 8);
+    }
+
+    void testSize(Tester t) {
+	t.checkExpect(n1.size(), 1);
+        t.checkExpect(n3.size(), 3);
+	t.checkExpect(n5.size(), 5);
+    }
+
+    void testSizeVisitor(Tester t) {
+	t.checkExpect(n1.accept(new Size<Integer>()), 1);
+        t.checkExpect(n3.accept(new Size<Integer>()), 3);
+	t.checkExpect(n5.accept(new Size<Integer>()), 5);
+    }
+}

### 2013-02-25 web/12.sp.reg/lect/bt/StringAppend.java

@@ -1,0 +1,11 @@
+class StringAppend implements IVisitorBT<String,String> {
+    public String visitNode(String val, 
+			    IBT<String> left, 
+			    IBT<String> right) {
+	return left.accept(this) + val + right.accept(this);
+    }
+
+    public String visitLeaf() {
+	return "";
+    }
+}

### 2013-02-25 exam4/NatIter.java
Exam 4 stuff.

@@ -1,0 +1,39 @@
+import java.util.*;
+import tester.*;
+
+class NaturalIterator implements Iterator<Integer>, Iterable<Integer> {
+    Integer i;
+    Integer n;
+    NaturalIterator(Integer n) {
+	this.i = 0;
+	this.n = n;
+    }
+
+    public boolean hasNext() {
+	return !this.i.equals(n);
+    }
+
+    public Integer next() {
+	Integer j = this.i;
+	this.i = i + 1;
+	return j;
+    }
+
+    public Iterator<Integer> iterator() {
+	return this;
+    }
+
+    public void remove() {}
+}
+
+class Examples {
+    void testFact(Tester t) {
+	Integer acc = 1;
+	Iterator<Integer> is = new NaturalIterator(5);
+	for(Integer i : is) {
+	    acc = i * acc;
+	    System.out.println(i);
+	}
+	t.checkExpect(acc, 125);
+    }
+}

### 2013-03-30 exam6/1.tex
start on exam

--- 2013-02-13 exam5/1.tex
@@
 {\bf Problem} & Points & out of \\ \hline
-1 & & 19\\ \hline
-2 & & 10\\ \hline
+1 & & 20\\ \hline
+2 & & 19\\ \hline
 Extra & & 5 \\ \hline
-{\bf Total} & & 28+5 \\ \hline
+{\bf Total} & & 39+5 \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 
-\begin{problem} \pts{??}
+\begin{problem} \pts{20}
 
@@
 \item Design an object-based data representation for JSON, Jr.
+
+\ifrubric
+\begin{alltt}
+;; A JSON is one of:  [5pt]
+;; - (new s% String)
+;; - Array
+;;
+;; An Array is one of
+;; - (new a0%)
+;; - (new a+% JSON Array)
+
+(define-class s%
+  (fields val))
+
+(define-class a0%)
+
+(define-class a+%
+  (fields hd tl))
+\end{alltt}
+\else
 \newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
 \item Design a method for counting the number of strings in a JSON,
   Jr.~object.
+\ifrubric
+\begin{alltt}
+;; JSON implements [1pts]
+;; - Count the number of strings in this object
+;;   count-strings : -> Natural
+
+;; in s% [1pt]
+(define (count-strings) 1)
+
+;; in a0% [1pt]
+(define (count-strings) 0)
+
+;; in a+% [1pt]
+(define (count-strings)
+  (+ (this . hd . count-strings)
+     (this . tl . count-strings)))
+
+;; [1pts]
+(define eg1 (new s% "This is JSON"))
+(define eg2 (new a+% eg1 (new a+% (new s% "Jr.") (new a0%))))
+(define eg3 (new a+% (new a0%)
+                 (new a+% (new a+% (new s% "So")
+                               (new a+% (new s% "is") (new a0%)))
+                      (new a+% (new s% "this")
+                           (new a0%)))))
+
+;; [1pts]
+(check-expect (eg1 . count-strings) 1)
+(check-expect (eg2 . count-strings) 2)
+(check-expect (eg3 . count-strings) 3)
+\end{alltt}
+\else
+\fi
 \newpage
@@
 \end{verbatim}
+\ifrubric
+\begin{alltt}
+;; in s%  [2pts]
+(define (find p)
+  (if (p . apply (this . val))
+      (this . val)
+      #f))
+
+;; in a0% [1pts]
+(define (find p) false)
+
+;; in a+% [3pts]
+(define (find p)
+  (local [(define r (this . hd . find p))]
+    (if (false? r)
+        (this . tl . find p)
+        r)))
+
+;; [2pts]
+;; A (new eq% String) implements StringPred
+(define-class eq%
+  (fields x)
+  (define (apply y)
+    (string=? (this . x) y)))
+
+;; [1pts]
+(check-expect (eg1 . find (new eq% "This is JSON")) "This is JSON")
+(check-expect (eg1 . find (new eq% "this")) false)
+(check-expect (eg2 . find (new eq% "This is JSON")) "This is JSON")
+(check-expect (eg2 . find (new eq% "this")) false)
+(check-expect (eg3 . find (new eq% "This is JSON")) false)
+(check-expect (eg3 . find (new eq% "this")) "this")
+\end{alltt}
+\else
+\fi
 \end{enumerate}
-
 \end{problem}
 
+\ifrubric
+\else
 \newpage
-\ifrubric\else
 [Here is some more space for the previous problem.]
+\fi
 \newpage
-\fi
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\newpage
 \ifrubric\else
-\newpage
 [Here is some more space for the previous problem.]

### 2013-03-30 exam6/2.tex
now it builds

--- 2013-03-30 exam6/1.tex
@@
 \vspace*{-1.5cm}
-\centerline{\Large CSU2510H Exam 1 \ifrubric (SOLUTION) \fi-- Spring 2013}
+\centerline{\Large CSU2510H Exam 2 \ifrubric (SOLUTION) \fi-- Spring 2013}
 
@@
 \begin{itemize}
-\item You may use the usual primitives and expression forms of any of
-  the \code{class} languages; for everything else, define it.
+
+\item All problems must be done in Java. You may use any Java we have
+  used in class and in lab; anything else must be defined.
 
 \item You may write {\tt {\slshape c} $\rightarrow$ {\slshape e}} as
-  shorthand for
-  {\tt (check-expect {\slshape c e})}.
+  shorthand for writing {\tt
+    Tester.checkExpect}; the {\tt Examples} class and {\tt test}
+  method around the tests are not required.
 
@@
 \end{itemize}
-
 \bigskip

### 2013-03-30 exam6/2.tex
outline the problems

--- 2013-02-13 exam5/1.tex
@@
 
-\begin{problem} \pts{??}
+\begin{problem} \pts{20}
 
-JSON is a lightweight data-interchange format.  It is widely used on
-the web and plays exactly the role of S-Expressions in universe
-programs, but for JavaScript programs.  It's used by folks like Google
-and there's tons of data on the web that's available in the JSON
-format.  So as to not be left out of the fun, your co-op employer has
-asked you to design an object representation of JSON to incorporate
-into their 2 million lines of {\tt class/2} code.  To start things off
-simply, here are some examples of what we'll call ``JSON, Jr.'':
+ Binary tree problem
 
-\begin{verbatim}
-"this is JSON"
-[ "this is JSON", "Jr." ]
-[ [], [ "So", "is" ], "this" ]
-\end{verbatim}
-The first example is just a string.  The second is an ``array'' of two
-JSON elements, which are both strings.  The third is another array,
-but of three elements: the first is an array of zero elements, the
-second is an array of two strings, and the third is a string.
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+\newpage
 
-\begin{enumerate}
-\item Design an object-based data representation for JSON, Jr.
-\newpage
-\item Design a method for counting the number of strings in a JSON,
-  Jr.~object.
-\newpage
-\item Among the 2M lines of code your employer maintains are 30,000
-  lines of code devoted to implementations of the {\tt
-    StringPred} interface, which is:
-\begin{verbatim}
-;; A StringPred implements
-;; - apply : String -> Boolean
-\end{verbatim}
-These objects represent predicates on strings.  Now that your JSON,
-Jr. library is in place, they've decided to put their predicates to
-use in searching a large corpus of JSON data.  To accomodate this,
-design the following method for JSON, Jr.~objects:
-\begin{verbatim}
-;; Find the first string in this JSON Jr object satisfying
-;; given predicate, or #f if there's no such string.
-;; find : StringPred -> String or #f
-\end{verbatim}
-\end{enumerate}
+\textbf{Extra Credit}
 
@@
 
-\newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
-\newpage
-\fi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{problem} \pts{19}
-
-% Set/Property problem
-
-In order to manage the distribution of ID cards, the University needs
-a data structure for managing sets of numbers.  Van Horn started on
-the implementation, resulting in what you see below.
-
-\begin{verbatim}
-;; A Set is one of:
-;; - (new empty%)
-;; - (new singelton% Number)
-;; - (new union% Set Set)
-;; Interpretation: a set that contains all the *unique* numbers in the data
-;; and implements
-;; - sum : -> Number
-;;   Produce the sum of all the unique elements of this set
-;; - sum/acc : Set -> Number
-;;   Produce the sum of all the unique elements of this set that are
-;;   *not* in the given set
-;; - contains : Number -> Boolean
-;;   Does this set contain the given number?
-;; - union : Set -> Set
-;;   Combine this set with the given set
-
-(define-class empty%
-  (define (sum) 0)
-  (define (sum/acc s) 0))
-
-(define-class singleton%
-  (fields n)
-  (define (sum) (this . n))
-  (define (sum/acc s) (cond [(s . contains (this . n)) 0]
-                            [else (this . n)])))
-
-(define-class union%
-  (fields l r)
-  (define (sum) (this . sum/acc (new empty%)))
-  (define (sum/acc s) (+ (this . l . sum/acc s)
-                         (this . r . sum/acc (s . union (this . l))))))
-\end{verbatim}
-
-\newpage
-
-Unfortunately, as you can see, Van Horn fell asleep before finishing,
-and the deadline is today.  Help him out by implementing the
-\verb|union| and \verb|contains| methods for all \verb|Set|s.
-
 \ifrubric
-\begin{alltt}
-\rubricpts{1 pt -- union method}
-\rubricpts{3 pts -- contains method}
-;; in empty%
-(define (union s) s)
-(define (contains n) #f)
-;; also ok:
-(define (union s) (new union% this s))
-;; in singleton%
-(define (union s) (new union% this s))
-(define (contains n) (= n (this . n)))
-;; in union%
-(define (union s) (new union% this s))
-(define (contains n) (or (this . l . contains n)
-                         (this . r . contains n)))
-(define e (new empty%))
-\rubricpts{2 pts -- tests}
-(check-expect (e . contains 5) #f)
-(check-expect ((new singleton% 5) . contains 5) #t)
-(check-expect ((new singleton% 6) . contains 5) #f)
-(check-expect ((new singleton% 6) . union  e . contains 5) #f)
-(check-expect (e . union (new singleton% 5) . contains 5) #t)
-(check-expect ((new singleton% 5) . union (new singleton% 5) . contains 5) #t)
-\end{alltt}
 \else
@@
 
-\verb|Set|s should satisfy the following property, called the
-``union-contains'' property, for any number $n$ and
-any two sets $s_1$ and $s_2$:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{quote}
-Given two sets $s_1$ and $s_2$, if $n$ is contained in the union of $s_1$ and $s_2$, then $n$ is either
-contained in $s_1$ or it is contained in $s_2$.
-\end{quote}
+\begin{problem} \pts{20}
 
-\begin{enumerate}
-\item Codify the ``union-contains'' property as a predicate.  Write a
-  test case that should pass (if your implementations work right) using the
-  predicate you have defined.
-\end{enumerate}
+Outline problem
 
 \ifrubric
-\begin{alltt}
-\rubricpts{1 pt -- contract & purpose}
-;; union-contains : Set Set Number -> Boolean
-\rubricpts{3 pts -- code}
-(define (union-contains s1 s2 n)
-  (check-expect (s1 . union s2 . contains n)
-                (or (s1 . contains n) (s2 . contains n))))
-\rubricpts{1 pt -- successful test}
-(union-contains e (new singleton% 17) 17)
-\end{alltt}
 \else
@@
 \fi
+\newpage
+\newpage
 
-\newpage
-\noindent
-The University doesn't really trust Van Horn, and they want to make
-sure that \verb|Set|s satisfy another property, called the
-``union-sum'' property:
-
-\begin{quote}
-Given two sets $s_1$ and $s_2$, computing the sum of $s_1$ and the sum
-of $s_2$ and then adding the two results is equal to the sum of the
-union of  $s_1$ and $s_2$.
-\end{quote}
-
-\noindent
-The University is considering adding this to the test suite as well.
-
-\begin{enumerate}
-\setcounter{enumi}{1}
-
-\item Codify this property as a predicate, and write a test case that
-  should pass using the predicate.
-
-\item Is the ``union-sum'' property true for all sets?  If so,
-  give an explanation.  If not, write a counter-example (i.e., a
-  failing test case) using your predicate.
-
-\end{enumerate}
-
-\begin{solution}
-\begin{alltt}
-\rubricpts{1 pt -- contract & purpose}
-;; union-sum : Set Set -> Boolean
-\rubricpts{3 pts -- code}
-(define (union-sum s1 s2 n)
-  (check-expect (s1 . union s2 . sum)
-                (+ (s1 . sum) (s2 . sum))))
-
-\rubricpts{1 pt -- successful test}
-;; successful test
-(union-sum (new singleton% 17) e)
-\rubricpts{3 pts -- failing test}
-;; It's false, counterexample:
-(union-sum (new singleton% 17) (new singleton% 17))
-\end{alltt}
-\end{solution}
-
-\newpage
-\ifrubric\else
-[Here is some more space for the previous problem.]
-\newpage
-\fi
 \end{problem}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
 \newpage
@@
 
-\begin{problem}\pts{5}
+\begin{problem} \pts{20}
 
-% Visitor on Binary Trees
+Computation/thunk problem
 
-\textbf{Extra Credit}
-\\
-\noindent
-Labich was unhappy to discover that the class languages don't have his
-favorite programming language feature---pattern matching.  After
-thinking hard about this problem, he came up with a solution, which he
-calls \emph{pattern objects}.
-
-Consider the following data defintion:
-\begin{verbatim}
-;; A BTree is one of:
-;; - (new leaf% Number)
-;; - (new node% BTree Number BTree)
-;; and implements
-;; - match : [Pattern X] -> X
-;;   applies the given pattern object to this tree
-
-;; A [Pattern X] implements:
-;; - leaf : Number -> X
-;; - node : BTree Number BTree -> X
-
-(define-class leaf%
-  (fields n)
-  ;; apply the appropriate method from the given pattern object
-  (define (match p) (p . leaf (this . n))))
-
-(define-class node%
-  (fields left n right)
-  ;; apply the appropriate method from the given pattern object
-  (define (match p) 
-    (p . node (this . left) (this . n) (this . right))))
-\end{verbatim}
-
-A pattern object is like the implementation of two different functions
-at once: one that implements a function for leaves from a binary tree,
-and one function that works on nodes.
-
-Write a pattern object that computes the product of all of the numbers
-in a binary tree.
-
-\ifrubric\else
+\ifrubric
+\else
 \newpage
@@
 \fi
+\newpage
+\newpage
 
-\begin{solution}
-\begin{alltt}
-\rubricpts{5 pts -- get it right}
-;; product% implements [Pattern Number]
-(define-class product%
-  ;; the product of a leaf
-  (define (leaf n) n)
-  ;; the product of a node
-  (define (node l n r) (* n (l . match this) (r . match this))))
+\end{problem}
 
-(define p (new product%))
-(check-expect ((new leaf% 5) . match p) 5)
-(check-expect ((new node%  (new leaf% 2) 5 (new leaf% 7)) . match p) 70)
-\end{alltt}
-\end{solution}
-\end{problem}
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2013-03-31 exam6/2.tex
a first draft of the section/outline problem

@@
 \begin{problem} \pts{20}
+\noindent
+Tobin-Hochstadt needs to automatically manage papers to make it
+easier to review them.  He wants to represent a paper in the following
+way:
 
-Outline problem
+A \emph{Section} has a title and a list of subsections, which are just
+sections themselves.  A paper is therefore just a big section.  Here
+is an example of a paper (and, of course, an example of a section):
+
+\begin{verbatim}
+   Fundies 2, Day 1
+   1. OO = struct + fun
+   1.1. Rocket
+   1.2. Moon
+   2. Design
+   2.1. Announce
+   2.1.1. Assign
+   3. End
+\end{verbatim}
+Note that the section numbers are \emph{NOT} part of the document!
+
+\bigskip
+%\vskip{2}
+
+\begin{enumerate}
+
+\item Develop data, class, and interface defintions for sections.
+
+\newpage
+
+\noindent
+\item Extend your defininition of \texttt{Section} to support an
+\texttt{accept} method that take \texttt{SectionVisitor}s over to
+compute over sections.  You will need to define the
+\texttt{SectionVisitor} interface.  You may need to define additional
+visitors as well, depending on your original data definition.
+
+\newpage
+
+\noindent
+\item Write a \texttt{SectionVisitor} that produces the ``outline'' of a section, with the
+section numbers.  The result should be an \texttt{ArrayList} of
+\texttt{String}s, as seen in the initial example for this problem.
+
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
+\noindent
+\item Here is a defintion for the \texttt{equals} and \texttt{hashCode}
+methods for \texttt{Section}s:
+
+\begin{verbatim}
+Boolean sameSec(Section s) { ...}
+
+boolean equals(Object o) {
+    if (o instanceof Section)
+        return this.sameSec((Section)o);
+    else
+        return false;
+}
+
+int hashCode() { return 42; }
+\end{verbatim}
+
+\noindent
+Develop the \texttt{sameSec} method, which \texttt{equals} relies on.
+
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
+\noindent
+\item The definition for \texttt{hashCode} that we have presented is very
+simple, but it will reduce the performance of hash tables of
+\texttt{Section}s.  Develop a better version of the \texttt{hashCode}
+method.  Make sure that it can distinguish (some) \texttt{Section}s
+with the same title but different contents, as well as (some)
+\texttt{Section}s with different titles but the same contents.
 
@@
 \newpage
-
+\end{enumerate}
 \end{problem}
-
-\ifrubric
-\else
-\newpage
-[Here is some more space for the previous problem.]
-\fi
-\newpage
 

### 2013-03-31 exam6/2.tex
Rewrote document problem.

--- 2013-03-30 exam6/2.tex
@@
 \begin{problem} \pts{20}
+\noindent
+Tobin-Hochstadt has been asked to serve on a review panel for the
+National Science Foundation, which requires he read and make notes on
+a set of grant proposals the government is considering to fund.
+Rather than read the proposals, he's decided to make a program for
+modeling his notes.  His notes take the form of an outline, which
+consists of a peice of text and a list of (sub)notes.  For example,
+here's one of his notes (thankfully not in his handwritting):
+\begin{verbatim}
+       NSF Proposal 1218390
+       |
+       +- Overview
+       |  |
+       |  +- WAT?
+       |  |
+       |  +- Motivations
+       |     |
+       |     +- unmotivating
+       |
+       +- Middle (fell asleep)
+       |
+       +- Conclusions
+       |
+       +- Budget
+          |
+          +- very large
+          |
+          +- lots of BU undergrads
+\end{verbatim}
 
-Outline problem
+%\bigskip
+%\vskip{2}
+\newpage
+
+\begin{enumerate}
+
+\item Develop data, class, and interface defintions for a note.
+
+\newpage
+
+\noindent
+\item Develop the visitor pattern for notes by designing
+  an \texttt{accept} method on notes that
+  consumes a note visitor and uses it to compute something of a note.
+  You will need to define the note visitor interface and may need to
+  define additional visitors as well, depending on your original data
+  definition.
+
+\newpage
+
+\noindent
+\item Write a note visitor that produces the numbered outline of a
+  note.  The result should be an \texttt{ArrayList} of
+  \texttt{String}s.  For example, the numbered outline of the note
+  note is an \texttt{ArrayList<String>} with the following elements:
+
+\begin{verbatim}
+"1. NSF Proposal 1218390"
+"1.1 Overview"
+"1.1.1 WAT?"
+"1.1.2 Motivations"
+"1.1.2.1 unmotivating"
+"1.2 Middle (fell asleep)"
+"1.3 Conclusions"
+"1.4 Budget"
+"1.4.1 very large"
+"1.4.2 lots of BU undergrads"
+\end{verbatim}
+
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
+\noindent
+\item Here is a defintion for the \texttt{equals} and \texttt{hashCode}
+methods for a note:
+
+\begin{verbatim}
+Boolean sameNote(Note s) { ...}
+
+boolean equals(Object o) {
+    return (o instanceof Note) 
+        && this.sameNote((Note)o);
+}
+
+int hashCode() { return 42; }
+\end{verbatim}
+
+\noindent
+Develop the \texttt{sameNote} method, which \texttt{equals} relies on.
+(Two notes are equal if they have the same text and subnotes.)
+
+\ifrubric
+\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+\newpage
+
+\noindent
+\item The definition for \texttt{hashCode} that we have presented is
+  very simple, but it will reduce the performance of hash tables of
+  notes.  Develop a better version of the \texttt{hashCode} method.
+  Make sure that it can distinguish (at least \emph{some}) notes with
+  the same text but different subnotes, as well as (some)
+  notes with different text but the same subnotes.
 

### 2013-03-31 exam6/2.tex
Small corrections.

@@
 "1. NSF Proposal 1218390"
-"1.1 Overview"
-"1.1.1 WAT?"
-"1.1.2 Motivations"
-"1.1.2.1 unmotivating"
-"1.2 Middle (fell asleep)"
-"1.3 Conclusions"
-"1.4 Budget"
-"1.4.1 very large"
-"1.4.2 lots of BU undergrads"
+"1.1. Overview"
+"1.1.1. WAT?"
+"1.1.2. Motivations"
+"1.1.2.1. unmotivating"
+"1.2. Middle (fell asleep)"
+"1.3. Conclusions"
+"1.4. Budget"
+"1.4.1. very large"
+"1.4.2. lots of BU undergrads"
 \end{verbatim}
@@
 Develop the \texttt{sameNote} method, which \texttt{equals} relies on.
-(Two notes are equal if they have the same text and subnotes.)
+(Two notes are the same if they have the same text and subnotes.)
 

### 2013-03-31 exam6/Note.java
Solutions for note problem.

@@ -1,0 +1,199 @@
+import tester.*;
+import java.util.*;
+
+class Note {
+    String txt;
+    List<Note> subnotes;
+    
+    Note(String txt) {
+	this(txt, new Empty<Note>());
+    }
+
+    Note(String txt, List<Note> subnotes) {
+	this.txt = txt;
+	this.subnotes = subnotes;
+    }
+
+    <R> R accept(NoteVisitor<R> v) {
+	return v.visitNote(this.txt, this.subnotes);
+    }
+}
+
+interface NoteVisitor<R> {
+    R visitNote(String txt, List<Note> subnotes);
+}
+
+interface List<X> {
+    <R> R accept(ListVisitor<X,R> v);
+}
+
+interface ListVisitor<X,R> {
+    R visitEmpty();
+    R visitCons(X first, List<X> rest);
+}
+
+class Empty<X> implements List<X> {
+    Empty() {}
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitEmpty();
+    }
+}
+
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+
+    Cons(X first, List<X> rest) {
+	this.first = first;
+	this.rest = rest;
+    }
+
+    public <R> R accept(ListVisitor<X,R> v) {
+	return v.visitCons(this.first, this.rest);
+    }
+}
+
+class Outline implements NoteVisitor<ArrayList<String>> {
+    String prefix;
+    Integer secnum;
+    ArrayList<String> output;
+
+    Outline() {
+	this("1.", 1, new ArrayList<String>());
+    }
+    Outline(String prefix, Integer secnum, ArrayList<String> output) {
+	this.prefix = prefix;
+	this.secnum = secnum;
+	this.output = output;
+    }
+
+    public ArrayList<String> visitNote(String txt, List<Note> subnotes) {
+	this.print(txt);
+	subnotes.accept(new OutlineList(this.prefix, 1, this.output));
+	return this.output;
+    }
+
+    public void print(String txt) {
+	this.output.add(this.prefix.concat(" ".concat(txt)));
+    }
+}
+
+class OutlineList implements ListVisitor<Note,Void> {
+    String prefix;
+    Integer secnum;
+    ArrayList<String> output;
+    OutlineList(String prefix, Integer secnum, ArrayList<String> output) {
+	this.prefix = prefix;
+	this.secnum = secnum;
+	this.output = output;
+    }
+    
[…]

### 2013-04-01 exam6/2.tex
BST problem.

@@
 
- Binary tree problem
+A \emph{binary search tree} is a lot like a binary tree in that a BST
+is either a node, which contains an element and a left and right
+subtree, or a leaf, which doesn't contain an element or subtrees.  The
+difference between a binary tree and a binary \emph{search} tree is
+that every leaf and node of a BST also contains a \texttt{Comparator},
+and nodes have the following invariant: every element in the left
+subtree is smaller (according to the comparator) than the element in
+the node, which in turn is smaller than or equal (according to the
+comparator) to every element in the right subtree.  (You can assume
+the comparator in every leaf and node is the same for any given BST.)
+
+\begin{enumerate}
+\item Design class and interface definitions for binary search trees
+  \texttt{BST<X>}.  Design the following method for binary search
+  trees:
+  \begin{verbatim}
+  // Insert x into this BST, preserving the BST invariant
+  BST<X> insert(X x);
+\end{verbatim}
+
+\ifrubric\else
+\newpage
+[Here is some more space for the previous problem.]
+\fi
+
+\newpage
+\item Here's an idea for sorting a series of elements: take all of the
+ elements and insert them into (an initially empty) binary search
+ tree.  Now, assuming \texttt{BST<X>} implemented the
+ \texttt{Iterable<X>} such that the iterator produced by
+ \texttt{iterator()} visited the elements of the BST in increasing
+ order (according to the comparator), then the elements could be retrieved
+in sorted order.
+
+Use this idea to implement a sorting method with the signature given
+ below.  It should sort a given iterable object in ascending order
+ according to a given comparator, producing the sorted list as an
+ array list of elements.
+
+You may assume \texttt{BST<X>} implements \texttt{Iterable<X>} correctly
+and do not need to define this yourself (see the extra credit).
+
+{\small
+\begin{verbatim}
+// Class of sorting methods
+class Sort {
+  // Produce an array list of elems sorted in ascending order by c
+  <X> ArrayList<X> treeSort(Comparator<X> c, Iterable<X> elems) { 
+    ... 
+  }
+}
+\end{verbatim}}
+\end{enumerate}
 
@@
 \newpage
-
-\textbf{Extra Credit}
-
 \end{problem}

### 2013-04-01 exam6/2.tex
Thunk problem.

@@
 
-Computation/thunk problem
+In most general terms, you can think of a \emph{computation} as something that
+\emph{runs} to compute an answer.  To solidify this idea as an interface, a
+computation could be represented by an object implementing:
+
+\begin{verbatim}
+interface Computation<R> {
+  // Run this computation
+  R run();
+}
+\end{verbatim}
+
+\noindent
+For example, this computation computes the square of five:
+
+\begin{verbatim}
+class TwentyFive implements Computation<Integer> {
+  Integer run() { return 5 * 5; }
+}
+\end{verbatim}
+
+\begin{enumerate}
+
+\item Design a computation that when constructed is given a
+ \texttt{Computation<Integer>} and a natural number \emph{n}, and when
+ run, runs the computation \emph{n} times and produces an array list
+of all the results.
+
+\newpage
+\item
+
+Computations can be expensive, so it's probably a good idea to avoid
+ running them if you can, or if you must run them, run them just once
+ and remember the result. Subsequent runs of the computation should
+ then just return the remembered value rather than doing any real
+ work.
+
+In contemplating the design of a library for such laziness, you should
+ realize that this idea of running something at most once and
+ remembering its result is itself a computation.  In other words, if
+ given a computation, you can make a new computation that will act
+ just like the given one, except it will run the given computation at
+ most one time.
+
+Design the following class:
+
+\begin{verbatim}
+class MaybeOnce<R> implements Computation<R> {
+  MaybeOnce(Computation<R> compute) { ... }
+
+  // If this computation has never been run, run
+  // compute and produce its result.  Remember
+  // this answer for future runs and just return
+  // it rather than re-run compute.
+  R run();
+}
+\end{verbatim}
+\end{enumerate}
+
+Take care to make sure that if a \texttt{MaybeOnce} computation is not
+run, it never runs the inner computation.  For example, 
+this program should not blow up (assuming this is the whole program):
+
+\begin{verbatim}
+class BlowUp implements Computation<Integer> {
+  // Blows up when run
+  Integer run() { throw new RuntimeException("KABOOM!"); }
+}
+
+// Never run, so never blows up.
+new MaybeOnce<Integer>(new BlowUp())
+\end{verbatim}
 

### 2013-04-01 exam6/2.tex
BST iterator extra credit.
abandon the outline problem

@@
 \noindent
-\item Write a note visitor that produces the numbered outline of a
-  note.  The result should be an \texttt{ArrayList} of
-  \texttt{String}s.  For example, the numbered outline of the note
-  note is an \texttt{ArrayList<String>} with the following elements:
+\item Write a note visitor that produces the maximum nesting depth of
+  a note.  For example, the maximum nesting depth of the given example
+  is 4.
 
-\begin{verbatim}
-"1. NSF Proposal 1218390"
-"1.1. Overview"
-"1.1.1. WAT?"
-"1.1.2. Motivations"
-"1.1.2.1. unmotivating"
-"1.2. Middle (fell asleep)"
-"1.3. Conclusions"
-"1.4. Budget"
-"1.4.1. very large"
-"1.4.2. lots of BU undergrads"
-\end{verbatim}
+%  numbered outline of a
+%   note.  The result should be an \texttt{ArrayList} of
+%   \texttt{String}s.  For example, the numbered outline of the note
+%   note is an \texttt{ArrayList<String>} with the following elements:
+
+% \begin{verbatim}
+% "1. NSF Proposal 1218390"
+% "1.1. Overview"
+% "1.1.1. WAT?"
+% "1.1.2. Motivations"
+% "1.1.2.1. unmotivating"
+% "1.2. Middle (fell asleep)"
+% "1.3. Conclusions"
+% "1.4. Budget"
+% "1.4.1. very large"
+% "1.4.2. lots of BU undergrads"
+% \end{verbatim}
 
@@
 
+\begin{problem}  \pts{5}
+
+{\bf Extra Credit}
+
+\noindent
+Revise your design of \texttt{BST<X>} to implement
+ \texttt{Iterable<X>} interface.  The iterator returned by
+ \texttt{iterator()} should iterate through the elements in sorted order.
+
+\end{problem}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-04-01 exam6/Note.java

@@ -96,6 +96,63 @@ class OutlineList implements ListVisitor<Note,Void> {
     }       
 }
 
+
+/////////////////////////
+
+// class Outline implements NoteVisitor<ArrayList<String>> {
+//     String prefix;
+//     Integer secnum;
+//     ArrayList<String> output;
+
+//     Outline() {
+// 	this("1.", 1, new ArrayList<String>());
+//     }
+//     Outline(String prefix, Integer secnum, ArrayList<String> output) {
+// 	this.prefix = prefix;
+// 	this.secnum = secnum;
+// 	this.output = output;
+//     }
+
+//     public ArrayList<String> visitNote(String txt, List<Note> subnotes) {
+// 	this.output.add(txt);
+// 	ArrayList<String> suboutline = subnotes.accept(new OutlineList(this.secnum));
+
+// 	for(String s : suboutline)
+// 	    this.output.add(this.prefix.concat(s));
+
+// 	return this.output;
+//     }
+
+// }
+
+// class OutlineList implements ListVisitor<Note,ArrayList<String>> {
+//     String prefix;
+//     Integer secnum;
+//     ArrayList<String> output;
+//     OutlineList(String prefix, Integer secnum, ArrayList<String> output) {
+// 	this.prefix = prefix;
+// 	this.secnum = secnum;
+// 	this.output = output;
+//     }
+
+//     OutlineList next() {
+// 	return new OutlineList(this.prefix, this.secnum+1, this.output);
+//     }
+    
+//     public Void visitEmpty() { return null; }
+//     public Void visitCons(Note first, List<Note> rest) {
+// 	ArrayList<String> firsts = first.accept(new Outline());
+// 	for (String s : first)
+// 	    this.output.add(this.prefix.concat(this.secnum.toString().concat("."))
+// 	first.accept(new Outline(this.prefix.concat(this.secnum.toString().concat(".")), this.secnum, this.output));
+// 	rest.accept(this.next());
+// 	return null;
+//     }       
+// }
+
+////////////////////////
+
+
 // Nesting Depth with recursion
 class NestDepthR implements NoteVisitor<Integer> {
     NestDepthR () {}

### 2013-04-01 exam6/2.tex
Point assignment and minor tweaks.

@@
 1 & & 20\\ \hline
-2 & & 19\\ \hline
+2 & & 25\\ \hline
+3 & & 18\\ \hline
 Extra & & 5 \\ \hline
-{\bf Total} & & 39+5 \\ \hline
+{\bf Total} & & 63+5 \\ \hline
 %{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
@@
 
-\begin{problem} \pts{20}
+\begin{problem} \pts{25}
 \noindent
@@
           |
-          +- lots of BU undergrads
+          +- underfund lots of BU undergrads
 \end{verbatim}
@@
 \noindent
-\item Write a note visitor that produces the maximum nesting depth of
+\item Design a note visitor that produces the maximum nesting depth of
   a note.  For example, the maximum nesting depth of the given example
@@
 % "1.4.1. very large"
-% "1.4.2. lots of BU undergrads"
+% "1.4.2. underfund lots of BU undergrads"
 % \end{verbatim}
@@
 
-\begin{problem} \pts{20}
+\begin{problem} \pts{18}
 

### 2013-04-01 exam6/2.tex
Tweak to thunk problem.
more tweaks.
Java interface reference.

@@
 
+\begin{center}
+{\bf Java Reference}
+\end{center}
+\noindent
+You may be asked to use or implement the following standard Java
+interfaces, which are described here for reference:
+
+\begin{enumerate}
+\item \texttt{Iterable<X>} -- Implementing this interface allows an object to be the target of the ``foreach'' statement.
+
+ Methods:
+ \begin{itemize}
+ \item \texttt{Iterator<X> iterator()} : Returns an iterator over a set of
+  elements of type \texttt{X}.
+  \end{itemize}
+
+\item \texttt{Iterator<X>} -- An iterator over a collection.
+
+  Methods:
+  \begin{itemize}
+    \item \texttt{boolean hasNext()} : Returns \texttt{true} if the iteration
+      has more elements.
+    \item \texttt{X next()} : Returns the next element in the iteration.
+  \end{itemize}
+
+\item \texttt{Comparator<X>} -- A comparison function, which imposes a total ordering on some collection of objects.
+
+  Methods:
+  \begin{itemize}
+    \item \texttt{int compare(X o1, X o2)} : Compares its two
+      arguments for order. Returns a negative integer, zero, or a
+      positive integer as the first argument is less than, equal to,
+      or greater than the second.
+  \end{itemize}
+
+\end{enumerate}
+
+\newpage
+
 \begin{problem} \pts{20}
 
-A \emph{binary search tree} is a lot like a binary tree in that a BST
-is either a node, which contains an element and a left and right
-subtree, or a leaf, which doesn't contain an element or subtrees.  The
+A \emph{binary search tree} (BST) is a lot like a binary tree in that
+a BST is either a node, which contains an element and a left and right
+subtree, or a leaf, which contains neither elements nor subtrees.  The
 difference between a binary tree and a binary \emph{search} tree is
@@
 subtree is smaller (according to the comparator) than the element in
-the node, which in turn is smaller than or equal (according to the
-comparator) to every element in the right subtree.  (You can assume
+the node, which in turn is smaller than or equal to (according to the
+comparator) every element in the right subtree.  (You can assume
 the comparator in every leaf and node is the same for any given BST.)
@@
  tree.  Now, assuming \texttt{BST<X>} implemented the
- \texttt{Iterable<X>} such that the iterator produced by
+ \texttt{Iterable<X>} interface such that the iterator produced by
  \texttt{iterator()} visited the elements of the BST in increasing
@@
 Use this idea to implement a sorting method with the signature given
- below.  It should sort a given iterable object in ascending order
- according to a given comparator, producing the sorted list as an
- array list of elements.
+ below.  It should sort the elements of a given iterable object in ascending order
+ according to a given comparator, producing the sorted elements as an
+ array list.
 
 You may assume \texttt{BST<X>} implements \texttt{Iterable<X>} correctly
-and do not need to define this yourself (see the extra credit).
+and do not need to define this yourself (see the extra credit problem).
 
@@
  a set of grant proposals the government is considering to fund.
-Rather than read the proposals, he's decided to make a program for
-modeling his notes.  His notes take the form of an outline, which
-consists of a peice of text and a list of (sub)notes.  For example,
-here's one of his notes (thankfully not in his handwritting):
+  Rather than read the proposals he's been assigned, he's decided to
+ make a program for modeling his notes.  His notes take the form of an
+ outline, which consists of a peice of text and a list of (sub)notes.
+  For example, here's one of his notes (thankfully not in his
+ handwritting):
 \begin{verbatim}
@@
 \begin{verbatim}
-Boolean sameNote(Note s) { ...}
+Boolean sameNote(Note s) { ... }
 
@@
   // it rather than re-run compute.
-  R run();
+  R run() { ... }
 }
 \end{verbatim}
-\end{enumerate}
 
@@
 \end{verbatim}
+\end{enumerate}
 
@@
  \texttt{iterator()} should iterate through the elements in sorted order.
+ It should do this efficiently by relying on the \texttt{BST} invariant.
 

### 2013-04-01 exam6/Note.java

@@ -53,105 +53,7 @@ class Cons<X> implements List<X> {
     }
 }
 
-class Outline implements NoteVisitor<ArrayList<String>> {
-    String prefix;
-    Integer secnum;
-    ArrayList<String> output;
-
-    Outline() {
-	this("1.", 1, new ArrayList<String>());
-    }
-    Outline(String prefix, Integer secnum, ArrayList<String> output) {
-	this.prefix = prefix;
-	this.secnum = secnum;
-	this.output = output;
-    }
-
-    public ArrayList<String> visitNote(String txt, List<Note> subnotes) {
-	this.print(txt);
-	subnotes.accept(new OutlineList(this.prefix, 1, this.output));
-	return this.output;
-    }
-
-    public void print(String txt) {
-	this.output.add(this.prefix.concat(" ".concat(txt)));
-    }
-}
-
-class OutlineList implements ListVisitor<Note,Void> {
-    String prefix;
-    Integer secnum;
-    ArrayList<String> output;
-    OutlineList(String prefix, Integer secnum, ArrayList<String> output) {
-	this.prefix = prefix;
-	this.secnum = secnum;
-	this.output = output;
-    }
-    
-    public Void visitEmpty() { return null; }
-    public Void visitCons(Note first, List<Note> rest) {
-	first.accept(new Outline(this.prefix.concat(this.secnum.toString().concat(".")), this.secnum, this.output));
-	rest.accept(new OutlineList(this.prefix, this.secnum+1, this.output));
-	return null;
-    }       
-}
-
-
-/////////////////////////
-
-// class Outline implements NoteVisitor<ArrayList<String>> {
-//     String prefix;
-//     Integer secnum;
-//     ArrayList<String> output;
-
-//     Outline() {
-// 	this("1.", 1, new ArrayList<String>());
-//     }
-//     Outline(String prefix, Integer secnum, ArrayList<String> output) {
-// 	this.prefix = prefix;
-// 	this.secnum = secnum;
-// 	this.output = output;
-//     }
-
-//     public ArrayList<String> visitNote(String txt, List<Note> subnotes) {
-// 	this.output.add(txt);
-// 	ArrayList<String> suboutline = subnotes.accept(new OutlineList(this.secnum));
-
-// 	for(String s : suboutline)
-// 	    this.output.add(this.prefix.concat(s));
-
-// 	return this.output;
-//     }
-
-// }
-
[…]

### 2013-04-09 exam6/2.tex
update ec problem

@@
 
+\begin{solution}
+\begin{verbatim}
+//> 5pts all correct
+//> Ok if they have another implementation if it's right
+
+// in BST<X> interface
+    // Add all elements of this BST to ls, in order.
+    void addTo(ArrayList<X> ls);
+
+// in ABST<X>
+    public Iterator<X> iterator() {
+        ArrayList<X> elems = new ArrayList<X>();
+        this.addTo(elems);
+        return elems.iterator();
+    }
+
+// in Leaf<X>
+    public void addTo(ArrayList<X> ls) {}
+
+// in Node<X>
+    public void addTo(ArrayList<X> ls) {
+        this.left.addTo(ls);
+        ls.add(this.x);
+        this.right.addTo(ls);
+    }
+\end{verbatim}
+\end{solution}
+
 \end{problem}

### 2013-04-09 exam6/2.tex
Rubric for problem 1.

@@
 
-\ifrubric\else
+\ifrubric
+
+\begin{solution}
+\begin{verbatim}
+// [1pt]
+interface BST<X> extends Iterable<X> {
+    // Insert given element into this BST.
+    BST<X> insert(X x);
+}
+
+// [optional]
+abstract class ABST<X> implements BST<X> {
+    Comparator<X> c;
+    ABST(Comparator<X> c) {
+        this.c = c;
+    }
+}
+
+// [1pt: fields, constructor, parameterized type]
+class Leaf<X> extends ABST<X> {
+    Leaf(Comparator<X> c) {
+        super(c);
+    }
+
+    // [1pt]
+    public BST<X> insert(X x) {
+        return new Node<X>(this.c, x, this, this);
+    }
+}
+
+// [2pt: fields, constructor, parameterized type]
+class Node<X> extends ABST<X> {
+    X x;
+    BST<X> left;
+    BST<X> right;
+    Node(Comparator<X> c, X x, BST<X> left, BST<X> right) {
+        super(c);
+        this.x = x;
+        this.left = left;
+        this.right = right;
+    }
+
+    // [3pt]
+    public BST<X> insert(X x) {
+        return (this.c.compare(x, this.x) < 0)
+        ? new Node<X>(this.c, this.x, this.left.insert(x), this.right)
+        : new Node<X>(this.c, this.x, this.left, this.right.insert(x));
+        }
+    }
+}
+
+// [2pt: tests]
+\end{verbatim}
+\end{solution}
+
+\else
 \newpage
@@
 \ifrubric
+\begin{solution}
+\begin{verbatim}
+// [5pt]
+class Sort {
+    <X> ArrayList<X> treeSort(Comparator<X> c, Iterable<X> elems) {
+        BST<X> bst = new Leaf<X>(c);
+        ArrayList<X> ls = new ArrayList<X>();
+        for (X x : elems) bst = bst.insert(x);
+        for (X x : bst)   ls.add(x);
+        return ls;
+    }
+}
+
+// [2pt]
+class LessThan implements Comparator<Integer> {
+    public int compare(Integer x, Integer y) {
+        return x - y;
+    }
+}
+
+// [3pt]
+class Examples {
+    void testTreeSort(Tester t) {
+        Sort s = new Sort();
+        Comparator<Integer> lt = new LessThan();
+        ArrayList<Integer> elems =
+            new ArrayList<Integer>(Arrays.asList(8,3,7,5));
+
+        ArrayList<Integer> sorted =
+            new ArrayList<Integer>(Arrays.asList(3,5,7,8));
+
+        t.checkExpect(s.treeSort(lt, new ArrayList<Integer>()),
+                      new ArrayList<Integer>());
+        t.checkExpect(s.treeSort(lt, elems), sorted);
+    }
+}
+\end{verbatim}
+\end{solution}
 \else

### 2013-04-09 exam6/2.tex
add prob 3

@@
 
+\begin{solution}
+\begin{verbatim}
+// 4 pts total
+// 1 pt class structure
+class Count implements Computation<Integer> {
+    Integer i;
+    Count() {
+        this.i = 0;
+    }
+
+// 2 pts for run
+    public Integer run() {
+        Integer j = this.i;
+        this.i = i+1;
+        return j;
+    }      
+}
+// 1 pt for tests
+\end{verbatim}
+\end{solution}
+
 \newpage
@@
 
+\begin{solution}
+\newpage
+\begin{verbatim}
+// 2 pt structure (including purpose etc)
+class Thunk<X> implements Computation<X> {
+    Delay<X> s;
+    Thunk(Computation<X> c) {
+        this.s = new Frozen<X>(c);
+    }
+// 2 pts 
+    public X run() {
+        this.s = this.s.force();
+        return this.s.val();
+    }
+}
+\end{verbatim}
+\newpage
+\begin{verbatim}
+// 8 pts for the code in general, may have different design
+interface Delay<X> {
+    Delay<X> force();
+    X val();
+}
+class Frozen<X> implements Delay<X> {
+    Computation<X> c;
+    Frozen(Computation<X> c) {
+        this.c = c;
+    }
+    public Delay<X> force() {
+        return new Thaw<X>(this.c.run());
+    }
+    public X val() {
+        return null;
+    }
+}
+class Thaw<X> implements Delay<X> {
+    X x;
+    Thaw(X x) {
+        this.x = x;
+    }
+    public Delay<X> force() {
+        return this;
+    }
+    public X val() {
+        return this.x;
+    }
+}
+\end{verbatim}
+\newpage
+\begin{verbatim}
+// 2 pts tests
+class Examples {
+    void testThunk(Tester t) {
+        Computation<Integer> count = new Count();
+        t.checkExpect(count.run(), 0);
+        t.checkExpect(count.run(), 1);
+        t.checkExpect(count.run(), 2);
+        t.checkExpect(count.run(), 3);
+        Computation<Integer> thunk = new Thunk<Integer>(new Count());
+        t.checkExpect(thunk.run(), 0);
+        t.checkExpect(thunk.run(), 0);
+    }
+
+}
+\end{verbatim}
+\end{solution}
+
 \ifrubric
@@
 \begin{verbatim}
-//> 5pts all correct
-//> Ok if they have another implementation if it's right
-
 // in BST<X> interface

### 2013-04-09 exam6/2.tex
fix prob 3

--- 2013-04-09 exam6/2.tex
@@
 
+\begin{solution}
+\begin{verbatim}
+// This class is an example
+class Count implements Computation<Integer> {
+    Integer i;
+    Count() {
+        this.i = 0;
+    }
+
+    public Integer run() {
+        Integer j = this.i;
+        this.i = i+1;
+        return j;
+    }      
+}
+\end{verbatim}
 \newpage
+\begin{verbatim}
+// 6 pts total
+// 2 pt class structure
+class Repeat implements Computation<ArrayList<Integer>> {
+    Computation<Integer> comp;
+    Integer n;
+    Repeat(Computation<Integer> comp, Integer n) {
+        this.comp = comp; this.n = n;
+    }
+
+// 2 pts for run
+    public ArrayList<Integer> run() {
+        ArrayList<Integer> as = new ArrayList<Integer>();
+        for(int i = 0; i < n; i++) 
+          as.add(comp.run());
+        return as;
+    }      
+}
+// 2 pt for tests (need an example Computation<Integer>)
+t.checkExpect((new Repeat(new Count(3))).run, new ArrayList(0,1,2));
+\end{verbatim}
+\newpage
+\end{solution}
+
 \item
@@
 
+\begin{solution}
+\newpage
+\begin{verbatim}
+// 2 pt structure (including purpose etc)
+class Thunk<X> implements Computation<X> {
+    Delay<X> s;
+    Thunk(Computation<X> c) {
+        this.s = new Frozen<X>(c);
+    }
+// 1 pts 
+    public X run() {
+        this.s = this.s.force();
+        return this.s.val();
+    }
+}
+\end{verbatim}
+\newpage
+\begin{verbatim}
+// 7 pts for the code in general, may have different design
+interface Delay<X> {
+    Delay<X> force();
+    X val();
+}
+class Frozen<X> implements Delay<X> {
+    Computation<X> c;
+    Frozen(Computation<X> c) {
+        this.c = c;
+    }
+    public Delay<X> force() {
+        return new Thaw<X>(this.c.run());
+    }
+    public X val() {
+        return null;
+    }
+}
+class Thaw<X> implements Delay<X> {
+    X x;
+    Thaw(X x) {
+        this.x = x;
+    }
+    public Delay<X> force() {
+        return this;
+    }
+    public X val() {
+        return this.x;
+    }
+}
+\end{verbatim}
+\newpage
+\begin{verbatim}
+// 2 pts tests
+class Examples {
+    void testThunk(Tester t) {
+        Computation<Integer> count = new Count();
+        t.checkExpect(count.run(), 0);
+        t.checkExpect(count.run(), 1);
+        t.checkExpect(count.run(), 2);
+        t.checkExpect(count.run(), 3);
+        Computation<Integer> thunk = new Thunk<Integer>(new Count());
+        t.checkExpect(thunk.run(), 0);
+        t.checkExpect(thunk.run(), 0);
+    }
+
+}
+\end{verbatim}
+\end{solution}
+
 \ifrubric

### 2013-04-09 exam6/2.tex
problem 2

@@
   For example, here's one of his notes (thankfully not in his
- handwritting):
+ handwriting):
 \begin{verbatim}
@@
 
+\begin{solution}
+\begin{verbatim}
+// 5 pts for Note
+class Note {
+    String txt;
+    List<Note> subnotes;
+
+    Note(String txt) {
+        this(txt, new Empty<Note>());
+    }
+
+    Note(String txt, List<Note> subnotes) {
+        this.txt = txt;
+        this.subnotes = subnotes;
+    }
+
+}
+
+// 4 pts for List
+interface List<X> {}
+
+class Empty<X> implements List<X> {
+    Empty() {}
+}
+
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+
+    Cons(X first, List<X> rest) {
+        this.first = first;
+        this.rest = rest;
+    }
+}
+
+\end{verbatim}
+\end{solution}
+
 \newpage
@@
 
+\begin{solution}
+\begin{verbatim}
+
+//> 4 pts for interface defintions
+
+interface ListVisitor<X,R> {
+    R visitEmpty();
+    R visitCons(X first, List<X> rest);
+}
+
+// in List
+    <R> R accept(ListVisitor<X,R> v);
+
+interface NoteVisitor<R> {
+    R visitNote(String txt, List<Note> subnotes);
+}
+
+//> 4 pts for accept methods
+// in Note
+    <R> R accept(NoteVisitor<R> v) {
+        return v.visitNote(this.txt, this.subnotes);
+    }
+
+// in Empty
+    public <R> R accept(ListVisitor<X,R> v) {
+        return v.visitEmpty();
+    }
+
+// in Cons
+    public <R> R accept(ListVisitor<X,R> v) {
+        return v.visitCons(this.first, this.rest);
+    }
+\end{verbatim}
+\end{solution}
+
 \newpage
@@
   is 4.
+
+\begin{solution}
+\begin{verbatim}
+//> 4 pts
+// Nesting Depth with recursion
+class NestDepthR implements NoteVisitor<Integer> {
+    NestDepthR () {}
+    public Integer visitNote(String txt, List<Note> subnotes) {
+        return 1 + subnotes.accept(new NestDepthListR());
+    }
+}
+
+//> 4 pts
+class NestDepthListR implements ListVisitor<Note,Integer> {
+    NestDepthListR () {}
+    public Integer visitEmpty() {
+        return 0;
+    }
+    public Integer visitCons(Note first, List<Note> rest) {
+        return Math.max(first.accept(new NestDepthR()),
+                        rest.accept(this));
+    }
+}
+\end{verbatim}
+\end{solution}
 
@@
     }
-
 }

### 2013-04-10 tfpie/example.tex
TFPIE start.

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{SOS 2007} % Name of the event you are submitting to
\usepackage{breakurl}             % Not needed if you use pdflatex only.

\title{An Example of a Paper with a Rather Large Title-to-Content Ratio}
\author{Rob van Glabbeek
\institute{NICTA\\ Sydney, Australia}
\institute{School of Computer Science and Engineering\\
University of New South Wales\thanks{A fine university.}\\
Sydney, Australia}
\email{rvg@cs.stanford.edu}
\and
Co Author \qquad\qquad Yet S. Else
\institute{Stanford Univeristy\\
California, USA}
\email{\quad is@gmail.com \quad\qquad somebody@else.org}
}
\def\titlerunning{A Longtitled Paper}
\def\authorrunning{R.J. van Glabbeek, C. Author \& Y.S. Else}
\begin{document}
\maketitle

\begin{abstract}
This is a sentence in the abstract.
This is another sentence in the abstract.
This is yet another sentence in the abstract.
This is the final sentence in the abstract.
\end{abstract}

\section{Introduction}

The optional arguments of {\tt $\backslash$documentclass$\{$eptcs$\}$} are
\begin{itemize}
\item at most one of
{\tt adraft},
{\tt submission} or
{\tt preliminary},
\item at most one of {\tt publicdomain} or {\tt copyright},
\item and optionally {\tt creativecommons},
  \begin{itemize}
  \item possibly augmented with
    \begin{itemize}
    \item {\tt noderivs}
    \item or {\tt sharealike},
    \end{itemize}
  \item and possibly augmented with {\tt noncommercial}.
  \end{itemize}
\end{itemize}
We use {\tt adraft} rather than {\tt draft} so as not to confuse hyperref.
The style-file option {\tt submission} is for papers that are
submitted to {\tt $\backslash$event}, where the value of the latter is
to be filled in in line 2 of the tex-file. Use {\tt preliminary} only
for papers that are accepted but not yet published. The final version
of your paper that is to be uploaded at the EPTCS website should have
none of these style-file options.

By means of the style-file option
\href{http://creativecommons.org/about/license/}{creativecommons}
authors equip their paper with a Creative Commons license that allows
everyone to copy, distribute, display, and perform their copyrighted
work and derivative works based upon it, but only if they give credit
the way you request. By invoking the additional style-file option {\tt
noderivs} you let others copy, distribute, display, and perform only
verbatim copies of your work, but not derivative works based upon
it. Alternatively, the {\tt sharealike} option allows others to
distribute derivative works only under a license identical to the
license that governs your work. Finally, you can invoke the option
{\tt noncommercial} that let others copy, distribute, display, and
perform your work and derivative works based upon it for
noncommercial purposes only.

Authors' (multiple) affiliations and emails use the commands
{\tt $\backslash$institute} and {\tt $\backslash$email}.
Both are optional.
Authors should moreover supply
{\tt $\backslash$titlerunning} and {\tt $\backslash$authorrunning},
and in case the copyrightholders are not the authors also
{\tt $\backslash$copyrightholders}.
As illustrated above, heuristic solutions may be called for to share
affiliations. Authors may apply their own creativity here.

Exactly 46 lines fit on a page.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill6\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill11\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.

Here starts a new paragraph. The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill16\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill21\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill26\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill31\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill36\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill41\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.\hfill46\\
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.
The rest is like any normal {\LaTeX} article.
We will spare you the details.

\section{Prefaces}

Volume editors may create prefaces using this very template,
with {\tt $\backslash$title$\{$Preface$\}$} and {\tt $\backslash$author$\{\}$}.

\section{Bibliography}

We request that you use
\href{http://www.cse.unsw.edu.au/~rvg/EPTCS/eptcs.bst}
{\tt $\backslash$bibliographystyle$\{$eptcs$\}$}
\cite{bibliographystylewebpage}. Compared to the original {\LaTeX}
{\tt $\backslash$biblio\-graphystyle$\{$plain$\}$},
it ignores the field {\tt month}, and uses the extra
bibtex fields {\tt eid}, {\tt doi}, {\tt ee} and {\tt url}.
The first is for electronic identifiers (typically the number $n$
indicating the $n^{\rm th}$ paper in an issue) of papers in electronic
journals that do not use page numbers. The other three are to refer,
with life links, to electronic incarnations of the paper.

Almost all publishers use digital object identifiers (DOIs) as a
persistent way to locate electronic publications. Prefixing the DOI of
any paper with {\tt http://dx.doi.org/} yields a URI that resolves to the
current location (URL) of the response page\footnote{Nowadays, papers
  that are published electronically tend
  to have a \emph{response page} that lists the title, authors and
  abstract of the paper, and links to the actual manifestations of
  the paper (e.g.\ as {\tt dvi}- or {\tt pdf}-file). Sometimes
  publishers charge money to access the paper itself, but the response
  page is always freely available.}
of that paper. When the location of the response page changes (for
instance through a merge of publishers), the DOI of the paper remains
the same and (through an update by the publisher) the corresponding
URI will then resolve to the new location. For that reason a reference
ought to contain the DOI of a paper, with a life link to corresponding
URI, rather than a direct reference or link to the current URL of
publisher's response page. This is the r\^ole of the bibtex field {\tt doi}.
DOIs of papers can often be found through
\url{http://www.crossref.org/guestquery};\footnote{For papers that will appear
  in EPTCS and use \href{http://www.cse.unsw.edu.au/~rvg/EPTCS/eptcs.bst}
  {\tt $\backslash$bibliographystyle$\{$eptcs$\}$} there is no need to
  find DOIs on this website, as EPTCS will look them up for you
  automatically upon submission of a first version of your paper;
  these DOIs can then be incorporated in the final version, together
  with the remaining DOIs that need to found at DBLP or publisher's webpages.}
the second method {\it Search on article title}, only using the {\bf
surname} of the first-listed author, works best.  
Other places to find DOIs are DBLP and the response pages for cited
papers (maintained by their publishers).
{\bf EPTCS requires the inclusion of a DOI in each cited paper, when available.}

Often an official publication is only available against payment, but
as a courtesy to readers that do not wish to pay, the authors also
make the paper available free of charge at a repository such as
\url{arXiv.org}. In such a case it is recommended to also refer and
link to the URL of the response page of the paper in such a
repository.  This can be done using the bibtex fields {\tt ee} or {\tt
url}, which are treated as synonyms.  These fields should not be used
to duplicate information that is already provided through the DOI of
the paper.
You can find archival-quality URL's for most recently published papers
in DBLP---they are in the bibtex-field {\tt ee}. In fact, it is often
useful to check your references against DBLP records anyway, or just find
them there in the first place.

When using {\LaTeX} rather than {\tt pdflatex} to typeset your paper, by
default no linebreaking within long URLs is allowed. This leads often
to very ugly output, that moreover is different from the output
generated when using {\tt pdflatex}. This problem is repaired when
invoking \href{http://www.cse.unsw.edu.au/~rvg/EPTCS/breakurl.sty}
{\tt $\backslash$usepackage$\{$breakurl$\}$}: it allows linebreaking
within links and yield the same output as obtained by default with
{\tt pdflatex}. 
When invoking {\tt pdflatex}, the package {\tt breakurl} is ignored.

\nocite{*}
\bibliographystyle{eptcs}
\bibliography{generic}
\end{document}

### 2013-04-11 tfpie/example.tex
First ideas for paper.

@@
-\documentclass[submission,copyright,creativecommons]{eptcs}
-\providecommand{\event}{SOS 2007} % Name of the event you are submitting to
+\documentclass[submission,copyright]{eptcs}
+\providecommand{\event}{TFPIE} % Name of the event you are submitting to
 \usepackage{breakurl}             % Not needed if you use pdflatex only.
 
-\title{An Example of a Paper with a Rather Large Title-to-Content Ratio}
-\author{Rob van Glabbeek
-\institute{NICTA\\ Sydney, Australia}
-\institute{School of Computer Science and Engineering\\
-University of New South Wales\thanks{A fine university.}\\
-Sydney, Australia}
-\email{rvg@cs.stanford.edu}
-\and
-Co Author \qquad\qquad Yet S. Else
-\institute{Stanford Univeristy\\
-California, USA}
-\email{\quad is@gmail.com \quad\qquad somebody@else.org}
+\title{How to Design Programs with Class}
+\author{Sam Tobin-Hochstadt \quad\qquad David Van Horn
+\institute{Northeastern University\\
+Boston, Massachusetts, USA}
+\email{\{samth,dvanhorn\}@ccs.neu.edu}
 }
 \def\titlerunning{A Longtitled Paper}
-\def\authorrunning{R.J. van Glabbeek, C. Author \& Y.S. Else}
+\def\authorrunning{S. Tobin-Hochstadt \& D. Van Horn}
 \begin{document}
@@
 \begin{abstract}
-This is a sentence in the abstract.
-This is another sentence in the abstract.
-This is yet another sentence in the abstract.
-This is the final sentence in the abstract.
+We propose a bridge between functional and object-oriented programming
+in the first-year cirriculum.
 \end{abstract}
@@
 
-The optional arguments of {\tt $\backslash$documentclass$\{$eptcs$\}$} are
+\emph{Thesis: ...}
+
+\section{Background: the first semester}
+
+\section{A small shift of focus}
+
+The semester starts with the minimal conceptual change from the
+previous semester: the addition of objects (same IDE, same syntax,
+same semantics [just a new kind of value], same interactive video
+games, etc.).
+
+\subsection{Side-by-side example of program design}
+
+[First semester vs second semester example programs.]
+
+\section{Distributed, Interactive Programs}
+
+\section{Representation-indepence and interface-oriented programming}
+
+\section{Unifying functions and objects}
+
+\section{From principles to fashions}
+
+Could work with any fashion, and should really lead into many
+fashionable languages.  The purpose of showing many is to demonstrate
+that the same principles apply in each setting.  Each language can also
+expose students to various mechanisms.
+
+\subsection{Java}
+
 \begin{itemize}
-\item at most one of
-{\tt adraft},
-{\tt submission} or
-{\tt preliminary},
-\item at most one of {\tt publicdomain} or {\tt copyright},
-\item and optionally {\tt creativecommons},
+\item Types
+\item Compiler
+\item ``Industrial'' features (collections library, ...)
+\end{itemize}
+
+\section{Comparison with HtDC}
+
+The noble goals of ProfJ:
+
 \begin{itemize}
-  \item possibly augmented with
+\item language levels
+\item continuity in IDE from first semester
+\end{itemize}
+
+Many of the shortcomings of HtDC and ProfJ are ``day 1'' problems.
+The transition is too abrupt and too large.
+
+The failures of  ProfJ:
+
 \begin{itemize}
-    \item {\tt noderivs}
-    \item or {\tt sharealike},
+\item changes languages from first semester rather than adding new
+  concept (causes students confusing deep issue (O.O. design) and
+  shallow ones (Java-like syntax) and orthogonal issues (static type
+  discipline).  (This also ultimately lead to the tool being abandoned
+  as it was too difficult to maintain.)
+
+\item commits to a particular fashion rather than enabling any.
+
+\item abandons the first semester approach and replaces it with a
+  (completely) object-oriented approach on day one.  There are
+  analogies between functional and object concepts (``a method is like
+  a function''), but the two never co-exist in a program.
+
+\item Java, even pared down pedagoical Java, is syntactically heavy; discourages
+  interactive programming
+
 \end{itemize}
-  \item and possibly augmented with {\tt noncommercial}.
+
+The changes that occur on Day 1 of HtDC w/ ProfJ:
+
+\begin{itemize}
+\item Radically different syntax
+\item Types
+\item Total committment to objects
+\item Subtly different base values
 \end{itemize}
+
+The changes that occur on Day 1 of HtDC w/o ProfJ:
+
+\begin{itemize}
+\item Different IDE
+\item Compilation
+\item Loss of interactivity
 \end{itemize}
-We use {\tt adraft} rather than {\tt draft} so as not to confuse hyperref.
-The style-file option {\tt submission} is for papers that are
-submitted to {\tt $\backslash$event}, where the value of the latter is
-to be filled in in line 2 of the tex-file. Use {\tt preliminary} only
-for papers that are accepted but not yet published. The final version
-of your paper that is to be uploaded at the EPTCS website should have
-none of these style-file options.
 
-By means of the style-file option
-\href{http://creativecommons.org/about/license/}{creativecommons}
-authors equip their paper with a Creative Commons license that allows
-everyone to copy, distribute, display, and perform their copyrighted
-work and derivative works based upon it, but only if they give credit
-the way you request. By invoking the additional style-file option {\tt
-noderivs} you let others copy, distribute, display, and perform only
-verbatim copies of your work, but not derivative works based upon
-it. Alternatively, the {\tt sharealike} option allows others to
-distribute derivative works only under a license identical to the
-license that governs your work. Finally, you can invoke the option
-{\tt noncommercial} that let others copy, distribute, display, and
-perform your work and derivative works based upon it for
-noncommercial purposes only.
+Some of these are premature, and better introduced later in the
+course.  Some are just irrelevant trivia (different syntax) that often
+favors ``experienced'' students while needlessly intimidating students
+with only the previous semester under their belt.  Many of these
+trivial matters, e.g. Eclipse configuration, javac command line
+arguments, the CLASSPATH, can be very frustrating to the uninitiated,
+causing students to believe struggling with such crap is \emph{the
+  essence of computer science}.  Bright, promising students
 
-Authors' (multiple) affiliations and emails use the commands
-{\tt $\backslash$institute} and {\tt $\backslash$email}.
-Both are optional.
-Authors should moreover supply
-{\tt $\backslash$titlerunning} and {\tt $\backslash$authorrunning},
-and in case the copyrightholders are not the authors also
-{\tt $\backslash$copyrightholders}.
-As illustrated above, heuristic solutions may be called for to share
-affiliations. Authors may apply their own creativity here.
+\section{Experience and outlook}
 
-Exactly 46 lines fit on a page.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill6\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill11\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
+Perspective into the upper-class cirriculum:
 
-Here starts a new paragraph. The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill16\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill21\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill26\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill31\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill36\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill41\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.\hfill46\\
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
-The rest is like any normal {\LaTeX} article.
-We will spare you the details.
+\begin{itemize}
+\item types, contract, invariants : formal methods, program verification,
 
-\section{Prefaces}
+\item representation independence : component-based software engineering,
+  data-structures, algorithms
 
-Volume editors may create prefaces using this very template,
-with {\tt $\backslash$title$\{$Preface$\}$} and {\tt $\backslash$author$\{\}$}.
-
-\section{Bibliography}
-
-We request that you use
-\href{http://www.cse.unsw.edu.au/~rvg/EPTCS/eptcs.bst}
-{\tt $\backslash$bibliographystyle$\{$eptcs$\}$}
-\cite{bibliographystylewebpage}. Compared to the original {\LaTeX}
-{\tt $\backslash$biblio\-graphystyle$\{$plain$\}$},
-it ignores the field {\tt month}, and uses the extra
-bibtex fields {\tt eid}, {\tt doi}, {\tt ee} and {\tt url}.
-The first is for electronic identifiers (typically the number $n$
-indicating the $n^{\rm th}$ paper in an issue) of papers in electronic
-journals that do not use page numbers. The other three are to refer,
-with life links, to electronic incarnations of the paper.
-
-Almost all publishers use digital object identifiers (DOIs) as a
-persistent way to locate electronic publications. Prefixing the DOI of
-any paper with {\tt http://dx.doi.org/} yields a URI that resolves to the
-current location (URL) of the response page\footnote{Nowadays, papers
-  that are published electronically tend
-  to have a \emph{response page} that lists the title, authors and
-  abstract of the paper, and links to the actual manifestations of
-  the paper (e.g.\ as {\tt dvi}- or {\tt pdf}-file). Sometimes
-  publishers charge money to access the paper itself, but the response
-  page is always freely available.}
-of that paper. When the location of the response page changes (for
-instance through a merge of publishers), the DOI of the paper remains
-the same and (through an update by the publisher) the corresponding
-URI will then resolve to the new location. For that reason a reference
-ought to contain the DOI of a paper, with a life link to corresponding
-URI, rather than a direct reference or link to the current URL of
-publisher's response page. This is the r\^ole of the bibtex field {\tt doi}.
-DOIs of papers can often be found through
-\url{http://www.crossref.org/guestquery};\footnote{For papers that will appear
-  in EPTCS and use \href{http://www.cse.unsw.edu.au/~rvg/EPTCS/eptcs.bst}
-  {\tt $\backslash$bibliographystyle$\{$eptcs$\}$} there is no need to
-  find DOIs on this website, as EPTCS will look them up for you
-  automatically upon submission of a first version of your paper;
-  these DOIs can then be incorporated in the final version, together
-  with the remaining DOIs that need to found at DBLP or publisher's webpages.}
-the second method {\it Search on article title}, only using the {\bf
-surname} of the first-listed author, works best.  
-Other places to find DOIs are DBLP and the response pages for cited
-papers (maintained by their publishers).
-{\bf EPTCS requires the inclusion of a DOI in each cited paper, when available.}
-
-Often an official publication is only available against payment, but
-as a courtesy to readers that do not wish to pay, the authors also
-make the paper available free of charge at a repository such as
-\url{arXiv.org}. In such a case it is recommended to also refer and
-link to the URL of the response page of the paper in such a
-repository.  This can be done using the bibtex fields {\tt ee} or {\tt
-url}, which are treated as synonyms.  These fields should not be used
-to duplicate information that is already provided through the DOI of
-the paper.
-You can find archival-quality URL's for most recently published papers
-in DBLP---they are in the bibtex-field {\tt ee}. In fact, it is often
-useful to check your references against DBLP records anyway, or just find
-them there in the first place.
-
-When using {\LaTeX} rather than {\tt pdflatex} to typeset your paper, by
-default no linebreaking within long URLs is allowed. This leads often
-to very ugly output, that moreover is different from the output
-generated when using {\tt pdflatex}. This problem is repaired when
-invoking \href{http://www.cse.unsw.edu.au/~rvg/EPTCS/breakurl.sty}
-{\tt $\backslash$usepackage$\{$breakurl$\}$}: it allows linebreaking
-within links and yield the same output as obtained by default with
-{\tt pdflatex}. 
-When invoking {\tt pdflatex}, the package {\tt breakurl} is ignored.
+\item interfaces, interactions, distribution : operating systems, networking
+\end{itemize}
 

### 2013-04-12 tfpie/tfpie.tex
Makefile and rename.
Some leg work on references.

--- 2013-04-11 tfpie/example.tex
@@
 \providecommand{\event}{TFPIE} % Name of the event you are submitting to
-\usepackage{breakurl}             % Not needed if you use pdflatex only.
 
-\title{How to Design Programs with Class}
+\title{From Principles to Fashion with Class}
 \author{Sam Tobin-Hochstadt \quad\qquad David Van Horn
@@
 
+\cite{dvanhorn:Felleisen2004Structure}
+
+\cite{dvanhorn:Findler2002DrScheme}
+
+\cite{dvanhorn:Felleisen2001How}
+
 \section{A small shift of focus}
@@
 
-\section{Distributed, Interactive Programs}
+\section{Our course}
 
-\section{Representation-indepence and interface-oriented programming}
+\subsection{Distributed, Interactive Programs}
 
-\section{Unifying functions and objects}
+\subsection{Representation-indepence and interface-oriented programming}
+
+\subsection{Unifying functions and objects}
 
@@
 \section{Comparison with HtDC}
+
+\cite{dvanhorn:Gray2003ProfessorJ}
+
+\cite{dvanhorn:Allen2002DrJava}
+
+\cite{dvanhorn:Kolling2003}
 
@@
 causing students to believe struggling with such crap is \emph{the
-  essence of computer science}.  Bright, promising students
+  essence of computer science}.  Bright, promising students may leave
+the program, understandably so, in order to pursue more intellectually
+fulfilling subjects.
 
@@
 
-\nocite{*}
+% \nocite{*}
 \bibliographystyle{eptcs}
-\bibliography{generic}
+\bibliography{dvh-bibliography}
 \end{document}

### 2013-04-12 tfpie/tfpie.tex
Progress.

@@
 
-\title{From Principles to Fashion with Class}
+\title{From Principles to Practice with Class in the First Year}
 \author{Sam Tobin-Hochstadt \quad\qquad David Van Horn
@@
 
+\begin{itemize}
+\item Lots of universities want their students to be proficient in an
+  industrial object-oriented programming language by the end of the first year.
+\item Teaching Java on Day 1 is a flawed approach.
+\item Teaching functional programming in Semester 1 and Java in Semester 2 is a flawed approach.
+\item We contribute a solution: a gradual transition from functional
+  programming to object-oriented programming that culminates in
+  industrial language proficiency.
+\end{itemize}
+
 \emph{Thesis: ...}
 
+Some of these are premature, and better introduced later in the
+course.  Some are just irrelevant trivia (different syntax) that often
+favors ``experienced'' students while needlessly intimidating students
+with only the previous semester under their belt.  Many of these
+trivial matters, e.g. Eclipse configuration, javac command line
+arguments, the CLASSPATH, can be very frustrating to the uninitiated,
+causing students to believe struggling with such crap is \emph{the
+  essence of computer science}.  Bright, promising students may leave
+the program, understandably so, in order to pursue more intellectually
+fulfilling subjects.
+
 \section{Background: the first semester}
+
+Context and contraints: HtDP vs. industrial co-op and subsequent Java courses.
+
+Universe and world \cite{dvanhorn:Felleisen2009Functional}.
 
@@
 
+\begin{enumerate}
+\item Lists of numbers
+\item World
+\end{enumerate}
+
 [First semester vs second semester example programs.]
 
-\section{Our course}
-
-\subsection{Distributed, Interactive Programs}
-
-\subsection{Representation-indepence and interface-oriented programming}
-
-\subsection{Unifying functions and objects}
-
-\section{From principles to fashions}
+\section{From principles to industrial languages}
 
@@
 
-\section{Comparison with HtDC}
+\section{Our course}
+
+\subsection{Representation-independence and interface-oriented programming}
+
+\subsection{Distributed, Interactive Programs}
+
+\subsection{Unifying functions and objects}
+
+\section{Related work}
 
@@
 
-\cite{dvanhorn:Allen2002DrJava}
+\cite{dvanhorn:Allen2002DrJava, dvanhorn:Hsia2005Taming}
 
 \cite{dvanhorn:Kolling2003}
+
+\cite{dvanhorn:Bruce2001Library}
+
+\cite{dvanhorn:Alphonce2003Using}
+
+\cite{dvanhorn:Chakravarty2004Risks}
+
+\cite{dvanhorn:Ragde2008Chilling}
+
+\cite{dvanhorn:Bloch2000Scheme}
 
@@
 \begin{itemize}
+\item dead.
 \item changes languages from first semester rather than adding new
@@
 
-Some of these are premature, and better introduced later in the
-course.  Some are just irrelevant trivia (different syntax) that often
-favors ``experienced'' students while needlessly intimidating students
-with only the previous semester under their belt.  Many of these
-trivial matters, e.g. Eclipse configuration, javac command line
-arguments, the CLASSPATH, can be very frustrating to the uninitiated,
-causing students to believe struggling with such crap is \emph{the
-  essence of computer science}.  Bright, promising students may leave
-the program, understandably so, in order to pursue more intellectually
-fulfilling subjects.
-
 \section{Experience and outlook}
@@
 
+Other steps:
+\begin{itemize}
+\item types in class language
+\item pedagogical Java environment
+\item ugly syntax for class language
+\item world in Java
+\end{itemize}
+
+\cite{dvanhorn:sicp}
+
 % \nocite{*}

### 2013-04-13 tfpie/tfpie.tex
Some work on intro.

@@
 
-\begin{itemize}
-\item Lots of universities want their students to be proficient in an
-  industrial object-oriented programming language by the end of the first year.
-\item Teaching Java on Day 1 is a flawed approach.
-\item Teaching functional programming in Semester 1 and Java in Semester 2 is a flawed approach.
-\item We contribute a solution: a gradual transition from functional
-  programming to object-oriented programming that culminates in
-  industrial language proficiency.
-\end{itemize}
+Many universities and colleges aim to teach their students proficiency
+in an industrial object-oriented programming language by the end of
+their first year.  Past approaches to achieve this aim have included
+teaching an industrial language, such as Java, starting on the first
+day.  Others have taken a more indirect route by teaching functional
+programming in the first semester, followed by a second semester that
+starts in Java.  The latter approach is an improvement over the first,
+but both suffer serious flaws.
+
+We contribute an alternative approach to the second semester that
+provides a gradual transition from functional programming to
+object-oriented programming that culminates in industrial language
+proficiency.
 

### 2013-04-13 tfpie/tfpie.tex
start on context

@@
 \section{Background: the first semester}
+
+At Northeastern, the College of Computer \& Information Science
+requires a four course introductory sequence in the first year.  The
+first semester features both a course on discrete mathematics as well
+as an introduction to programming following the \emph{How to Design
+  Programs} curriculum.  The second semester follows with a course on
+object-oriented programming and one featuring formal reasoning about
+programs, both on paper and with the ACL2 theorem prover.
+
+Subsequent to the first year, students take a wide variety of
+follow-up courses, ranging from a required course in ``Object-oriented
+design'' to architecture, operating systems, robotics, and programming
+languages.  No standard language is used in these courses.
+
+More significantly, Northeastern distinctively emphasizes experiential
+education, with almost all Computer Science majors participating in a
+6 month ``co-op'' internship after their third semester.  These co-ops
+take place at a wide variety of companies, and while most students do
+some software development, there is almost no uniformity beyond that.
 

### 2013-04-13 tfpie/tfpie.tex
Fix to local bib.

@@
 
-\title{From Principles to Practice with Class in the First Year}
+\newcommand\thetitle{From Principles to Practice with Class in the First Year}
+
+\title{\thetitle}
 \author{Sam Tobin-Hochstadt \quad\qquad David Van Horn
@@
 }
-\def\titlerunning{A Longtitled Paper}
+\def\titlerunning{\thetitle}
 \def\authorrunning{S. Tobin-Hochstadt \& D. Van Horn}
@@
 
+As an example, Northeastern University teaches functional programming
+in the first semester using \emph{How to Design Programs}
+\cite{dvanhorn:Felleisen2001How}, followed by object-oriented
+programming using \emph{How to Design Classes} \cite{local:htdc} in
+the second semester.  This sequence is one
+
+These two courses are tightly integrated
+and have been taught by the designers of this cirriculum.
+
 We contribute an alternative approach to the second semester that
@@
 \emph{Thesis: ...}
+
+\cite{local:dpc}
 
@@
 
+\cite{local:htdc}
+
 % \nocite{*}
 \bibliographystyle{eptcs}
-\bibliography{dvh-bibliography}
+\bibliography{dvh-bibliography,local}
 \end{document}

### 2013-04-13 tfpie/tfpie.tex
some more

@@
 
-Context and contraints: HtDP vs. industrial co-op and subsequent Java courses.
+This combination sets the constraints in which we designed our
+approach.  Our students begin the course with a firm grasp of
+data-driven program design, as well as experience with the tools used
+in the first course:
 
-Universe and world \cite{dvanhorn:Felleisen2009Functional}.
+\begin{itemize}
+  \item the student languages introduced in \emph{How to Design
+    Programs}~\cite{dvanhorn:Felleisen2004Structure,
+    dvanhorn:Felleisen2001How} as well as the idea of ``language levels'';
 
-\cite{dvanhorn:Felleisen2004Structure}
+  \item the DrRacket programming environment~\cite{dvanhorn:Findler2002DrScheme};
 
-\cite{dvanhorn:Findler2002DrScheme}
+  \item the \texttt{check-expect} testing framework~\cite{local:check-expect};
 
-\cite{dvanhorn:Felleisen2001How}
+  \item the ``World'' approach to developing interactive animations
+    and games using functional
+    programming~\cite{dvanhorn:Felleisen2009Functional}.
+\end{itemize}
+
+\noindent
+We reuses all of these elements in our approach.
+
+After our course, our students should both (a) be prepared for
+subsequent courses in the curriculum, which expect familiarity with
+Java and standard Java libraries, (b) be prepared for co-ops in which
+they will use industrial-strength languages and tools which will
+almost certainly be object-oriented.  More significantly, we aim to
+teach the key insights behind the object-oriented approach to program
+design.
 

### 2013-04-13 tfpie/tfpie.tex
acks

@@
 
+\subsection*{Acknowledgements}
+
+Matthias Felleisen's approach to pedagody and passion for
+undergraduate teaching has inspired this work from the beginning.
+CCIS Dean Larry Finkelstein allowed two postdocs to redesign a key
+undergraduate course.  Our teaching assistants, Dan Brown, Asumu
+Takikawa, and Nicholas Labich, as well as the tutors and graders,
+contributed enormously to the success of our courses. Finally, and
+most importantly, our students at Northeastern for the last three
+years have put up with a curriculum in progress, and the opportunity
+to teach them has been truly rewarding.
+
 % \nocite{*}

### 2013-04-13 tfpie/tfpie.tex
finish a draft of 2

@@
 
-\section{Background: the first semester}
+\section{Background: the context at Northeastern}
 
@@
 
-  \item the DrRacket programming environment~\cite{dvanhorn:Findler2002DrScheme};
+  \item the DrRacket (formerly DrScheme) 
+    programming environment~\cite{dvanhorn:Findler2002DrScheme};
 
@@
 \noindent
-We reuses all of these elements in our approach.
+We reuse all of these elements in our approach.
 
@@
 design.
+
+These constraints, while in detail specific to Northeastern and the
+CCIS curriculum, are broadly similar to the requirements for the first
+year at many universities.  Our course also attends to smaller and
+more idiosyncratic elements of our curriculum, ranging from formal
+reasoning to algorithmic analysis, as described in
+section~\ref{sec:our-course}.
 
@@
 \section{Our course}
+\label{sec:our-course}
 

### 2013-04-13 tfpie/tfpie.tex
First draft of intro.

@@
 As an example, Northeastern University teaches functional programming
-in the first semester using \emph{How to Design Programs}
-\cite{dvanhorn:Felleisen2001How}, followed by object-oriented
-programming using \emph{How to Design Classes} \cite{local:htdc} in
-the second semester.  This sequence is one
+in the first semester using \emph{How to Design
+  Programs}~\cite{dvanhorn:Felleisen2001How}, followed by
+object-oriented programming using \emph{How to Design
+  Classes}~\cite{local:htdc} in the second semester.  This sequence
+was designed to provide a smooth path for incoming students assumed
+only to have a competence in high-school level algebra to reach
+proficiency in Java by the end of their first year
+\cite{dvanhorn:Felleisen2004Structure}.  It was a major improvement
+over the previous Java-first cirriculum in terms of student sucess,
+attrition, and preparation for subsequent
+courses~\cite{dvanhorn:Proulx2006Design}.  However significant
+problems remain; in particular, the second semester course violates
+the designers' own principles (as recalled
+in~\cite{dvanhorn:Bloch2000Scheme}):
+%
+\begin{enumerate}
+\item \emph{introduce only those language constructs that are necessary to
+  teach programming principles}, and
+\item \emph{choose a language with as
+  few language constructs as possible, and one in which they can be
+  introduced one at a time}.
+\end{enumerate}
 
-These two courses are tightly integrated
-and have been taught by the designers of this cirriculum.
+The problem is that the first semester ends with an advanced
+pedagogical functional language and the second semester starts with a
+beginning pedagogical subset of Java.  Despite this slimmed-down
+subset, this transition is too abrupt to meaningfully bridge the gap
+between functional and object-oriented programming, because there are
+several other significant transitions happening in concert such as the
+transitions:
+\begin{itemize}
+\item from a highly regular and minimal syntax to a complicated
+  irregular syntax,
+
+\item from an untyped language to a typed language,
+
+\item from a pedagogical programming environment to an industrial
+programming environment (Eclipse),
+
+\item from a language with numeric values corresponding to
+  mathematical objects to a language with numeric values corresponding
+  to common machine representations,
+
+\item from a language with image literals and graphical libraries to
+  one in which graphical programming is tedious,
+
+\item from a language with a simple structural-equality-based testing
+  framework to a language with an industrial strength testing library,
+
+\item from an interaction-oriented language to a compiled,
+  batch-oriented language.  
+\end{itemize}
+
+This abrupt transition has several negative consequences: the
+principles of object-oriented programming are obscured and
+de-emphasized, struggling with the programming environment is
+frustrating and can cause potentially good students to leave the
+program, it favors students with prior exposure to the particular
+tools, it inhibits students from experimenting by relying upon their
+past skills, and it creates the false impression that courses are
+discrete units of instruction that can be discarded after successful
+completion rather than being part of a continuous and culmative
+educational experience.
 
 We contribute an alternative approach to the second semester that
-provides a gradual transition from functional programming to
-object-oriented programming that culminates in industrial language
-proficiency.
+overcomes these problems and provides a gradual transition from
+functional programming to object-oriented programming.  Our approach
+is able to start the second semester by introducing \emph{only} the
+concept of programming with objects, while all other aspects of course
+remain where they were left off in the previous semester.  This
+approach allows the other concepts to be introduced at the point at
+which they are relevant and motivated.  Despite this more gradual
+approach, the course accomplishes the goal of teaching industrial
+language proficiency by the end of the semester, covering a super-set
+of the concepts and topics covered in the \emph{How to Design Classes}
+course.
 
-\emph{Thesis: ...}
+\paragraph{Outline} The remainder of this paper is organized as follows\dots
 
-\cite{local:dpc}
+%% \cite{local:dpc}
 
-Some of these are premature, and better introduced later in the
-course.  Some are just irrelevant trivia (different syntax) that often
-favors ``experienced'' students while needlessly intimidating students
-with only the previous semester under their belt.  Many of these
-trivial matters, e.g. Eclipse configuration, javac command line
-arguments, the CLASSPATH, can be very frustrating to the uninitiated,
-causing students to believe struggling with such crap is \emph{the
-  essence of computer science}.  Bright, promising students may leave
-the program, understandably so, in order to pursue more intellectually
-fulfilling subjects.
+%% Some of these are premature, and better introduced later in the
+%% course.  Some are just irrelevant trivia (different syntax) that often
+%% favors ``experienced'' students while needlessly intimidating students
+%% with only the previous semester under their belt.  Many of these
+%% trivial matters, e.g. Eclipse configuration, javac command line
+%% arguments, the CLASSPATH, can be very frustrating to the uninitiated,
+%% causing students to believe struggling with such crap is \emph{the
+%%   essence of computer science}.  Bright, promising students may leave
+%% the program, understandably so, in order to pursue more intellectually
+%% fulfilling subjects.
 

### 2013-04-13 tfpie/tfpie.tex
Section labels.

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 \section{Background: the context at Northeastern}
+\label{sec:background}
 
@@
 object-oriented programming and one featuring formal reasoning about
-programs, both on paper and with the ACL2 theorem prover.
+programs, both on paper and with the ACL2 theorem
+prover~\cite{dvanhorn:Kaufmann2000ComputerAided}.
 
@@
 \section{A small shift of focus}
+\label{sec:shift}
 
@@
 \section{From principles to industrial languages}
+\label{sec:industrial}
 
@@
 \section{Related work}
+\label{sec:related-work}
 
@@
 \section{Experience and outlook}
+\label{sec:conclusion}
 

### 2013-04-13 tfpie/tfpie.tex
code comparison

@@
 
+In intermediate student:
+
+\begin{figure}
+\begin{minipage}{3.5in}
+\begin{verbatim}
+
+;; A Tree is one of:
+;; - (make-leaf Number)
+;; - (make-node Tree Number Tree)
+(define-struct leaf (v))
+(define-struct node (left v right))
+
+;; sum : Tree -> Number
+;; sums the elements of the given tree
+(define (sum a-tree)
+  (cond [(leaf? a-tree) (leaf-v a-tree)]
+        [else
+        (+ (sum (node-left  a-tree))
+           (node-v a-tree)
+           (sum (node-right a-tree)))]))
+
+(check-expect (sum (make-leaf 7)) 7)
+(check-expect 
+  (sum (make-node
+         (make-leaf 1)
+         5 
+         (make-node (make-leaf 0)
+                    10
+                    (make-leaf 0))))
+  16)
+\end{verbatim}
+\end{minipage}
+\begin{minipage}{3in}
+\begin{verbatim}
+#lang class/1
+;; A Tree is one of:
+;; - (new leaf% Number)
+;; - (new node% Tree Number Tree)
+;; and implements
+;; sum : -> Number
+;; sums the elements of this tree
+
+(define-class leaf%
+  (fields v)
+  (define (sum) (this . v)))
+
+(define-class node%
+  (fields left v right)
+  (define (sum)
+    (+ (this . left . sum)
+       (this . v)
+       (this . right .sum))))
+
+(check-expect (sum (new leaf% 7)) 7)
+(check-expect 
+  (sum (new node%
+         (new leaf% 1)
+         5 
+         (new node% (new leaf% 0) 
+                    10
+                    (new leaf% 0))))
+  16)
+\end{verbatim}
+\end{minipage}
+\caption{Binary tree sum in the first course, and in our approach}
+\end{figure}
+
+In Java:
+
+\begin{figure}
+\begin{verbatim}
+import tester.*;
+
+interface Tree {
+    // sums the elements of this tree
+    Integer sum();
+}
+
+class Leaf {
+    Integer v;
+    Leaf(Integer v) { this.v = v; }
+    public Integer sum() { return this.v; }
+}
+
+class Node {
+    Tree left; Integer v; Tree right;
+    Node(Tree l, Integer v, Tree r) {
+        this.left = l;
+        this.v = v;
+        this.right = r;
+    }
+
+    public Integer sum() {
+        return this.left.sum() + this.v + this.right.sum();
+    }
+}
+
+class Examples {
+    void test_tree(Tester t) {
+        t.checkExpect((new Leaf(7)).sum(), 7);
+        t.checkExpect((new Node(new Leaf(1),
+                                5, 
+                                new Node(new Leaf(0), 10, new Leaf(0)))).sum(),
+                      16);
+    }
+}
+\end{verbatim}
+\caption{Binary tree sum in How to Design Classes}
+\end{figure}
+
 \begin{enumerate}

### 2013-04-13 tfpie/tfpie.tex
Outline.

@@
 
-\paragraph{Outline} The remainder of this paper is organized as follows\dots
+\paragraph{Outline} The remainder of this paper is organized as follows:
+in section~\ref{sec:background} provides background on \emph{How to
+  Design Programs} and the context and constraints involved in the
+first-year at Northeastern.  Section~\ref{sec:shift} describes our
+approach to the second semester, which starts with a small shift in
+perspective to bridge the gap between functional programming and
+object oriented programming.  Section~\ref{sec:industrial} describes
+the path toward programming in an industrial object-oriented
+programming language.  Section~\ref{sec:related-work} discusses the
+relation to existing work and section~\ref{sec:conclusion} concludes.
 

### 2013-04-13 tfpie/tfpie.tex
fixes for formatting

@@
 \begin{figure}
-\begin{minipage}{3.5in}
+\begin{minipage}[t]{3.5in}
 \begin{verbatim}
-
+#lang bsl
 ;; A Tree is one of:
@@
 \end{minipage}
-\begin{minipage}{3in}
+\begin{minipage}[t]{3in}
 \begin{verbatim}
@@
 \end{minipage}
-\caption{Binary tree sum in the first course, and in our approach}
+\caption{Binary tree sum in Beginning Student and in our language}
 \end{figure}
-
-In Java:
 

### 2013-04-13 tfpie/tfpie.tex
down with ()

@@
     void test_tree(Tester t) {
-        t.checkExpect((new Leaf(7)).sum(), 7);
-        t.checkExpect((new Node(new Leaf(1),
+        t.checkExpect(new Leaf(7).sum(), 7);
+        t.checkExpect(new Node(new Leaf(1),
                                5, 
-                                new Node(new Leaf(0), 10, new Leaf(0)))).sum(),
+                               new Node(new Leaf(0), 10, new Leaf(0))).sum(),
                       16);

### 2013-04-13 tfpie/tfpie.tex
Spelling errors.

@@
 We propose a bridge between functional and object-oriented programming
-in the first-year cirriculum.
+in the first-year curriculum.
 \end{abstract}
@@
 in an industrial object-oriented programming language by the end of
-their first year.  Past approaches to achieve this aim have included
-teaching an industrial language, such as Java, starting on the first
-day.  Others have taken a more indirect route by teaching functional
-programming in the first semester, followed by a second semester that
-starts in Java.  The latter approach is an improvement over the first,
-but both suffer serious flaws.
+the students' first year.  Past approaches to achieve this aim have
+included teaching an industrial language, such as Java, starting on
+the first day.  Others have taken a more indirect route by teaching
+functional programming in the first semester, followed by a second
+semester that starts in Java.  The latter approach is an improvement
+over the first, but both suffer serious flaws.
 
@@
 \cite{dvanhorn:Felleisen2004Structure}.  It was a major improvement
-over the previous Java-first cirriculum in terms of student sucess,
+over the previous Java-first curriculum in terms of student success,
 attrition, and preparation for subsequent
@@
 
-\item from a language with a simple structural-equality-based testing
-  framework to a language with an industrial strength testing library,
-
 \item from an interaction-oriented language to a compiled,
@@
 discrete units of instruction that can be discarded after successful
-completion rather than being part of a continuous and culmative
+completion rather than being part of a continuous and cumulative
 educational experience.
@@
 
-\item Java, even pared down pedagoical Java, is syntactically heavy; discourages
+\item Java, even pared down pedagogical Java, is syntactically heavy; discourages
   interactive programming
@@
 \item Types
-\item Total committment to objects
+\item Total commitment to objects
 \item Subtly different base values
@@
 
-Perspective into the upper-class cirriculum:
+Perspective into the upper-class curriculum:
 
@@
 
-Matthias Felleisen's approach to pedagody and passion for
+Matthias Felleisen's approach to pedagogy and passion for
 undergraduate teaching has inspired this work from the beginning.
-CCIS Dean Larry Finkelstein allowed two postdocs to redesign a key
-undergraduate course.  Our teaching assistants, Dan Brown, Asumu
-Takikawa, and Nicholas Labich, as well as the tutors and graders,
-contributed enormously to the success of our courses. Finally, and
-most importantly, our students at Northeastern for the last three
-years have put up with a curriculum in progress, and the opportunity
-to teach them has been truly rewarding.
+CCIS Dean Larry Finkelstein entrusted two postdocs with the redesign
+of a key undergraduate course, which made this experiment possible.
+Our teaching assistants, Dan Brown, Asumu Takikawa, and Nicholas
+Labich, as well as the tutors and graders, contributed enormously to
+the success of our courses. Finally, and most importantly, our
+students at Northeastern for the last three years have put up with a
+curriculum in progress, and the opportunity to teach them has been
+truly rewarding.
 

### 2013-04-13 tfpie/tfpie.tex
Intro CCIS acronym.
Elim some industrials.

@@
 
-\item from a pedagogical programming environment to an industrial
+\item from a pedagogical programming environment to a professional
 programming environment (Eclipse),
@@
 
-At Northeastern, the College of Computer \& Information Science
+At Northeastern, the College of Computer \& Information Science (CCIS)
 requires a four course introductory sequence in the first year.  The
@@
 Java and standard Java libraries, (b) be prepared for co-ops in which
-they will use industrial-strength languages and tools which will
+they will use professional-grade languages and tools which will
 almost certainly be object-oriented.  More significantly, we aim to

### 2013-04-13 tfpie/tfpie.tex
world programs

@@
 
-\begin{enumerate}
-\item Lists of numbers
-\item World
-\end{enumerate}
+\begin{figure}
+\begin{minipage}[t]{3.4in}
+\begin{verbatim}
+#lang bsl
+(require 2htdp/image 2htdp/universe)
+
+;; A World is a Number
+
+;; on-tick : World -> World
+(define (tick w) (add1 w))
+
+;; to-draw : World -> Image
+(define (draw w)
+  (circle w "solid" "red"))
+
+;; on-key : KeyEvent World -> World
+(define (on-key k w) 10)
+
+(big-bang 10
+          [to-draw draw]
+          [on-tick tick])
+\end{verbatim}
+\end{minipage}
+\begin{minipage}[t]{3in}
+\begin{verbatim}
+#lang class/1
+(require 2htdp/image class/universe)
+
+;; A World (new world% Number)
+(define-class world%
+  (fields n)
+  ;; on-tick : -> World
+  (define (on-tick)
+    (new world% (add1 (this . n))))
+  ;; to-draw : -> Image
+  (define (on-tick) 
+    (circle (this . n) "solid" "red"))
+  ;; on-key : KeyEvent -> World
+  (define (on-key k) (new world% 10)))
+
+(big-bang (new world% 10))
+\end{verbatim}
+\end{minipage}
+\caption{World programs}
+\label{fig:world}
+\end{figure}
 

### 2013-04-13 tfpie/tfpie.tex
formatting

@@
 ;; on-tick : World -> World
-(define (tick w) (add1 w))
+(define (tick w)
+  (add1 w))
 
@@
 
-;; A World (new world% Number)
+;; A World is a (new world% Number)
 (define-class world%
   (fields n)
+
   ;; on-tick : -> World
@@
     (new world% (add1 (this . n))))
+
   ;; to-draw : -> Image
@@
     (circle (this . n) "solid" "red"))
+
   ;; on-key : KeyEvent -> World

### 2013-04-15 tfpie/tfpie.tex
Small fix to OO tests.
Fix bug in world program.

@@
 
-(check-expect (sum (new leaf% 7)) 7)
+(check-expect ((new leaf% 7) . sum) 7)
 (check-expect 
-  (sum (new node%
+  ((new node%
         (new leaf% 1)
@@
                    (new leaf% 0))))
+    . sum)
   16)
@@
   ;; to-draw : -> Image
-  (define (on-tick) 
+  (define (to-draw) 
     (circle (this . n) "solid" "red"))

### 2013-04-17 tfpie/tfpie.tex
small edits from Katie

@@
 attrition, and preparation for subsequent
-courses~\cite{dvanhorn:Proulx2006Design}.  However significant
+courses~\cite{dvanhorn:Proulx2006Design}.  However, significant
 problems remain; in particular, the second semester course violates
@@
 requires a four course introductory sequence in the first year.  The
-first semester features both a course on discrete mathematics as well
-as an introduction to programming following the \emph{How to Design
+first semester features both a course on discrete mathematics and
+ an introduction to programming following the \emph{How to Design
   Programs} curriculum.  The second semester follows with a course on
@@
 
-This combination sets the constraints in which we designed our
+This combination sets the constraints under which we designed our
 approach.  Our students begin the course with a firm grasp of

### 2013-04-17 tfpie/tfpie.tex
reorg
start on related

@@
 
-\subsection{Side-by-side example of program design}
+\subsection{Day 1}
+
+[First semester vs second semester example programs.]
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
 \begin{minipage}[t]{3.5in}
@@
 
-\begin{figure}
-\begin{verbatim}
-import tester.*;
-
-interface Tree {
-    // sums the elements of this tree
-    Integer sum();
-}
-
-class Leaf {
-    Integer v;
-    Leaf(Integer v) { this.v = v; }
-    public Integer sum() { return this.v; }
-}
-
-class Node {
-    Tree left; Integer v; Tree right;
-    Node(Tree l, Integer v, Tree r) {
-        this.left = l;
-        this.v = v;
-        this.right = r;
-    }
-
-    public Integer sum() {
-        return this.left.sum() + this.v + this.right.sum();
-    }
-}
-
-class Examples {
-    void test_tree(Tester t) {
-        t.checkExpect(new Leaf(7).sum(), 7);
-        t.checkExpect(new Node(new Leaf(1),
-                               5, 
-                               new Node(new Leaf(0), 10, new Leaf(0))).sum(),
-                      16);
-    }
-}
-\end{verbatim}
-\caption{Binary tree sum in How to Design Classes}
-\end{figure}
+\subsection{Worlds and animations}
 
@@
 
-[First semester vs second semester example programs.]
+\subsection{Language levels}
+
+No SE tools/mechanism built into the language.
+
+\subsection{Universe}
+
+\subsection{Programming to interfaces}
 
@@
 
+\begin{figure}
+\begin{verbatim}
+import tester.*;
+
+interface Tree {
+    // sums the elements of this tree
+    Integer sum();
+}
+
+class Leaf {
+    Integer v;
+    Leaf(Integer v) { this.v = v; }
+    public Integer sum() { return this.v; }
+}
+
+class Node {
+    Tree left; Integer v; Tree right;
+    Node(Tree l, Integer v, Tree r) {
+        this.left = l;
+        this.v = v;
+        this.right = r;
+    }
+
+    public Integer sum() {
+        return this.left.sum() + this.v + this.right.sum();
+    }
+}
+
+class Examples {
+    void test_tree(Tester t) {
+        t.checkExpect(new Leaf(7).sum(), 7);
+        t.checkExpect(new Node(new Leaf(1),
+                               5, 
+                               new Node(new Leaf(0), 10, new Leaf(0))).sum(),
+                      16);
+    }
+}
+\end{verbatim}
+\caption{Binary tree sum in How to Design Classes}
+\end{figure}
+
 \subsection{Java}
@@
 
-\section{Our course}
-\label{sec:our-course}
-
-\subsection{Representation-independence and interface-oriented programming}
-
-\subsection{Distributed, Interactive Programs}
-
 \subsection{Unifying functions and objects}
@@
 
-\cite{dvanhorn:Gray2003ProfessorJ}
+Teaching programming principles in a functional style has a long
+history, with Abelson and Sussman's \emph{Structure and Interpretation
+of Computer Programs}~\cite{dvanhorn:sicp} a prominent example.  Our
+work follows in the tradition of the \emph{TeachScheme!} project, now
+\emph{Program by Design} (PbD), which emphasizes a systematic
+approach to program construction.
 
-\cite{dvanhorn:Allen2002DrJava, dvanhorn:Hsia2005Taming}
+Since the introduction of functional-first curricula, and more
+specifically in the Program by Design framework, numerous courses have
+tackled the problem of transition.  Typically they, as we, transition
+to Java in the second course.  We discuss first the approach developed
+by some of the principal creators of PbD, and then other approaches.
 
-\cite{dvanhorn:Kolling2003}
-
-\cite{dvanhorn:Bruce2001Library}
-
-\cite{dvanhorn:Alphonce2003Using}
-
-\cite{dvanhorn:Chakravarty2004Risks}
-
-\cite{dvanhorn:Ragde2008Chilling}
+\subsection{ProfessorJ}
 
 \cite{dvanhorn:Bloch2000Scheme}
+
+\cite{dvanhorn:Gray2003ProfessorJ}
 
@@
 
+\subsection{Other transitions}
+
+\cite{dvanhorn:Ragde2008Chilling}
+\cite{dvanhorn:Chakravarty2004Risks}
+
+\subsection{Other approaches to Java}
+
+\cite{dvanhorn:Allen2002DrJava, dvanhorn:Hsia2005Taming}
+
+\cite{dvanhorn:Kolling2003}
+
+\cite{dvanhorn:Bruce2001Library}
+
+\cite{dvanhorn:Alphonce2003Using}
+
 \section{Experience and outlook}

### 2013-04-17 tfpie/tfpie.tex
Start on section 3.

@@
 
-The semester starts with the minimal conceptual change from the
-previous semester: the addition of objects (same IDE, same syntax,
-same semantics [just a new kind of value], same interactive video
-games, etc.).
+On the first day of the second semester, we introduce a single
+linguistic concept to an otherwise unchanged context of the previous
+semester.
+
+An object is a new kind of value that can, as a first cut, be
+understood as a pairing together of two familiar concepts: data and
+functionality.
+
+\begin{itemize}
+\item An object is like a structure in that it has a fixed number of
+fields, thus an object (again, like a structure) can represent
+compound data. But unlike a structure, an object contains not just
+data, but functionality too;
+
+\item An object is like a (set of) function(s) in that it has behavior—it computes; it is
+not just inert data.
+\end{itemize}
+
+This suggests that objects are a natural ﬁt for well-designed programs
+since good programs are organized around data definitions and
+functions that operate over such data. An object, in essence, packages
+these two things together into a single programming apparatus. This
+has two important consequences:
+
+\begin{enumerate}
+
+\item Students already know how to design programs oriented around objects.
+
+Since objects are just the combination of two familiar concepts that
+students already use to design programs, they already know how to
+design programs around objects, even if they have never heard the term
+``object'' before. Our message is the better you are at programming
+with functions, the better you will be at programming with objects,
+and \emph{vice versa}.
+
+\item Objects enable new kinds of abstraction and composition.
+
+Although the combination of data and function may seem simple, objects
+enable new forms of abstraction and composition. That is, objects open
+up new approaches to the construction of computations. By studying
+these new approaches, we can distill new design principles. Because we
+understand objects are just the combination of data and function, we
+can understand how all of these principles apply in the familiar
+context of programming with functions. In short, the better you are at
+programming with objects, the better you will be at programming with
+functions.
+
+\end{enumerate}
+
+\subsection{The basics of objects}
+
+To begin with, we introduce the notion of a \emph{class definition},
+which can be thought of at first as a structure definition in that
+it defines a new class of compound data.  A class is defined using the
+{\tt define-class} form:
+\begin{verbatim}
+    (define-class posn% (fields x y))
+\end{verbatim}
+An \emph{object} is a value that is a member of this class of data,
+which can be constructed with the {\tt new} keyword, a class name, and
+the appropriate number of arguments for the fields of the object:
+\begin{verbatim}
+    (new posn% 3 4)
+\end{verbatim}
+An object understands some set of \emph{messages}.  Simple
+structure-like objects understand messages for accessing their fields
+and message are sent by using the {\tt send} keyword, followed by an object,
+a message name, and some number of arguments:
+\begin{verbatim}
+    (send (new posn% 3 4) x) ;=> 3
+    (send (new posn% 3 4) y) ;=> 4
+\end{verbatim}
+The {\tt send} notation is simple, but syntactically heavy.  Once
+students are comfortable with the {\tt send} form, we introduce a
+shorthand to make it more convenient by writing {\tt (x . m)} for {\tt
+  (send x m)}.  The \emph{dot notation} can be nested, so {\tt (x . m
+  . n)} is shorthand for {\tt (send (send x m) n)}.  (The approach of
+introducing a simple, uniform syntax and later introducing a
+convenient shorthand that would have been confusing to start with
+follows the approach of introducing {\tt cons} and later {\tt list}
+and {\tt quote} in the first semester.)
+
+It is possible to endow objects with functionality by defining \emph{methods},
+which extend the set of messages an object understands.  A method definition
+follows the same syntax as a function definition, but is located inside of a
+class definition.  Here is a more complete development of the {\tt posn\%} class
+that includes a couple of methods:
+\begin{verbatim}
+    ;; A Posn is a (new posn% Number Number),
+    ;; which represents a point on the Cartesian plane
+    (define-class posn% (fields x y)
+
+      ;; Posn -> Number
+      ;; Distance between this posn and that posn
+      (check-expect ((new posn% 0 0) . dist (new posn% 3 4)) 5)
+      (define (dist that)
+        (sqrt (+ (sqr (- (this . x) (that . x)))
+                 (sqr (- (this . y) (that . y))))))
+
+      ;; -> Number
+      ;; Distance of this posn from the origin
+      (check-expect ((new posn% 0 0) . dist-origin) 0)
+      (check-expect ((new posn% 3 4) . dist-origin) 5)
+      (define (dist-origin)
+        (this . dist (new posn% 0 0))))
+\end{verbatim}
+
+This class definition defines a new class of values which are {\tt
+  posn\%} objects.  Such objects are comprised of two numeric values
+and understand the messages {\tt x}, {\tt y}, {\tt dist}, and {\tt
+  dist-origin}.  Unit tests have been included with each method
+definition, following the principles of the design recipe studied in
+the first semester.  In fact, the {\tt check-expect} mechanism works
+exactly as it did before.
+
+Methods can be defined to consume any number of arguments, but they
+are implicitly parameterized over {\tt this}, the object that was sent
+the message.
+
+\subsection{Dispatch}
+
+Unions, and recursive unions in particular, are a fundamental kind of
+data definition that students are well-versed in from the previous
+semester.  A fundamental early lesson is how to represent (recursive)
+unions using classes and how to write recursive methods.  As an
+example, here is a simple class definition for lists of numbers
+(an archetypical recursive union data definition) that includes
+a method for computing the length of a list:
+\begin{verbatim}
+    ;; A list of numbers (LoN) is one of:
+    ;; - (new empty%)
+    ;; - (new cons% Number LoN)
+
+    (define-class empty%
+      ;; Length of this empty list
+      (check-expect ((new empty%) . len) 0)
+      (define (len) 0))
+
+    (define-class cons% (fields first rest)
+      ;; Length of this cons list
+      (check-expect ((new cons% 7 (new empty%)) . len) 1)
+      (define (len)
+        (add1 (this . rest . len))))
+\end{verbatim}
+
+The structure of this data definition is analagous to the approach of
+the previous semester but this example brings to light an important
+difference with the functional approach.
+
+%% The semester starts with the minimal conceptual change from the
+%% previous semester: the addition of objects (same IDE, same syntax,
+%% same semantics [just a new kind of value], same interactive video
+%% games, etc.).
 

### 2013-04-17 tfpie/tfpie.tex
up

@@
 \providecommand{\event}{TFPIE} % Name of the event you are submitting to
+
+\usepackage[numbers]{natbib}
+\usepackage{xspace}
 
@@
 
-\subsection{ProfessorJ}
+\subsection{TeachScheme! and ProfessorJ}
 
-\cite{dvanhorn:Bloch2000Scheme}
+The TeachScheme! project initially focused only on the first course,
+with the second course typically taught in Java in
+institution-specific ways. Subsequently, the pedagogical approach was
+extended to Java, but without the tool support and textbook of the
+first course.  An example of this approach is described by
+\citet{dvanhorn:Bloch2000Scheme}, who presents the experience
+integrating these courses at Adelphi.  He reports that ``many of
+Java's concepts could be introduced more easily in a second course
+than a first''.
 
-\cite{dvanhorn:Gray2003ProfessorJ}
+With these lessons in mind, the TeachScheme! project set out to apply
+the lessons of teaching languages and IDE support to Java, as well as
+to present the approach to object-oriented programming in textbook
+form. ProfessorJ~\cite{dvanhorn:Gray2003ProfessorJ} is the resulting
+system, accompanying the draft textbook \emph{How to Design
+  Classes}~\cite{local:htdc}.  In parallel to our course, Northeastern
+teaches the remainder of its computer science majors following this
+approach.
 
-The noble goals of ProfJ:
+ProfessorJ and How to Design Classes maintain many of the excellent
+ideas of the first course.  In particular, ProfessorJ brings language
+levels to Java, in an attempt to smooth the transition for students
+from the first course and provide more helpful feedback.  ProfessorJ
+is also embedded in the DrRacket IDE, increasing familiarity for the
+students and supporting tools such as an interactive read-eval-print
+loop.
 
-\begin{itemize}
-\item language levels
-\item continuity in IDE from first semester
-\end{itemize}
+\def\htdp{\emph{How to Design Programs}\xspace}
+\def\htdc{\emph{How to Design Classes}\xspace}
+\def\profj{{ProfessorJ}\xspace}
 
-Many of the shortcomings of HtDC and ProfJ are ``day 1'' problems.
-The transition is too abrupt and too large.
+However, the ``day 1'' transition from the student languages used with
+\htdp to \profj is too abrupt and too large.
+%
+Most significantly, changing languages from the first semester
+immediately rather than simply adding a new concept confuses too many
+issues for students.  On the first day of a \htdc-based course,
+students see object-orientation, a new programming paridigm; Java, a
+new language with new syntax, and static types, a crucial but
+orthogonal concept.
+
+This also ultimately lead to the tool being abandoned
+  as it was too difficult to maintain.)
 
@@
 \item dead.
-\item changes languages from first semester rather than adding new
-  concept (causes students confusing deep issue (O.O. design) and
-  shallow ones (Java-like syntax) and orthogonal issues (static type
-  discipline).  (This also ultimately lead to the tool being abandoned
-  as it was too difficult to maintain.)
+\item
 
@@
 % \nocite{*}
-\bibliographystyle{eptcs}
+\bibliographystyle{plainnat}
 \bibliography{dvh-bibliography,local}

### 2013-04-17 tfpie/tfpie.tex
Some crap on dispatch.

@@
 the previous semester but this example brings to light an important
-difference with the functional approach.
+difference with the functional approach.  The method for computing the
+length of the empty list is defined in the {\tt empty\%} class, while
+the method for computing the length of a non-empty list is in the {\tt
+  cons\%} class.  When a list object is sent the {\tt len} method,
+there is no function with a conditional to determine whether the
+object is empty or not---instead, the object itself takes care of
+computing length based on the {\tt len} method it contains.  This
+shift in perspective is at the core of object-orientation: objects
+contain their own behavior and the case analysis previously done in
+functions is replaced by dispatch.
 

### 2013-04-17 tfpie/tfpie.tex
Elim dispatch.

@@
 
-\subsection{Dispatch}
+\subsection{Where did the {\tt cond} go?}
 
@@
 contain their own behavior and the case analysis previously done in
-functions is replaced by dispatch.
+functions is eliminated.
 

### 2013-04-17 tfpie/tfpie.tex
more on profj
Use tree example.

--- 2013-04-17 tfpie/tfpie.tex
@@
 unions using classes and how to write recursive methods.  As an
-example, here is a simple class definition for lists of numbers
-(an archetypical recursive union data definition) that includes
-a method for computing the length of a list:
-\begin{verbatim}
-    ;; A list of numbers (LoN) is one of:
-    ;; - (new empty%)
-    ;; - (new cons% Number LoN)
+example, figure~\ref{fig:tree} defines binary trees of numbers
+(an archetypical recursive union data definition)
+using the BSL language and the Class language.
+%% \begin{verbatim}
+%%     ;; A list of numbers (LoN) is one of:
+%%     ;; - (new empty%)
+%%     ;; - (new cons% Number LoN)
 
-    (define-class empty%
-      ;; Length of this empty list
-      (check-expect ((new empty%) . len) 0)
-      (define (len) 0))
+%%     (define-class empty%
+%%       ;; Length of this empty list
+%%       (check-expect ((new empty%) . len) 0)
+%%       (define (len) 0))
 
-    (define-class cons% (fields first rest)
-      ;; Length of this cons list
-      (check-expect ((new cons% 7 (new empty%)) . len) 1)
-      (define (len)
-        (add1 (this . rest . len))))
-\end{verbatim}
-
-The structure of this data definition is analagous to the approach of
-the previous semester but this example brings to light an important
-difference with the functional approach.
-
-%% The semester starts with the minimal conceptual change from the
-%% previous semester: the addition of objects (same IDE, same syntax,
-%% same semantics [just a new kind of value], same interactive video
-%% games, etc.).
-
-\subsection{Day 1}
-
-[First semester vs second semester example programs.]
-
-In intermediate student:
+%%     (define-class cons% (fields first rest)
+%%       ;; Length of this cons list
+%%       (check-expect ((new cons% 7 (new empty%)) . len) 1)
+%%       (define (len)
+%%         (add1 (this . rest . len))))
+%% \end{verbatim}
 
@@
 \end{minipage}
-\caption{Binary tree sum in Beginning Student and in our language}
+\caption{Binary tree sum in Beginning Student and in the Class language}
+\label{fig:tree}
 \end{figure}
+
+The structure of this data definition is analagous to the approach of
+the previous semester but this example brings to light an important
+difference with the functional approach.  The method for computing the
+sum of the empty tree is defined in the {\tt leaf\%} class, while the
+method for computing the sum of a node is in the {\tt node\%} class.
+When a tree object is sent the {\tt sum} method, there is no function
+with a conditional to determine whether the object is a
+leaf---instead, the object itself takes care of computing the sum
+based on the {\tt sum} method it contains.  This shift in perspective
+is at the core of object-orientation: objects contain their own
+behavior and the case analysis previously done in functions is
+eliminated.
+
+%% The semester starts with the minimal conceptual change from the
+%% previous semester: the addition of objects (same IDE, same syntax,
+%% same semantics [just a new kind of value], same interactive video
+%% games, etc.).
 
@@
 orthogonal concept.
+%
+In contrast, our course presents just one of these concepts on the
+first day, but  covers all of them by the end of the semester.
 

### 2013-04-17 tfpie/tfpie.tex
more about profj

@@
 
-This also ultimately lead to the tool being abandoned
-  as it was too difficult to maintain.)
+\profj also takes on the the dual challenges of implementing Java as
+well as subsetting it.  This ultimately resulted in both a limited
+Java environment as well as the eventual abandoning of the tool since
+it was too difficult to maintain, let alone keep up with advances in
+Java.
 
-The failures of  ProfJ:
+Finally, committing to Java on the first day, regardless of the
+environment provided to students, has significant limitations.  First,
+the syntactic and semantic heaviness of Java is a burden for beginning
+students, and discourages interactive and experimental programming.
+The very first chapter of \htdc discusses the fixed size of Java integers,
+a topic avoided entirely in the first course.
+Second, by committing to a particular industrial-strength language, it
+closes off possibilities in the curriculum.  Third, it commits
+entirely to the new paradigm, making it more difficult for students to
+compare the approaches.
 
-\begin{itemize}
-\item dead.
-\item
-
-\item commits to a particular fashion rather than enabling any.
-
-\item abandons the first semester approach and replaces it with a
-  (completely) object-oriented approach on day one.  There are
-  analogies between functional and object concepts (``a method is like
-  a function''), but the two never co-exist in a program.
-
-\item Java, even pared down pedagogical Java, is syntactically heavy; discourages
-  interactive programming
-
-\end{itemize}
-
-The changes that occur on Day 1 of HtDC w/ ProfJ:
-
-\begin{itemize}
-\item Radically different syntax
-\item Types
-\item Total commitment to objects
-\item Subtly different base values
-\end{itemize}
-
-The changes that occur on Day 1 of HtDC w/o ProfJ:
-
-\begin{itemize}
-\item Different IDE
-\item Compilation
-\item Loss of interactivity
-\end{itemize}
+Since \profj is no longer available, students are faced with an even
+starker change on the first day.  Even with a student-oriented
+environment such as DrJava or BlueJ~\cite{dvanhorn:Allen2002DrJava,
+  dvanhorn:Hsia2005Taming}, students must learn an entirely new tool,
+along with new libraries.
+If the course uses a typical professional development environment such
+as Eclipse, students must also contend with compilation, loss of
+interactivity, and subtle issues such as classpaths, none of which are
+fundamental to the concepts that the course focuses on.
 

### 2013-04-17 tfpie/tfpie.tex
Small touch ups.

@@
 linguistic concept to an otherwise unchanged context of the previous
-semester.
-
+semester: the idea of an object.
+%
 An object is a new kind of value that can, as a first cut, be
@@
 
-\item An object is like a (set of) function(s) in that it has behavior—it computes; it is
+\item An object is like a (set of) function(s) in that it has behavior---it computes; it is
 not just inert data.
@@
 
-This suggests that objects are a natural ﬁt for well-designed programs
+This suggests that objects are a natural fit for well-designed programs
 since good programs are organized around data definitions and
@@
 design programs around objects, even if they have never heard the term
-``object'' before. Our message is the better you are at programming
-with functions, the better you will be at programming with objects,
-and \emph{vice versa}.
+``object'' before.
 
@@
 
-Although the combination of data and function may seem simple, objects
-enable new forms of abstraction and composition. That is, objects open
-up new approaches to the construction of computations. By studying
-these new approaches, we can distill new design principles. Because we
-understand objects are just the combination of data and function, we
-can understand how all of these principles apply in the familiar
-context of programming with functions. In short, the better you are at
-programming with objects, the better you will be at programming with
-functions.
+Although the combination of data and functionality may seem simple,
+objects enable new forms of abstraction and composition. That is,
+objects open up new approaches to the construction of computations. By
+studying these new approaches, we can distill new design
+principles. Because we understand objects are just the combination of
+data and functionality, we can understand how all of these principles
+apply in the familiar context of programming with functions. In short,
+the better you are at programming with objects, the better you will be
+at programming with functions, and \emph{vice versa}.
 

### 2013-04-17 tfpie/tfpie.tex
No % convention.

@@
 \begin{verbatim}
-    (define-class posn% (fields x y))
+    (define-class posn (fields x y))
 \end{verbatim}
@@
 \begin{verbatim}
-    (new posn% 3 4)
+    (new posn 3 4)
 \end{verbatim}
@@
 \begin{verbatim}
-    (send (new posn% 3 4) x) ;=> 3
-    (send (new posn% 3 4) y) ;=> 4
+    (send (new posn 3 4) x) ;=> 3
+    (send (new posn 3 4) y) ;=> 4
 \end{verbatim}
@@
 follows the same syntax as a function definition, but is located inside of a
-class definition.  Here is a more complete development of the {\tt posn\%} class
+class definition.  Here is a more complete development of the {\tt posn} class
 that includes a couple of methods:
 \begin{verbatim}
-    ;; A Posn is a (new posn% Number Number),
+    ;; A Posn is a (new posn Number Number),
     ;; which represents a point on the Cartesian plane
-    (define-class posn% (fields x y)
+    (define-class posn (fields x y)
 
@@
       ;; Distance between this posn and that posn
-      (check-expect ((new posn% 0 0) . dist (new posn% 3 4)) 5)
+      (check-expect ((new posn 0 0) . dist (new posn 3 4)) 5)
       (define (dist that)
@@
       ;; Distance of this posn from the origin
-      (check-expect ((new posn% 0 0) . dist-origin) 0)
-      (check-expect ((new posn% 3 4) . dist-origin) 5)
+      (check-expect ((new posn 0 0) . dist-origin) 0)
+      (check-expect ((new posn 3 4) . dist-origin) 5)
       (define (dist-origin)
-        (this . dist (new posn% 0 0))))
+        (this . dist (new posn 0 0))))
 \end{verbatim}
@@
 This class definition defines a new class of values which are {\tt
-  posn\%} objects.  Such objects are comprised of two numeric values
+  posn} objects.  Such objects are comprised of two numeric values
 and understand the messages {\tt x}, {\tt y}, {\tt dist}, and {\tt
@@
 ;; A Tree is one of:
-;; - (new leaf% Number)
-;; - (new node% Tree Number Tree)
+;; - (new leaf Number)
+;; - (new node Tree Number Tree)
 ;; and implements
@@
 
-(define-class leaf%
+(define-class leaf
   (fields v)
@@
 
-(define-class node%
+(define-class node
   (fields left v right)
@@
 
-(check-expect ((new leaf% 7) . sum) 7)
+(check-expect ((new leaf 7) . sum) 7)
 (check-expect 
-  ((new node%
-        (new leaf% 1)
+  ((new node
+        (new leaf 1)
         5 
-        (new node% (new leaf% 0) 
+        (new node (new leaf 0) 
                   10
-                   (new leaf% 0))))
+                  (new leaf 0))))
     . sum)
@@
 difference with the functional approach.  The method for computing the
-sum of the empty tree is defined in the {\tt leaf\%} class, while the
-method for computing the sum of a node is in the {\tt node\%} class.
+sum of the empty tree is defined in the {\tt leaf} class, while the
+method for computing the sum of a node is in the {\tt node} class.
 When a tree object is sent the {\tt sum} method, there is no function
@@
 
-;; A World is a (new world% Number)
-(define-class world%
+;; A World is a (new world Number)
+(define-class world
   (fields n)
@@
   (define (on-tick)
-    (new world% (add1 (this . n))))
+    (new world (add1 (this . n))))
 
@@
   ;; on-key : KeyEvent -> World
-  (define (on-key k) (new world% 10)))
+  (define (on-key k) (new world 10)))
 
-(big-bang (new world% 10))
+(big-bang (new world 10))
 \end{verbatim}

### 2013-04-18 tfpie/tfpie.tex
start on more transitions
sec 5.2

@@
 
-\cite{dvanhorn:Ragde2008Chilling}
-\cite{dvanhorn:Chakravarty2004Risks}
+Of course, not every curriculum that begins with \htdp transitions to
+Java after the first course.  \citet{dvanhorn:Ragde2008Chilling}
+describes a second course that includes both more advanced work in
+Scheme beyond teaching-oriented languages as well as low-level
+programming in C, taught to computer science majors at University of
+Waterloo.  Radge's course intentionally does not use student-oriented
+languages, although the recently-developed C0 language~\cite{local:c0}
+could provide this. Other discussions of functional programming in the
+first year~\cite{dvanhorn:Chakravarty2004Risks} do not discuss the
+problems of transition.
 

### 2013-04-18 tfpie/tfpie.tex
Some writing on world.

@@
 data and functionality, we can understand how all of these principles
-apply in the familiar context of programming with functions. In short,
-the better you are at programming with objects, the better you will be
-at programming with functions, and \emph{vice versa}.
-
+apply in the familiar context of programming with functions. 
 \end{enumerate}
@@
 \subsection{Worlds and animations}
+
+Programming in the first semester is often oriented around interactive
+event-driven video games.  The basic design of a video game involves
+defining a data representation for states of the game and functions
+for transitioning between states based on events such as clock ticks,
+keyboard input, or mouse events.  The design of a game thus involves
+the design of data and functions on that data; in other words, the
+game involves the design of objects.  We therefore continue in the
+second semester with the use of programming video games but supplement
+the course with a library for doing so in an object-oriented style.
+Figure~\ref{fig:world} gives an example written in both the functional
+style and object-oriented style.
 

### 2013-04-18 tfpie/tfpie.tex
finish related work

@@
 
-\cite{dvanhorn:Allen2002DrJava, dvanhorn:Hsia2005Taming}
-
-\cite{dvanhorn:Kolling2003}
-
-\cite{dvanhorn:Bruce2001Library}
-
-\cite{dvanhorn:Alphonce2003Using}
+The problems of teaching Java in introductory courses have been
+well-explored; we mention only a few related directions here.
+DrJava~\cite{dvanhorn:Allen2002DrJava} and
+BlueJ~\cite{dvanhorn:Hsia2005Taming,dvanhorn:Kolling2003} are
+introductory environments for Java, which alleviate some but not all
+of the drawbacks we have outlined.
+%
+Several teaching-oriented graphics libraries for Java have been
+proposed~\cite{dvanhorn:Bruce2001Library,dvanhorn:Alphonce2003Using},
+but these are significantly more complex than the graphics and
+interaction libraries we are able to use in the introductory language
+we present.
 

### 2013-04-18 tfpie/tfpie.tex
fix secref

@@
 more idiosyncratic elements of our curriculum, ranging from formal
-reasoning to algorithmic analysis, as described in
-section~\ref{sec:our-course}.
+reasoning to algorithmic analysis, as described in the following
+sections.
 

### 2013-04-18 tfpie/tfpie.tex
Citation for image library.
start on conclusion

@@
   \item the ``World'' approach to developing interactive animations
-    and games using functional
-    programming~\cite{dvanhorn:Felleisen2009Functional}.
+    and games using functional programming and functional
+    graphics~\cite{dvanhorn:Felleisen2009Functional,
+      local:barland-sfp10}.
 \end{itemize}
@@
 
+We have now completed the third iteration of this course, teaching
+approximately 35 students each time.  Our experience has been
+uniformly positive, and the students have gone on to significant
+success in the subsequent courses, despite the curriculum differing
+from what the bulk of Northeastern University computer science majors
+take.
+
 Perspective into the upper-class curriculum:

### 2013-04-18 tfpie/tfpie.tex
finish conclusion

@@
 from what the bulk of Northeastern University computer science majors
-take.
+take.  The class has also had notable success in the recruitment and
+retention of female students, as compared to the other versions of the
+second-semester course.
 
-Perspective into the upper-class curriculum:
+The course has also provided a vantage point to introduce topics that
+will be taken up later in the curriculum.  We present types,
+contracts, invariants, and properties of functions, all of which tie
+into both the concurrent course on logic and computation, as well as
+later classes on formal methods.  The emphasis on
+representation-independence and interfaces  both ties into later
+classes on software engineering, as well as preparing the ground for
+algorithms and data structures courses.  Finally, the use of
+interactive and distributed systems connects to later courses on
+operating systems and networks.
 
-\begin{itemize}
-\item types, contract, invariants : formal methods, program verification,
+Despite our success, much remains to be done.  Type systems are a
+fundamental concept, but their introduction accompanies the rest of
+Java.  Developing a typed version of our introductory languages would
+allow a smoother introduction of this idea.
 
-\item representation independence : component-based software engineering,
-  data-structures, algorithms
+Our class's use of Eclipse could also be improved by first
+transitioning to a pedagogically-oriented Java environment, but we
+have not evaluated the specific options.  Alternatively, introducing
+Java-like syntax for the teaching languages we have developed would
+help tease apart the difficult transitions still present in the course.
 
-\item interfaces, interactions, distribution : operating systems, networking
-\end{itemize}
-
-Other steps:
-\begin{itemize}
-\item types in class language
-\item pedagogical Java environment
-\item ugly syntax for class language
-\item world in Java
-\end{itemize}
-
-\cite{dvanhorn:sicp}
-
-\cite{local:htdc}
+Finally, the Java portion of the class does not continue the use of
+``World''-style interactive graphical
+programming. \citet{local:java-world}.  Instead, our course focuses on
+coverage of standard Java libraries, as well as introductory
+algorithmic and data structure topics.  Continuing to use World-style
+programming in motivating examples might be valuable for continutity
+between the two halves of the course.
 

### 2013-04-19 tfpie/tfpie.tex
progress on 4

@@
 \usepackage{xspace}
+
+\def\htdp{\emph{How to Design Programs}\xspace}
+\def\htdc{\emph{How to Design Classes}\xspace}
+\def\profj{{ProfessorJ}\xspace}
 
@@
 
-Could work with any fashion, and should really lead into many
-fashionable languages.  The purpose of showing many is to demonstrate
-that the same principles apply in each setting.  Each language can also
-expose students to various mechanisms.
+The transition from custom teaching languages to a professional
+language takes place about half-way through the course.  At this
+point, students have already seen many of the essentials of
+object-oriented programming. In particular: object, classes, fields
+and methods, dynamic dispatch, inheritance, and overriding.
+
+From this point, almost any language that students might use in future
+co-op positions and courses would be an appropriate follow-up.  Our
+course transitions to Java, but C\#, Python, Ruby, Eiffel, or
+JavaScript would all work naturally.  The key lesson of the transition
+is that the fundamental principles underlying object-oriented
+programming remain the same between languages, and that learning a new
+language is primarily a matter of mapping these concepts to specific
+constructs in the new language.  Of course, particular languages also
+use unique specific mechanisms which need to be taught to use the
+language effectively, but these are rarely as vital as the
+cross-language principles.
+
+\subsection{Functional Java}
+
+The transition begins with replicating the object-oriented style of
+our teaching languages in Java.  In particular, we do not introduce
+mutation, for loops, or mutable data structures such as arrays or
+\texttt{ArrayList}s until later in the semester.  Instead, students
+design data representations using classes, with \texttt{interface}s
+representing unions of data.  Additionally, we avoid mention of the
+distinction between primitive and other values in Java, which is made
+easier by not using standard libraries early.  An example of this
+style of programming is presented in figure~\ref{fig:java}, repeating
+the binary tree sum from the previous section.
 
@@
 \caption{Binary tree sum in How to Design Classes}
+\label{fig:java}
 \end{figure}
 
-\subsection{Java}
+Comparing this figure to the previous example illustrates a number of the
+differences that students are exposed to upon transition to Java.
 
-\begin{itemize}
-\item Types
-\item Compiler
-\item ``Industrial'' features (collections library, ...)
-\end{itemize}
+\begin{enumerate}
+\item Explicit representation of unions and interfaces in the
+  language.  Previously, interfaces were simply described in stylized
+  comments, following the \htdp approach.
+\item Types are now specified as part of the program, and are now
+  enforced.
+\item Java syntax is substantially different and more verbose. For
+  example, constructors must be defined explicitly.
+\item The testing environment is somewhat different, and requires
+  additional boilerplate, although we are
+  able to use the JavaLib framework~\cite{local:java-world} to support simple testing by
+  structural equality.
+\end{enumerate}
 
-\subsection{Unifying functions and objects}
+\noindent
+Of course, there are other differences which cannot be seen from a
+code snippet.
+
+\begin{enumerate}
+  \setcounter{enumi}{4}
+\item Students must use a new development environment and compiler.
+  In class, we primarily develop in a text editor and run the Java
+  compiler at the command line.  In labs and on homeworks, students
+  typically use the Eclipse IDE.
+\item Installing and configuring libraries is now required.  Because
+  we use a custom library for testing, students must cope with library
+  installation and class paths on the first day.
+\end{enumerate}
+
+Of course, all but the first two of these changes are unrelated to the
+fundamental lessons we hope to teach---the rest merely present
+additional hurdles for students.
+
+\subsection{Traditional Java}
+
+Thanks to the the preparation in the first half of the course, we can
+cover functional OO programming in Java in a just a few lectures.  We
+then increase the subset of the language we use to encompass mutation,
+loops, and mutable data structures. We present \texttt{ArrayList}s,
+followed briefly by arrays. Students use, and then implement, hash
+tables as well as other mutable and immutable data structures.
+Conventional input and output are treated only very briefly, as we
+focus instead of both fundamentals and exercises making use of real
+APIs such as hashing functions or Twitter posting.  Finally, while,
+for, and for-each loops are presented, following the methodology of
+\htdc which connects loops to stylized use of recursive functions with
+accumulators, a technique the students now have two semsesters of
+practice with.
+
+\subsection{Beyond Traditional Java}
 
@@
 loop.
-
-\def\htdp{\emph{How to Design Programs}\xspace}
-\def\htdc{\emph{How to Design Classes}\xspace}
-\def\profj{{ProfessorJ}\xspace}
 

### 2013-04-19 tfpie/tfpie.tex
something for all of 4

@@
 
+Finally, at the end of the course, we are able to build on the two
+major segments to examine less-well-explored topics in object-oriented
+programming.  Typically, we cover the basics of implementing OO
+programming in a functional language, advanced OO techniques such as
+mixins and prototypes, and a new OO language such as Ruby or
+JavaScript.  Additionally, we emphasize the ability to embed
+functional programming in an OO context, using techniques such as the
+command pattern and the visitor patterns.  Again, the key message is
+the transferability of concepts across languages.
+
 \section{Related work}

### 2013-04-19 tfpie/tfpie.tex
abstract

@@
 We propose a bridge between functional and object-oriented programming
-in the first-year curriculum.
+in the first-year curriculum.  Traditionally, curricula that begin
+with functional programming transition to a professional, usually
+object-oriented, language in the second course.  This transition poses
+numerous obstacles to students, and often results in confusing the
+details of development environments, syntax, and libraries with the
+fundamentals that the course focuses on.  Our proposal instead begins
+the second course with a sequence of custom teaching languages which
+minimize the transition from the first course, and allow students to
+focus on the core ideas.  We then transition to Java half-way through
+the course, at which point students have a strong command of the basic
+ideas.  We have 3 years of experience with this course, and it has
+produced notable success.
 \end{abstract}

### 2013-04-21 tfpie/tfpie.tex
More on world.

@@
 
+The key difference between these two programs is that the functional
+program uses the {\tt 2htdp/universe} library, which provides a {\tt
+  big-bang} form that consumes the initial state of the world and has
+a declarative form of associating event-handler functions.  The
+object-oriented program uses an alternative library developed for the
+class language: {\tt class/universe}.  It also provides a {\tt
+  big-bang} form but it consumes a single argument, the initial state
+of the world represented as an object.  Event handlers are just
+methods of this object.
+
+The program on the left is the first program of the first semester,
+while the one on the right is the first program of the second
+semester.  Our approach is able to make the conceptual connection
+between functional and object-oriented programming quite clear while
+appealing the familiar event-driven interactive programs developed
+throughout the year.
+
+The move to object-oriented style immediately and naturally leads to
+design principles that are enabled by organizing programs around
+objects.  For example, the state-pattern becomes useful almost
+immediately.  The programs in figure~\ref{fig:world} animate a rocket
+(rendered as a circle in this example) taking off.  An illustrative
+follow-up exercise is to animate a rocket that \emph{lands}.  The
+natural design is to have two variants for states of the rocket: one
+for descending rockets, one for landed rockets.  While in the
+functional approach it is easy to use the state-pattern for the
+\emph{data} representing a rocket, it is more difficult to have states
+of \emph{behavior}.  Of course in the object approach, states of
+behavior are just as natural as data.  Therefore it is straightforward
+to design programs with easy to observe invariants such as ``a landed
+rocket never changes position.''  In the functional approach, even such
+simple properties are more involved to establish.
+
 \begin{figure}

### 2013-04-21 tfpie/tfpie.tex
something submittable

@@
 
-No SE tools/mechanism built into the language.
+Our introduction to object-oriented programming is built on a series
+of ``language levels'', each of which introduces additional features,
+adding complexity to the programming model and expressiveness to the
+programs.  Each language is  \texttt{class/}\textit{N}
+for some \textit{N}, with features appearing in the following order.
+\begin{enumerate}
+  \setcounter{enumi}{-1}
+\item Classes and Objects
+\item Abbreviated notation for method calls
+\item Super classes
+\item Overriding
+\item Constructors
+\end{enumerate}
 
-\subsection{Universe}
+Several commonalities run through all of these languages.  First, they
+are all purely functional; we do not introduce imperative I/O or
+side-effects until after transitioning to Java in the second half of
+the course.  Second, they all are a super set of the
+\emph{Intermediate Student} language from \htdp, meaning that they
+support higher-order functional programming and lists.
 
-\subsection{Programming to interfaces}
+One key principle that we adhere to in the design of the language
+levels is that no features of the language are added purely to support
+``software engineering'' concerns such as specification mechanisms.
+Not only does that language not support declaring types or contracts,
+but interfaces are described purely in comments.
+
+This is not to say that interfaces and contracts are optional; in
+fact, they are mandatory.  But the focus of the first part of the
+course is on the fundamentals of object-orientation.  Teaching the use
+of software engineering tools such as type systems, while vital, is a
+topic which we defer to the second half of the course when we
+transition to Java.
+
+We made this decision after experience in which students were confused
+about the relationship between explicit interface specifications, type
+systems, and the informal data definitions and contracts which
+students are required to write for all methods.  After removing
+interfaces from the language and making them purely a specification
+construct, this confusion disappeared.
+
+%% Fixme: add these sections back later
+
+%\subsection{Universe}
+
+%\subsection{Programming to interfaces}
 

### 2013-04-21 tfpie/tfpie.tex
spellcheck

@@
 example, figure~\ref{fig:tree} defines binary trees of numbers
-(an archetypical recursive union data definition)
+(an archetypal recursive union data definition)
 using the BSL language and the Class language.
@@
 
-The structure of this data definition is analagous to the approach of
+The structure of this data definition is analogous to the approach of
 the previous semester but this example brings to light an important
@@
 \htdc which connects loops to stylized use of recursive functions with
-accumulators, a technique the students now have two semsesters of
+accumulators, a technique the students now have two semesters of
 practice with.
@@
 issues for students.  On the first day of a \htdc-based course,
-students see object-orientation, a new programming paridigm; Java, a
+students see object-orientation, a new programming paradigm; Java, a
 new language with new syntax, and static types, a crucial but
@@
 algorithmic and data structure topics.  Continuing to use World-style
-programming in motivating examples might be valuable for continutity
+programming in motivating examples might be valuable for continuity
 between the two halves of the course.
 
-\subsection*{Acknowledgements}
+\subsection*{Acknowledgments}
 

### 2013-06-18 tfpie/tfpie.tex
fix some bugs DVH found

@@
 
-class Leaf {
+class Leaf implements Tree {
     Integer v;
@@
 
-class Node {
+class Node implements Tree {
     Tree left; Integer v; Tree right;
@@
 Finally, the Java portion of the class does not continue the use of
-``World''-style interactive graphical
-programming. \citet{local:java-world}.  Instead, our course focuses on
-coverage of standard Java libraries, as well as introductory
-algorithmic and data structure topics.  Continuing to use World-style
-programming in motivating examples might be valuable for continuity
-between the two halves of the course.
+``World''-style interactive graphical programming, although a version
+of the ``World'' library has been developed for Java
+\cite{local:java-world}.  Instead, our course focuses on coverage of
+standard Java libraries, as well as introductory algorithmic and data
+structure topics.  Continuing to use World-style programming in
+motivating examples might be valuable for continuity between the two
+halves of the course.
 

### 2013-06-19 tfpie/tfpie.tex
A few edits.

@@
 language takes place about half-way through the course.  At this
-point, students have already seen many of the essentials of
-object-oriented programming. In particular: object, classes, fields
-and methods, dynamic dispatch, inheritance, and overriding.
+point, students already have experience with many of the essential
+concepts of object-oriented programming. In particular: objects,
+classes, fields and methods, dynamic dispatch, inheritance, and
+overriding.
 
-From this point, almost any language that students might use in future
-co-op positions and courses would be an appropriate follow-up.  Our
-course transitions to Java, but C\#, Python, Ruby, Eiffel, or
-JavaScript would all work naturally.  The key lesson of the transition
-is that the fundamental principles underlying object-oriented
-programming remain the same between languages, and that learning a new
-language is primarily a matter of mapping these concepts to specific
-constructs in the new language.  Of course, particular languages also
-use unique specific mechanisms which need to be taught to use the
-language effectively, but these are rarely as vital as the
-cross-language principles.
+From this point, almost any language that students might encounter in
+future co-op positions, summer internships, or upper-level courses
+would be an appropriate follow-up.  Our course transitions to Java,
+but C\#, Python, Ruby, Eiffel, or JavaScript would all work naturally.
+The key lesson of the transition is that the fundamental principles
+underlying object-oriented programming remain the same between
+languages, and that learning a new language is primarily a matter of
+mapping these concepts to specific constructs in the new language.  Of
+course, particular languages also use unique specific mechanisms which
+need to be taught to use the language effectively, but these are
+rarely as vital as the cross-language principles.
 
@@
   comments, following the \htdp approach.
-\item Types are now specified as part of the program, and are now
+\item Types are now specified as part of the program and are (statically)
   enforced.
@@
   additional boilerplate, although we are
-  able to use the JavaLib framework~\cite{local:java-world} to support simple testing by
+  able to use the JavaLib framework~\cite{local:java-world} to support testing with
   structural equality.
@@
 \noindent
-Of course, there are other differences which cannot be seen from a
-code snippet.
+There are other differences which cannot be seen from a code snippet.
 
@@
 
-Of course, all but the first two of these changes are unrelated to the
+All but the first two of these changes are unrelated to the
 fundamental lessons we hope to teach---the rest merely present

### 2013-06-19 tfpie/tfpie.tex
End of my edits.

@@
 \end{verbatim}
-\caption{Binary tree sum in How to Design Classes}
+\caption{Binary tree sum in the style of \emph{How to Design Classes}}
 \label{fig:java}
@@
 then increase the subset of the language we use to encompass mutation,
-loops, and mutable data structures. We present \texttt{ArrayList}s,
-followed briefly by arrays. Students use, and then implement, hash
-tables as well as other mutable and immutable data structures.
-Conventional input and output are treated only very briefly, as we
-focus instead of both fundamentals and exercises making use of real
-APIs such as hashing functions or Twitter posting.  Finally, while,
-for, and for-each loops are presented, following the methodology of
-\htdc which connects loops to stylized use of recursive functions with
-accumulators, a technique the students now have two semesters of
-practice with.
+loops, and mutable data structures and introduce the underlying design
+principles. We present \texttt{ArrayList}s, followed briefly by
+arrays. Students use, and then implement, hash tables as well as other
+mutable and immutable data structures.  Conventional input and output
+are treated only very briefly, as we focus instead on both
+fundamentals and exercises making use of real APIs such as hashing
+functions or Twitter posting.  Finally, \texttt{while}, \texttt{for},
+and for-each loops are presented, following the methodology of \htdc
+which connects loops to stylized use of recursive functions with
+accumulators, a technique the students now have used for two semesters.
 
@@
 major segments to examine less-well-explored topics in object-oriented
-programming.  Typically, we cover the basics of implementing OO
-programming in a functional language, advanced OO techniques such as
-mixins and prototypes, and a new OO language such as Ruby or
-JavaScript.  Additionally, we emphasize the ability to embed
-functional programming in an OO context, using techniques such as the
-command pattern and the visitor patterns.  Again, the key message is
-the transferability of concepts across languages.
+programming.  Typically, we cover the basics of representing objects
+in a functional language, advanced OO techniques such as mixins and
+prototypes, and a new OO language such as Ruby or JavaScript.
+Additionally, we emphasize the ability to embed functional programming
+in an OO context, using techniques such as the command pattern and the
+visitor patterns.  Again, the key message is the transferability of
+concepts across languages.
 
@@
 history, with Abelson and Sussman's \emph{Structure and Interpretation
-of Computer Programs}~\cite{dvanhorn:sicp} a prominent example.  Our
-work follows in the tradition of the \emph{TeachScheme!} project, now
-\emph{Program by Design} (PbD), which emphasizes a systematic
-approach to program construction.
+  of Computer Programs}~\cite{dvanhorn:sicp} being a prominent
+example.  Our work follows in the tradition of the \emph{Program by
+  Design} (PbD)
+project\footnote{\url{http://www.programbydesign.org/}} (previously
+known as the \emph{TeachScheme!} project), which emphasizes a
+systematic approach to program construction.
 
@@
 
-\subsection{TeachScheme! and ProfessorJ}
+\subsection{Program by Design and ProfessorJ}
 
-The TeachScheme! project initially focused only on the first course,
+The Program by Design project initially focused only on the first course,
 with the second course typically taught in Java in
@@
 Java's concepts could be introduced more easily in a second course
-than a first''.
+than a first.''
 
-With these lessons in mind, the TeachScheme! project set out to apply
+With these lessons in mind, the PbD project set out to apply
 the lessons of teaching languages and IDE support to Java, as well as
@@
 
-ProfessorJ and How to Design Classes maintain many of the excellent
-ideas of the first course.  In particular, ProfessorJ brings language
-levels to Java, in an attempt to smooth the transition for students
-from the first course and provide more helpful feedback.  ProfessorJ
-is also embedded in the DrRacket IDE, increasing familiarity for the
-students and supporting tools such as an interactive read-eval-print
-loop.
+ProfessorJ and \emph{How to Design Classes} maintain many of the ideas
+of the first course.  In particular, ProfessorJ brings language levels
+to Java, in an attempt to smooth the transition for students from the
+first course and provide more helpful feedback.  ProfessorJ is also
+embedded in the DrRacket IDE, increasing familiarity for the students
+and supporting tools such as an interactive read-eval-print loop.
 
@@
 students see object-orientation, a new programming paradigm; Java, a
-new language with new syntax, and static types, a crucial but
+new language with new syntax, and a static type system, a crucial but
 orthogonal concept.
@@
 
-Finally, committing to Java on the first day, regardless of the
-environment provided to students, has significant limitations.  First,
-the syntactic and semantic heaviness of Java is a burden for beginning
+Committing to Java on the first day, regardless of the environment
+provided to students, has significant limitations.  First, the
+syntactic and semantic heaviness of Java is a burden for beginning
 students, and discourages interactive and experimental programming.
-The very first chapter of \htdc discusses the fixed size of Java integers,
-a topic avoided entirely in the first course.
-Second, by committing to a particular industrial-strength language, it
-closes off possibilities in the curriculum.  Third, it commits
-entirely to the new paradigm, making it more difficult for students to
-compare the approaches.
+The very first chapter of \htdc discusses the fixed size of Java
+integers, a topic avoided entirely in the first course.  Second, by
+committing to a particular industrial-strength language, it closes off
+possibilities in the curriculum.  Third, it commits entirely to the
+new paradigm, making it more difficult for students to compare the
+approaches.
 
@@
 
-Of course, not every curriculum that begins with \htdp transitions to
-Java after the first course.  \citet{dvanhorn:Ragde2008Chilling}
-describes a second course that includes both more advanced work in
-Scheme beyond teaching-oriented languages as well as low-level
-programming in C, taught to computer science majors at University of
-Waterloo.  Radge's course intentionally does not use student-oriented
-languages, although the recently-developed C0 language~\cite{local:c0}
-could provide this. Other discussions of functional programming in the
-first year~\cite{dvanhorn:Chakravarty2004Risks} do not discuss the
-problems of transition.
+Not every curriculum that begins with \htdp transitions to Java after
+the first course.  \citet{dvanhorn:Ragde2008Chilling} describes a
+second course that includes both more advanced work in Scheme beyond
+teaching-oriented languages as well as low-level programming in C,
+taught to computer science majors at University of Waterloo.  Radge's
+course intentionally does not use student-oriented languages, although
+the recently-developed C0 language~\cite{local:c0} could provide such
+a language. Other discussions of functional programming in the first
+year~\cite{dvanhorn:Chakravarty2004Risks} do not discuss the problems
+of transition.
 
@@
 
-The course has also provided a vantage point to introduce topics that
-will be taken up later in the curriculum.  We present types,
-contracts, invariants, and properties of functions, all of which tie
-into both the concurrent course on logic and computation, as well as
-later classes on formal methods.  The emphasis on
-representation-independence and interfaces  both ties into later
-classes on software engineering, as well as preparing the ground for
-algorithms and data structures courses.  Finally, the use of
-interactive and distributed systems connects to later courses on
-operating systems and networks.
+The course has provided a vantage point to introduce topics that will
+be taken up later in the curriculum.  We present types, contracts,
+invariants, and properties of functions, all of which tie into both
+the concurrent course on logic and computation, as well as later
+classes on formal methods.  The emphasis on
+representation-independence and interfaces both tie into later classes
+on software engineering, as well as preparing students for algorithms
+and data structures courses.  Finally, the use of interactive and
+distributed systems connects to later courses on operating systems and
+networks.
 

### 2013-06-19 tfpie/tfpie.tex
Some of Sam's edits.

@@
 
-      ;; Posn -> Number
+      ;; dist : Posn -> Number
       ;; Distance between this posn and that posn
@@
 
-      ;; -> Number
+      ;; dist-origin : -> Number
       ;; Distance of this posn from the origin
@@
 This class definition defines a new class of values which are {\tt
-  posn} objects.  Such objects are comprised of two numeric values
-and understand the messages {\tt x}, {\tt y}, {\tt dist}, and {\tt
+  posn} objects.  Such objects are comprised of two numeric values and
+understand the messages {\tt x}, {\tt y}, {\tt dist}, and {\tt
   dist-origin}.  Unit tests have been included with each method
 definition, following the principles of the design recipe studied in
-the first semester.  In fact, the {\tt check-expect} mechanism works
-exactly as it did before.
+the first semester.  Although {\tt check-expect} forms can appear
+within class definitions, they are lifted to the top-level when a
+program is run.
 
 Methods can be defined to consume any number of arguments, but they
-are implicitly parameterized over {\tt this}, the object that was sent
+are implicitly parameterized over {\tt this}, the object that received
 the message.
@@
 difference with the functional approach.  The method for computing the
-sum of the empty tree is defined in the {\tt leaf} class, while the
+sum of a leaf is defined in the {\tt leaf} class, while the
 method for computing the sum of a node is in the {\tt node} class.
@@
 leaf---instead, the object itself takes care of computing the sum
-based on the {\tt sum} method it contains.  This shift in perspective
+based on its own {\tt sum} method.  This shift in perspective
 is at the core of object-orientation: objects contain their own
@@
 
-Programming in the first semester is often oriented around interactive
-event-driven video games.  The basic design of a video game involves
-defining a data representation for states of the game and functions
-for transitioning between states based on events such as clock ticks,
-keyboard input, or mouse events.  The design of a game thus involves
-the design of data and functions on that data; in other words, the
-game involves the design of objects.  We therefore continue in the
-second semester with the use of programming video games but supplement
-the course with a library for doing so in an object-oriented style.
-Figure~\ref{fig:world} gives an example written in both the functional
-style and object-oriented style.
+At Northeastern, Programming in the first semester is often oriented
+around interactive event-driven video games.  The basic design of a
+video game involves defining a data representation for states of the
+game and functions for transitioning between states based on events
+such as clock ticks, keyboard input, or mouse events.  The design of a
+game thus involves the design of data and operations on that data; in
+other words, the game involves the design of objects.  We therefore
+continue in the second semester with the use of programming video
+games but supplement the course with a library for doing so in an
+object-oriented style.  Figure~\ref{fig:world} gives an example
+written in both the functional style and object-oriented style.
 
@@
   big-bang} form that consumes the initial state of the world and has
-a declarative form of associating event-handler functions.  The
-object-oriented program uses an alternative library developed for the
-class language: {\tt class/universe}.  It also provides a {\tt
-  big-bang} form but it consumes a single argument, the initial state
-of the world represented as an object.  Event handlers are just
-methods of this object.
+a declarative form of associating event-handler functions, while the
+object-oriented program uses an alternative library developed for our
+class: {\tt class/universe}.  It also provides a {\tt big-bang} form
+but it consumes a single argument, the initial state of the world
+represented as an object.  Event handlers are simply methods of this
+object.
 
@@
 between functional and object-oriented programming quite clear while
-appealing the familiar event-driven interactive programs developed
+appealing to the familiar event-driven interactive programs developed
 throughout the year.
@@
 The move to object-oriented style immediately and naturally leads to
-design principles that are enabled by organizing programs around
-objects.  For example, the state-pattern becomes useful almost
+designs that are enabled by organizing programs around objects.  For
+example, the state pattern~\cite{samth:GOF} becomes useful almost
 immediately.  The programs in figure~\ref{fig:world} animate a rocket
@@
 \emph{data} representing a rocket, it is more difficult to have states
-of \emph{behavior}.  Of course in the object approach, states of
-behavior are just as natural as data.  Therefore it is straightforward
+of \emph{behavior}.  The typical solution adds conditionals to all of
+the event handlers.  In the object-oriented approach, states of
+behavior are just as natural as data.  It is therefore straightforward
 to design programs with easy to observe invariants such as ``a landed
-rocket never changes position.''  In the functional approach, even such
-simple properties are more involved to establish.
+rocket never changes position.''  In the functional approach, even
+such simple properties are more involved to establish.
 
@@
 \bibliographystyle{plainnat}
-\bibliography{dvh-bibliography,local}
+\bibliography{dvh-bibliography,local,sth-bibliography}
 \end{document}

### 2013-06-19 tfpie/tfpie.tex
up

@@
 object-oriented, language in the second course.  This transition poses
-numerous obstacles to students, and often results in confusing the
+ obstacles for students, and often results in confusing the
 details of development environments, syntax, and libraries with the
-fundamentals that the course focuses on.  Our proposal instead begins
+fundamentals of OO programming 
+that the course focuses on.  Our proposal instead begins
 the second course with a sequence of custom teaching languages which
 minimize the transition from the first course, and allow students to
-focus on the core ideas.  We then transition to Java half-way through
+focus on core ideas.  We then transition to Java half-way through
 the course, at which point students have a strong command of the basic
-ideas.  We have 3 years of experience with this course, and it has
-produced notable success.
+principles.  We have 3 years of experience with this course, with
+ notable success.
 \end{abstract}
@@
 in an industrial object-oriented programming language by the end of
-the students' first year.  Past approaches to achieve this aim have
-included teaching an industrial language, such as Java, starting on
-the first day.  Others have taken a more indirect route by teaching
-functional programming in the first semester, followed by a second
-semester that starts in Java.  The latter approach is an improvement
-over the first, but both suffer serious flaws.
+the students' first year.  The most common approache to achieve this
+aim is to teach an industrial language in the first course, often
+Java, starting on the first day.  Other curricula take a more indirect
+route by teaching functional programming in the first semester,
+followed by a second semester in Java.  The latter approach is an
+improvement over the first, as pointed out by numerous
+observers~\cite{dvanhorn:Felleisen2004Structure, local:spolsky}, but
+both suffer serious flaws.
 
-As an example, Northeastern University teaches functional programming
+As an example, Northeastern University teaches introductory programming
 in the first semester using \emph{How to Design
@@
   Classes}~\cite{local:htdc} in the second semester.  This sequence
-was designed to provide a smooth path for incoming students assumed
-only to have a competence in high-school level algebra to reach
+was designed to provide a smooth path for incoming students with
+ a competence in high-school level algebra to reach
 proficiency in Java by the end of their first year
@@
 The problem is that the first semester ends with an advanced
-pedagogical functional language and the second semester starts with a
-beginning pedagogical subset of Java.  Despite this slimmed-down
+pedagogical functional language and the second semester starts with
+Java, although it focuses on a side-effect free subset.  Despite this focused
 subset, this transition is too abrupt to meaningfully bridge the gap
-between functional and object-oriented programming, because there are
-several other significant transitions happening in concert such as the
-transitions:
+between functional and object-oriented programming, because
+several other significant transitions happen in concert:
 \begin{itemize}
@@
 
-\item from a pedagogical programming environment to a professional
+\item from a pedagogical programming environment (DrRacket) to a professional
 programming environment (Eclipse),
@@
   mathematical objects to a language with numeric values corresponding
-  to common machine representations,
+  to common machine representations,\footnote{While this may seem like
+    a minor point, details of numeric representation can crop up
+    quickly in a Java-based course.}
 
@@
 program, it favors students with prior exposure to the particular
-tools, it inhibits students from experimenting by relying upon their
-past skills, and it creates the false impression that courses are
-discrete units of instruction that can be discarded after successful
-completion rather than being part of a continuous and cumulative
-educational experience.
+tools (a set that is demographically skewed), it inhibits students
+from experimenting by allowing them to rely upon past skills, and it creates
+the false impression that courses are discrete units of instruction
+that can be discarded after successful completion rather than being
+part of a continuous and cumulative educational experience.
 
 We contribute an alternative approach to the second semester that
-overcomes these problems and provides a gradual transition from
-functional programming to object-oriented programming.  Our approach
-is able to start the second semester by introducing \emph{only} the
+overcomes these problems and provides a gradual introduction 
+to object-oriented programming.  Our approach
+starts the second semester by introducing \emph{only} the
 concept of programming with objects, while all other aspects of course
 remain where they were left off in the previous semester.  This
-approach allows the other concepts to be introduced at the point at
+ allows other concepts to be introduced at the point at
 which they are relevant and motivated.  Despite this more gradual
-approach, the course accomplishes the goal of teaching industrial
-language proficiency by the end of the semester, covering a super-set
-of the concepts and topics covered in the \emph{How to Design Classes}
+approach, the course accomplishes the goal of reaching industrial
+language competence by the end of the semester, covering a super-set
+of the concepts and topics covered in the \emph{How to Design Classes}-based
 course.
@@
   Design Programs} and the context and constraints involved in the
-first-year at Northeastern.  Section~\ref{sec:shift} describes our
+first year at Northeastern.  Section~\ref{sec:shift} describes our
 approach to the second semester, which starts with a small shift in
@@
 object oriented programming.  Section~\ref{sec:industrial} describes
-the path toward programming in an industrial object-oriented
+the path from our pedagogical languages to an industrial object-oriented
 programming language.  Section~\ref{sec:related-work} discusses the

### 2013-06-19 tfpie/tfpie.tex
section 2

@@
 At Northeastern, the College of Computer \& Information Science (CCIS)
-requires a four course introductory sequence in the first year.  The
+requires a four course introductory sequence for Computer Science majors in the first year.  The
 first semester features both a course on discrete mathematics and
@@
 
-Subsequent to the first year, students take a wide variety of
+After the first year, students take a wide variety of
 follow-up courses, ranging from a required course in ``Object-oriented
@@
 some software development, there is almost no uniformity beyond that.
+In particular, there is wide variation in the languages students use
+while on co-op.
 
@@
 \noindent
-We reuse all of these elements in our approach.
+We deliberately reuse all of these elements in our approach.
 

### 2013-06-19 tfpie/tfpie.tex
my part of sec 3

@@
 \end{verbatim}
+This is similar to the {\tt define-struct} form of the first semester,
+used as follows:
+\begin{verbatim}
+    (define-struct posn (x y))
+\end{verbatim}
+
 An \emph{object} is a value that is a member of this class of data,
@@
 The {\tt send} notation is simple, but syntactically heavy.  Once
-students are comfortable with the {\tt send} form, we introduce a
-shorthand to make it more convenient by writing {\tt (x . m)} for {\tt
+students are comfortable with {\tt send}, we introduce 
+shorthand to make it more convenient, writing {\tt (x . m)} for {\tt
   (send x m)}.  The \emph{dot notation} can be nested, so {\tt (x . m
@@
 convenient shorthand that would have been confusing to start with
-follows the approach of introducing {\tt cons} and later {\tt list}
+follows the approach of first introducing {\tt cons} and then later {\tt list}
 and {\tt quote} in the first semester.)
@@
 class definition.  Here is a more complete development of the {\tt posn} class
-that includes a couple of methods:
+with two methods:
 \begin{verbatim}

### 2013-06-19 tfpie/tfpie.tex
Tweaks to abstract.

@@
 object-oriented, language in the second course.  This transition poses
- obstacles for students, and often results in confusing the
-details of development environments, syntax, and libraries with the
-fundamentals of OO programming 
-that the course focuses on.  Our proposal instead begins
-the second course with a sequence of custom teaching languages which
-minimize the transition from the first course, and allow students to
-focus on core ideas.  We then transition to Java half-way through
-the course, at which point students have a strong command of the basic
-principles.  We have 3 years of experience with this course, with
- notable success.
+obstacles for students, and often results in confusing the details of
+development environments, syntax, and libraries with the fundamentals
+of OO programming that the course should focus on.  Instead, we
+propose to begin the second course with a sequence of custom teaching
+languages which minimize the transition from the first course, and
+allow students to focus on core ideas.  After working through the
+sequence of pedagogical languages, we then transition to Java, at
+which point students have a strong command of the basic principles.
+We have 3 years of experience with this course, with notable success.
 \end{abstract}

### 2013-09-29 tfpie/response.txt
progress

----------------------- REVIEW 1 ---------------------

I use the term "work-in-progress," because the authors have
side-stepped the methodology used to introduce types and the
transition from their tower of student languages to an industrial
language. Granted, this is likely something the authors are still
polishing. Nonetheless, it is important to provide a vision for how to
do these things and to outline what they have done highlighting both
succeses and failures. In this manner, interested readers may join
their efforts by reproducing the authors' methodology in their schools
using the languages developed by the authors. Let's face it,
introducing types is necessary to make the move to an
industrial-strength language and without providing a vision for it
many may balk. I suspect that the transition builds on the informal
contracts students are familiar with. Thus, adding this vision should
not be to difficult. Outlining this vision may obviate doubts readers
may have about making the transition to an industrial-strength
language. This is the primary reason that I am unable to put this
article in the strong accept column.

RESPONSE: We have added a paragraph in section 4.2 on this subject.

Page 2

Spell out an example of the details of numeric representation that
crop up quickly.

RESPONSE: We have added an example.

Implicit in the discussion is the assumption that reaching industrial
language competence by the end of the semester is a desirable
goal. The authors should address this head on. An obvious alternative
is emphasize and reinforce OO concepts and leave the industrial
language for latter (one semester?) in the curriculum. I am aware that
at NE students are expected to get a co-op/intership over the
summer. That, however, is a NE pecularity and is not a pedagogical
argument.

RESPONSE: We make explicit in section 2 the aspects of Northeastern
that inform our course.  It is not only co-op, but downstream courses,
that require Java.

Page 3

"sent by using the send keyword, followed by an object, a message
name, and some number of arguments"

        What are these "some number of arguments?"
        They are not part of the examples.

RESPONSE: These particular methods take no arguments, but the `dist`
method, on the same page, takes an argument.

Page 6

Spell out for the reader how students are explained the world
parameter is not needed and, therefore, removed from the methods in
the class definition.

RESPONSE: This is already covered in the discussion of OO
fundamentals.

In the class version, how does the runtime know which method to use?
Are the names of the handlers restricted to on-tick, on-key, etc.? Do
the methods have to appear in a certain order? Does the mechanism used
to make the association of events with the proper handler pose any
difficulty for the students? One nice feature of the functional
approach is that such associations are made explicitly and this allows
students to name their handlers whatever makes sense to them.

RESPONSE: We made this more explicit in the paper.

Do you mean the first interactive event-driven program?

RESPONSE: No, this is the first program we show.

The programs in Figure 2 do not animate a rocket. Please reformulate
the description of the programs in Figure 2.

RESPONSE: We have revised the program.

Show the reader how you use with students in class the state pattern
to animate a rocket that lands. There is plenty of room within the
20-page limit.

RESPONSE: We added an appendix with this program.

Page 11

Spell out the drawbacks that are alleviated and those that are not
addressed by DrJava and BlueJ. As a reader, I want to clearly see the
contrast of your approach with these.

Page 12

Can you quantify the recruitment/retention impact on female students?
Give the reader a sense of what you mean. It must be significant given
that you mention it!

Minor comments

Overall: Capitalize major words in section/subsection headers (not
just the first word).

Page 3
        "and message are sent" --> "and messages are sent"

Page 5
        "figure 1" --> "Figure 1"

Page 6

        "the structure of this data definition" --> "the structure of
        this class definition"

        "design of a game thus involves" --> "design of a game, thus,
        involves"

        "2htdp/universe" over the margin

        "figure 2" --> "Figure 2"

        "rockets, one" --> "rockets and one"

        "language levels", --> "language levels,"
Page 8
        "figure 3" --> "Figure 3"

----------------------- REVIEW 2 ---------------------
PAPER: 1
TITLE: From Principles to Practice with Class in the First Year
AUTHORS: Sam Tobin-Hochstadt and David Van Horn

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 4 (high)

----------- REVIEW -----------

I enjoyed reading this paper. It describes the approach to teaching
programming at Northeastern University, which begins by teaching
Racket in the first semester, followed by an introduction to
programming with classes in the second semester, initially by
extending Racket with classes, then moving on to a functional subset
of Java, and finally "traditional" Java.  The aim is to introduce
important programming concepts step by step, while still providing
students with enough practical knowledge to participate in an
internship. In particular, it is a nice observation that it is only a
small shift in focus to move from HDTP style to object oriented style
by adding a new language concept which the students already (almost)
understand.

The paper sets out the background and the constraints under which the
course has been designed, and does this well. It describes informally
the experience of both students and teachers on the course, and
suggests that the results have been positive. It is easy to read and
proceeds at a good pace, and I think it deserves to be published in
order to encourage more institutions to consider designing courses in
this way.

Some specific questions/comments/typos:

p1 "approache" => approach

p2 "a set that is demographically skewed" - in what way?  Also, is
this first paragraph written from your own experience of past versions
of the course?

p5 Would "figure 1", "figure 2" etc be better capitalised?

p5 I think this is the first time the names "BSL" and "Class" are
mentioned.  As far as I know the name BSL is taken from HTDP, and
there does seem to be a few places where it is assumed the reader is
familiar with the content of HTDP. Maybe it would be good to give a
brief summary of the HTDP approach and what these languages are
earlier.

p6 Interesting (and somewhat surprising to me) comment just before
section 3.4 that some invariants are easier to establish in the object
oriented style.

p7 Is this the first mention of "Intermediate Student" language?
Again, maybe in section two where the idea of language levels is
mentioned, maybe it would be good to go into a bit more detail for the
benefit of those who haven't read HTDP.

p9 Is it possible to include the Racket version alongside this
program? I found it interesting to compare them, and there does seem
to be enough space on the page for it.

p10 "functional OO programming" reads a bit strangely to me, since it
could mean "functional" in a non-technical sense. Maybe "a functional
subset of ..."?

p12 It's an interesting comment about retaining female students, but
it's not clear how the design of the course might be relevant to
this. Do you have any suggestions why? Could it just be a random
fluctuation?

p13 Some of the capitalisation in the references has gone ('java'
appears a few times for example)

----------------------- REVIEW 3 ---------------------
PAPER: 1
TITLE: From Principles to Practice with Class in the First Year
AUTHORS: Sam Tobin-Hochstadt and David Van Horn

OVERALL EVALUATION: 1 (weak accept)
REVIEWER'S CONFIDENCE: 4 (high)

----------- REVIEW -----------

In this paper the authors present their approach to smoothen the
transition during the first year programming course from semester 1:
FP to semester 2: OO programming. The key contribution of their work
is to demonstrate how a pure FP approach can be extended step by step
with OO concepts. In the end students acquire the same (and even
higher) level of proficiency of programming OO in Java as comparable
other courses.

The project's success relies on DrRacket, used in semester 1, which
can be extended with the desired features step by step. Halfway during
semester 2, when a sufficient 'critical mass' of OO concepts has been
introduced, the course and the students switch to Java and Eclipse.

The paper is well written. The paper is clearly in scope of TFPIE.

My main motivation for marking the paper as 'weak accept' is that in
my opinion it is not (yet) complete to deserve publication in
EPTCS. My main criticism plus advice for improvement is:

(1) in the introduction, page 1, the bullet-list enumerates
'significant transitions happen in concert'. In the paper we learn how
to overcome bullet 1, but not bullets 2, 3, 4, and 5. On page 8, first
paragraph: "In particular: objects, classes, fields and methods,
dynamic dispatch, inheritance, and overriding." are enumerated as
being essential concepts of object-oriented programming that students
have had experience with in the new set-up. The paper explains this
for the first half ('objects, classes, fields and methods') but not
for the second half ('dynamic dispatch, inheritance, and overriding').

Suggestion for improvement: extend section 3 with material how dynamic
dispatch, inheritance, and overriding is presented to students.

(2) Section 5.1 is too long to my taste: a lot of the criticism has
been discussed in the motivation of the introduction of this paper.

Suggestion for improvement: shorten section 5.1. The extra space can
be used to elaborate on issue (1) above. I don't think this section
requires sub sections, so all text can be organized into a single
section.

(3) In the introduction, it is said that 'an alternative approach to
the second semester [that] overcomes these problems and provides a
gradual introduction to OO programming'. However, acting as the
devil's advocate, I can argue that this is not true: halfway semester
2 there is still the abrupt switch to both another language, Java,
*and* another tool set, Eclipse. Stated bluntly, the switch is only
postponed instead of solved.

Suggestion for improvement: if issue (1) is resolved, it will become
clearer how the 'switch-gap' is made smaller. Section 4.1 enumerates
the differences of this gap, but does not explain how the students are
taught to jump over them. This should be added in this
section. Sections 4.2 and 4.3 are too brief in my opinion. Are the
topics in these sections taken immediately from 'How to Design
Classes'? If so, please state this. If not, please include material
that illustrates how these topics are taught and how this fits in the
philosophy of the chosen approach.

Minor remarks:
page 1:
"approache" --> "approach"

"numerous observers [9, 23]": only two references, of which [23] is
not scientific is not particularly 'numerous'. Please provide more
scientific references or remove the word 'numerous' (I prefer the
first option)

page 2:
"as follows: in section 2" --> "as follows: section 2"

page 4: the use of ";=> 3" and ";=> 4" in the same teletype font is
confusing because they are not part of the expression, but probably
intended to denote 'evaluates to'? Please use $math$ here.

page 5: "the BSL language and the Class language." these languages
have not been introduced yet, please do.

----------------------- REVIEW 4 ---------------------
PAPER: 1
TITLE: From Principles to Practice with Class in the First Year
AUTHORS: Sam Tobin-Hochstadt and David Van Horn

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 4 (high)

----------- REVIEW -----------

The authors describe a new approach to tackle difficulties in
transitioning from a functional semester one course to an
object-oriented semester two course. Students are reportedly confused
by the circumstantial differences (compilation model / environment /
syntax) which makes grasping the fundamental concepts being taught
harder to grasp. Instead of transitioning on day one of the second
semester to Java in Eclipse (or similar), the authors present an
approach that continues the progression of student languages to
include objects and other OO concepts before transitioning to a "real"
OO environment and language.

The paper is well-written, is well-positioned in the context of the
status quo and balances the pedagogical approach with the material
used. The idea of language levels is properly motivated and the
continuation of this approach beyond the first semester seems - after
reading the paper - the Obvious Choice. What is lacking, however, is a
motivation for the point at which the transition to Java is made. Why
aren't e.g. iteration, mutability & typing introduced in the similar
vein as objects, class-hierarchies, etc.? Was this a choice from
practicality, or is there a reason to introduce them post-transition?

A few specific comments and minor corrections:

 - (Sec.1) "common approache to" --> "common approach to"

 - (Sec.1) "interaction-oriented language" / "batch-oriented language"
   --> it seems the authors are refering to the environments/tooling
   rather than the language. If not, this terminology warrents
   clarification.

 - (Sec.1) "aspects of course remain" --> "aspects of *the* course
   remain"

 - (Sec.2) "while most students do some software development, there is
   almost no uniformity beyond that" & "in which they will use
   professional-grade languages and tools which will almost certainly
   be object-oriented" are, at least, paradoxical (if not
   contradictory). This conflict in the motivation should be addressed
   for the final version of the paper.

 - (Sec.3.1) "and message are sent" --> "and message*s* are sent"

 - (Sec.4.2) "Thanks to the the preparation" --> omit one "the"

- (Sec.6) "The class has also had notable success in the recruitment
  and retention of female students, as compared to the other versions
  of the second-semester course." --> I find this a rather hefty
  statement. It would - in my mind - at least warrent further
  research. If this consistently holds true, there are valuable
  insights to be gained about pedagogical gender-script in the field
  of CS. The statement by itself seems a bit hefty to be included
  without at least some concrete data and/or qualitative evaluation.

### 2013-09-29 tfpie/tfpie.tex
progress

@@
 \usepackage{xspace}
+\usepackage{alltt}
 
@@
     a minor point, details of numeric representation can crop up
-    quickly in a Java-based course.}
+    quickly in a Java-based course---for example, $1/3$ cannot be
+    represented by any Java numeric types.}
 
@@
 represented as an object.  Event handlers are simply methods of this
-object.
+object; for example, clock ticks trigger the {\tt on-tick} method.
 
@@
 natural design is to have two variants for states of the rocket: one
-for descending rockets, one for landed rockets.  While in the
-functional approach it is easy to use the state-pattern for the
-\emph{data} representing a rocket, it is more difficult to have states
-of \emph{behavior}.  The typical solution adds conditionals to all of
-the event handlers.  In the object-oriented approach, states of
-behavior are just as natural as data.  It is therefore straightforward
-to design programs with easy to observe invariants such as ``a landed
-rocket never changes position.''  In the functional approach, even
-such simple properties are more involved to establish.
+for descending rockets, one for landed rockets (an example is given in
+appendix~\ref{sec:appendix}).  While in the functional approach it
+is easy to use the state-pattern for the \emph{data} representing a
+rocket, it is more difficult to have states of \emph{behavior}.  The
+typical solution adds conditionals to all of the event handlers.  In
+the object-oriented approach, states of behavior are just as natural
+as data.  It is therefore straightforward to design programs with easy
+to observe invariants such as ``a landed rocket never changes
+position.''  In the functional approach, even such simple properties
+are more involved to establish.
 
@@
 (define (draw w)
-  (circle w "solid" "red"))
+  (place-image 
+    (circle 10 "solid" "red")
+    w 200 (empty-scene 400 400)))
 
@@
   (define (to-draw) 
-    (circle (this . n) "solid" "red"))
+    (place-image 
+     (circle 10 "solid" "red")
+     (this . w) 200 (empty-scene 400 400)))
 
@@
 \bibliography{dvh-bibliography,local,sth-bibliography}
+
+\newpage
+\appendix
+\section{Worlds and the State pattern}
+\label{sec:appendix}
+
+\begin{alltt}
+#lang class/1
+(require 2htdp/image class/universe)
+
+;; A World is one of 
+;; - (new upworld Number)
+;; - (new downworld Number)
+(define-class upworld
+  (fields n)
+
+  ;; on-tick : -> World
+  (define (on-tick)
+    (new world (add1 (this . n))))
+
+  ;; to-draw : -> Image
+  (define (to-draw) 
+    (place-image 
+     (circle 10 "solid" "red")
+     (this . w) 200 (empty-scene 400 400)))
+
+  ;; on-key : KeyEvent -> World
+  (define (on-key k) (new world 10)))
+
+(define-class downworld
+  (fields n)
+
+  ;; on-tick : -> World
+  (define (on-tick)
+    (cond [(zero? (this . n))
+           (new upworld (this . n))]
+          [else
+           (new world (sub1 (this . n)))]))
+
+  ;; to-draw : -> Image
+  (define (to-draw) 
+    (place-image 
+     (circle 10 "solid" "red")
+     (this . w) 200 (empty-scene 400 400)))
+
+  ;; on-key : KeyEvent -> World
+  (define (on-key k) (new world 10)))
+
+(big-bang (new downworld 400))
+\end{alltt}
+
 \end{document}

### 2013-09-29 tfpie/response.txt
review 1

@@
 
+RESPONSE: We now make this explicit.
+
 Page 12
@@
 
-Minor comments
-
-Overall: Capitalize major words in section/subsection headers (not
-just the first word).
-
-Page 3
-        "and message are sent" --> "and messages are sent"
-
-Page 5
-        "figure 1" --> "Figure 1"
-
-Page 6
-
-        "the structure of this data definition" --> "the structure of
-        this class definition"
-
-        "design of a game thus involves" --> "design of a game, thus,
-        involves"
-
-        "2htdp/universe" over the margin
-
-        "figure 2" --> "Figure 2"
-
-        "rockets, one" --> "rockets and one"
-
-        "language levels", --> "language levels,"
-Page 8
-        "figure 3" --> "Figure 3"
+RESPONSE: A quantitative evaluation would be inappropriate in this
+situation. We have clarified the point.
 
 ----------------------- REVIEW 2 ---------------------
-PAPER: 1
-TITLE: From Principles to Practice with Class in the First Year
-AUTHORS: Sam Tobin-Hochstadt and David Van Horn
-
-OVERALL EVALUATION: 2 (accept)
-REVIEWER'S CONFIDENCE: 4 (high)
-
------------ REVIEW -----------
 

### 2013-09-29 tfpie/tfpie.tex
review 1

@@
 structure-like objects understand messages for accessing their fields
-and message are sent by using the {\tt send} keyword, followed by an object,
+and messages are sent by using the {\tt send} keyword, followed by an object,
 a message name, and some number of arguments:
@@
 
-The structure of this data definition is analogous to the approach of
+The structure of this program is analogous to the approach of
 the previous semester but this example brings to light an important
@@
 natural design is to have two variants for states of the rocket: one
-for descending rockets, one for landed rockets (an example is given in
+for descending rockets and one for landed rockets (an example is given in
 appendix~\ref{sec:appendix}).  While in the functional approach it
@@
 %
+For example, both of these systems immediately present students with
+(1) type systems and (2) Java syntax, and (3) do not support the image
+values and exact numeric values that we rely on in our course.
+
 Several teaching-oriented graphics libraries for Java have been
@@
 from what the bulk of Northeastern University computer science majors
-take.  The class has also had notable success in the recruitment and
-retention of female students, as compared to the other versions of the
-second-semester course.
+take.  Anectodally, the class has also had notable success in the
+recruitment and retention of female students, as compared to the other
+versions of the second-semester course. However, the classes are
+sufficiently different as to make a precise comparison impossible.
 

### 2013-09-29 tfpie/tfpie.tex
Blah about types.

@@
   comments, following the \htdp approach.
-\item Types are now specified as part of the program and are (statically)
-  enforced.
+\item Types are now specified as part of the program and are
+  (statically) enforced.  Data definitions and interfaces can be
+  transformed from the stylized comments into {\tt interface}
+  definitions and method signatures annotated with types.  Students
+  are taught the benefits of type systems, which impose syntactic
+  restrictions sufficient to prove objects meet (the structural
+  aspects of) their interface definitions.  Students also encounter
+  the downside of types when they discover the type system cannot
+  always follow valid reasoning about program invariants and may reject
+  perfectly good programs.
 \item Java syntax is substantially different and more verbose. For

### 2013-09-29 tfpie/response.txt
sync w/ david

@@
 
-RESPONSE: We have added a paragraph in section 4.2 on this subject.
+RESPONSE: We have added a paragraph in section 4.1 on this subject.
 
@@
 
-I enjoyed reading this paper. It describes the approach to teaching
-programming at Northeastern University, which begins by teaching
-Racket in the first semester, followed by an introduction to
-programming with classes in the second semester, initially by
-extending Racket with classes, then moving on to a functional subset
-of Java, and finally "traditional" Java.  The aim is to introduce
-important programming concepts step by step, while still providing
-students with enough practical knowledge to participate in an
-internship. In particular, it is a nice observation that it is only a
-small shift in focus to move from HDTP style to object oriented style
-by adding a new language concept which the students already (almost)
-understand.
-
-The paper sets out the background and the constraints under which the
-course has been designed, and does this well. It describes informally
-the experience of both students and teachers on the course, and
-suggests that the results have been positive. It is easy to read and
-proceeds at a good pace, and I think it deserves to be published in
-order to encourage more institutions to consider designing courses in
-this way.
-
-Some specific questions/comments/typos:
-
-p1 "approache" => approach
-
 p2 "a set that is demographically skewed" - in what way?  Also, is
@@
 
-p5 Would "figure 1", "figure 2" etc be better capitalised?
+RESPONSE: We commented on our prior experience.
 
@@
 earlier.
+
+RESPONSE: We clarify what these languages are.
 

### 2013-09-29 tfpie/tfpie.tex
sync w/ david

@@
 in an industrial object-oriented programming language by the end of
-the students' first year.  The most common approache to achieve this
+the students' first year.  The most common approach to achieve this
 aim is to teach an industrial language in the first course, often
@@
 
-This abrupt transition has several negative consequences: the
-principles of object-oriented programming are obscured and
-de-emphasized, struggling with the programming environment is
-frustrating and can cause potentially good students to leave the
-program, it favors students with prior exposure to the particular
-tools (a set that is demographically skewed), it inhibits students
-from experimenting by allowing them to rely upon past skills, and it creates
-the false impression that courses are discrete units of instruction
-that can be discarded after successful completion rather than being
-part of a continuous and cumulative educational experience.
+This abrupt transition has several negative consequences which we have
+experienced first-hand: the principles of object-oriented programming
+are obscured and de-emphasized, struggling with the programming
+environment is frustrating and can cause potentially good students to
+leave the program, it favors students with prior exposure to the
+particular tools (a set that is demographically skewed), it inhibits
+students from experimenting by allowing them to rely upon past skills,
+and it creates the false impression that courses are discrete units of
+instruction that can be discarded after successful completion rather
+than being part of a continuous and cumulative educational experience.
 
@@
 (an archetypal recursive union data definition)
-using the BSL language and the Class language.
+using the Beginning Student language (BSL) as used at the start of the
+first semester, and also using the Class language of our course.
 %% \begin{verbatim}

### 2013-09-29 tfpie/response.txt
review 2

@@
 
-p6 Interesting (and somewhat surprising to me) comment just before
-section 3.4 that some invariants are easier to establish in the object
-oriented style.
-
 p7 Is this the first mention of "Intermediate Student" language?
@@
 
+RESPONSE: We extended section 2 to cover more HTDP background.
+
 p9 Is it possible to include the Racket version alongside this
@@
 
+RESPONSE: The BSL and Class versions of this program are in figure 1.
+
 p10 "functional OO programming" reads a bit strangely to me, since it
@@
 subset of ..."?
+
+RESPONSE: Clarified.
 
@@
 
-p13 Some of the capitalisation in the references has gone ('java'
-appears a few times for example)
+RESPONSE: We clarified the anecdotal nature of this.
 

### 2013-09-29 tfpie/tfpie.tex
review 2

@@
 Thanks to the the preparation in the first half of the course, we can
-cover functional OO programming in Java in a just a few lectures.  We
-then increase the subset of the language we use to encompass mutation,
-loops, and mutable data structures and introduce the underlying design
-principles. We present \texttt{ArrayList}s, followed briefly by
-arrays. Students use, and then implement, hash tables as well as other
-mutable and immutable data structures.  Conventional input and output
-are treated only very briefly, as we focus instead on both
-fundamentals and exercises making use of real APIs such as hashing
-functions or Twitter posting.  Finally, \texttt{while}, \texttt{for},
-and for-each loops are presented, following the methodology of \htdc
-which connects loops to stylized use of recursive functions with
-accumulators, a technique the students now have used for two semesters.
+cover OO programming in a functional subset of Java in a just a few
+lectures.  We then increase the subset of the language we use to
+encompass mutation, loops, and mutable data structures and introduce
+the underlying design principles. We present \texttt{ArrayList}s,
+followed briefly by arrays. Students use, and then implement, hash
+tables as well as other mutable and immutable data structures.
+Conventional input and output are treated only very briefly, as we
+focus instead on both fundamentals and exercises making use of real
+APIs such as hashing functions or Twitter posting.  Finally,
+\texttt{while}, \texttt{for}, and for-each loops are presented,
+following the methodology of \htdc which connects loops to stylized
+use of recursive functions with accumulators, a technique the students
+now have used for two semesters.
 

### 2013-09-29 tfpie/response.txt
review 3

@@
 ----------------------- REVIEW 3 ---------------------
-PAPER: 1
-TITLE: From Principles to Practice with Class in the First Year
-AUTHORS: Sam Tobin-Hochstadt and David Van Horn
-
-OVERALL EVALUATION: 1 (weak accept)
-REVIEWER'S CONFIDENCE: 4 (high)
-
------------ REVIEW -----------
-
-In this paper the authors present their approach to smoothen the
-transition during the first year programming course from semester 1:
-FP to semester 2: OO programming. The key contribution of their work
-is to demonstrate how a pure FP approach can be extended step by step
-with OO concepts. In the end students acquire the same (and even
-higher) level of proficiency of programming OO in Java as comparable
-other courses.
-
-The project's success relies on DrRacket, used in semester 1, which
-can be extended with the desired features step by step. Halfway during
-semester 2, when a sufficient 'critical mass' of OO concepts has been
-introduced, the course and the students switch to Java and Eclipse.
-
-The paper is well written. The paper is clearly in scope of TFPIE.
-
-My main motivation for marking the paper as 'weak accept' is that in
-my opinion it is not (yet) complete to deserve publication in
-EPTCS. My main criticism plus advice for improvement is:
 
@@
 'significant transitions happen in concert'. In the paper we learn how
-to overcome bullet 1, but not bullets 2, 3, 4, and 5. On page 8, first
+to overcome bullet 1, but not bullets 2, 3, 4, and 5.
+
+RESPONSE: Our approach presents OO programming in a context that
+avoids _all 5_ of these significant transitions until after the
+fundamental concepts of OO programming are taught, as we emphasize.
+
+On page 8, first
 paragraph: "In particular: objects, classes, fields and methods,
@@
 
+RESPONSE: We discuss dynamic dispatch extensively in the paper.  This
+is the entire point of section 3.2.  We added a paragraph discussion
+inheritance and overriding.
+
 (2) Section 5.1 is too long to my taste: a lot of the criticism has
@@
 section.
+
+RESPONSE: We feel that this discussion is needed.
 
@@
 
-Suggestion for improvement: if issue (1) is resolved, it will become
-clearer how the 'switch-gap' is made smaller. Section 4.1 enumerates
-the differences of this gap, but does not explain how the students are
-taught to jump over them. This should be added in this
-section. Sections 4.2 and 4.3 are too brief in my opinion. Are the
-topics in these sections taken immediately from 'How to Design
-Classes'? If so, please state this. If not, please include material
-that illustrates how these topics are taught and how this fits in the
-philosophy of the chosen approach.
-
-Minor remarks:
-page 1:
-"approache" --> "approach"
+RESPONSE: We clarify how the transition is less troublesome here.  We
+also describe future work that would further alleviate these
+difficulties.
 
@@
 
-page 2:
-"as follows: in section 2" --> "as follows: section 2"
-
-page 4: the use of ";=> 3" and ";=> 4" in the same teletype font is
-confusing because they are not part of the expression, but probably
-intended to denote 'evaluates to'? Please use $math$ here.
-
-page 5: "the BSL language and the Class language." these languages
-have not been introduced yet, please do.
+RESPONSE: We have added additional references.
 

### 2013-09-29 tfpie/tfpie.tex
review 3

@@
 improvement over the first, as pointed out by numerous
-observers~\cite{dvanhorn:Felleisen2004Structure, local:spolsky}, but
+observers~\cite{dvanhorn:Felleisen2004Structure, local:spolsky,
+  dvanhorn:Chakravarty2004Risks, dvanhorn:Ragde2008Chilling}, but
 both suffer serious flaws.
@@
 \paragraph{Outline} The remainder of this paper is organized as follows:
-in section~\ref{sec:background} provides background on \emph{How to
+ section~\ref{sec:background} provides background on \emph{How to
   Design Programs} and the context and constraints involved in the
@@
 natural design is to have two variants for states of the rocket: one
-for descending rockets and one for landed rockets (an example is given in
-appendix~\ref{sec:appendix}).  While in the functional approach it
+for descending rockets and one for landed rockets (an example is given
+in appendix~\ref{sec:appendix}).  While in the functional approach it
 is easy to use the state-pattern for the \emph{data} representing a
@@
 position.''  In the functional approach, even such simple properties
-are more involved to establish.
+are more involved to establish, because all event handlers must be
+inspected.
+
+This approach also leads naturally to discussion of inheritance.
+Often programs with multiple states wish to share the implementation
+of some methods.  We first show that this can be accomplished at the
+cost of minor boilerplate with delegation, and then show how
+inheritance allows the programmer to avoid duplication and boilerplate
+entirely.  Once inheritance is able to group identical methods,
+overriding is a natural next step when some but not all of the
+implementations are identical across the variants.
 
@@
 fundamental lessons we hope to teach---the rest merely present
-additional hurdles for students.
+additional hurdles for students.  However, at this point in the
+semester, the students are far better equipped to meet these
+challenges.  They are already familiar with objects, classes, and the
+other concepts we have covered.  They are also fully engaged in the
+class, instead of making the transition in the midst of the transition
+between semesters.  Finally, they have now been programming for 50\%
+longer than they had at the start of the semester.
 
@@
 ;; A World is one of 
-;; - (new upworld Number)
+;; - (new landed-world)
 ;; - (new downworld Number)
-(define-class upworld
-  (fields n)
-
-  ;; on-tick : -> World
-  (define (on-tick)
-    (new world (add1 (this . n))))
+(define-class landed-world
 
@@
      (circle 10 "solid" "red")
-     (this . w) 200 (empty-scene 400 400)))
-
-  ;; on-key : KeyEvent -> World
-  (define (on-key k) (new world 10)))
+     390 200 (empty-scene 400 400))))
 
@@
     (cond [(zero? (this . n))
-           (new upworld (this . n))]
+           (new landed-world)]
           [else
@@
   ;; on-key : KeyEvent -> World
-  (define (on-key k) (new world 10)))
+  (define (on-key k) (new world 400)))
 

### 2013-09-29 tfpie/tfpie.tex
Some background on HtDP.

@@
 In particular, there is wide variation in the languages students use
-while on co-op.
-
-This combination sets the constraints under which we designed our
-approach.  Our students begin the course with a firm grasp of
-data-driven program design, as well as experience with the tools used
-in the first course:
-
-\begin{itemize}
-  \item the student languages introduced in \emph{How to Design
-    Programs}~\cite{dvanhorn:Felleisen2004Structure,
-    dvanhorn:Felleisen2001How} as well as the idea of ``language levels'';
-
-  \item the DrRacket (formerly DrScheme) 
-    programming environment~\cite{dvanhorn:Findler2002DrScheme};
-
-  \item the \texttt{check-expect} testing framework~\cite{local:check-expect};
-
-  \item the ``World'' approach to developing interactive animations
-    and games using functional programming and functional
-    graphics~\cite{dvanhorn:Felleisen2009Functional,
-      local:barland-sfp10}.
-\end{itemize}
-
-\noindent
-We deliberately reuse all of these elements in our approach.
-
-After our course, our students should both (a) be prepared for
+while on co-op. This combination sets the constraints under which we
+designed our approach.
+
+\subsection{A first course on \emph{How to Design Programs}}
+
+In the first semester, students are introduced to the ``design
+recipe'', a step-by-step process for going from English descriptions
+of problems to working programs.  The design recipe involves six steps:
+\begin{enumerate}
+\item Analyze the information involved in the problem and express
+  how to represent it as data.
+\item Write down a function signature, a summary of the purporse
+  of the function, and a function stub.
+\item Illustrate the signature and the purpose statement with some
+  functional examples.
+\item Take an inventory of the input data that can be used to compute
+  an answer.
+\item Write the code for the function.
+\item Verify the behavior of the program against the functional
+  examples given earlier.
+\end{enumerate}
+
+Students explore program design using this process in the context of a
+series of successively richer pedagogical programming language
+levels~\cite{dvanhorn:Felleisen2004Structure,
+  dvanhorn:Felleisen2001How} that are included in the DrRacket
+(formerly DrScheme) programming
+environment~\cite{dvanhorn:Findler2002DrScheme}.
+%
+The language and environment include several tools in support of the
+design recipe.  For example, functional examples can be written as
+executable tests by writting \texttt{check-expect}
+expression~\cite{local:check-expect}; an algebraic stepper and REPL
+are provided to interact with programs at each step of the design
+process.
+
+Finally, the first semester course makes extensive use of a library
+for developing interactive animations and games using functional
+programming and functional
+graphics~\cite{dvanhorn:Felleisen2009Functional, local:barland-sfp10}.
+
+\subsection{The goal of the second course}
+
+After the second course, students should both (a) be prepared for
 subsequent courses in the curriculum, which expect familiarity with
 Java and standard Java libraries, (b) be prepared for co-ops in which
-they will use professional-grade languages and tools which will
-almost certainly be object-oriented.  More significantly, we aim to
-teach the key insights behind the object-oriented approach to program
-design.
+they will use professional-grade languages and tools which will almost
+certainly be object-oriented.  More significantly, we aim to teach the
+key insights behind the object-oriented approach to program design.
 

### 2013-09-29 tfpie/tfpie.tex
done

@@
 
-\item from an interaction-oriented language to a compiled,
-  batch-oriented language.  
+\item from an interaction-oriented language and tool suite to a
+  compiled, batch-oriented language.
 \end{itemize}
@@
 starts the second semester by introducing \emph{only} the
-concept of programming with objects, while all other aspects of course
+concept of programming with objects, while all other aspects of the course
 remain where they were left off in the previous semester.  This
@@
 take place at a wide variety of companies, and while most students do
-some software development, there is almost no uniformity beyond that.
+some software development, there is no single platform or
+set of tools that commands majority use.
 In particular, there is wide variation in the languages students use
@@
 
+We chose the half-way point as the time for transition based on
+experience with earlier versions of this course. In particular, we
+found that a later transition, while allowing us to present additional
+concepts in a controlled environment, did not give students sufficient
+time and experience with Java.  Subsequent classes found that students
+were strong on fundamentals but weak on Java practice.  The other
+alternative, transitioning earlier, would not provide sufficient time
+to cover the fundamental topics before the transition.
+
 \subsection{Functional Java}
@@
 
-Thanks to the the preparation in the first half of the course, we can
+Thanks to the preparation in the first half of the course, we can
 cover OO programming in a functional subset of Java in a just a few

### 2013-11-06 tfpie/eptcs/tfpie.tex
Everything for final EPTCS version.
Copy over eptcs edits.

--- 2013-09-29 tfpie/tfpie.tex
@@
   how to represent it as data.
-\item Write down a function signature, a summary of the purporse
+\item Write down a function signature, a summary of the purpose
   of the function, and a function stub.
@@
 %
+These languages are called the beginning student language (BSL), the
+intermediate student language (ISL), and the advance student language
+(ASL).
+%
 The language and environment include several tools in support of the
 design recipe.  For example, functional examples can be written as
-executable tests by writting \texttt{check-expect}
+executable tests by writing \texttt{check-expect}
 expression~\cite{local:check-expect}; an algebraic stepper and REPL
@@
 
-\profj also takes on the the dual challenges of implementing Java as
+\profj also takes on the dual challenges of implementing Java as
 well as subsetting it.  This ultimately resulted in both a limited
@@
 from what the bulk of Northeastern University computer science majors
-take.  Anectodally, the class has also had notable success in the
+take.  Anecdotally, the class has also had notable success in the
 recruitment and retention of female students, as compared to the other
@@
 % \nocite{*}
-\bibliographystyle{plainnat}
+\bibliographystyle{eptcs}
 \bibliography{dvh-bibliography,local,sth-bibliography}
 
-\newpage
 \appendix

### 2017-09-26 1/exams/m1.tex
Practice exam.

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{times}
\usepackage{xcolor,censor}
\usepackage{fancyvrb}

\censorruleheight=0ex
\StopCensoring

\begin{document}

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 1 (Practice)}\\
    \ \\
    Fall 2017}
\end{center}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 10 \\ \hline
  4 & 20 \\ \hline
  5 & 10 \\ \hline
  6 & 20 \\ \hline \hline
  7 & 20 \\ \hline \hline
  Total: & 100\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 50 minutes to complete the test.

\vskip 1em

\noindent
The phrase ``design a program'' or ``design a function'' means follow
the steps of the design recipe.  Unless specifically asked for, you do
not need to provide intermediate products like templates or stubs,
though they may be useful to help you construct correct solutions.

\vskip 1em

\noindent
You may use any of the data definitions given to you within this exam
and do not need to repeat their definitions.

\vskip 1em

\noindent
When writing tests, you may use a shorthand for writing check-expects
by drawing an arrow between two expressions to mean you expect the
first to evaluate to same result as the second.  For example, you may
write \verb|(add1 3)| $\rightarrow$ \verb|4| instead of
\verb|(check-expect (add1 3) 4)|.

\newpage

\noindent
{\bf Problem 1 (10 points).}
%
For each of the following programs, write out each step of
computation. At each step, underline the expression being
simplified. Label each step as being ``arithmetic'' (meaning any
built-in operation), ``conditional'', ``plug'' (for plugging in an
argument for a function parameter), or ``constant'' for replacing a
constant with its value.

\noindent
{\bf Problem 1(a).}

\begin{verbatim}
(define R 9)
(define (w z) (* R z))
(add1 (cond [(= (w 2) 12) 4]
            [else 9]))
\end{verbatim}

\vskip 2.5in

\noindent
{\bf Problem 1(b).}

\begin{verbatim}
(posn-x (cond [(= (string-length "hi") 2) (make-posn 3 4)]
              [else (make-posn 8 0)]))
\end{verbatim}

\newpage

\noindent
{\bf Problem 2 (10 points).}
%
The graphical text editor you developed in a recent assignment has
garned major attention by Bay-area venture capitalists who want to
base a start-up company around your editor.  They're ready to invest
to help you develop a full-featured prototype.  But before they write
a check, they want to make sure you can handle fancy operations for
expert text editor users.  They ask if you can implement a
\emph{transpose} operation, which will transpose the character to the
left of a cursor with the character to the right of the cursor.  To
satisfy your potential investors, you define the following function
for transposing characters at a given position in a string that has at
least 2 characters in it:
\begin{verbatim}
;; transpose : String Index -> String
;; Transpose characters at index i and i+1.
;; Assumes string has length >= 2 and 0<=i, i+1<=length.
(check-expect (transpose "ab" 0) "ba")
(check-expect (transpose "peice" 1) "piece")
(check-expect (transpose "student's" 7) "students'")
(define (transpose s i) ...)
\end{verbatim}

\noindent
Give a correct definition for \verb|transpose|.  (You only need to
provide code, not the design steps.)

\newpage 
\noindent
{\bf Problem 3 (10 points).}
%
A point in three dimensional space can be represented by three numbers
specifying the points distance from the $x$-axis, $y$-axis, and
$z$-axis.  The distance of a point $(x,y,z)$ can be computed as:
\[\sqrt(x^2+y^2+z^2)\]
For example, the distance of $(2,6,9)$ to the origin is $11$; the
distance of $(3,2,6)$ is $7$.  Assume the following data definition
and design a program for computing the distance of a given 3D
point.

\begin{verbatim}
;; A 3D is a (make-3d Number Number Number)
;; Interp: a point in three dimension space
(define-struct 3d (x y z))
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; dist3d : 3D -> Number
;; Compute distance to origin from given 3-d point
(check-expect (dist3d (make-3d 3 2 6)) 7)
(define (dist3d p)
  (sqrt (+ (sqr (3d-x p)) (sqr (3d-y p)) (sqr (3d-z p)))))  
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage
\noindent
{\bf Problem 4 (20 points).}  Growing tired of Space Invaders, you
decide to build a new video game.  Part of the game consists of flying
billiard balls that move in straight lines at varying velocities until
hitting other balls or bouncing off walls and other obstacles.  After
making some sketches, you decide on the following data representation
for billiard balls:

\begin{verbatim}
;; A BB is a:
;; (make-bb Color 
;;          (make-posn Integer Integer) 
;;          (make-vel Integer Integer))

;; A Color is one of:
;; - "red"
;; - "yellow"
;; - "green"

(define-struct bb (color center vel))
(define-struct vel (deltax deltay))
\end{verbatim}
The interpretation of a \verb|BB| is that the color is the color of
the ball, the \verb|posn| is the location of the center of the ball,
and the \verb|vel| structure describes the ball's velocity as a change
along the $x$-axis (\verb|deltax|) and $y$-axis (\verb|deltay|) in one
clock tick.

\vskip 1em

\noindent
Design a function called \verb|tock : BB -> BB| that calculates where
a given billiard ball will be, based on its velocity, after one tick
of the clock and assuming it does not encounter any obstacle.

\newpage

\noindent
{\bf Problem 5 (10 points).} You've been hired by UPS, a package
delivery service, to manage and improve their truck routing software.
The software directs their truck drivers on how to proceed at each
intersection (we assume streets are laid out on a simple grid for this
problem): either continue straight, turn left, turn right, or take a
u-turn (i.e. turn around and proceed in the opposite direction.

You decide on the following data definition for representing these
instructions:

\begin{verbatim}
;; A Turn is one of:
;; - "S"
;; - "L"
;; - "R"
;; - "U"
;; Interp: directions for truck turns: straight (S), left (L), 
;; right (R), or u-turn (U).
\end{verbatim}

\noindent
{\bf Problem 5(a).} Write a template for \verb|Turn| functions.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(define (turn-temp t)
  (cond [(string=? t "S") ...]
        [(string=? t "L") ...]
        [(string=? t "R") ...]
        [(string=? t "U") ...]))

\end{SaveVerbatim}

\vskip 15em

\censor{\vspace{-15em}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\noindent
{\bf Problem 5(b).} Write a stub for the following function:

\begin{verbatim}
;; backtrack : Turn -> Turn
;; Compute the turn to go backward on route
(check-expect (backtrack "S") "S")
(check-expect (backtrack "L") "R")
(check-expect (backtrack "R") "L")
(check-expect (backtrack "U") "U")
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(define (backtrack t) "R")

\end{SaveVerbatim}

\vskip 15em

\censor{\vspace{-15em}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 6 (20 points).}
%
Part of your new UPS job is to compute useful metrics about trucking
routes, which will be used by the business office to optimize routes
for improving gasoline consumption.  A route consists of any number of
segments and each segment consists of a turn (as in problem 2) and a
distance to drive (measured in miles) after making that turn.

\vskip 1em
\noindent
Define a data definition for routes and give two examples of values
which are routes.

\newpage

\noindent
{\bf Problem 7 (20 points).}
%
Students seem obsessed with grades, which are calculated as numbers
between 0 and 100, inclusive.  At the end of the year, the professor
needs to figure out how to convert a numeric grade into a letter
grade, so they decide to write a small program using the following
data definition:
\begin{verbatim}
;; A Grade is one of:
;; - [90,100]  ;; Interp: A
;; - [80,90)   ;;         B
;; - [70,80)   ;;         C
;; - [60,70)   ;;         D
;; - [0,60)    ;;         F
\end{verbatim}

While students are obsessed with their individual grades, the
university is more interested in knowing statistics about the class as
a whole.  They ask the professor questions like ``how many `C'
students were in your class?,'' so they decide to write another
program using the following data definition:
\begin{verbatim}
;; A ClassGrades is one of:
;; - '()
;; - (cons Grade ClassGrades)
\end{verbatim}
But then, because the professor is lazy, they outsource the problem to
you.

\vskip 1em

\noindent
Design a program that calculates the number of `C' grades in a
\verb|ClassGrades| collection.

\end{document}

### 2017-09-27 1/exams/m1.tex

@@
 \censorruleheight=0ex
-\StopCensoring
+%\StopCensoring
 
@@
   {\Large {\bf CMSC 131A, Midterm 1 (Practice)}\\
+    \censor{{\bf SOLUTION}} \ \\
     \ \\
-    Fall 2017}
+    Fall 2017 \ \\
+  }
 \end{center}
@@
 
+\begin{SaveVerbatim}{VerbEnv}  
+SOLUTION:
+(add1 (cond [(= (w 2) 12) 4] [else 9]))   -->[plug]
+                ^^^^^                                    
+(add1 (cond [(= (* R 2) 12) 4] [else 9])) -->[const]
+                   ^                          
+(add1 (cond [(= (* 9 2) 12) 4] [else 9])) -->[arith]
+                ^^^^^^^           
+(add1 (cond [(= 18 12) 4] [else 9]))      -->[arith]
+             ^^^^^^^^^            
+(add1 (cond [#false 4] [else 9]))         -->[cond]
+      ^^^^^^^^^^^^^^^^^^^^^^^^^^            
+(add1 (cond [else 9]))                    -->[cond]
+      ^^^^^^^^^^^^^^^             
+(add1 9)                                  -->[arith]
+^^^^^^^^                          
+10
+\end{SaveVerbatim}
+
+\censor{\vspace{-2.5in}}
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
+
 \noindent
@@
 \end{verbatim}
+
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+(posn-x (cond [(= (string-length "hi") 2) (make-posn 3 4)]
+                  ^^^^^^^^^^^^^^^^^^^^                     -->[arith]
+              [else (make-posn 8 0)]))
+(posn-x (cond [(= 2 2) (make-posn 3 4)]
+               ^^^^^^^                                     -->[arith]
+              [else (make-posn 8 0)]))
+(posn-x (cond [#true (make-posn 3 4)]
+        ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^       
+              [else (make-posn 8 0)]))                     -->[arith]
+              ^^^^^^^^^^^^^^^^^^^^^^^         
+(posn-x (make-posn 3 4))              
+^^^^^^^^^^^^^^^^^^^^^^^^                                   -->[arith]
+3
+\end{SaveVerbatim}
+
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
 
@@
 ;; transpose : String Index -> String
-;; Transpose characters at index i and i+1.
-;; Assumes string has length >= 2 and 0<=i, i+1<=length.
-(check-expect (transpose "ab" 0) "ba")
-(check-expect (transpose "peice" 1) "piece")
-(check-expect (transpose "student's" 7) "students'")
+;; Transpose characters at left and right of index i.
+;; Assumes string has length >= 2 and 1<=i, i+1<=length.
+(check-expect (transpose "ab" 1) "ba")
+(check-expect (transpose "peice" 2) "piece")
+(check-expect (transpose "student's" 8) "students'")
 (define (transpose s i) ...)
@@
 provide code, not the design steps.)
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (transpose s i)
+  (string-append (substring s 0 (sub1 i))
+                 (substring s i (add1 i))
+                 (substring s (sub1 i) i)
+                 (substring s (add1 i))))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; tock : BB -> BB
+;; Move ball one tick based on velocity
+(check-expect (tock (make-bb "red" (make-posn 2 5) (make-vel 1 -2)))
+              (make-bb "red" (make-posn 3 3) (make-vel 1 -2)))
+(define (tock bb)
+  (make-bb (bb-color bb)
+           (make-posn (+ (posn-x (bb-center bb))
+                         (vel-deltax (bb-vel bb)))
+                      (+ (posn-y (bb-center bb))
+                         (vel-deltay (bb-vel bb))))
+           (bb-vel bb)))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \newpage
@@
 
-\vskip 15em
-
-\censor{\vspace{-15em}}
 \censor{%
@@
 which are routes.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION (there are many, this is just one):
+
+;; A Route is one of:
+;; - '()
+;; - (cons Seg Route)
+;; Interp: list of segments to follow in a truck route
+
+;; A Seg is a (make-seg Turn Number)
+(define-struct seg (turn dur))
+;; Interp: turn to make immediately,
+;; followed by distance dur going straight
+
+'()
+(cons (make-seg "L" 50)
+      (cons (make-seg "R" 20)
+            '()))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; cs : ClassGrades -> Natural
+;; Count the number of Cs in the class, i.e. numbers in [70,80).
+(check-expect (cs '()) 0)
+(check-expect (cs (cons 99 (cons 72 (cons 74 (cons 50 '()))))) 2)
+(define (cs gs)
+  (cond [(empty? gs) 0]
+        [(cons? gs)
+         (+ (cond [(c? (first gs)) 1]
+                  [else 0])
+            (cs (rest gs)))]))
+
+;; c? : Grade -> Boolean
+;; is the given grade a C (i.e. in [70,80))?
+(check-expect (c? 70) #true)
+(check-expect (c? 80) #false)
+(define (c? g)
+  (and (<= 70 g) (< g 80)))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \end{document}

### 2017-10-02 1/exams/m1.tex

--- 2017-09-26 1/exams/m1.tex
@@
 \begin{center}
-  {\Large {\bf CMSC 131A, Midterm 1 (Practice)}\\
+  {\Large {\bf CMSC 131A, Midterm 1}\\
+    \censor{{\bf SOLUTION}} \ \\
     \ \\
-    Fall 2017}
+    Fall 2017 \ \\
+  }
 \end{center}
+
+\vspace{2em}
+
+\noindent
+NAME:\verb|________________________________________________________|
+
+\vspace{2em}
+
+\noindent
+UID: \verb|__________________________________________|
+
+\vspace{2em}
 
@@
   3 & 10 \\ \hline
-  4 & 20 \\ \hline
-  5 & 10 \\ \hline
-  6 & 20 \\ \hline \hline
+  4 & 15 \\ \hline
+  5 & 15 \\ \hline
+  6 & 10 \\ \hline
   7 & 20 \\ \hline \hline
-  Total: & 100\\
+  Total: & 90\\
   \hline
@@
 \noindent
+Unless specifically instructed otherwise, you may use any built-in BSL
+functions or data types.
+
+\vskip 1em
+
+\noindent
 When writing tests, you may use a shorthand for writing check-expects
@@
 %
-For each of the following programs, write out each step of
-computation. At each step, underline the expression being
-simplified. Label each step as being ``arithmetic'' (meaning any
-built-in operation), ``conditional'', ``plug'' (for plugging in an
-argument for a function parameter), or ``constant'' for replacing a
-constant with its value.
-
-\noindent
-{\bf Problem 1(a).}
+For the following program, write out each step of computation. At each
+step, underline the expression being simplified. Label each step as
+being ``arithmetic'' (meaning any built-in operation),
+``conditional'', ``plug'' (for plugging in an argument for a function
+parameter), or ``constant'' for replacing a constant with its value.
 
@@
 (define R 9)
-(define (w z) (* R z))
-(add1 (cond [(= (w 2) 12) 4]
-            [else 9]))
+(define (w z) (= R z))
+(cond [(w 2) 4]
+      [else 9])
 \end{verbatim}
@@
 
-\noindent
-{\bf Problem 1(b).}
+\begin{SaveVerbatim}{VerbEnv}  
+SOLUTION:
+\end{SaveVerbatim}
 
-\begin{verbatim}
-(posn-x (cond [(= (string-length "hi") 2) (make-posn 3 4)]
-              [else (make-posn 8 0)]))
-\end{verbatim}
+%\censor{\vspace{-2.5in}}
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
 
@@
 %
+For the following structure definition, list the names of every
+function it creates.  For each function, classify it as being either a
+constructor, accessor, or predicate.
+
+\begin{verbatim}
+(define-struct run (time up score))
+\end{verbatim}
+
+\newpage
+
+\noindent
+{\bf Problem 3 (10 points).}
+%
 The graphical text editor you developed in a recent assignment has
-garned major attention by Bay-area venture capitalists who want to
-base a start-up company around your editor.  They're ready to invest
-to help you develop a full-featured prototype.  But before they write
-a check, they want to make sure you can handle fancy operations for
-expert text editor users.  They ask if you can implement a
-\emph{transpose} operation, which will transpose the character to the
-left of a cursor with the character to the right of the cursor.  To
-satisfy your potential investors, you define the following function
-for transposing characters at a given position in a string that has at
-least 2 characters in it:
+attracted the interest of investors.  But before they write you a
+check, they want to make sure you can handle fancy operations for
+expert text editor users.  They ask if you can implement a \emph{swap}
+operation, which will swap content to the left of the cursor with the
+content to the right..  To satisfy your potential investors, you
+define the following function for swapping strings at a given
+position:
 \begin{verbatim}
-;; transpose : String Index -> String
-;; Transpose characters at index i and i+1.
-;; Assumes string has length >= 2 and 0<=i, i+1<=length.
-(check-expect (transpose "ab" 0) "ba")
-(check-expect (transpose "peice" 1) "piece")
-(check-expect (transpose "student's" 7) "students'")
-(define (transpose s i) ...)
+;; swap : String Integer -> String
+;; Swap string to the left of i with string to the right.
+;; Assume (<= 0 i (string-length s)).
+(check-expect (swap "" 0) "")
+(check-expect (swap "ab" 1) "ba")
+(check-expect (swap "bannana" 3) "nanaban")
+(define (swap s i) ...)
 \end{verbatim}
@@
 \noindent
-Give a correct definition for \verb|transpose|.  (You only need to
+Give a correct definition for \verb|swap|.  (You only need to
 provide code, not the design steps.)
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (swap s i)
+  (string-append (substring s i)
+                 (substring s 0 i)))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 \noindent
-{\bf Problem 3 (10 points).}
+{\bf Problem 4 (15 points).}
 %
-A point in three dimensional space can be represented by three numbers
-specifying the points distance from the $x$-axis, $y$-axis, and
-$z$-axis.  The distance of a point $(x,y,z)$ can be computed as:
-\[\sqrt(x^2+y^2+z^2)\]
-For example, the distance of $(2,6,9)$ to the origin is $11$; the
-distance of $(3,2,6)$ is $7$.  Assume the following data definition
-and design a program for computing the distance of a given 3D
-point.
+Use the following data definition for representing names:
+\begin{verbatim}
+;; A Name is a (make-name String String)
+;; Interp: a person's full (first and last) name.
+;; Both strings must contain at least one letter.
+(define-struct name (first last))
+\end{verbatim}
+
+\noindent
+Write a program, \verb|name-width|, that determines the width needed
+to print someone's name on a letter using a mono-typed font (meaning
+each letter has the same width).  The program can calculate this width
+based on the length of the person's first and last name, plus the
+width of a space to go in between.  Assume each letter is 10 pixels
+wide.  For example, the name Bob Harper requires 100 pixels; the name
+Bo Jack requires 70 pixels.  You may start from the following header:
 
 \begin{verbatim}
-;; A 3D is a (make-3d Number Number Number)
-;; Interp: a point in three dimension space
-(define-struct 3d (x y z))
+;; name-width : Name -> Number
+;; Computes width needed to print name in mono-type font
 \end{verbatim}
@@
 \noindent
-{\bf Problem 4 (20 points).}  Growing tired of Space Invaders, you
-decide to build a new video game.  Part of the game consists of flying
-billiard balls that move in straight lines at varying velocities until
-hitting other balls or bouncing off walls and other obstacles.  After
-making some sketches, you decide on the following data representation
-for billiard balls:
+{\bf Problem 5 (15 points).}  
+%
+Design a program that takes a list of numbers and produces the count
+of non-negative numbers in the list.  For example, if the list
+contains 3, 4, -1, 0, -2, and 5, the count of non-negative numbers is
+4.  You may use the following data definition:
 
 \begin{verbatim}
-;; A BB is a:
-;; (make-bb Color 
-;;          (make-posn Integer Integer) 
-;;          (make-vel Integer Integer))
-
-;; A Color is one of:
-;; - "red"
-;; - "yellow"
-;; - "green"
-
-(define-struct bb (color center vel))
-(define-struct vel (deltax deltay))
+;; A LoN (list of numbers) is one of:
+;; - '()
+;; - (cons Number LoN)
+;; Interp: an arbitrarily long list of numbers
 \end{verbatim}
-The interpretation of a \verb|BB| is that the color is the color of
-the ball, the \verb|posn| is the location of the center of the ball,
-and the \verb|vel| structure describes the ball's velocity as a change
-along the $x$-axis (\verb|deltax|) and $y$-axis (\verb|deltay|) in one
-clock tick.
-
-\vskip 1em
-
-\noindent
-Design a function called \verb|tock : BB -> BB| that calculates where
-a given billiard ball will be, based on its velocity, after one tick
-of the clock and assuming it does not encounter any obstacle.
-
-\newpage
-
-\noindent
-{\bf Problem 5 (10 points).} You've been hired by UPS, a package
-delivery service, to manage and improve their truck routing software.
-The software directs their truck drivers on how to proceed at each
-intersection (we assume streets are laid out on a simple grid for this
-problem): either continue straight, turn left, turn right, or take a
-u-turn (i.e. turn around and proceed in the opposite direction.
-
-You decide on the following data definition for representing these
-instructions:
-
-\begin{verbatim}
-;; A Turn is one of:
-;; - "S"
-;; - "L"
-;; - "R"
-;; - "U"
-;; Interp: directions for truck turns: straight (S), left (L), 
-;; right (R), or u-turn (U).
-\end{verbatim}
-
-\noindent
-{\bf Problem 5(a).} Write a template for \verb|Turn| functions.
 
@@
 
-(define (turn-temp t)
-  (cond [(string=? t "S") ...]
-        [(string=? t "L") ...]
-        [(string=? t "R") ...]
-        [(string=? t "U") ...]))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
+\newpage
+
+\noindent
+{\bf Problem 6 (10 points).} You've been hired by the US Department of
+Transportation to design traffic simulation software, which includes
+simulating traffic lights.  Discovering the simple design from 131A is
+insufficient, you come up with the following representation, which
+makes it possible to model not just the color, but also how much
+longer the current light has before changing:
+
+\begin{verbatim}
+;; A Light is one of:
+;; - (make-red Natural)
+;; - (make-yellow Natural)
+;; - (make-green Natural)
+;; Interp: color and remaining duration (in ticks) of a light 
+(define-struct red (count))
+(define-struct yellow (count))
+(define-struct green (count))
+\end{verbatim}
+(Recall that a \verb|Natural| is a natural number, i.e. a non-negative
+integer.)
+
+\vspace{1em}
+
+\noindent
+{\bf Problem 6(a).} Write a template for \verb|Light| functions.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (light-temp l)
+  (cond [(red? l) (... (red-count l) ...)]
+        [(yellow? l) (... (yellow-count l) ...)]
+        [(green? l) (... (green-count l) ...)]
 
@@
 \noindent
-{\bf Problem 5(b).} Write a stub for the following function:
+{\bf Problem 6(b).} Write a stub for the following function:
 
 \begin{verbatim}
-;; backtrack : Turn -> Turn
-;; Compute the turn to go backward on route
-(check-expect (backtrack "S") "S")
-(check-expect (backtrack "L") "R")
-(check-expect (backtrack "R") "L")
-(check-expect (backtrack "U") "U")
+;; tock : Light -> Light
+;; Advance the light by one tick of time
+(check-expect (tock (make-red 10)) (make-red 9))
+(check-expect (tock (make-red 0)) (make-green 10))
 \end{verbatim}
@@
 
-(define (backtrack t) "R")
+(define (tock l) (make-red 0))
 
@@
 \noindent
-{\bf Problem 6 (20 points).}
+{\bf Problem 7 (20 points).}
 %
-Part of your new UPS job is to compute useful metrics about trucking
-routes, which will be used by the business office to optimize routes
-for improving gasoline consumption.  A route consists of any number of
-segments and each segment consists of a turn (as in problem 2) and a
-distance to drive (measured in miles) after making that turn.
+Over the summer, you take an internship at the University of
+Maryland's Center for Bioinformatics and Computational Biology (CBCB).
+Your first task deals with computing information about \emph{DNA
+  strands}.  A DNA strand is an arbitrarily long sequence of
+\emph{nucleotides}, which are either: C, G, A, or T.  Design a program
+that, given a DNA strand, computes the number of times the G
+nucleotide occurs in it.  For example, if the strand is AGCTTTGA, your
+program should produce 2.
 
-\vskip 1em
+\begin{SaveVerbatim}{VerbEnv}
+
+;; SOLUTION:
+
+;; A Nucleotide is one of:
+;; - "C"
+;; - "G"
+;; - "A"
+;; - "T"
+;; Interp: DNA nucleotide C,G,A,T.
+
+;; A DNA is one of:
+;; - '()
+;; - (cons Nucleotide DNA)
+;; Interp: a DNA strand of nucleotides
+
+;; count-g : DNA -> Natural
+;; Count the number of occurrence of G in given DNA strand
+(check-expect (count-g '()) 0)
+(check-expect (count-g (cons "A" (cons "G" (cons "G" '())))) 2)
+(define (count-g dna)
+  (cond [(empty? dna) 0]
+        [(cons? dna)
+         (+ (if (string=? (first dna) "G") 1 0)
+            (count-g (rest dna)))]))
+
+\end{SaveVerbatim}
+
+\censor{%
 \noindent
-Define a data definition for routes and give two examples of values
-which are routes.
+\BUseVerbatim{VerbEnv}}
 
@@
 \noindent
-{\bf Problem 7 (20 points).}
-%
-Students seem obsessed with grades, which are calculated as numbers
-between 0 and 100, inclusive.  At the end of the year, the professor
-needs to figure out how to convert a numeric grade into a letter
-grade, so they decide to write a small program using the following
-data definition:
-\begin{verbatim}
-;; A Grade is one of:
-;; - [90,100]  ;; Interp: A
-;; - [80,90)   ;;         B
-;; - [70,80)   ;;         C
-;; - [60,70)   ;;         D
-;; - [0,60)    ;;         F
-\end{verbatim}
+[Extra space, should you need it.]
 
-While students are obsessed with their individual grades, the
-university is more interested in knowing statistics about the class as
-a whole.  They ask the professor questions like ``how many `C'
-students were in your class?,'' so they decide to write another
-program using the following data definition:
-\begin{verbatim}
-;; A ClassGrades is one of:
-;; - '()
-;; - (cons Grade ClassGrades)
-\end{verbatim}
-But then, because the professor is lazy, they outsource the problem to
-you.
+%% Design a program that consumes an arbitrarily large collection of
+%% numbers and produces the count of non-negative numbers in the
+%% collection.
 
-\vskip 1em
+%% \begin{SaveVerbatim}{VerbEnv}
 
-\noindent
-Design a program that calculates the number of `C' grades in a
-\verb|ClassGrades| collection.
+%% SOLUTION:
+
+%% ;; A LoN (list of numbers) is one of:
+%% ;; - '()
+%% ;; - (cons Number LoN)
+
+%% ;; count-non-neg : LoN -> Natural
+%% ;; Count the number of non-negative numbers in the given list.
+%% (check-expect (count-non-neg '()) 0)
+%% (check-expect (count-non-neg (cons -1 (cons 1 (cons -2 '())))) 1)
+%% (define (count-non-neg lon)
+%%   (cond [(empty? lon) 0]
+%%         [(cons? lon) 
+%%          (+ (if (<= 0 (first lon)) 1 0)
+%%             (count-non-neg (rest lon)))]))
+%% \end{SaveVerbatim}
+
+%% \censor{%
+%% \noindent
+%% \BUseVerbatim{VerbEnv}}
 

### 2017-10-04 1/exams/m1-practice.tex
Rename exam source.
Typos.

--- 2017-09-27 1/exams/m1.tex
@@
   5 & 10 \\ \hline
-  6 & 20 \\ \hline \hline
+  6 & 20 \\ \hline
   7 & 20 \\ \hline \hline
@@
         ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^       
-              [else (make-posn 8 0)]))                     -->[arith]
+              [else (make-posn 8 0)]))                     -->[cond]
               ^^^^^^^^^^^^^^^^^^^^^^^         
@@
 The graphical text editor you developed in a recent assignment has
-garned major attention by Bay-area venture capitalists who want to
+garnered major attention by Bay-area venture capitalists who want to
 base a start-up company around your editor.  They're ready to invest

### 2017-10-28 1/exams/m2.tex
m2 practice.

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{times}
\usepackage{xcolor,censor}
\usepackage{fancyvrb}

\censorruleheight=0ex
%\StopCensoring

\begin{document}

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2 (Practice)}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

\vspace{2em}

\noindent
NAME:\verb|________________________________________________________|

\vspace{2em}

\noindent
UID: \verb|__________________________________________|

\vspace{2em}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 12 \\ \hline
  2 & 12 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 10 \\ \hline
  7 & 15 \\ \hline \hline
  Total: & 94\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 50 minutes to complete the test.

\vskip 1em

\noindent
The phrase ``design a program'' or ``design a function'' means follow
the steps of the design recipe.  Unless specifically asked for, you do
not need to provide intermediate products like templates or stubs,
though they may be useful to help you construct correct solutions.

\vskip 1em

\noindent
You may use any of the data definitions given to you within this exam
and do not need to repeat their definitions.

\vskip 1em

\noindent
Unless specifically instructed otherwise, you may use any built-in ISL+
functions or data types.

\vskip 1em

\noindent
When writing tests, you may use a shorthand for writing check-expects
by drawing an arrow between two expressions to mean you expect the
first to evaluate to same result as the second.  For example, you may
write \verb|(add1 3)| $\rightarrow$ \verb|4| instead of
\verb|(check-expect (add1 3) 4)|.

\newpage

\noindent
{\bf Problem 1 (12 points).}
%
Each of the following signatures describes a class of functions:

\begin{verbatim}
Boolean String -> String
[Boolean -> String] -> String
[Listof Boolean] [Listof Number] -> Number
[X] . [Number -> X] -> X
[X Y] . [X -> Y] [Listof X] -> [Listof Y]
[X Y Z] . [X -> Y] [Y -> Z] X -> Z
\end{verbatim}
Define one example of a function in each class.  (You only need to
provide the code, not a full design.)

\vskip 2.5in

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 2 (12 points).}
%
For each of the following functions, provide the most general signature
that correctly describes the function:

\begin{verbatim}
;; f :

(define (f x)
  (+ (sqr x) 2))

;; g :

(define (g x)
  (map add1 x))

;; h :

(define (h x y)
  (cond [(empty? x) y]
        [(cons? x)
         (cons (sqr (first x))
               (h (rest x) y))]))

;; i :

(define (i x)
  (sqr (x 5)))

;; j :

(define (j x)
  (lambda (y)
    (+ x y)))

;; k :

(define (k x y)
  (cond [(= y 0) (list (x 0))]
        [else
          (cons (x y) (k x (sub1 y)))]))
\end{verbatim}

\newpage

\noindent
{\bf Problem 3 (15 points).}
%
Design a program called {\tt avg} that computes the average of a list of numbers.

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Here is a parametric data definition for a non-empty list of items:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
;; Interp: an arbitrarily long but non-empty sequence of items.
\end{verbatim}
Here are two similar programs that operate over non-empty lists:
\begin{verbatim}
;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;

;; smallest : [NEListof Number] -> Number
;; Produce the smallest number in the list
(check-expect (smallest (list 6 3 2 4)) 2)
(define (smallest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (min (first xs)
               (smallest (rest xs)))]))

;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;

;; longest : [NEListof String] -> String
;; Produce the first longest string in the list
(check-expect (longest (list "hi" "there")) "there")
(check-expect (longest (list "same" "size")) "same")
(define (longest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (longer (first xs)
                  (longest (rest xs)))]))

;; longer : String String -> String
;; Select the longer of the two strings (s1 if tied)
(check-expect (longer "hi" "there") "there")
(check-expect (longer "same" "size") "same")
(define (longer s1 s2)
  (cond [(< (string-length s1) (string-length s2)) s2]
        [else s1]))
\end{verbatim}

Design an abstraction of the {\tt smallest} and {\tt longest}
functions and redefine {\tt smallest} and {\tt longest} in terms of
it.  Be sure to give the most general parametric signature for your
abstraction function.

\newpage

\noindent
[Space for problem 4.]

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of numbers and produces the count
of positive numbers in the list.  For example, if the list contains 3,
4, -1, 0, -2, and 5, the count of positive numbers is 3.  You may
assume the {\tt{[Listof Number]}} data definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Listof Number]}}.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 6 (10 points).}
%
It's been said that {\tt foldr} is \emph{the} fundamental list
abstraction function because any function that follows the list
template can be expressed in terms of {\tt foldr}.  Put that theory to
the test by providing an equivalent definition for the following
function in terms of {\tt foldr}:

\begin{verbatim}
;; map : [X Y] . [X -> Y] [Listof X] -> [Listof Y]
;; Apply f to each element of list, collect a list of results
(define (map f ls)
  (cond [(empty? ls) '()]
        [(cons? ls)
         (cons (f (first ls))
               (map f (rest ls)))]))
\end{verbatim}

\noindent
(Just provide code, not a full design.)

\newpage

\noindent
{\bf Problem 7 (15 points).}
%
A sports tournament in which teams pair off against eachother and
the winner advances to the next level in the tournament can be modelled
with the following data definition:

\begin{verbatim}
;; A Tournament is one of:
;; - String
;; - (make-bracket Tournament Tournament)
;; Interp: a sports tournament where a String is a team name
;; that is still in the tournament and a bracket represents
;; a match between the winner of two tournaments. 
\end{verbatim}

Design a function {\tt most-games} that computes the maximum number of
games any team in the tournament would have to play in order to win
the tournament.

\end{document}

### 2017-10-31 1/exams/m2-practice.tex
Moving README.
Start of practice exam soln.

--- 2017-10-28 1/exams/m2.tex
@@
 
-\vskip 2.5in
-
 \begin{SaveVerbatim}{VerbEnv}
 SOLUTION:
+
+(define (f b s) s)
+(define (g fs) (fs #false))
+(define (h bs ns) 2)
+(define (i f) (f 2))
+(define (j xs) (map f xs))
+(define (k f g x) (g (f x)))
 \end{SaveVerbatim}
@@
 \begin{verbatim}
-;; f :
+;; f : \censor{Number -> Number}
 
@@
 
-;; g :
+;; g : \censor{[Listof Number] -> [Listof Number]}
 
@@
 
-;; h :
+;; h : \censor{[Listof Number] [Listof Number] -> [Listof Number]}
 
@@
 
-;; i :
+;; i : \censor{(Number -> Number) -> Number}
 
@@
 
-;; j :
+;; j : \censor{Number -> (Number -> Number)}
 
@@
 
-;; k :
+;; k : \censor{[X] (Natural -> X) Natural -> [Listof X]}
 
@@
 Design a program called {\tt avg} that computes the average of a list of numbers.
+
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+
+;; avg : [Listof Number] -> Number
+;; Compute the average of a list of numbers
+;; Assume: the list is non-empty
+(check-expect (avg (list 3 4 5)) 4)
+(check-expect (avg (list 10 20)) 15)
+(define (avg lon)
+  (/ (sum lon) 
+     (length lon)))
+
+;; sum : [Listof Number] -> Number
+;; Sum the list of numbers
+(check-expect (sum '()) 0)
+(check-expect (sum (list 3 4 5)) 12)
+(define (sum lon)
+  (foldr + 0 lon))
+
+;; Alt: follow [Listof Number] template
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+
+;; mostest : [X] [NEListof X] [X X -> X] -> X
+;; Produce the first most-est element in the list
+(define (mostest xs most)
+  (cond [(empty? (rest xs)) (first xs)]
+        [else
+          (most (first xs)
+                (mostest (rest xs) most))]))
+
+(define (longest los)
+  (mostest los longer))
+
+(define (smallest lon)
+  (mostest lon min))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \newpage
@@
 \begin{SaveVerbatim}{VerbEnv}
-
 SOLUTION:
 
+;; count-pos : [Listof Number] -> Number
+;; Count the number of non-negative numbers in given list
+(check-expect (count-pos '()) 0)
+(check-expect (count-pos (list 3 -1 0 2)) 2)
+(define (count-pos lon)
+  (foldr (lambda (n c) (if (positive? n) (add1 c) c)) 0 lon))
+
+;; Alt:
+(define (count-pos lon)
+  (length (filter positive? lon)))
 \end{SaveVerbatim}
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+
+(define (map f xs)
+  (foldr (lambda (x ys) (cons (f x) ys)) '() xs))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \newpage
@@
 ;; a match between the winner of two tournaments. 
+(define-struct bracket (t1 t2))
 \end{verbatim}
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+
+;; most-games : Tournament -> Number
+;; Compute most games any team has to win to win tournament
+(check-expect (most-games "NY") 0)
+(check-expect (most-games (make-bracket "LA" "HOU")) 1)
+(check-expect (most-games (make-bracket (make-bracket "LA" "DC") "HOU")) 2)
+(define (most-games t)
+  (cond [(string? t) 0]
+        [(bracket? t)
+         (add1 (max (most-games(bracket-t1 t))
+                    (most-games(bracket-t2 t))))]))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \end{document}

### 2017-11-01 1/exams/m2-practice.tex
Fix bug.

@@
 \usepackage{fullpage}
-\usepackage{times}
 \usepackage{xcolor,censor}
 \usepackage{fancyvrb}
+\usepackage{fontspec}
+\usepackage{alltt}
+
+\setmainfont{Times New Roman}
 
@@
 \begin{SaveVerbatim}{VerbEnv}
+
 SOLUTION:
@@
 (define (i f) (f 2))
-(define (j xs) (map f xs))
+(define (j f xs) (map f xs))
 (define (k f g x) (g (f x)))
@@
 
-\begin{verbatim}
+\begin{alltt}
 ;; f : \censor{Number -> Number}
@@
           (cons (x y) (k x (sub1 y)))]))
-\end{verbatim}
+\end{alltt}
 
@@
         [(bracket? t)
-         (add1 (max (most-games(bracket-t1 t))
-                    (most-games(bracket-t2 t))))]))
+         (add1 (max (most-games (bracket-t1 t))
+                    (most-games (bracket-t2 t))))]))
 \end{SaveVerbatim}

### 2017-11-06 1/exams/m2.tex
Progress on m2.

@@
 \begin{center}
-  {\Large {\bf CMSC 131A, Midterm 2 (Practice)}\\
+  {\Large {\bf CMSC 131A, Midterm 2}\\
     \censor{{\bf SOLUTION}} \ \\
@@
 
-\vskip 2.5in
+\begin{SaveVerbatim}{VerbEnv}
 
-\begin{SaveVerbatim}{VerbEnv}  
 SOLUTION:
+
+(define (f b s) s)
+(define (g fs) (fs #false))
+(define (h bs ns) 2)
+(define (i f) (f 2))
+(define (j f xs) (map f xs))
+(define (k f g x) (g (f x)))
 \end{SaveVerbatim}
@@
 
-\begin{verbatim}
-;; f :
+\begin{alltt}
+;; method : \censor{Number -> Number}
 
-(define (f x)
+(define (method x)
   (+ (sqr x) 2))
 
-;; g :
+;; raekwon : \censor{[Listof Number] -> Boolean}
 
-(define (g x)
-  (map add1 x))
+(define (raekwon x)
+  (ormap positive? x))
 
-;; h :
+;; rza : \censor{[Listof String] [Listof Number] -> [Listof Number]}
 
-(define (h x y)
+(define (rza x y)
   (cond [(empty? x) y]
         [(cons? x)
-         (cons (sqr (first x))
-               (h (rest x) y))]))
+         (cons (string-length (first x))
+               (rza (rest x) y))]))
 
-;; i :
+;; gza : \censor{(String Number -> Number) -> Number}
 
-(define (i x)
-  (sqr (x 5)))
+(define (gza x)
+  (foldr x 0 (list "a" "b" "c")))
 
-;; j :
+;; odb : \censor{(Number -> Number) -> (Number -> Number)}
 
-(define (j x)
+(define (odb x)
   (lambda (y)   
-    (+ x y)))
+    (/ (- (x (+ y 0.001))
+          (x (- y 0.001)))
+       (* 2 0.001))))
 
-;; k :
+;; ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]}
 
-(define (k x y)
-  (cond [(= y 0) (list (x 0))]
-        [else
-          (cons (x y) (k x (sub1 y)))]))
-\end{verbatim}
+(define (ghostface x y)
+  (cond [(empty? x) '()]
+        [(cons? x)
+         (cons (y (first x)) 
+               (ghostface (rest x) y))]))
+\end{alltt}
 
@@
 %
-Design a program called {\tt avg} that computes the average of a list of numbers.
+Design a program called {\tt w-avg} that computes the weighted average
+of a list of numbers and a list of weights.
+
+For example, let's say a class grade is based on two midterms and a
+project where the project is worth twice as much as the midterms.  A
+student who gets an 80 and a 70 on the midterms and a 90 on the
+project would have a weighted average of 82.5: $((1 \times 80) + (1
+\times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
+  (w-avg (list 80 70 90) (list 1 1 2))}.  You should assume the two
+lists have the same length.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; avg : [Listof Number] -> Number
+;; Compute the average of a list of numbers
+;; Assume: the list is non-empty
+(check-expect (avg (list 3 4 5)) 4)
+(check-expect (avg (list 10 20)) 15)
+(define (avg lon)
+  (/ (sum lon) 
+     (length lon)))
+
+;; sum : [Listof Number] -> Number
+;; Sum the list of numbers
+(check-expect (sum '()) 0)
+(check-expect (sum (list 3 4 5)) 12)
+(define (sum lon)
+  (foldr + 0 lon))
+
+;; Alt: follow [Listof Number] template
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 \noindent
-{\bf Problem 4 (15 points).}
+{\bf Problem 4 (20 points).}
 %
-Here is a parametric data definition for a non-empty list of items:
+Here is a parametric data definition for a tree of elements:
 \begin{verbatim}
-;; A [NEListof X] is one of:
-;; - (cons X '())
-;; - (cons X [NEListof X])
-;; Interp: an arbitrarily long but non-empty sequence of items.
+;; A [Tree X] is one of:
+;; - (make-leaf)
+;; - (make-node X [Tree X] [Tree X])
+;; Interp: a binary tree that is either empty (a leaf), or non-empty (a node)
+;; with an element and two sub-trees.
+(define-struct leaf ())
+(define-struct node (elem left right))
 \end{verbatim}
-Here are two similar programs that operate over non-empty lists:
-\begin{verbatim}
-;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;
+{\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
+that operates on a binary tree of numbers, {\tt [Tree Number]}, and
+produces the sum of all the numbers in the tree.
 
-;; smallest : [NEListof Number] -> Number
-;; Produce the smallest number in the list
-(check-expect (smallest (list 6 3 2 4)) 2)
-(define (smallest xs)
-  (cond [(empty? (rest xs)) (first xs)]
-        [else
-          (min (first xs)
-               (smallest (rest xs)))]))
+\begin{SaveVerbatim}{VerbEnv}
 
-;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;
+SOLUTION:
 
-;; longest : [NEListof String] -> String
-;; Produce the first longest string in the list
-(check-expect (longest (list "hi" "there")) "there")
-(check-expect (longest (list "same" "size")) "same")
-(define (longest xs)
-  (cond [(empty? (rest xs)) (first xs)]
-        [else
-          (longer (first xs)
-                  (longest (rest xs)))]))
+;; tree-sum : [Tree Number] -> Number
+;; Sum all the numbers in the tree
+(check-expect (tree-sum (make-leaf)) 0)
+(check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
+                                     (make-leaf)))
+              8)
+(define (tree-sum bt)
+  (cond [(leaf? bt) 0]
+        [(node? bt)
+         (+ (node-elem bt)
+            (tree-sum (node-left bt))
+            (tree-sum (node-right bt)))]))
 
-;; longer : String String -> String
-;; Select the longer of the two strings (s1 if tied)
-(check-expect (longer "hi" "there") "there")
-(check-expect (longer "same" "size") "same")
-(define (longer s1 s2)
-  (cond [(< (string-length s1) (string-length s2)) s2]
-        [else s1]))
-\end{verbatim}
+\end{SaveVerbatim}
 
-Design an abstraction of the {\tt smallest} and {\tt longest}
-functions and redefine {\tt smallest} and {\tt longest} in terms of
-it.  Be sure to give the most general parametric signature for your
-abstraction function.
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 \noindent
-[Space for problem 4.]
+{\bf Problem 4(b) (10 points).}
+Here is a function that produces a list of all the elements in a tree:
+\begin{verbatim}
+;; tree-elems : [X] . [Tree X] -> [Listof X]
+;; Produce a list of all elements in tree going top-down, left-to-right
+(check-expect (tree-elems (make-leaf)) '())
+(check-expect (tree-elems (make-node 7 
+                                     (make-node 9 (make-leaf) (make-leaf))
+                                     (make-node 2 (make-leaf) (make-leaf))))
+              (list 7 9 2))
+(define (tree-elems bt)
+  (cond [(leaf? bt) '()]
+        [(node? bt)
+         (cons (node-elem bt)
+               (append (tree-elems (node-left bt))
+                       (tree-elems (node-right bt))))]))
+\end{verbatim}
+Here is an abstraction function for trees that is similar to
+{\tt foldr} for lists, but works on trees:
+\begin{verbatim}
+;; tree-fold : [X Y] . [X Y Y -> Y] Y [Tree X] -> Y
+;; The fundamental abstraction function for trees
+(define (tree-fold f b bt)
+  (cond [(leaf? bt) b]
+        [(node? bt)
+         (f (node-elem bt)
+            (tree-fold f b (node-left bt))
+            (tree-fold f b (node-right bt)))]))
+\end{verbatim}
+Give an equivalent definition of {\tt tree-elems} in terms of {\tt
+  tree-fold}.  (Just provide the code.)
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (tree-elems bt)
+  (tree-fold (lambda (x l r) (cons x (append l r))) '() bt))
+
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 %
-Design a program that takes a list of numbers and produces the count
-of positive numbers in the list.  For example, if the list contains 3,
-4, -1, 0, -2, and 5, the count of positive numbers is 3.  You may
-assume the {\tt{[Listof Number]}} data definition is defined.
+Design a program that takes a list of strings and produces the count
+of strings with length longer than 3 in the list.  For example, if the
+list contains {\tt "a"}, {\tt "dave"}, {\tt "abc"}, and {\tt "fred"},
+the count is 2.  You may assume the {\tt{[Listof String]}} data
+definition is defined.
 %
 For full credit, use list abstraction functions.  For partial credit,
-follow the template for {\tt{[Listof Number]}}.
+follow the template for {\tt{[Listof String]}}.
 
@@
 
+;; count-3+ : [Listof String] -> Number
+;; Count the number of strings with length more than 3 in given list
+(check-expect (count-3+ '()) 0)
+(check-expect (count-3+ (list "a" "abcd" "fred" "c")) 2)
+(define (count-3+ los)
+  (foldr (lambda (s c) (if (> (string-length s) 3) (add1 c) c)) 0 los))
+
+;; Alts:
+(define (count-3+ los)
+  (local [;; count+ : String Number -> Number
+          ;; Bump count if string has length > 3
+          (define (count+ s c) 
+            (if (> (string-length s) 3) (add1 c) c))]
+    (foldr count+ 0 los))
+
+(define (count-3+ los)
+  (local [;; 3+? : String -> Boolean
+          ;; Does the string have length > 3?
+          (define (3+? s) 
+            (> (string-length s) 3))]
+    (length (filter 3+? los))))
+
+(define (count-3+ los)
+  (length (filter (lambda (s) (> (string-length s) 3)) los)))
 \end{SaveVerbatim}
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (map f xs)
+  (foldr (lambda (x ys) (cons (f x) ys)) '() xs))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \newpage
@@
 %
-A sports tournament in which teams pair off against eachother and
-the winner advances to the next level in the tournament can be modelled
-with the following data definition:
-
+Here's a data definition for choose-your-own-adventure books:
 \begin{verbatim}
-;; A Tournament is one of:
+;; A CYOA is one:
 ;; - String
-;; - (make-bracket Tournament Tournament)
-;; Interp: a sports tournament where a String is a team name
-;; that is still in the tournament and a bracket represents
-;; a match between the winner of two tournaments. 
+;; - (make-choice String CYOA CYOA)
+(define-struct choice (q yes no))
+;; Interp: a choose-your-own-adventure book where a string
+;; is a conclusion and a choice is a yes/no question and
+;; the two adventures that follow as a consequnce of the answer.
 \end{verbatim}
 
-Design a function {\tt most-games} that computes the maximum number of
-games any team in the tournament would have to play in order to win
-the tournament.
+\noindent
+Design a function called {\tt ans} that consumes a {\tt CYOA} and a
+{\tt [Listof Boolean]} that represents answers to the yes/no
+questions: {\tt \#true} means yes, {\tt \#false} means no.  The {\tt
+  ans} function produces the {\tt CYOA} adventure that results after
+giving all of the answers in the list.  If the list contains too many
+answers, {\tt ans} just produces the conclusion of the adventure.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; most-games : Tournament -> Number
+;; Compute most games any team has to win to win tournament
+(check-expect (most-games "NY") 0)
+(check-expect (most-games (make-bracket "LA" "HOU")) 1)
+(check-expect (most-games (make-bracket (make-bracket "LA" "DC") "HOU")) 2)
+(define (most-games t)
+  (cond [(string? t) 0]
+        [(bracket? t)
+         (add1 (max (most-games (bracket-t1 t))
+                    (most-games (bracket-t2 t))))]))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 

### 2017-11-06 1/exams/m2.tex
m2 draft.

@@
   Question & Points \\ \hline \hline
-  1 & 12 \\ \hline
+  1 & 15 \\ \hline
   2 & 12 \\ \hline
-  3 & 15 \\ \hline
-  4 & 15 \\ \hline
+  3 & 20 \\ \hline
+  4 & 20 \\ \hline
   5 & 15 \\ \hline
-  6 & 10 \\ \hline
-  7 & 15 \\ \hline \hline
-  Total: & 94\\
+  6 & 20 \\ \hline\hline
+  Total: & 102\\
   \hline
@@
 \noindent
-{\bf Problem 1 (12 points).}
+{\bf Problem 1 (15 points).}
 %
-Each of the following signatures describes a class of functions:
+Here are some data definition relevant to representing a dictionary,
+which associates words with the their definitions:
 
 \begin{verbatim}
-Boolean String -> String
-[Boolean -> String] -> String
-[Listof Boolean] [Listof Number] -> Number
-[X] . [Number -> X] -> X
-[X Y] . [X -> Y] [Listof X] -> [Listof Y]
-[X Y Z] . [X -> Y] [Y -> Z] X -> Z
+;; A Dict is one of:
+;; - '()
+;; - (cons (list String String) Dict)
+;; Interp: represents a collection of definitions where each
+;; element is a two-element list consisting of a word (first) and its
+;; meaning (second).
+
+;; A Result is one of:
+;; - #false
+;; - String
+;; Interp: represents the result of looking up a word in the
+;; dictionary; #false means not found.
 \end{verbatim}
-Define one example of a function in each class.  (You only need to
-provide the code, not a full design.)
+
+\noindent
+Design the following function:
+
+\begin{verbatim}
+;; lookup : String Dict -> Result
+;; Lookup the (first) definition of a word in the dictionary,
+;; produces #false if not found
+\end{verbatim}
 
@@
 \noindent
-{\bf Problem 3 (15 points).}
+{\bf Problem 3 (20 points).}
 %
@@
 \noindent
-{\bf Problem 6 (10 points).}
-%
-It's been said that {\tt foldr} is \emph{the} fundamental list
-abstraction function because any function that follows the list
-template can be expressed in terms of {\tt foldr}.  Put that theory to
-the test by providing an equivalent definition for the following
-function in terms of {\tt foldr}:
-
-\begin{verbatim}
-;; map : [X Y] . [X -> Y] [Listof X] -> [Listof Y]
-;; Apply f to each element of list, collect a list of results
-(define (map f ls)
-  (cond [(empty? ls) '()]
-        [(cons? ls)
-         (cons (f (first ls))
-               (map f (rest ls)))]))
-\end{verbatim}
-
-\noindent
-(Just provide code, not a full design.)
-
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
-(define (map f xs)
-  (foldr (lambda (x ys) (cons (f x) ys)) '() xs))
-\end{SaveVerbatim}
-
-\censor{%
-\noindent
-\BUseVerbatim{VerbEnv}}
-
-\newpage
-
-\noindent
-{\bf Problem 7 (15 points).}
+{\bf Problem 6 (20 points).}
 %
@@
 
-;; most-games : Tournament -> Number
-;; Compute most games any team has to win to win tournament
-(check-expect (most-games "NY") 0)
-(check-expect (most-games (make-bracket "LA" "HOU")) 1)
-(check-expect (most-games (make-bracket (make-bracket "LA" "DC") "HOU")) 2)
-(define (most-games t)
-  (cond [(string? t) 0]
-        [(bracket? t)
-         (add1 (max (most-games (bracket-t1 t))
-                    (most-games (bracket-t2 t))))]))
+(define a0 "A0")
+(define a1 (make-choice "Q1" "A1" "A2"))
+(define a2 (make-choice "Q2" a0 a1))
+
+;; ans : CYOA [Listof Boolean] -> CYOA
+;; ...
+(check-expect (adv a0 (list #true #false)) "A0")
+(check-expect (adv a1 (list #true #false)) "A1")
+(check-expect (adv a2 (list #false #true)) "A1")
+(check-expect (adv a2 (list #false)) a1)
+(define (ans adv loa)
+  (cond [(string? adv) adv]
+        [(empty? loa) adv]
+        [(first loa)
+         (ans (choice-yes adv) (rest adv))]
+        [else
+         (ans (choice-no adv) (rest adv))]))
 \end{SaveVerbatim}

### 2017-11-06 1/exams/m2.tex
Tweaks.

@@
   3 & 20 \\ \hline
-  4 & 20 \\ \hline
+  4 & 15 \\ \hline
   5 & 15 \\ \hline
   6 & 20 \\ \hline\hline
-  Total: & 102\\
+  Total: & 97\\
   \hline
@@
 ;; - (cons (list String String) Dict)
-;; Interp: represents a collection of definitions where each
-;; element is a two-element list consisting of a word (first) and its
-;; meaning (second).
+;; Interp: a collection of definitions where each element is a 
+;; two-element list of a word (first) and its meaning (second).
 
@@
 \noindent
-{\bf Problem 4 (20 points).}
+{\bf Problem 4 (15 points).}
 %
@@
 \end{verbatim}
-{\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
-that operates on a binary tree of numbers, {\tt [Tree Number]}, and
-produces the sum of all the numbers in the tree.
+%% {\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
+%% that operates on a binary tree of numbers, {\tt [Tree Number]}, and
+%% produces the sum of all the numbers in the tree.
 
-\begin{SaveVerbatim}{VerbEnv}
+%% \begin{SaveVerbatim}{VerbEnv}
 
-SOLUTION:
+%% SOLUTION:
 
-;; tree-sum : [Tree Number] -> Number
-;; Sum all the numbers in the tree
-(check-expect (tree-sum (make-leaf)) 0)
-(check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
-                                     (make-leaf)))
-              8)
-(define (tree-sum bt)
-  (cond [(leaf? bt) 0]
-        [(node? bt)
-         (+ (node-elem bt)
-            (tree-sum (node-left bt))
-            (tree-sum (node-right bt)))]))
+%% ;; tree-sum : [Tree Number] -> Number
+%% ;; Sum all the numbers in the tree
+%% (check-expect (tree-sum (make-leaf)) 0)
+%% (check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
+%%                                      (make-leaf)))
+%%               8)
+%% (define (tree-sum bt)
+%%   (cond [(leaf? bt) 0]
+%%         [(node? bt)
+%%          (+ (node-elem bt)
+%%             (tree-sum (node-left bt))
+%%             (tree-sum (node-right bt)))]))
 
-\end{SaveVerbatim}
+%% \end{SaveVerbatim}
 
-\censor{%
-\noindent
-\BUseVerbatim{VerbEnv}}
+%% \censor{%
+%% \noindent
+%% \BUseVerbatim{VerbEnv}}
 
-\newpage
+%% \newpage
 
 \noindent
-{\bf Problem 4(b) (10 points).}
+%{\bf Problem 4(b) (10 points).}
 Here is a function that produces a list of all the elements in a tree:

### 2017-11-06 1/exams/m2.tex
Solution to m2.

@@
 \censorruleheight=0ex
-%\StopCensoring
+\StopCensoring
 
@@
 
-(define (f b s) s)
-(define (g fs) (fs #false))
-(define (h bs ns) 2)
-(define (i f) (f 2))
-(define (j f xs) (map f xs))
-(define (k f g x) (g (f x)))
+(check-expect (lookup "a" '()) #false)
+(check-expect (lookup "a" (list (list "a" "b"))) "b")
+(check-expect (lookup "a" (list (list "c" "b") (list "a" "d"))) "d")
+(define (lookup w d)
+  (cond [(empty? d) #false]
+        [(cons? d)
+         (if (string=? (first (first d)) w)
+             (second (first d))
+             (lookup w (rest d)))]))
+
 \end{SaveVerbatim}
@@
 
-;; avg : [Listof Number] -> Number
-;; Compute the average of a list of numbers
-;; Assume: the list is non-empty
-(check-expect (avg (list 3 4 5)) 4)
-(check-expect (avg (list 10 20)) 15)
-(define (avg lon)
-  (/ (sum lon) 
-     (length lon)))
+;; w-avg : [Listof Number] [Listof Number] -> Number
+;; Compute the weighted average of a list of numbers and weights
+;; Assume: lists have the same length (and non-empty)
+(check-expect (w-avg (list 70 80 90) (list 1 1 2)) 82.5)
+(define (w-avg lon ws)
+  (/ (w-sum lon ws) 
+     (foldr + 0 ws)))
 
-;; sum : [Listof Number] -> Number
-;; Sum the list of numbers
-(check-expect (sum '()) 0)
-(check-expect (sum (list 3 4 5)) 12)
-(define (sum lon)
-  (foldr + 0 lon))
+;; w-sum : [Listof Number] [Listof Number] -> Number
+;; Sum the list of numbers according to given weights
+;; Assume: lists have same length
+(check-expect (w-sum (list 70 80 90) (list 1 1 2)) (+ 70 80 180))
+(define (w-sum lon ws)
+  (cond [(empty? lon) 0]
+        [(cons? lon)
+         (+ (* (first lon) (first ws))
+            (w-sum (rest lon) (rest ws)))]))
 
-;; Alt: follow [Listof Number] template
+;; Alt: define w-total function for denominator
+
+;; w-total : [Listof Number] -> Number
+;; Total a list of weights
+(define (w-total ws)
+  (cond [(empty? ws) 0]
+        [(cons? ws)
+         (+ (first ws) 
+            (w-total (rest ws)))]))
 \end{SaveVerbatim}
@@
 
+\vspace{1em}
+\noindent
+[Provide your answer on the next page.]
+
 \begin{SaveVerbatim}{VerbEnv}
@@
 \BUseVerbatim{VerbEnv}}
+
+\newpage
+
+\noindent
+[Space for problem 4.]
 
@@
 ;; ans : CYOA [Listof Boolean] -> CYOA
-;; ...
+;; Compute the remaining adventure after using answers from list
 (check-expect (adv a0 (list #true #false)) "A0")
@@
         [(empty? loa) adv]
-        [(first loa)
-         (ans (choice-yes adv) (rest adv))]
-        [else
-         (ans (choice-no adv) (rest adv))]))
+        [(cons? loa)
+         (if (first loa)
+             (ans (choice-yes adv) (rest adv))
+             (ans (choice-yes adv) (rest adv)))]))
 \end{SaveVerbatim}

### 2017-11-30 exams/exam6/2.tex
Housekeeping.

--- 2013-04-09 exam6/2.tex
@@
 \begin{verbatim}
-// 5 pts for Note
+// 3 pts for Note
 class Note {
@@
 
-// 4 pts for List
+// 2 pts for List
 interface List<X> {}
@@
 
-//> 4 pts for interface defintions
+//> 3 pts for interface defintions
 
@@
 
-//> 4 pts for accept methods
+//> 2 pts for accept methods
 // in Note
@@
 \begin{verbatim}
-//> 4 pts
+//> 6 pts overall
 // Nesting Depth with recursion
@@
 
-//> 4 pts
 class NestDepthListR implements ListVisitor<Note,Integer> {
@@
 \fi
+
+\begin{solution}
+\begin{verbatim}
+// 5 pts (including tests)
+// in Note
+    Boolean sameNote(Note s) {
+        return this.txt.equals(s.txt)
+            && this.subnotes.sameList(s.subnotes)
+    }
+
+// in List<X>
+    Boolean sameList(List<X> l);
+    Boolean sameEmpty(Empty<X> l);
+    Boolean sameCons(Cons<X> l);
+
+// in Empty<X>
+    public Boolean sameList(List<X> l) { return l.sameEmpty(this); }
+    public Boolean sameEmpty(Empty<X> l) { return true; }
+    public Boolean sameCons(Cons<X> l) { return false; }
+
+// in Cons<X>
+    public Boolean sameList(List<X> l) { return l.sameCons(this); }
+    public Boolean sameEmpty(Empty<X> l) { return false; }
+    public Boolean sameCons(Cons<X> l) { 
+        return this.first.equals(l.first)
+            && this.rest.sameList(l.rest);
+    }
+\end{verbatim}
+\end{solution}
+
 \newpage
@@
 \fi
+
+\begin{solution}
+\begin{verbatim}
+// 4 pts (including tests)
+// in Note
+    int hashCode() { return this.txt.hashCode() + this.subnotes.hashCode(); }
+
+// in Empty
+    int hashCode() { return 17; }
+
+// in Cons
+    int hashCode() { return this.first.hashCode() + this.rest.hashCode(); }
+\end{verbatim}
+\end{solution}
+
 \newpage

### 2017-12-01 2/tfpie/response.txt
Move Fundamentals II content into own dir.

--- 2013-09-29 tfpie/response.txt
@@
 ----------------------- REVIEW 4 ---------------------
-PAPER: 1
-TITLE: From Principles to Practice with Class in the First Year
-AUTHORS: Sam Tobin-Hochstadt and David Van Horn
-
-OVERALL EVALUATION: 2 (accept)
-REVIEWER'S CONFIDENCE: 4 (high)
-
------------ REVIEW -----------
-
-The authors describe a new approach to tackle difficulties in
-transitioning from a functional semester one course to an
-object-oriented semester two course. Students are reportedly confused
-by the circumstantial differences (compilation model / environment /
-syntax) which makes grasping the fundamental concepts being taught
-harder to grasp. Instead of transitioning on day one of the second
-semester to Java in Eclipse (or similar), the authors present an
-approach that continues the progression of student languages to
-include objects and other OO concepts before transitioning to a "real"
-OO environment and language.
 
@@
 
-A few specific comments and minor corrections:
-
- - (Sec.1) "common approache to" --> "common approach to"
+RESPONSE: We clarify why the transition point was chosen.
 
@@
 
- - (Sec.1) "aspects of course remain" --> "aspects of *the* course
-   remain"
+RESPONSE: That Java has no semantics for single-expression evaluation
+is a fact about the language, not its tools.
 
@@
 
- - (Sec.3.1) "and message are sent" --> "and message*s* are sent"
-
- - (Sec.4.2) "Thanks to the the preparation" --> omit one "the"
+RESPONSE: We edited this sentence.
 
@@
   without at least some concrete data and/or qualitative evaluation.
+
+RESPONSE: We agree that this merits further study.  However, our
+remark is anecdotal, and we have clarified it.

### 2017-12-01 2/admin/charter/CS3500.tex
Move Fundamentals II content into own dir.

\documentclass[11pt]{article}

\usepackage{fullpage}

\title{CS~3500 Course Charter \& Syllabus}

\author{Jay Aslam, Will Clinger, David Van Horn \\
        College of Computer and Information Science \\
        Northeastern University}

\date{February 21, 2012}

\begin{document}

\maketitle

\section{CS~3500 Course Charter}

\paragraph{Course Title:}
%
CS~3500 \emph{Object-Oriented Design} (4 SH)

\paragraph{Current Catalog Description:}
%
Presents a comparative approach to object-oriented programming
and design. Discusses the concepts of object, class, meta-class,
message, method, inheritance, and genericity. Reviews forms of
polymorphism in object-oriented languages. Contrasts the use of
inheritance and composition as dual techniques for software reuse:
forwarding vs. delegation and subclassing vs. subtyping. Fosters
a deeper understanding of the principles of object-oriented
programming and design including software components,
object-oriented design patterns, and the use of graphical design
notations such as UML (unified modeling language). Basic concepts
in object-oriented design are illustrated with case studies in
application frameworks and by writing programs in one or more
object-oriented languages.

\paragraph{Course Description As Stated In Obsolete Course Charter:}
%
Presents a comparative approach to object-oriented
programming and design, with an emphasis on techniques for programming
large-scale systems. Discusses the concepts of object, class,
meta-class, message, method, inheritance, and genericity. Reviews
forms of polymorphism in object-oriented languages. Contrasts the use
of inheritance and composition as dual techniques for software reuse:
forwarding versus delegation and sub-classing versus
sub-typing. Studies main programming language tools and techniques to 
facilitate the creation of applications: modularity, encapsulation,
code reuse techniques, design patterns. Basic concepts in
object-oriented design are illustrated by writing programs in one or
more object-oriented languages.

\paragraph{Course Description As Stated in Online Course Charter:}
%
The course presents the theory and practice of object-oriented
programming. The course enhances students understanding of the
concepts of object, class, message, method, inheritance, and
genericity. The course covers a basic model for objects; the
principles of types and polymorphism in object-oriented programming
languages; different forms of abstraction; and theory and practice
of reuse. The course also introduces students to some
object-oriented design patterns that practitioners have found useful.

\paragraph{Proposed Course Description:}
%
Presents the theory and practice of object-oriented programming,
emphasizing abstractions and ideas that scale to large systems.
Introduces representation independence and related concepts such as
information hiding, encapsulation, modularity, and interchangeability
of parts.  Analyzes strengths and weaknesses of alternative
representations by considering asymptotic efficiency for the best,
worst, average, and amortized case.  Illustrates general paradigms
such as divide-and-conquer algorithms, precomputation, caching,
memoization, and dynamic programming.  Covers static types and
explains how subtyping and several forms of polymorphism
(parametric, ad hoc, and inclusion) support design patterns that
contribute to reuse and extensibility.

\paragraph{Prerequisites:}
%
CS~2510.

\paragraph{Textbooks:}
%
\noindent
Barbara Liskov and John Guttag.
\emph{Program Development in Java: Abstraction, Specification,
and Object-Oriented Design}.
Addison-Wesley, 2001.

\noindent
Peter Sestoft.  \emph{Java Precisely}, Second Edition.
MIT Press, 2005.

\paragraph{Topics Covered:}
%
Object-oriented programming:
%
classes and inheritance,
dynamic method dispatch,
class diagrams for inheritance and composition,
abstract classes and protocols (e.g. Java interfaces),
overriding versus overloading,
immutable versus mutable,
organized approaches to designing objects and classes,
refactoring.
%
Abstraction: 
%
the abstraction barrier,
client perspective versus implementor perspective,
specifications as contracts,
black-box testing against specifications,
substitutability of implementations,
representation independence,
information hiding,
encapsulation,
tradeoffs between information hiding, encapsulation, and reuse.
%
Algorithmic thinking:
%
asymptotic notation,
worst case, best case, average case,
amortized efficiency,
divide-and-conquer algorithms,
precomputation,
caching,
memoization and/or dynamic programming.
%
Static types:
%
compile-time versus run-time,
interactions with reuse,
subtyping,
ad hoc (sum) polymorphism, 
inclusion (product) polymorphism,
parametric polymorphism.
%
Basic design patterns:
%
Factory Method,
Singleton,
Wrapper/Adapter,
Composite,
Iterator,
Functional Visitor.

\paragraph{Course Outcomes:}
%
Upon completion of this course, a student should be able to: 
%
\begin{enumerate}
\item
List and describe fundamental mechanisms of object-oriented
programming, including inheritance, subtyping, polymorphism,
and dynamic method dispatch.
\item
Use at least one object-oriented programming language well
enough to implement abstract data types.
\item
Evaluate implementations of abstract data types with respect
to correctness, abstraction, efficiency, and code reuse.
%\item
%Transfer his or her programming knowledge to other OO languages and
%other programming paradigms.
\end{enumerate}

\paragraph{Measurement of Course Outcomes:}
%
Outcomes are measured and verified through: (1) regular programming
homeworks (five to ten recommended), which may combine to develop
a larger piece of software by the end of the semester; (2) two to four
major exams; (3) inclusion of student's work in an electronic portfolio.

\section{CS~3500 Sample Syllabus}

The theme of the course is representation independence and
its contribution to algorithmic creativity.

That theme presents a clean break from Fundies 2, in which
students are taught how to write code that mirrors the
structure of their data.  The representation-dependent
approach taken in Fundies 2 is appropriate for implementors
of an abstract data type, but is inappropriate for the ADT's
clients.

CS 3500 introduces the abstraction barrier that separates
clients from implementors.  Implementors must know how values
of an ADT are representated, but clients should not.  Making
client code independent of the representation makes it possible
for implementors to change representations without breaking
client code.  Representation independence allows implementors
to replace naive implementations by implementations that use
more sophisticated representations and algorithms.

CS 3500 should be taught using a programming langauge that
provides adequate support for class-based inheritance and for
information hiding.  Although Java has been used to teach most
sections of CS 3500, mainly because students are taught Java in
Fundies 2, some instructors have taught CS 3500 using Smalltalk
or Scala.  For concreteness, the sample syllabus below will
assume the language is Java.

\paragraph{MODULE 1:}
%
ADTs and Representation Independence (3 weeks)
%
\begin{itemize}
\item the abstraction barrier
\item ADTs, signatures, specifications
\item algebraic specifications of immutable ADTs
\item the Factory Method pattern hides representations from clients
\item UML class diagrams
\item the recipe representation
\item translation of algebraic specifications into idiomatic OO code
\item help clients out by throwing exceptions for unspecified cases
\item axiomatic specifications of equals(Object) and hashCode()
      may hide some representation details
\item substitution principle:
      all implementations that satisfy the specification are interchangeable
\item black-box testing against specification
\end{itemize}

\paragraph{MODULE 2:}
%
Taking Advantage of Representation Independence (4 weeks)
%
\begin{itemize}
\item parametric polymorphism
\item asymptotic notation (big-O, Omega, Theta)
\item best, average, and worst case
\item general rules for improving performance
\begin{itemize}
\item Don't.
\item Don't yet.
\item Don't optimize any more than necessary.
\end{itemize}
\item general paradigms of algorithm design
\begin{itemize}
\item Don't compute things you don't need to compute.
\item Don't recompute things you can avoid recomputing.
\begin{itemize}
\item divide and conquer
\item precomputation
\item caching
\item memoization and/or dynamic programming
\end{itemize}
\end{itemize}
\item refactoring transformations
\item invariants
\item more efficient representations
\end{itemize}

\paragraph{MODULE 3:}
%
Software Reuse (2 weeks)
%
\begin{itemize}
\item the Functional Visitor pattern
\item the Iterator pattern
\item mutable state versus concurrency
\item reuse by inheritance versus reuse by composition
\item protocols (which Java calls interfaces)
\item the fragile base class problem
\end{itemize}

\paragraph{MODULE 4:}
%
Static Types (2 weeks)
%
\begin{itemize}
\item safety properties and type errors
\item interaction with reuse
\item subtyping
\item sum and product types
\item ad hoc and inclusion polymorphism
\item bounded parametric polymorphism
\item covariant, contravariant, and monovariant type constructors
\end{itemize}

\end{document}

### 2017-12-01 2/admin/charter/CS3500_Report.tex
Move Fundamentals II content into own dir.

--- 2017-12-01 2/admin/charter/CS3500.tex
@@
 
-\title{CS~3500 Course Charter \& Syllabus}
+\title{Report from the CS3500 Committee}
 
-\author{Jay Aslam, Will Clinger, David Van Horn \\
+\author{Jay Aslam, Will Clinger, David Van Horn\\
         College of Computer and Information Science \\
@@
 
-\date{February 21, 2012}
+\date{DRAFT: May 27, 2012}
 
@@
 
-\section{CS~3500 Course Charter}
+The committee was charged with reviewing the CS3500 (\emph{Object-oriented Design}) curriculum, with the end goal of revising CS3500 in such a way that
+%
+\begin{itemize}
+\item it retains the best aspects of the current object-oriented design course,
+\item it does not excessively and unnecessarily overlap with CS2510 (\emph{Fundamentals~2}), and
+\item it contains substantially more algorithmic content.
+\end{itemize}
+%
+Specific concerns included:
+%
+\begin{enumerate}
+\item That CS2510 is teaching substantial amounts of object-oriented design.  Therefore, there is a concern that there is currently a fair amount of overlap between CS2510 and CS3500.
+\item That students have reported that in co-op interviews they are asked about algorithms and data structures, but do not know how to answer these questions.
+\item That CS3500 is serving, in part, as a review and consolidation course, due to the variability in how CS2510 is taught: Students entering CS3500 have mixed backgrounds in various programming topics, depending on the specific instance of CS2510 that they took, necessitating that CS3500 spend a fair amount of time bringing everyone up to speed.
+\end{enumerate}
+
+Given the above charge and specific concerns, the committee chose to review the three course introductory programming sequence (CS~2500 \emph{Fundamentals~1}, CS2510, and CS3500) in its entirety for content, redundancy, and the possible inclusion of algorithmic concepts.
+
+Our committee began by (1) studying the extant course charters and sample syllabi for the three course programming sequence, noting differences with the courses as currently taught, (2) identifying redundancies and inconsistencies in the programming sequence as currently taught, and (3) identifying those object-oriented programming concepts that could reasonably be moved from CS3500 to CS2510.
+
+The end result of eliminating inconsistencies and excessive redundancies, as well as moving a handful of programming concepts from CS3500 to CS2510 where natural and practicable, was to free up approximately three weeks of time within CS3500 that could be devoted to algorithmic concepts.
+Upon conferring with Deans Finkelstein and Rasala, it was decided that this would be sufficient for the time being.
+
+Two options were considered for adding algorithmic concepts to CS3500: (1) devoting a three week block to algorithms or (2) integrating algorithmic concepts throughout the course.
+It was decided that, pedagogically, Option~(2) was preferable to Option~(1).
+CS3500, as currently taught, already covers a number of topics in algorithms such as big-O notation and some advanced data structures (e.g., red-black trees), and there are natural opportunities to introduce other data structures and algorithmic concepts such as the divide-and-conquer, greedy, and dynamic programming paradigms.
+These additional algorithmic concepts should be introduced and taught, where appropriate, within the course.
+
+In what follows, we provide updated course charters and sample syllabi for CS2500, CS2510, and CS3500, reflecting our recommendations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% CS2500
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\pagebreak
+
+\section*{CS~2500 Course Charter}
+
+\paragraph{Course Title:}
+%
+CS~2500 \emph{Fundamentals of Computer Science 1} (4 SH)
+
+\paragraph{Course Description:}
+%
+Introduces the fundamentals of program design via study of increasingly
+complex data definitions: structures, containment, unions, self-referential,
+mutually referential. In addition, the course studies design via generative
+recursion, accumulator-style functions, and mutation of variables and
+structures. The course also introduces models of computations, which explain
+what programs do when they are applied to inputs.
+
+\paragraph{Prerequisites:}
+%
+The course requires no previous experience in writing computer programs, and
+only assumes mathematics at the level of high-school Algebra. Students are
+expected to have an interest in learning about Computer Science in depth.
+
+\paragraph{Textbooks:}
+%
+\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, 
+MIT Press, 2000.
+
+\paragraph{Topics Covered:}
+%
+Computers and the nature of computation. (Formal) models of computation.
+Functions; function definitions. Conditional computation structures. Structure
+definitions. Modeling information with data. The structure of data:
+composite, containment, union, self-referential, mutually referential,
+cyclically defined. Program design and design recipes: design based on the
+structure of data. Extensional equality of structured data. (Automated) testing.
+Accumulators. Designing abstractions. Generative recursion.
+
+\paragraph{Course Outcomes:}
+%
+Upon completion of this course, a student should be able to \emph{design programs},
+and, consequently, have some understanding of the fundamentals of computation.
+In particular, the student should be able to: Complete data analysis and
+definition for problems that require the use of composition, containment,
+union, and self-reference. Design programs to solve problems that consume and
+produce data structures based on composition, containment, union, and
+self-reference. Design test suites for these programs and evaluate the
+outcomes. Design simple abstractions and test the design. Reason about the
+performance characteristics of programs in a basic way.
+
+\paragraph{Measurement of Course Outcomes:}
+%
+Outcomes are measured and verified through: Weekly programming homeworks.
+Daily short paper-and-pencil quizzes. Two to four major exams. Inclusion of
+student's work in an electronic portfolio.
+
+\pagebreak
+
+\section*{CS~2500 Sample Syllabus}
+
+The course sequences through the following topics:
+\begin{itemize}
+\item overview, intro to simple arithmetic calculation
+\item atomic \& compound data: 
+  \begin{itemize}
+    \item integers
+    \item booleans \& conditional computation
+    \item records
+  \end{itemize}
+\item representation: information vs. data
+\item data-driven programming
+\item testing
+\item unions
+\item self-referential unions 1---recursion
+\item self-referential unions 2---list processing
+\item designing larger programs: top-down design; iterative refinement
+\item recursion in all forms: trees, s-expressions, forests
+\item abstraction: interfaces and implementations
+\item abstraction: parametric data definitions and contracts
+\item notions of equality
+\item local name scope; simple asymptotic complexity
+\item abstracting control with higher-order functions
+\item constructing higher-order functions with lambda
+\item infinite data structures and computations
+\item generative recursion: root finding, sorting
+\item generative recursion: graphs \& search; accumulators
+\item review and wrap up
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% CS2510
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\pagebreak
+
+\section*{CS~2510 Course Charter}
+
+\paragraph{Course Title:}
+%
+CS~2510 \emph{Fundamentals of Computer Science 2}	(4 SH)
+
+\paragraph{Course Description:}
+%
+
+Introduces the fundamentals of the design of class hierarchies in
+object-oriented programming with emphasis on representation-dependent
+program design with objects. The course also covers basic reasoning
+techniques concerning the correctness and algorithmic behavior of
+computations.
+
+\paragraph{Prerequisites:}
+%
+CS~2500. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2510 or concurrently.
+
+\paragraph{Textbooks:}
+%
+\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000. \emph{How to Design Classes}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.
+
+\paragraph{Topics Covered:}
+%
+Objects, classes, class hierarchies.  Data definitions and methods.
+%
+Class diagrams.
+%
+Interfaces, abstract classes, subclasses.  Types.  Polymorphism.
+%
+Accumulators, iterators, loops.
+%
+Function objects, comparators, visitors.  Sorting: quick sort, merge
+sort, heap sort.  Notions of equality.
+%
+Java collections library.
+%
+Overriding {\tt equals} and {\tt hashCode}.
+%
+Testing: unit testing, stress testing, and property-based random testing.
+%
+Binary trees, binary search, heaps, hash tables.
+%
+Invariants.
+%
+Graphs, graph traversal, generative recursion.  Depth-first,
+breadth-first, and priority-queue based search.
+%
+Design patterns.
+%
+Mutation and stateful programming.
+%
+Indexed data structures.
+
+\paragraph{Course Outcomes:}
+%
+Upon completion of this course, a student should be able to: 
+(1) Design, implement, and test class hierarchies with appropriate methods. 
+(2) Read the documentation of class libraries and write programs that use them appropriately. 
+(3) Understand the algorithmic behavior of individual methods.
+
+\paragraph{Measurement of Course Outcomes:}
+%
+Weekly programming homeworks. Final project. Weekly short paper and pencil quizzes. Two to four major exams. Inclusion of student's work in an electronic portfolio.
+
+\pagebreak
+
+\section*{CS~2510 Sample Syllabus}
+
+The course studies the class-based program design and the design of abstractions that support the design of reusable software and libraries.
+It covers the principles of object oriented program design, the basic rules of program evaluation, the use of abstractions in support of the design of reusable code, and examines the relationship between algorithms and data structures, as well as basic techniques for analyzing algorithm complexity.
+
+The course is suitable for both CS majors and non-majors. It assumes that student has been introduced to the basic principles of program design and computation.
+
+\paragraph{Course Objectives:}
+%
+The goal is to help the student understand the principles of class
+based program design using an object-oriented programming
+language. Java is used to illustrate how the program design principles
+are used in practical applications. Using a concrete language also
+presents an opportunity to discuss the strengths and weaknesses of
+languages and language paradigms.
+
+\paragraph{MODULE 1:}
+%
+Review of Fundies 1 in the context of a class-based statically typed language (2 weeks)
+%
+\begin{itemize}
+\item Review of the program design and abstractions in functional style;
+  design recipe, abstractions, loops with accumulators
+\item Defining classes and class hierarchies to represent complex information:
+  simple classes, classes with fields that are instances of another class, 
+  unions of classes, self-referential and mutually referential class   
+  hierarchies: lists, trees, graphs
+\item Designing methods for classes: static type system, dynamic dispatch,
+  delegation (chain of responsibility)
+\item Design Recipe for data definitions, for method definitions
+\end{itemize}
+
+\paragraph{MODULE 2:}
+%
+Designing Abstractions (3 weeks)
+%
+\begin{itemize}
+\item Review of the Design Recipe for Abstractions
+\item Abstracting over the data: abstract classes; subclasses
+\item Abstracting over the functional behavior: function objects, functional iterator
+\item Understanding equality of complex data
+\end{itemize}
+
+\paragraph{MODULE 3:}
+%
+Stateful Programming (3 weeks)
+%
+\begin{itemize}
+\item Designing methods with effects; Designing tests for effects
+\item Understanding equality in the presence of mutation
+\item Direct access data structures: (e.g. Java ArrayList)
+\item Loops
+\end{itemize}
+
+\paragraph{MODULE 4:}
+%
+Abstracting over the Data Type; Designing and using libraries (6 weeks)
+%
+\begin{itemize}
+\item Polymorphism: parametrized classes, methods
+\item Abstract Data Type --- stack, queue, list, set
+\item Abstract Data Type --- map, hashmap
+\item HashCode method and overriding equals method
+\item Algorithm complexity --- estimates; Stress tests
+\item Java Collection Framework
+\end{itemize}
+
+\paragraph{Notes:}
+%
+\begin{itemize}
+
+\item A number of algorithms (filter, andmap, ormap, linear search, binary search, binary search trees, insertion sort, selection sort, quicksort, heap-based priority queue, heapsort, table lookup, use of hashmap, and some graph traversal algorithms (BFS, DFS, and possibly shortest path) are used to illustrate the program design, the need for appropriate data representation, and the need for understanding the complexity of algorithms.
+
+\item Stress tests are used to reinforce the need for understanding the algorithm complexity.
+
+\item Data structures (stacks, queues, lists, trees, arraylist, hashmap, lookup tables) are presented in the context of problems that motivate their use.
+
+\item The goal is to make sure the students begin to understand the design of class based libraries, know how to look up the needed information, and can begin using libraries in their programs.
+
+\item A comprehensive final project challenges the students to apply the design principles and the knowledge of the libraries to a creative larger program with up to 1000 LOC.
+
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% CS3500
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\pagebreak
+
+\section*{CS~3500 Course Charter}
 
@@
 
-\paragraph{Current Catalog Description:}
-%
-Presents a comparative approach to object-oriented programming
-and design. Discusses the concepts of object, class, meta-class,
-message, method, inheritance, and genericity. Reviews forms of
-polymorphism in object-oriented languages. Contrasts the use of
-inheritance and composition as dual techniques for software reuse:
-forwarding vs. delegation and subclassing vs. subtyping. Fosters
-a deeper understanding of the principles of object-oriented
-programming and design including software components,
-object-oriented design patterns, and the use of graphical design
-notations such as UML (unified modeling language). Basic concepts
-in object-oriented design are illustrated with case studies in
-application frameworks and by writing programs in one or more
-object-oriented languages.
-
-\paragraph{Course Description As Stated In Obsolete Course Charter:}
-%
-Presents a comparative approach to object-oriented
-programming and design, with an emphasis on techniques for programming
-large-scale systems. Discusses the concepts of object, class,
-meta-class, message, method, inheritance, and genericity. Reviews
-forms of polymorphism in object-oriented languages. Contrasts the use
-of inheritance and composition as dual techniques for software reuse:
-forwarding versus delegation and sub-classing versus
-sub-typing. Studies main programming language tools and techniques to 
-facilitate the creation of applications: modularity, encapsulation,
-code reuse techniques, design patterns. Basic concepts in
-object-oriented design are illustrated by writing programs in one or
-more object-oriented languages.
-
-\paragraph{Course Description As Stated in Online Course Charter:}
-%
-The course presents the theory and practice of object-oriented
-programming. The course enhances students understanding of the
-concepts of object, class, message, method, inheritance, and
-genericity. The course covers a basic model for objects; the
-principles of types and polymorphism in object-oriented programming
-languages; different forms of abstraction; and theory and practice
-of reuse. The course also introduces students to some
-object-oriented design patterns that practitioners have found useful.
-
-\paragraph{Proposed Course Description:}
+\paragraph{Course Description:}
 %
@@
 
-\section{CS~3500 Sample Syllabus}
+\pagebreak
+
+\paragraph{For reference\ldots}
+
+\paragraph{Current Catalog Description:}
+%
+Presents a comparative approach to object-oriented programming
+and design. Discusses the concepts of object, class, meta-class,
+message, method, inheritance, and genericity. Reviews forms of
+polymorphism in object-oriented languages. Contrasts the use of
+inheritance and composition as dual techniques for software reuse:
+forwarding vs. delegation and subclassing vs. subtyping. Fosters
+a deeper understanding of the principles of object-oriented
+programming and design including software components,
+object-oriented design patterns, and the use of graphical design
+notations such as UML (unified modeling language). Basic concepts
+in object-oriented design are illustrated with case studies in
+application frameworks and by writing programs in one or more
+object-oriented languages.
+
+\paragraph{Course Description As Stated In Obsolete Course Charter:}
+%
+Presents a comparative approach to object-oriented
+programming and design, with an emphasis on techniques for programming
+large-scale systems. Discusses the concepts of object, class,
+meta-class, message, method, inheritance, and genericity. Reviews
+forms of polymorphism in object-oriented languages. Contrasts the use
+of inheritance and composition as dual techniques for software reuse:
+forwarding versus delegation and sub-classing versus
+sub-typing. Studies main programming language tools and techniques to 
+facilitate the creation of applications: modularity, encapsulation,
+code reuse techniques, design patterns. Basic concepts in
+object-oriented design are illustrated by writing programs in one or
+more object-oriented languages.
+
+\paragraph{Course Description As Stated in Online Course Charter:}
+%
+The course presents the theory and practice of object-oriented
+programming. The course enhances students understanding of the
+concepts of object, class, message, method, inheritance, and
+genericity. The course covers a basic model for objects; the
+principles of types and polymorphism in object-oriented programming
+languages; different forms of abstraction; and theory and practice
+of reuse. The course also introduces students to some
+object-oriented design patterns that practitioners have found useful.
+
+\pagebreak
+
+\section*{CS~3500 Sample Syllabus}
 

### 2017-12-01 2/exams/exam4/2.tex
Move Fundamentals II content into own dir.

--- 2012-03-27 exam4/2.tex
@@
 \item Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
-\verb|visit| method in all of your classes from the previous problem.
+\verb|accept| method in all of your classes from the previous problem.
 Make sure you design the interfaces and methods so that a visitor can
@@
 
+\begin{solution}
+\begin{verbatim}
+public boolean same(UnPair<X> p) {
+  return (this.left.equals(p.left) &&
+          this.right.equals(p.right))
+      || (this.left.equals(p.right) &&
+          this.right.equals(p.left));
+}
+
+{5,7}.equals({5,7}) --> true
+{5,7}.equals({7,5}) --> true
+{5,5}.equals({5,7}) --> false
+\end{verbatim}
+\end{solution}
+
 \newpage
@@
 {\tt hashCode}.
+
+\begin{solution}
+\begin{verbatim}
+It is not valid.
+
+Counterexample:
+{5,7}.equals({7,5}) but 5.hashCode() does not equal
+7.hashCode()
+\end{verbatim}
+\end{solution}
 \end{enumerate}
@@
 {\tt setTail} on an empty list, you may raise an exception.)
+
+\begin{solution}
+\begin{verbatim}
+interface List<X> {
+    void setTail(List<X> ls);
+    void setTailAcc(List<X> ls, Cons<X> last);
+}
+
+class MT<X> implements List<X> {
+    public void setTail(List<X> ls) {
+        throw new RuntimeException("setTail on MT");
+    }
+    public void setTailAcc(List<X> ls, Cons<X> last) {
+        last.rest = ls;
+    }
+}
+
+class Cons<X> implements List<X> {
+    X first;
+    List<X> rest;
+    Cons(X first, List<X> rest) {
+        this.first = first;
+        this.rest = rest;
+    }
+    public void setTail(List<X> ls) {
+        this.rest.setTailAcc(ls, this);
+    }
+    public void setTailAcc(List<X> ls, Cons<X> last) {
+        this.rest.setTailAcc(ls, this);
+    }
+}
+
+// Plus tests
+\end{verbatim}
+\end{solution}
 

### 2017-12-01 2/web/12.sp.reg/exams/exam1bx.tex
Move Fundamentals II content into own dir.

\documentclass[11pt]{article}

%% ------------------------------------------------------------------
%% SOLUTIONS:
\def\thel{\noindent\rule{2.5cm}{.5pt}}
\long\def\begsol#1 #2\endsol{#1}
%\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
%%uncomment the above line to get the solutions printed

\newcommand\code[1]{\texttt{#1}}
\newcommand\bcode[1]{\texttt{\textbf{#1}}}

%% PROBLEMS:
\def\pts#1{\marginpar{\footnotesize \raggedright  \fbox{#1 {\sc Points}}}}
\newcounter{Pctr}
\newenvironment{problem}{\stepcounter{Pctr}%
\begin{description}
\item[\noindent{\bf Problem} \arabic{Pctr}] 
\end{description}}{\relax}
%% ------------------------------------------------------------------

%% SUBPARTS:
\newcounter{parts}
\renewcommand{\theparts}{\Alph}
%% ------------------------------------------------------------------

\begin{document}

\renewcommand{\theenumi}{\Alph{enumi}}
\setcounter{Pctr}{0}

%% ------------------------------------------------------------------

\vfill
\centerline{\Large CS 2510 Exam 1 -- Spring 2012}

~\\[2cm]

\begin{center}
\begin{tabular}{l@{\qquad}l}
Name:                        & \rule{200pt}{.1pt} \\[.5cm]
Student Id (last 4 digits):  & \rule{200pt}{.1pt} \\[.5cm]
\end{tabular}
\end{center}

\noindent\begin{minipage}{7.5cm} $\bullet$ Write down the answers in the
space provided.

$\bullet$ You may use all syntax that you know from \textit{FunJava}
(that is, the parts of Java we have studied in class),
although there are several features you will \emph{not} need.

$\bullet$ When defining methods, you do not need to give a complete
class definition---just indicate in which class your method definition
should be placed.

$\bullet$ For tests you only need to provide the expression that
computes the actual value, connecting it with an arrow to the expected
value. For example \code{s.method() -> true} is sufficient.

$\bullet$ Remember that the phrase ``design a class'' or ``design a
method'' means more than just providing a definition. It means to
design them according to the \textbf{design recipe}.  You are
\textit{not} required to provide a method template unless the problem
specifically asks for one.  However, be prepared to struggle if you
choose to skip the template step.

$\bullet$ We will not answer \textit{any} questions during the exam.

\bigskip

\textit{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{4.5cm}
\begin{tabular}{|c|l|@/r|}
\hline
\textbf{Problem} & Points & \\ \hline
A & & 3\\ \hline
B & & 9\\ \hline
C & & 9\\ \hline
D & & 9\\ \hline
E & & 4\\ \hline
\textbf{Total} &  & 34 \\ \hline
\end{tabular}
\end{minipage}

\vfill\thispagestyle{empty}
\newpage

%% -----------------------------------------------------------------------------
%% 
\pts{34}
\begin{problem}  Tonight is another meeting of the ACM, and they
  always serve pizza. So, we want to know a bit about what they serve.

Here is a class diagram that describes a pizza:

\begin{verbatim}
                    +--------------+
                    |    IPizza    |<-------------------+
                    +--------------+                    |
                     / \   / \   / \                    |
                     ---   ---   ---                    |
                      |     |     |                     |
        +-------------+     |     +-----------+         |
        |                   |                 |         |
        |                   |                 |         |
+-------------+      +-------------+   +-------------+  |
| Crust       |<--+  | Cheese      |   | Topping     |  |
+-------------+   |  +-------------+   +-------------+  |
| String name |   |  | String kind |   | IPizza base |--+ 
+-------------+   +--| Crust crust |   +-------------+
                     +-------------+
\end{verbatim}

(It may seem strange, but the toppings have no names.)

\begin{enumerate}

\newpage
%-----------------------------------------------
\item {\em{(3 points)}}

 Make examples of data (i.e., instances) for this class
    hierarchy. Include an example of a thin crust pizza with mozarella
    cheese, and three toppings.

\begsol{\vspace{0.5in}}
[POINTS 3: 1 point for the \texttt{Crust} example, 1 point for the a 
\texttt{Cheese}, 1 point for the the specified example of
\texttt{Topping}]

%{\vspace{0.5in}}
\begin{verbatim}
IBang p0 = new Pop(0);
IBang p1 = new Pop(1);
IBang p2 = new Pop(2);
IBang p3 = new Pop(3);

IBang b0 = new Fizz(4, this.p0, this.p0);
IBang b1 = new Fizz(4, this.p2, this.p3);
IBang b1t1 = new Fizz(5, this.p3, this.p3);
IBang b1t2 = new Fizz(6, this.p3, this.b0);
IBang b1t7 = new Fizz(7, this.b0, this.b0);
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(9 points)}}

Design a method, \code{weight}, that computes the weight of the pizza
represented by a \code{IPizza}.

      The weight of \code{crust} is always \code{5}.  The weight 
      of a \code{Cheese} is \code{3}. Each
      topping contributes one to the total size.

\begsol{\vspace{0.5in}}
[POINTS 9: 1 point for the purpose and header in the interface
\texttt{IGift}; in the class \texttt{Jewel}: 1 point for the body, 1
point for tests; in the class \texttt{Box}: 1 point for the body, 1
point for tests; in the class \texttt{Wrap}: 2 
points for the body, 2 points for examples]

%{\vspace{0.5in}}
\begin{verbatim}
// in the interface IBang:

// compute the age of this pop
int howOld();

// in the class Pop:
// compute the age of this pop
int howOld(){
  return this.age;
}

// in the class Fizz:
// compute the age of this fizz
int howOld(){
  return this.age + 
         Math.max(this.bang1.howOld(), 
                  this.bang2.howOld());
}

// in the Examples class
// test the method howOld for the IBang classes
boolean testHowOld(Tester t){
  return
  t.checkExpect(this.p0.howOld(), 0) &&
  t.checkExpect(this.p3.howOld(), 3) &&
  t.checkExpect(this.b1.howOld(), 7) &&
  t.checkExpect(this.b1t1.howOld(), 8) &&
  t.checkExpect(this.b1t7.howOld(), 11);
}	
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(9 points)}}

Design the method \code{replaceCheese} that produces an
    \code{IPizza} with the same toppings as this one,
    having the same crust but with the cheese replaced by the
    cheese of the given kind.

\begsol{\vspace{0.5in}}
[POINTS 9: 1 point for the purpose and header in the interface
\texttt{IGift}; in the class \texttt{Jewel}: 1 point for the body, 1
point for tests; in the class \texttt{Box}: 1 point for the body, 1
point for tests; in the class \texttt{Wrap}: 2 
points for the body, 2 points for examples]

%{\vspace{0.5in}}
\begin{verbatim}
// in the interface IBang:

// produce a new bang from this Bang after one tick
IBang nextYear();

// in the class Pop:
// produce a new bang from this Pop after one tick
IBang nextYear(){
  if (this.age == 3){
    return new Fizz(4, new Pop(0), new Pop(0));
  } else {
    return new Pop(this.age + 1);
  }
}

// in the class Fizz:
// produce a new bang from this Fizz after one tick
IBang nextYear(){
  if (this.bang2.howOld() > this.bang1.howOld()){
    return new Fizz(this.age + 1, 
             this.bang1.nextYear(), this.bang2);
  } else {
    return new Fizz(this.age + 1, 
             this.bang1, this.bang2.nextYear());
  }
}

// in the Examples class:
// test the method nextYear for the IBang classes
boolean testNextYear(Tester t){
  return
  t.checkExpect(this.p0.nextYear(), this.p1) &&
  t.checkExpect(this.p3.nextYear(), this.b0) &&
  t.checkExpect(this.b1.nextYear(), this.b1t1) &&
  t.checkExpect(this.b1t1.nextYear(), this.b1t2) &&
 t.checkExpect(this.b1t2.nextYear(), this.b1t7);
}
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
\item {\em{(9 points)}}

Design the method \code{sameCheeseCrust} that checks whether this pizza
has the crust with the given name and the cheese of the given kind.

\begsol{\vspace{0.5in}}
[POINTS 9: 1 point for the purpose and header in the interface
\texttt{IGift}; in the class \texttt{Jewel}: 1 point for the body, 1
point for tests; in the class \texttt{Box}: 1 point for the body, 1
point for tests; in the class \texttt{Wrap}: 2 
points for the body, 2 points for examples]

%{\vspace{0.5in}}
\begin{verbatim}
// in the interface IBang:

// produce a new bang from this Bang after one tick
IBang nextYear();

// in the class Pop:
// produce a new bang from this Pop after one tick
IBang nextYear(){
  if (this.age == 3){
    return new Fizz(4, new Pop(0), new Pop(0));
  } else {
    return new Pop(this.age + 1);
  }
}

// in the class Fizz:
// produce a new bang from this Fizz after one tick
IBang nextYear(){
  if (this.bang2.howOld() > this.bang1.howOld()){
    return new Fizz(this.age + 1, 
             this.bang1.nextYear(), this.bang2);
  } else {
    return new Fizz(this.age + 1, 
             this.bang1, this.bang2.nextYear());
  }
}

// in the Examples class:
// test the method nextYear for the IBang classes
boolean testNextYear(Tester t){
  return
  t.checkExpect(this.p0.nextYear(), this.p1) &&
  t.checkExpect(this.p3.nextYear(), this.b0) &&
  t.checkExpect(this.b1.nextYear(), this.b1t1) &&
  t.checkExpect(this.b1t1.nextYear(), this.b1t2) &&
 t.checkExpect(this.b1t2.nextYear(), this.b1t7);
}
\end{verbatim}
\endsol

\newpage
%-----------------------------------------------
{\center{... This page intentionally left blank ...}}

\newpage
%-----------------------------------------------
\item {\em{(4 points)}}

Show the resulting templates for the \code{Cheese} class.

\begsol{\vspace{0.5in}}
[POINTS 4:
  Pop -- field and two methods with types  -- 1 point;
  Bing -- fields with types                -- 1 point;
          methods defined here with types  -- 1 point;
          methods for fields with types    -- 1 point]]

%{\vspace{0.5in}}
\begin{verbatim}
// in the class Pop:
/* Template:
   Fields:
   ... this.age ...     -- int

   Methods:
   ... this.howOld() ...   -- int
   ... this.nextYear() ... -- IBang
*/

// in the class Fizz:
/* Template:
   Fields:
   ... this.age ...     -- int
   ... this.bang1 ...     -- IBang
   ... this.bang2 ...     -- IBang

   Methods:
   ... this.howOld() ...   -- int
   ... this.nextYear() ... -- IBang

   Methods for fields:
   ... this.bang1.howOld() ...   -- int
   ... this.bang2.howOld() ...   -- int
   ... this.bang1.nextYear() ... -- IBang
   ... this.bang2.nextYear() ... -- IBang   
 */	
\end{verbatim}
\endsol

\end{enumerate}
\end{problem}
\end{document}

### 2017-12-01 2/web/12.su.reg/exams/exam1a.tex
Move Fundamentals II content into own dir.

--- 2017-12-01 2/web/12.sp.reg/exams/exam1bx.tex
@@
 $\bullet$ You may use all syntax that you know from \textit{FunJava}
-(that is, the parts of Java we have studied in class),
-although there are several features you will \emph{not} need.
+(that is, the parts of Java we have studied in class), although there
+are several features you will \emph{not} need.
 
@@
 \pts{34}
-\begin{problem}  Tonight is another meeting of the ACM, and they
-  always serve pizza. So, we want to know a bit about what they serve.
+\begin{problem}  Valentine's day is coming, so here is a nice gift you
+  can give or receive. Of course, it needs to be in a nice box, and
+  wrapped carefully.
 
-Here is a class diagram that describes a pizza:
+Here is a class diagram that describes this gift:
 
@@
                      +--------------+
-                    |    IPizza    |<-------------------+
+                     |     IGift    |<------------------+
                      +--------------+                   |
@@
         |                   |                 |         |
-+-------------+      +-------------+   +-------------+  |
-| Crust       |<--+  | Cheese      |   | Topping     |  |
-+-------------+   |  +-------------+   +-------------+  |
-| String name |   |  | String kind |   | IPizza base |--+ 
-+-------------+   +--| Crust crust |   +-------------+
-                     +-------------+
+ +-------------+      +--------------+  +------------+  |
+ | Jewel       |<--+  | Box          |  | Wrap       |  |
+ +-------------+   |  +--------------+  +------------+  |
+ | String name |   |  | String kind  |  | IGift gift |--+ 
+ +-------------+   +--| Jewel inside |  +------------+
+                      +--------------+ 
 \end{verbatim}
-
-(It may seem strange, but the toppings have no names.)
 
@@
  Make examples of data (i.e., instances) for this class
-    hierarchy. Include an example of a thin crust pizza with mozarella
-    cheese, and three toppings.
+    hierarchy. Include an example of a ruby in a red box with three
+    wrappings around it.
 
 \begsol{\vspace{0.5in}}
-[POINTS 3: 1 point for the \texttt{Crust} example, 1 point for the a 
-\texttt{Cheese}, 1 point for the the specified example of
-\texttt{Topping}]
+[POINTS 3: 1 point for the \texttt{Jewel} example, 1 point for the a 
+\texttt{Box}, 1 point for the the specified example of \texttt{Wrap}]
 
@@
 
-Design a method, \code{weight}, that computes the weight of the pizza
-represented by a \code{IPizza}.
+Design a method, \code{size}, that computes the size of the gift
+represented by a \code{IGift}.
 
-      The weight of \code{crust} is always \code{5}.  The weight 
-      of a \code{Cheese} is \code{3}. Each
-      topping contributes one to the total size.
+      The size of \code{Jewel} is always \code{5}.  The size
+      of a \code{Box} is \code{8} regardless of what it contains. Each
+      wrap contributes one to the total size.
 
@@
 
-Design the method \code{replaceCheese} that produces an
-    \code{IPizza} with the same toppings as this one,
-    having the same crust but with the cheese replaced by the
-    cheese of the given kind.
+Design the method \code{replaceBox} that produces an
+    \code{IGift} wrapped in the same wrappings as this one,
+    containing the same jewel  inside but with the box replaced by the
+    box of the given kind.
 
@@
 
-Design the method \code{sameCheeseCrust} that checks whether this pizza
-has the crust with the given name and the cheese of the given kind.
+Design the method \code{sameJewelBox} that checks whether this gift
+contains the jewel with the given name and is inside of the box of the
+given kind.
 
@@
 
-Show the resulting templates for the \code{Cheese} class.
+Show the resulting templates for the \code{Box} class.
 

### 2017-12-01 2/web/12.su.reg/exams/exam4.su.12.tex
Move Fundamentals II content into own dir.

--- 2012-05-31 web/12.su.reg/exams/exam4.su.12.tex
@@
 An association list is a way of representing an association between
-values.  For example, you might you use an association list to
+values.  For example, you might use an association list to
 associate students with their grades, or battleships with their

### 2017-12-01 2/web/12.su.reg/exams/exam5.su.12.tex
Move Fundamentals II content into own dir.

--- 2012-05-17 web/12.su.reg/exams/exam2.tex
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 5 -- Summer 2012}
 
@@
 %% 
-\pts{30}
+\pts{45}
 \begin{problem}
+Develop an implementation of {\tt Comparator<Posn>} that orders
+Cartesian-coordinates by their distance from the origin in such a way
+that the further away from the origin, the smaller the position is
+considered.  So for example, there is no position larger than (0,0).
 
-Binary trees are one of the most widely used data structures in
-computer science.  Here are three examples of binary trees of numbers:
+You may rely on the following definition of {\tt Posn}:
 
 \begin{verbatim}
-         5        1           3
-                   \         / \
-                    2       4  18
-                              /  \
-                             9    12
-                            /    /
-                           7    3
+// Represents (x,y) in Cartesian-coordinate system
+class Posn {
+  Integer x;
+  Integer y;
+  Posn(Integer x, Integer y) {
+    this.x = x;
+    this.y = y;
+  }
+}
 \end{verbatim}
+\end{problem}
 
-Nodes in a binary tree carry data, which in this case are numbers, and
-a node has a left and right subtree.  Binary trees, like lists, may
-contain \emph{any} number of elements, including 5, 8, 19, and 0.
-
-\begin{enumerate}
-\item Design a data definition for binary trees.  It must adequately
-  represent any possible binary tree.
-
-\item Give class diagrams corresponding to the data definition from
-  part A.
-
-\item Design a {\tt mirror} method that computes the mirror image of a
-  binary tree.  For example, the mirror image of the above trees would
-  be:
+\newpage
+\begin{problem}
+Design a method:
+\begin{center}
+{\tt <T> Boolean isSorted(ArrayList<T> ls, Comparator<T> c)}
+\end{center}
+that determines if the given array list is in ascending sorted
+order according to the given comparator.
+\end{problem}
+\newpage
+\begin{problem}
+Here is a data definition for pairs:
 
 \begin{verbatim}
-     5        1           3
-             /           / \
-            2           18  4
-                       /  \
-                     12    9
-                       \    \
-                        3    7
+// Represents a pair of A and B.
+class Pair<A,B> {
+  A left;
+  B right;
+  Pair(A left, B right) {
+    this.left = left;
+    this.right = right;
+  }
+}
 \end{verbatim}
-\end{enumerate}
-
+Design a method:
+\begin{verbatim}
+<A,B> Comparator<Pair<A,B>> lexi(Comparator<A> ca,
+                                 Comparator<B> cb)
+\end{verbatim}
+that consumes a comparator for {\tt A}s and a comparator for {\tt B}s
+and produces a comparator for pairs of {\tt A}s and {\tt B}s that is
+the \emph{lexicographic order}.  Mathematically speaking, the
+lexicographic order of a pair is defined as follows:
+$$
+(a,b) \leq (a',b')\text{ if and only if }a < a'\text{ or }(a = a'\text{ and }b \leq b')
+$$
 \end{problem}
+\ 
+\newpage
+\ 
 \newpage

### 2017-12-01 2/web/12.su.reg/exams/exam6.su.12.tex
Move Fundamentals II content into own dir.

--- 2012-05-31 web/12.su.reg/exams/exam4.su.12.tex
@@
 \vfill
-\centerline{\Large CS 2510 Exam 4 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 6 -- Summer 2012}
 
@@
 %% 
+\begin{center}
+A useful utility
+\end{center}
+You may assume, \textbf{for the purposes of examples only}, the existence of a
+class {\tt Iter<X>} that implements {\tt Iterable<X>}.  It has a
+somewhat magical constructor in that it can take \emph{any} number of
+{\tt X}s as arguments.  Its {\tt iterator} method produces an iterator
+that iterates through those arguments.
+
+For example:
+\begin{verbatim}
+new Iter<Integer>().iterator().hasNext()  --> false
+new Iter<Integer>(1).iterator().hasNext() --> true
+new Iter<Integer>(1).iterator().next()    --> 1
+Iterator<Integer> i =
+  new Iter<Integer>(1,2,3).iterator();
+i.next() --> 1
+i.next() --> 2
+i.next() --> 3
+\end{verbatim}
+\newpage
 \pts{45}
 \begin{problem}
-An association list is a way of representing an association between
-values.  For example, you might you use an association list to
-associate students with their grades, or battleships with their
-position, or bank account numbers with their account balances.
+Design a method
+\begin{verbatim}
+    Double avg(Iterable<Double> nums)
+\end{verbatim}
+that computes the average of a series of numbers.  It should raise an
+exception if the series is empty.
+\end{problem}
+\newpage
+\begin{problem}
+Design a class
+\begin{verbatim}
+   class SumSquares implements Iterable<Integer>
+\end{verbatim}
+which is given an {\tt Iterable<Integer>} when constructed and whose
+{\tt iterator} method produces an iterator that produces the sum of
+the squares the given iterator produces.
 
-The representation of an association list is much like a typical list,
-except that every ``cons'' has \emph{two} elements and the rest of the
-list.  One element is a ``key'', the other is its association.
+For example, if given a {\tt new Nat()}, the iterable you wrote in lab
+that produces the natural numbers {\tt 0, 1, 2, 3, 4}, \dots, then
+iterator should produce {\tt 0, 1, 5, 14, 30}, etc.
 
-Here is a data definition for association lists parameterized over the
-types of keys and their values:
+\end{problem}
+\newpage
+\begin{problem}
+Design a class
+\begin{verbatim}
+   class Zip<X,Y> implements Iterable<Pair<X,Y>>
+\end{verbatim}
+which is given an {\tt Iterable<X>} and an {\tt Iterable<Y>} when
+constructed and whose {\tt iterator} method produces an iterator
+that produces pairs of values: one from the {\tt X} series, one from
+the {\tt Y} series.  The iterator should have no more elements
+whenever either of the given iterators have no more elements.
 
+You may rely on the following definition of {\tt Pair}:
 \begin{verbatim}
-interface Assoc<K,V> {}
-
-class EmptyAssoc<K,V> implements Assoc<K,V> {}
-
-class ConsAssoc<K,V> implements Assoc<K,V> {
-  K key;
-  V val;
-  Assoc<K,V> rest;
-  ConsAssoc(K key, V val, Assoc<K,V> rest) {
-    this.key = key;
-    this.val = val;
-    this.rest = rest;
+   class Pair<X,Y> {
+       X left;
+       Y right;
+       Pair(X left, Y right) {
+           this.left = left;
+           this.right = right;
        }
@@
 
-\begin{enumerate}
-\item Develop the visitor pattern for association lists.
-\item Design an implementation of an association list visitor that
-  looks up the (first) value associated to a given key.  If the key is
-  not in the list, the visitor shoud throw an exception.
-\item Using the double dispatch approach, design a {\tt same} method
-  that determines if this association list is structurally equal to a
-  given association list.
-\end{enumerate}
 \end{problem}

### 2017-12-01 2/web/12.su.reg/exams/exam7.su.12.tex
Move Fundamentals II content into own dir.

--- 2012-05-17 web/12.su.reg/exams/exam2.tex
@@
 \vfill
-\centerline{\Large CS 2510 Exam 2 -- Summer 2012}
+\centerline{\Large CS 2510 Exam 7 -- Summer 2012}
 
@@
 %% 
-\pts{30}
+\pts{45}
 \begin{problem}
 
-Binary trees are one of the most widely used data structures in
-computer science.  Here are three examples of binary trees of numbers:
-
-\begin{verbatim}
-         5        1           3
-                   \         / \
-                    2       4  18
-                              /  \
-                             9    12
-                            /    /
-                           7    3
-\end{verbatim}
-
-Nodes in a binary tree carry data, which in this case are numbers, and
-a node has a left and right subtree.  Binary trees, like lists, may
-contain \emph{any} number of elements, including 5, 8, 19, and 0.
+A matryoshka doll, or a Russian nesting doll, refers to a set of
+wooden dolls of decreasing size placed one inside the other.  The
+innermost doll is usually a baby, and thus we can describe matryoshka
+dolls as either a baby, or a layer with a matryoshka doll inside.
 
 \begin{enumerate}
-\item Design a data definition for binary trees.  It must adequately
-  represent any possible binary tree.
-
-\item Give class diagrams corresponding to the data definition from
-  part A.
-
-\item Design a {\tt mirror} method that computes the mirror image of a
-  binary tree.  For example, the mirror image of the above trees would
-  be:
-
-\begin{verbatim}
-     5        1           3
-             /           / \
-            2           18  4
-                       /  \
-                     12    9
-                       \    \
-                        3    7
-\end{verbatim}
+\item Design a data definition for matryoshka dolls.
+\item Draw the class diagram corresponding to your data definition.
+\item Develop the visitor pattern for matryoshka dolls.
+\item Design a visitor for matryoshka dolls that computes the nesting
+  depth of the doll.  The nesting depth of the baby is 0.
+\item Have a great summer.
 \end{enumerate}
-
 \end{problem}
@@
 \
+\newpage
+\
 \end{document}

### 2017-12-02 1/exams/m1.tex

@@
 \usepackage{xcolor,censor}
+\usepackage{fontspec}
 \usepackage{fancyvrb}
+
+\setmainfont{Times New Roman}
 
@@
 
-\vskip 2.5in
+\begin{SaveVerbatim}{VerbEnv}
 
-\begin{SaveVerbatim}{VerbEnv}  
 SOLUTION:
+
+(cond [(w 2) 4] [else 9])     -->[plug]
+       ^^^^^
+(cond [(= R 2) 4] [else 9])   -->[const]
+          ^
+(cond [(= 9 2) 4] [else 9])   -->[arith]
+       ^^^^^^^
+(cond [#false 4] [else 9])    -->[cond]
+^^^^^^^^^^^^^^^^^^^^^^^^^^
+(cond [else 9])               -->[cond]
+^^^^^^^^^^^^^^^
+9
 \end{SaveVerbatim}
@@
 \end{verbatim}
+
+\begin{SaveVerbatim}{VerbEnv}  
+SOLUTION:
+
+- make-run  : Constructor
+- run-time  : Accessor
+- run-up    : Accessor
+- run-score : Accessor
+- run?      : Predicate
+\end{SaveVerbatim}
+
+%\censor{\vspace{-2.5in}}
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
 
@@
 
-;; dist3d : 3D -> Number
-;; Compute distance to origin from given 3-d point
-(check-expect (dist3d (make-3d 3 2 6)) 7)
-(define (dist3d p)
-  (sqrt (+ (sqr (3d-x p)) (sqr (3d-y p)) (sqr (3d-z p)))))  
+(check-expect (name-width (make-name "Bob" "Harper")) 100)
+(check-expect (name-width (make-name "Bo" "Jack")) 70)
+(define (name-width n)
+  (* 10 (+ 1 (string-length (name-first n))
+             (string-length (name-last n)))))
 \end{SaveVerbatim}
@@
 
+;; count : LoN -> Number
+;; Count the number of non-negative numbers in list
+(check-expect (count-non-neg '()) 0)
+(check-expect (count-non-neg (list 1 -2 8 0)) 3)
+(define (count-non-neg lon)
+  (cond [(empty? lon) 0]
+        [(cons? lon)
+         (if (< (first lon) 0)
+             (count-non-neg (rest lon))
+             (add1 (count-non-neg (rest lon))))]))
 \end{SaveVerbatim}

### 2017-12-08 1/exams/f-practice.tex
Practice final.

--- 2017-11-06 1/exams/m2.tex
@@
 \begin{center}
-  {\Large {\bf CMSC 131A, Midterm 2}\\
+  {\Large {\bf CMSC 131A, Final Exam (Practice)}\\
     \censor{{\bf SOLUTION}} \ \\
@@
   Question & Points \\ \hline \hline
-  1 & 15 \\ \hline
-  2 & 12 \\ \hline
-  3 & 20 \\ \hline
-  4 & 15 \\ \hline
+  1 & 10 \\ \hline
+  2 & 10 \\ \hline
+  3 & 15 \\ \hline
+  4 & 12 \\ \hline
   5 & 15 \\ \hline
-  6 & 20 \\ \hline\hline
-  Total: & 97\\
+  6 & 15 \\ \hline
+  7 & 15 \\ \hline
+  8 & 20 \\ \hline\hline
+  Total: & 112\\
   \hline
@@
 \noindent
-{\bf Problem 1 (15 points).}
+{\bf Problem 1 (10 points).}
+%
+For the following program, write out each step of computation. At each
+step, underline the expression being simplified. Label each step as
+being ``arithmetic'' (meaning any built-in operation),
+``conditional'', ``plug'' (for plugging in an argument for a function
+parameter), or ``constant'' for replacing a constant with its value.
+
+\begin{verbatim}
+(define T 7)
+(define (q z) (sqr z))
+(cond [(> T 3) (q 4)]
+      [else 9])
+\end{verbatim}
+
+\newpage
+
+\noindent
+{\bf Problem 2 (10 points).}
+%
+For the following structure definition, list the names of every
+function it creates.  For each function, classify it as being either a
+constructor, accessor, or predicate.
+
+\begin{verbatim}
+(define-struct zip (pad msg))
+\end{verbatim}
+
+\begin{SaveVerbatim}{VerbEnv}  
+SOLUTION:
+
+- make-zip  : Constructor
+- zip-pad  : Accessor
+- zip-msg   : Accessor
+- zip?      : Predicate
+\end{SaveVerbatim}
+
+\newpage
+
+\noindent
+{\bf Problem 3 (15 points).}
 %
@@
 ;; two-element list of a word (first) and its meaning (second).
-
-;; A Result is one of:
-;; - #false
-;; - String
-;; Interp: represents the result of looking up a word in the
-;; dictionary; #false means not found.
 \end{verbatim}
@@
 \begin{verbatim}
-;; lookup : String Dict -> Result
-;; Lookup the (first) definition of a word in the dictionary,
-;; produces #false if not found
+;; update : String String Dict -> Dict
+;; Update entry for given word and meaning, if it exists
+;; Add the entry if it does not
 \end{verbatim}
@@
 
-(check-expect (lookup "a" '()) #false)
-(check-expect (lookup "a" (list (list "a" "b"))) "b")
-(check-expect (lookup "a" (list (list "c" "b") (list "a" "d"))) "d")
-(define (lookup w d)
-  (cond [(empty? d) #false]
+(check-expect (update "a" "b" '()) (list (list "a" "b")))
+(check-expect (update "a" "c" (list (list "a" "b")))
+              (list (list "a" "c")))
+(check-expect (update "a" "e" (list (list "c" "b") (list "a" "d"))) 
+              (list (list "c" "b") (list "a" "e")))
+(define (update w m d)
+  (cond [(empty? d) (list (list w m))]
         [(cons? d)
          (if (string=? (first (first d)) w)
-             (second (first d))
-             (lookup w (rest d)))]))
+             (cons (list w m) d)
+             (cons (first d) (lookup w (rest d))))]))
 
@@
 \noindent
-{\bf Problem 2 (12 points).}
+{\bf Problem 4 (12 points).}
 %
-For each of the following functions, provide the most general signature
-that correctly describes the function:
+For each of the following functions, provide the most general {\bf type} signature
+in Typed Racket that correctly describes the function:
 
 \begin{alltt}
-;; method : \censor{Number -> Number}
+(: method : \censor{Number -> Number                                    })
 
@@
 
-;; raekwon : \censor{[Listof Number] -> Boolean}
+(: raekwon : \censor{[Listof Number] -> Boolean                         })
 
@@
 
-;; rza : \censor{[Listof String] [Listof Number] -> [Listof Number]}
+(: rza : \censor{[Listof String] [Listof Number] -> [Listof Number]     })
 
@@
 
-;; gza : \censor{(String Number -> Number) -> Number}
+(: gza : \censor{(String Number -> Number) -> Number                    })
 
@@
 
-;; odb : \censor{(Number -> Number) -> (Number -> Number)}
+(: odb : \censor{(Number -> Number) -> (Number -> Number)               })
 
@@
 
-;; ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]}
+(: ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]          })
 
@@
 \noindent
-{\bf Problem 3 (20 points).}
+{\bf Problem 5 (15 points).}
 %
-Design a program called {\tt w-avg} that computes the weighted average
-of a list of numbers and a list of weights.
+Here is the design of a program {\tt w-avg} that computes the weighted
+average of a list of numbers and a list of weights.
+
+Design an accumulator variant of {\tt w-avg} that maintains two
+accumulators: the weighted sum of scores and the sum of weights.
+Redefine {\tt w-avg} in terms of your accumulator design.
 
@@
 \times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
-  (w-avg (list 80 70 90) (list 1 1 2))}.  You should assume the two
-lists have the same length.
+  (w-avg (list 80 70 90) (list 1 1 2))}.
 
-\begin{SaveVerbatim}{VerbEnv}
+In the accumulator version, after all elements of the list have been
+seen, the first accumulator would be $330 = 80+70+2\times 90$) and
+the second accumulator would be $4 =1+1+2$.
 
-SOLUTION:
-
+\begin{verbatim}
 ;; w-avg : [Listof Number] [Listof Number] -> Number
@@
             (w-sum (rest lon) (rest ws)))]))
+\end{verbatim}
+
+\noindent
+Your solution should not need a helper function like {\tt w-sum}.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
 
@@
 \newpage
+
+[Space for problem 5.]
+
+\newpage 
 \noindent
-{\bf Problem 4 (15 points).}
+{\bf Problem 6 (15 points).}
 %
@@
 \end{verbatim}
-%% {\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
-%% that operates on a binary tree of numbers, {\tt [Tree Number]}, and
-%% produces the sum of all the numbers in the tree.
-
-%% \begin{SaveVerbatim}{VerbEnv}
-
-%% SOLUTION:
-
-%% ;; tree-sum : [Tree Number] -> Number
-%% ;; Sum all the numbers in the tree
-%% (check-expect (tree-sum (make-leaf)) 0)
-%% (check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
-%%                                      (make-leaf)))
-%%               8)
-%% (define (tree-sum bt)
-%%   (cond [(leaf? bt) 0]
-%%         [(node? bt)
-%%          (+ (node-elem bt)
-%%             (tree-sum (node-left bt))
-%%             (tree-sum (node-right bt)))]))
-
-%% \end{SaveVerbatim}
-
-%% \censor{%
-%% \noindent
-%% \BUseVerbatim{VerbEnv}}
-
-%% \newpage
 
 \noindent
-%{\bf Problem 4(b) (10 points).}
-Here is a function that produces a list of all the elements in a tree:
+Here is a function that counts the elements in a tree:
 \begin{verbatim}
-;; tree-elems : [X] . [Tree X] -> [Listof X]
-;; Produce a list of all elements in tree going top-down, left-to-right
-(check-expect (tree-elems (make-leaf)) '())
-(check-expect (tree-elems (make-node 7 
+;; tree-count : [X] . [Tree X] -> Natural
+;; Count all the elements in a tree
+(check-expect (tree-count (make-leaf)) 0)
+(check-expect (tree-count (make-node 7 
                                      (make-node 9 (make-leaf) (make-leaf))
                                      (make-node 2 (make-leaf) (make-leaf))))
-              (list 7 9 2))
-(define (tree-elems bt)
-  (cond [(leaf? bt) '()]
+              3)
+(define (tree-count bt)
+  (cond [(leaf? bt) 0]
         [(node? bt)
-         (cons (node-elem bt)
-               (append (tree-elems (node-left bt))
-                       (tree-elems (node-right bt))))]))
+         (add1 (+ (tree-count (node-left bt))
+                  (tree-count (node-right bt))))]))
 \end{verbatim}
@@
 \end{verbatim}
-Give an equivalent definition of {\tt tree-elems} in terms of {\tt
+Give an equivalent definition of {\tt tree-count} in terms of {\tt
   tree-fold}.  (Just provide the code.)
@@
 \noindent
-[Space for problem 4.]
+[Space for problem 6.]
 
@@
 \noindent
-{\bf Problem 5 (15 points).}  
+{\bf Problem 7 (15 points).}  
 %
-Design a program that takes a list of strings and produces the count
-of strings with length longer than 3 in the list.  For example, if the
-list contains {\tt "a"}, {\tt "dave"}, {\tt "abc"}, and {\tt "fred"},
-the count is 2.  You may assume the {\tt{[Listof String]}} data
-definition is defined.
+During your internship at the University of Maryland's Center for
+Bioinformatics and Computational Biology (CBCB), you're asked to
+design a program for transcribing DNA strands into RNA strands.
 %
-For full credit, use list abstraction functions.  For partial credit,
-follow the template for {\tt{[Listof String]}}.
+A DNA strand is an arbitrarily long sequence of bases, which are
+either: C, G, A, or T; an RNA strand is an arbitrarily long sequences
+of either: C, G, A, or U.  The \emph{transcription} of DNA to RNA
+replaces each occurrence of T with U.
 
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
-;; count-3+ : [Listof String] -> Number
-;; Count the number of strings with length more than 3 in given list
-(check-expect (count-3+ '()) 0)
-(check-expect (count-3+ (list "a" "abcd" "fred" "c")) 2)
-(define (count-3+ los)
-  (foldr (lambda (s c) (if (> (string-length s) 3) (add1 c) c)) 0 los))
-
-;; Alts:
-(define (count-3+ los)
-  (local [;; count+ : String Number -> Number
-          ;; Bump count if string has length > 3
-          (define (count+ s c) 
-            (if (> (string-length s) 3) (add1 c) c))]
-    (foldr count+ 0 los))
-
-(define (count-3+ los)
-  (local [;; 3+? : String -> Boolean
-          ;; Does the string have length > 3?
-          (define (3+? s) 
-            (> (string-length s) 3))]
-    (length (filter 3+? los))))
-
-(define (count-3+ los)
-  (length (filter (lambda (s) (> (string-length s) 3)) los)))
-\end{SaveVerbatim}
-
-\censor{%
-\noindent
-\BUseVerbatim{VerbEnv}}
+Design a program that, given a DNA strand, computes its transcription
+to RNA.
 
@@
 \noindent
-{\bf Problem 6 (20 points).}
-%
-Here's a data definition for choose-your-own-adventure books:
+{\bf Problem 8 (20 points).} Here is the data definition for graphs we studied in class:
 \begin{verbatim}
-;; A CYOA is one:
-;; - String
-;; - (make-choice String CYOA CYOA)
-(define-struct choice (q yes no))
-;; Interp: a choose-your-own-adventure book where a string
-;; is a conclusion and a choice is a yes/no question and
-;; the two adventures that follow as a consequnce of the answer.
+;; A Graph is [Listof (cons String [Listof String])]
+;; Interp: a graph is a list of nodes and neighbors reachable from the node
+
+;; A Path is a [Listof String]
+;; Interp: series of nodes neighboring each other that connect a source
+;; to destination in some graph.
 \end{verbatim}
@@
 \noindent
-Design a function called {\tt ans} that consumes a {\tt CYOA} and a
-{\tt [Listof Boolean]} that represents answers to the yes/no
-questions: {\tt \#true} means yes, {\tt \#false} means no.  The {\tt
-  ans} function produces the {\tt CYOA} adventure that results after
-giving all of the answers in the list.  If the list contains too many
-answers, {\tt ans} just produces the conclusion of the adventure.
+Design the following function:
 
-\begin{SaveVerbatim}{VerbEnv}
+\begin{verbatim}
+;; all-paths : Graph String String -> [Listof Path]
+;; Produce all non-cyclic paths from src to dst in g.
+;; Assume: both nodes exist in graph.
+(define (all-paths g src dst) '())
+\end{verbatim}
 
-SOLUTION:
+\noindent
+You may assume the following function is already defined and works correctly:
 
-(define a0 "A0")
-(define a1 (make-choice "Q1" "A1" "A2"))
-(define a2 (make-choice "Q2" a0 a1))
+\begin{verbatim}
+;; neighbors : Graph String -> [Listof String]
+;; Produce the list of immediate neighbors of given node name.
+;; Assume: node exists in graph.
+\end{verbatim}
 
-;; ans : CYOA [Listof Boolean] -> CYOA
-;; Compute the remaining adventure after using answers from list
-(check-expect (adv a0 (list #true #false)) "A0")
-(check-expect (adv a1 (list #true #false)) "A1")
-(check-expect (adv a2 (list #false #true)) "A1")
-(check-expect (adv a2 (list #false)) a1)
-(define (ans adv loa)
-  (cond [(string? adv) adv]
-        [(empty? loa) adv]
-        [(cons? loa)
-         (if (first loa)
-             (ans (choice-yes adv) (rest adv))
-             (ans (choice-yes adv) (rest adv)))]))
-\end{SaveVerbatim}
+\noindent
+Note: this function is tricky to test since the tests shouldn't
+specify the order of the paths in the list.  You may ignore this
+subtly and just write tests as though a {\tt check-expect} will pass
+if the checked and expected expressions produce lists with the same
+elements in any order, e.g. {\tt (check-expect (list 1 2) (list 2 1))}
+would pass.
 
-\censor{%
-\noindent
-\BUseVerbatim{VerbEnv}}
-
+\newpage
+[Space for problem 8.]
 \end{document}

### 2017-12-08 1/exams/f-practice.tex
Solution for practice final.

@@
 
+\vskip 2.5in
+
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+(cond [(> T 3) (q 4)] [else 9])           -->[const]
+          ^
+(cond [(> 7 3) (q 4)] [else 9])           -->[arith]
+       ^^^^^^^
+(cond [#true (q 4)] [else 9])             -->[cond]
+^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+(q 4)                                     -->[plug]
+^^^^^
+(sqr 4)                                   -->[arith]
+^^^^^^^
+16
+\end{SaveVerbatim}
+
+\censor{\vspace{-2.5in}}
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
+
 \newpage
@@
 \end{SaveVerbatim}
+
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
 
@@
          (if (string=? (first (first d)) w)
-             (cons (list w m) d)
-             (cons (first d) (lookup w (rest d))))]))
-
+             (cons (list w m) (rest d))
+             (cons (first d) (update w m (rest d))))]))
 \end{SaveVerbatim}
@@
 
-(: raekwon : \censor{[Listof Number] -> Boolean                         })
+(: raekwon : \censor{[Listof Real] -> Boolean                           })
 
@@
 
-(: ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]          })
+(: ghostface : \censor{(All [X Y] [Listof X] [X -> Y] -> [Listof Y])    })
 
@@
 In the accumulator version, after all elements of the list have been
-seen, the first accumulator would be $330 = 80+70+2\times 90$) and
+seen, the first accumulator would be $330 = 80+70+2\times 90$ and
 the second accumulator would be $4 =1+1+2$.
@@
 
+\newpage
+
+[Space for problem 5.]
+
 \begin{SaveVerbatim}{VerbEnv}
@@
 
-;; Alt: define w-total function for denominator
+(define (w-avg xs ws) (w-avg/a xs ws 0 0))
 
-;; w-total : [Listof Number] -> Number
-;; Total a list of weights
-(define (w-total ws)
-  (cond [(empty? ws) 0]
-        [(cons? ws)
-         (+ (first ws) 
-            (w-total (rest ws)))]))
+;; w-avg/a : [Listof Number] [Listof Number] Number Number -> Number
+;; Compute weighted average with accumulators:
+;;  - sum-wxs: sum of weighted elements seen so far
+;;  - sum-ws: sum of weights
+(define (w-avg/a xs ws sum-wxs sum-ws)
+  (cond [(empty? xs) (/ sum-wxs sum-ws)]
+        [(cons? xs)
+         (w-avg/a (rest xs) 
+                  (rest ws)
+                  (+ (* (first xs) (first ws)) sum-wxs)
+                  (+ (first ws) sum-ws))]))
 \end{SaveVerbatim}
@@
 \BUseVerbatim{VerbEnv}}
-
-\newpage
-
-[Space for problem 5.]
 
@@
 
+\newpage
+
+\noindent
+[Space for problem 6.]
+
 \begin{SaveVerbatim}{VerbEnv}
@@
 
-(define (tree-elems bt)
-  (tree-fold (lambda (x l r) (cons x (append l r))) '() bt))
+(define (tree-count bt)
+  (tree-fold (lambda (x l r) (add1 (+ l r))) 0 bt))
 
@@
 \BUseVerbatim{VerbEnv}}
-
-\newpage
-
-\noindent
-[Space for problem 6.]
 
@@
 to RNA.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; A DNA is a [Listof DNA-Base]
+;; A RNA is a [Listof RNA-Base]
+;; A DNA-Base is one of: "C", "G", "A", "T"
+;; A RNA-Base is one of: "C", "G", "A", "U"
+
+;; trans : DNA -> RNA
+;; Transcribe DNA into RNA
+(check-expect (trans '()) '())
+(check-expect (trans (list "T" "A" "C" "T" "G"))
+              (list "U" "A" "C" "U" "G"))
+(define (trans dna)
+  (map (lambda (db) (if (string=? db "T") "U" db)) dna))    
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
 
@@
 (define (all-paths g src dst) '())
+
+(define G
+  (list (list "a" (list "b" "c"))
+        (list "b" (list "a" "d"))
+        (list "c" (list "d"))))
+
+(check-expect (all-paths G "a" "a")
+              (list (list "a")))
+(check-expect (all-paths G "a" "d")
+              (list (list "a" "b" "d")
+                    (list "a" "c" "d")))
+(check-expect (all-paths G "a" "d")
+              (list (list "a" "b" "d")
+                    (list "a" "c" "d")))
 \end{verbatim}
@@
 elements in any order, e.g. {\tt (check-expect (list 1 2) (list 2 1))}
-would pass.
+would pass.  (Tests are provided above, you don't need to add more unless
+it's useful to you.)
 
@@
 [Space for problem 8.]
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (all-paths g src dst)
+  (all-paths/seen g src dst '()))
+
+;; all-paths/seen : Graph String String [Listof String] -> [Listof Path]
+;; Accumulator: list of nodes visited so far on path
+;; Termination: each recursive call adds an item to seen list not previously
+;; seen.  Since there are finite nodes, we get closer to trivial case:
+;; src is in seen.
+(define (all-paths/seen g src dst seen)
+  (cond [(string=? src dst) (list (list src))]
+        [(member? src seen) '()]
+        [else
+         (local [;; all-paths-from : String -> [Listof Path]
+                 ;; All paths from given node to dst
+                 (define (all-paths-from n)
+                   (map (lambda (p) (cons n p))
+                        (all-paths/seen g n dst (cons src seen))))]
+
+         (map-append all-paths-from (neighbors g src)))]))
+
+;; map-append : [X Y] . [X -> [Listof Y]] [Listof X] -> [Listof Y]
+;; map a function and append all the list of results
+(define (map-append f xs) (foldr append '() (map f xs)))
+;; NB: could be more efficient
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \end{document}

### 2017-12-15 1/exams/f.tex

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{xcolor,censor}
\usepackage{fancyvrb}
\usepackage{fontspec}
\usepackage{alltt}

\setmainfont{Times New Roman}

\censorruleheight=0ex
\StopCensoring

\begin{document}

\begin{center}
  {\Large {\bf CMSC 131A, Final Exam}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

\vspace{2em}

\noindent
NAME:\verb|________________________________________________________|

\vspace{2em}

\noindent
UID: \verb|__________________________________________|

\vspace{2em}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 15 \\ \hline
  7 & 15 \\ \hline\hline
  Total: & 95\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 120 minutes to complete the test.

\vskip 1em

\noindent
The phrase ``design a program'' or ``design a function'' means follow
the steps of the design recipe.  Unless specifically asked for, you do
not need to provide intermediate products like templates or stubs,
though they may be useful to help you construct correct solutions.

\vskip 1em

\noindent
You may use any of the data definitions given to you within this exam
and do not need to repeat their definitions.

\vskip 1em

\noindent
Unless specifically instructed otherwise, you may use any built-in ISL+
functions or data types.

\vskip 1em

\noindent
When writing tests, you may use a shorthand for writing check-expects
by drawing an arrow between two expressions to mean you expect the
first to evaluate to same result as the second.  For example, you may
write \verb|(add1 3)| $\rightarrow$ \verb|4| instead of
\verb|(check-expect (add1 3) 4)|.

\newpage

\noindent
{\bf Problem 1 (10 points).}
%
For the following program, write out each step of computation. At each
step, underline the expression being simplified. Label each step as
being ``arithmetic'' (meaning any built-in operation),
``conditional'', ``plug'' (for plugging in an argument for a function
parameter), or ``constant'' for replacing a constant with its value.

\begin{verbatim}
(define S 1)
(define (q z)
  (cond [(< 4 z) (+ z S)]
        [else 9]))
(q 8)
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(q 8)                                --[plug]-->
^^^^^
(cond [(< 4 8) (+ 8 S)] [else 9])    --[arith]-->
       ^^^^^^^
(cond [#true (+ 8 S)] [else 9])      --[cond]-->
^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
(+ 8 S)                              --[const]-->
     ^
(+ 8 1)                              --[arith]-->
^^^^^^^
9
\end{SaveVerbatim}

\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 2 (10 points).}
%
For the following structure definition, list the names of every
function it creates.  For each function, classify it as being either a
constructor, accessor, or predicate.

\begin{verbatim}
(define-struct qt (nw ne se sw))
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:

- make-qt  : Constructor
- qt-nw    : Accessor
- qt-ne    : Accessor
- qt-se    : Accessor
- qt-sw    : Accessor
- qt?      : Predicate
\end{SaveVerbatim}

\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 3 (15 points).}
%
Consider the following data definition (and examples) for representing sentences:
\begin{verbatim}
;; A Sentence is a [NEListof String]

(define sent1 (list "a" "dog" "barked"))
(define sent2 (list "a" "cat" "meowed"))
\end{verbatim}
Recall the following data definition of {\tt [NEListof X]}:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
\end{verbatim}
Given a {\tt Sentence}, a \emph{bigram} is a list of two consecutive
elements of the sentence. For example, {\tt sent1} has exactly two
bigrams:
\begin{verbatim}
(list "a" "dog")
(list "dog" "barked")
\end{verbatim}
Using the following data definition,
\begin{verbatim}
;; a Bigram is a (list String String)
\end{verbatim}
design a function {\tt bigrams} which takes a {\tt Sentence} and
returns a list of {\tt Bigrams}.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; bigrams : Sentence -> [Listof Bigram]
;; List of all adjacent words in a sentence
(check-expect (bigrams (list "one")) '())
(check-expect (bigrams sent1)
              (list (list "a" "dog") (list "dog" "barked")))
(define (bigrams sent)
  (cond [(empty? (rest sent)) '()]
        [else
         (cons (list (first sent) (second sent))
               (bigrams (rest sent)))]))
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 4 (15 points).}
%
Here is a program that, given a sentence (see problem 3), computes a
two element list consisting of the minimal and maximal words according
to alphabetic ordering ({\tt string<?}).

\begin{verbatim}
;; min-and-max : Sentence -> (list String String)
;; Produces the min and max words in the sentence.
(check-expect (min-and-max (list "the" "zoo" "has" "aardvarks"))
              (list "aardvarks" "zoo"))
(define (min-and-max s)
  (cond [(empty? (rest s)) (list (first s) (first s))]
        [else
         (local [(define w (first s))
                 (define m+m (min-and-max (rest s)))]
            (cond [(string<? w (first m+m))
                   (list w (second m+m))]
                  [(string<? (second m+m) w)
                   (list (first m+m) w)]
                  [else m+m]))]))
\end{verbatim}
Redesign this program to use an accumulator-based helper function with
two accumulators: the mininmal and maximal elements of the sentence
seen so far.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(define (min-and-max s)
  (min-and-max/a (rest s) (first s) (first s)))

;; min-and-max/a : [Listof String] String String -> (list String String)
;; Accumulator: min and max element seen so far
(define (min-and-max/a s min max)
  (cond [(empty? s) (list min max)]
        [(cons? s)
         (local [(define w (first s))]
           (cond [(string<? w min) 
                  (min-and-max/a (rest s) w max)]
                 [(string<? max w)
                  (min-and-max/a (rest s) min w)]
                 [else
                  (min-and-max/a (rest s) min max)]))]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

[Space for problem 4.]

\newpage 
\noindent
{\bf Problem 5 (15 points).}
%
Design the following program for determining if two sentences (see problem 3) are equal,
meaning they are lists of the same length with the same elements (in the same order):
\begin{verbatim}
;; sent=? : Sentence Sentence -> Boolean
;; Are the two sentences the same?
\end{verbatim}
(You may not use {\tt equal?} in the design of this program.)

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; sent=? : Sentence Sentence -> Boolean
;; Are the two sentences equal?
(check-expect (sent=? (list "one") (list "one")) #true)
(check-expect (sent=? (list "one") (list "two")) #false)
(check-expect (sent=? (list "one") (list "one" "two")) #false)
(check-expect (sent=? (list "one" "two") (list "one" "two")) #true)
(define (sent=? s1 s2)
  (cond [(empty? (rest s1))
         (and (empty? (rest s2))
              (string=? (first s1) (first s2)))]
        [else
         (and (not (empty? (rest s2)))
              (string=? (first s1) (first s2))
              (sent=? (rest s1) (rest s2)))]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage
\noindent
{\bf Problem 6 (15 points).}  
%
During your internship at the University of Maryland's Center for
Bioinformatics and Computational Biology (CBCB), you're asked to
design a program for computing the \emph{reverse complement} of DNA strands.
%
A DNA strand is an arbitrarily long sequence of bases, which are
either: C, G, A, or T.  A and T are complements of each other, as are
C and G.  The reverse complement of a DNA strand is formed by
reversing the strand and taking the complement of each symbol.  For
example, the reverse complement of AAAACCCGGT is ACCGGGTTTT.
%
Design a program that, given a DNA strand, computes its reverse complement.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; A DNA is a [Listof Base]
;; A Base is one of: "A", "C", "G", "T"
;; Interp: DNA is a sequence of nucleotide bases

;; rev-comp : DNA -> DNA
;; Compute the reverse complement of the given strand
(check-expect (rev-comp (list "A" "A" "C" "G" "T"))
              (list "A" "C" "G" "T" "T"))
(define (rev-comp dna)
  (reverse (map comp dna)))

;; comp : Base -> Base
;; Compute the complement of the given base
(check-expect (comp "A") "T")
(check-expect (comp "T") "A")
(check-expect (comp "C") "G")
(check-expect (comp "G") "C")
(define (comp b)
  (cond [(string=? b "A") "T"]
        [(string=? b "T") "A"]
        [(string=? b "C") "G"]
        [(string=? b "G") "C"]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 7 (15 points).} Design a program that takes two
strings and counts how many times the first string occurs within the
second.
For example,
\begin{itemize}
\item {\tt "ho"} occurs three times in {\tt "ho ho ho"}, while
\item {\tt "ho ho"} occurs once in {\tt "ho ho ho"} (notice that occurences
  cannot overlap).
\end{itemize}
You may assume the first string is non-empty.
%
It may be helpful to use the {\tt substring} function.  Recall: {\tt
  (substring "hello world" 1 5)} produces {\tt "ello"} and {\tt
  (substring "hello world" 4)} produces {\tt "o world"}.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; occurs : String String -> Natural
;; Count (non-overlapping) occurences of first string in second
;; Assume: first string is non-empty
;; Termination: because s1 is non-empty, each recursive call is on a strictly
;; smaller s2, therefore eventually reaching first cond clause (trivial case).
(check-expect (occurs "ho" "ho ho ho") 3)
(check-expect (occurs "ho ho" "ho ho ho") 1)
(define (occurs s1 s2)
  (local [(define s1-len (string-length s1))]
    (cond [(< (string-length s2) s1-len) 0]
          [(string=? s1 (substring s2 0 s1-len))
           (add1 (occurs s1 (substring s2 s1-len)))]
          [else
           (occurs s1 (substring s2 1))])))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

%% \newpage

%% \noindent
%% {\bf Problem 8 (15 points).} An association list is a way of
%% representing associations between pairs of elements.  For example, a
%% dictionary is an association between words (strings) and their
%% definition (also strings).  The final grades for this class might be
%% represented as an association between student IDs (numbers) and letter
%% grades (strings).  Here is a parameterized data definition for
%% associations:
%% \begin{verbatim}
%% ;; An [Assoc X Y] is one of:
%% ;; - '()
%% ;; - (cons (list X Y) [Assoc X Y])
%% ;; Interp: a collection of associations where each element is a
%% ;; two-element list of associated elements.
%% \end{verbatim}
%% Here are two functions for looking up values in different kinds of associations:
%% \begin{verbatim}
%% ;; lookup-dict : [Assoc String String] String -> [Maybe String]
%% ;; Lookup meaning of a word in a dictionary
%% (check-expect (lookup-dict (list (list "dog" "goodboy")) "dog") "goodboy")
%% (check-expect (lookup-dict (list (list "dog" "goodboy")) "cat") #false)
%% (define (lookup-dict d w)
%%   (cond [(empty? d) #false]
%%         [(string=? (first (first d)) w) (second (first d))]
%%         [else (lookup-dict (rest d) w)]))

%% ;; lookup-grade : [Assoc Number String] Number -> [Maybe String]
%% ;; Lookup letter grade for given student ID
%% (check-expect (lookup-grade (list (list 1112 "A++")) 1112) "A++")
%% (check-expect (lookup-grade (list (list 1112 "A++")) 9999) #false)
%% (define (lookup-grade d w)
%%   (cond [(empty? d) #false]
%%         [(= (first (first d)) w) (second (first d))]
%%         [else (lookup-grade (rest d) w)]))
%% \end{verbatim}
%% Design an abstraction of these two functions and redefine these functions in terms of this abstraction.
%% (You do not need to write additional tests; your solution should not use {\tt equal?}.)

%% \newpage
%% [Space for problem 8.]

\end{document}

### 2018-10-18 1/exams/m1.tex
Midterm 1 source.

@@
     \ \\
-    Fall 2017 \ \\
+    Fall 2018 \ \\
   }
@@
   2 & 10 \\ \hline
-  3 & 10 \\ \hline
+  3 & 15 \\ \hline
   4 & 15 \\ \hline
-  5 & 15 \\ \hline
-  6 & 10 \\ \hline
-  7 & 20 \\ \hline \hline
+  5 & 40 \\ \hline
   Total: & 90\\
@@
 \begin{verbatim}
-(define R 9)
-(define (w z) (= R z))
-(cond [(w 2) 4]
-      [else 9])
+(define Q 1)
+(define (w z) (< (string-length z) 20))
+(w (cond [(= 1 Q) "fred"]
+         [else 9]))
 \end{verbatim}
@@
 
-(cond [(w 2) 4] [else 9])     -->[plug]
-       ^^^^^
-(cond [(= R 2) 4] [else 9])   -->[const]
+(w (cond [(= 1 Q) "fred"] [else 9])) -->[const]
                ^
-(cond [(= 9 2) 4] [else 9])   -->[arith]
+(w (cond [(= 1 1) "fred"] [else 9])) -->[arith]
           ^^^^^^^
-(cond [#false 4] [else 9])    -->[cond]
-^^^^^^^^^^^^^^^^^^^^^^^^^^
-(cond [else 9])               -->[cond]
-^^^^^^^^^^^^^^^
-9
+(w (cond [#true "fred"] [else 9]))   -->[cond]
+   ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
+(w "fred")                           -->[plug]
+^^^^^^^^^^
+(< (string-length "fred") 20))       -->[arith]
+   ^^^^^^^^^^^^^^^^^^^^^^
+(< 5 20)                             -->[arith]
+#true
 \end{SaveVerbatim}
@@
 \begin{verbatim}
-(define-struct run (time up score))
+(define-struct dr (who strange))
 \end{verbatim}
@@
 
-- make-run  : Constructor
-- run-time  : Accessor
-- run-up    : Accessor
-- run-score : Accessor
-- run?      : Predicate
+- make-dr    : Constructor
+- dr-who     : Accessor
+- dr-strange : Accessor
+- dr?        : Predicate
 \end{SaveVerbatim}
@@
 \noindent
-{\bf Problem 3 (10 points).}
+{\bf Problem 3 (15 points).}
 %
-The graphical text editor you developed in a recent assignment has
-attracted the interest of investors.  But before they write you a
-check, they want to make sure you can handle fancy operations for
-expert text editor users.  They ask if you can implement a \emph{swap}
-operation, which will swap content to the left of the cursor with the
-content to the right..  To satisfy your potential investors, you
-define the following function for swapping strings at a given
-position:
+You've been hired by the CMNS development office and been given their
+existing software for spamming potential donors.  It contains the following
+data definition:
 \begin{verbatim}
-;; swap : String Integer -> String
-;; Swap string to the left of i with string to the right.
-;; Assume (<= 0 i (string-length s)).
-(check-expect (swap "" 0) "")
-(check-expect (swap "ab" 1) "ba")
-(check-expect (swap "bannana" 3) "nanaban")
-(define (swap s i) ...)
+;; A Person is a (make-name Title String String)
+;; Interp: a person's title and first and last names.
+(define-struct name (title first last))
+
+;; A Title is one of:
+;; - "r"      Interp: Mr.
+;; - "s"              Ms.
+;; - "d"              Dr.
 \end{verbatim}
@@
 \noindent
-Give a correct definition for \verb|swap|.  (You only need to
-provide code, not the design steps.)
+Finish the design of the following function for creating letter
+openings:
+\begin{verbatim}
+;; dear : Person -> String
+;; Create a letter opening for the given person.
+(check-expect (dear (make-name "d" "Minnie" "Maisy"))
+              "Dear Dr. Maisy:")
+(check-expect (dear (make-name "r" "Fred" "Rogers"))
+              "Dear Mr. Rogers:")
+(check-expect (dear (make-name "s" "Miriam" "Maisel"))
+              "Dear Ms. Maisel:")
+\end{verbatim}
 
@@
 %
-Use the following data definition for representing names:
-\begin{verbatim}
-;; A Name is a (make-name String String)
-;; Interp: a person's full (first and last) name.
-;; Both strings must contain at least one letter.
-(define-struct name (first last))
-\end{verbatim}
-
-\noindent
-Write a program, \verb|name-width|, that determines the width needed
-to print someone's name on a letter using a mono-typed font (meaning
-each letter has the same width).  The program can calculate this width
-based on the length of the person's first and last name, plus the
-width of a space to go in between.  Assume each letter is 10 pixels
-wide.  For example, the name Bob Harper requires 100 pixels; the name
-Bo Jack requires 70 pixels.  You may start from the following header:
-
-\begin{verbatim}
-;; name-width : Name -> Number
-;; Computes width needed to print name in mono-type font
-\end{verbatim}
-
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
-(check-expect (name-width (make-name "Bob" "Harper")) 100)
-(check-expect (name-width (make-name "Bo" "Jack")) 70)
-(define (name-width n)
-  (* 10 (+ 1 (string-length (name-first n))
-             (string-length (name-last n)))))
-\end{SaveVerbatim}
-
-\censor{%
-\noindent
-\BUseVerbatim{VerbEnv}}
-
-\newpage
-\noindent
-{\bf Problem 5 (15 points).}  
-%
-Design a program that takes a list of numbers and produces the count
-of non-negative numbers in the list.  For example, if the list
-contains 3, 4, -1, 0, -2, and 5, the count of non-negative numbers is
-4.  You may use the following data definition:
+Design a program that takes a list of numbers and produces the sum of
+positive numbers in the list.  For example, if the list contains 3, 4,
+-1, 0, -2, and 5, the sum of positive numbers (3+4+5) is 12.  You may
+use the following data definition:
 
@@
 \noindent
-{\bf Problem 6 (10 points).} You've been hired by the US Department of
-Transportation to design traffic simulation software, which includes
-simulating traffic lights.  Discovering the simple design from 131A is
-insufficient, you come up with the following representation, which
-makes it possible to model not just the color, but also how much
-longer the current light has before changing:
+{\bf Problem 5 (40 points).} You've been hired by Nintendo to design
+their next blockbuster game: Dots.  It consists of, well, dots: lots
+and lots of dots, each of which occupies a coordinate in space.  You
+come up with the following data definitions for representing a single
+dot and a collection of dots:
 
 \begin{verbatim}
-;; A Light is one of:
-;; - (make-red Natural)
-;; - (make-yellow Natural)
-;; - (make-green Natural)
-;; Interp: color and remaining duration (in ticks) of a light 
-(define-struct red (count))
-(define-struct yellow (count))
-(define-struct green (count))
+;; A Dots is one of:
+;; - empty
+;; - (cons Dot Dots)
+;; Interp: collection of an arbitrary number of dots
+
+;; A Dot is a (make-posn Number Number)
+;; Interp: a dot's position in space
 \end{verbatim}
-(Recall that a \verb|Natural| is a natural number, i.e. a non-negative
-integer.)
 
@@
 \noindent
-{\bf Problem 6(a).} Write a template for \verb|Light| functions.
+{\bf Problem 5(a) [10 points].} Write templates for \verb|Dot| and
+\verb|Dots| functions.
 
@@
 
-(define (light-temp l)
-  (cond [(red? l) (... (red-count l) ...)]
-        [(yellow? l) (... (yellow-count l) ...)]
-        [(green? l) (... (green-count l) ...)]
+(define (dot-template d)
+  (... (posn-x d)
+       (posn-y d)
+       ...))
 
+(define (dots-template ds)
+  (cond [(empty? ds) ...]
+        [(cons? ds) (... (dot-template (first ds))
+                         (dots-template (rest ds)) ...)]))
 \end{SaveVerbatim}
 
-\vskip 15em
+\vskip 18em
 
@@
 \noindent
-{\bf Problem 6(b).} Write a stub for the following function:
+{\bf Problem 5(b) [10 points].} Write a stub for the following function:
 
 \begin{verbatim}
-;; tock : Light -> Light
-;; Advance the light by one tick of time
-(check-expect (tock (make-red 10)) (make-red 9))
-(check-expect (tock (make-red 0)) (make-green 10))
+;; nw : Dots -> Dots
+;; Move given dots in the northwest direction one unit
+(check-expect (nw empty) empty)
+(check-expect (nw (cons (make-posn 3 4) empty))
+              (cons (make-posn 4 5) empty))
 \end{verbatim}
@@
 
-(define (tock l) (make-red 0))
+(define (nw ds) empty)
 
@@
 \noindent
-{\bf Problem 7 (20 points).}
+{\bf Problem 5(c) [20 points].}
 %
-Over the summer, you take an internship at the University of
-Maryland's Center for Bioinformatics and Computational Biology (CBCB).
-Your first task deals with computing information about \emph{DNA
-  strands}.  A DNA strand is an arbitrarily long sequence of
-\emph{nucleotides}, which are either: C, G, A, or T.  Design a program
-that, given a DNA strand, computes the number of times the G
-nucleotide occurs in it.  For example, if the strand is AGCTTTGA, your
-program should produce 2.
-
+Design two functions: {\tt dot-flip} reflects a given dot over the
+y-axis, for example, $(3,4)$ becomes $(-3,4)$, and {\tt dots-flip}
+reflects a collection of dots over the y-axis.
 \begin{SaveVerbatim}{VerbEnv}
@@
 
-;; A Nucleotide is one of:
-;; - "C"
-;; - "G"
-;; - "A"
-;; - "T"
-;; Interp: DNA nucleotide C,G,A,T.
+;; dot-flip : Dot -> Dot
+;; Flip dot over y-axis
+(check-expect (dot-flip (make-posn 3 4)) (make-posn -3 4))
+(define (dot-flip d)
+  (make-posn (- (posn-x d)) (posn-y d)))
 
-;; A DNA is one of:
-;; - '()
-;; - (cons Nucleotide DNA)
-;; Interp: a DNA strand of nucleotides
-
-;; count-g : DNA -> Natural
-;; Count the number of occurrence of G in given DNA strand
-(check-expect (count-g '()) 0)
-(check-expect (count-g (cons "A" (cons "G" (cons "G" '())))) 2)
-(define (count-g dna)
-  (cond [(empty? dna) 0]
-        [(cons? dna)
-         (+ (if (string=? (first dna) "G") 1 0)
-            (count-g (rest dna)))]))
-
+;; dots-flip : Dots -> Dots
+;; Flip dots over y-axis
+(check-expect (dots-flip empty) empty)
+(check-expect (dots-flip (cons (make-posn 3 4) empty))
+              (cons (make-posn -3 4) empty))
+(define (dots-flip ds)
+  (cond [(empty? ds) empty]
+        [(cons? ds) 
+         (cons (dot-flip (first ds))
+               (dots-flip (rest ds)))]))                   
 \end{SaveVerbatim}

### 2018-10-18 1/exams/m1.tex
Fix up solution to m1.

@@
 
-;; count : LoN -> Number
-;; Count the number of non-negative numbers in list
-(check-expect (count-non-neg '()) 0)
-(check-expect (count-non-neg (list 1 -2 8 0)) 3)
-(define (count-non-neg lon)
+;; sum-pos : LoN -> Number
+;; Count the number of positive numbers in list
+(check-expect (sum-pos '()) 0)
+(check-expect (sum-pos (list 1 -2 8 0)) 9)
+(define (sum-pos lon)
   (cond [(empty? lon) 0]
@@
          (if (< (first lon) 0)
-             (count-non-neg (rest lon))
-             (add1 (count-non-neg (rest lon))))]))
+             (sum-pos (rest lon))
+             (+ (first lon) (sum-pos (rest lon))))]))
 \end{SaveVerbatim}

### 2018-11-05 1/exams/m2.tex
Midterm 2.

--- 2017-11-06 1/exams/m2.tex
@@
   Question & Points \\ \hline \hline
-  1 & 15 \\ \hline
-  2 & 12 \\ \hline
-  3 & 20 \\ \hline
+  1 & 10 \\ \hline
+  2 & 15 \\ \hline
+  3 & 15 \\ \hline
   4 & 20 \\ \hline
-  5 & 15 \\ \hline
-  6 & 20 \\ \hline\hline
-  Total: & 102\\
+  5 & 20 \\ \hline\hline
+  Total: & 80\\
   \hline
@@
 \noindent
-{\bf Problem 1 (15 points).}
+{\bf Problem 1 (10 points).}
 %
-Here are some data definition relevant to representing a dictionary,
-which associates words with the their definitions:
-
+Here is a data definition for representing definitions in a
+dictionary:
 \begin{verbatim}
-;; A Dict is one of:
-;; - '()
-;; - (cons (list String String) Dict)
-;; Interp: represents a collection of definitions where each
-;; element is a two-element list consisting of a word (first) and its
-;; meaning (second).
-
-;; A Result is one of:
-;; - #false
-;; - String
-;; Interp: represents the result of looking up a word in the
-;; dictionary; #false means not found.
+;; A Defn is a (make-defn String String)
+(define-struct defn (word meaning))
+;; Interp: a word and its definition
+;; Example:
+(make-defn "pair" 
+           "a set of two things used together or regarded as a unit")
 \end{verbatim}
 
-\noindent
-Design the following function:
-
+Design the following function for determining whether one definition
+should come before another in a dictionary.  (Hint: you can use the
+function {\tt string<?} to help.)
 \begin{verbatim}
-;; lookup : String Dict -> Result
-;; Lookup the (first) definition of a word in the dictionary,
-;; produces #false if not found
+;; defn<? : Defn Defn -> Boolean
+;; Does the word defined in d1 come before d2, alphabetically?
+(define (defn<? d1 d2) ...)
 \end{verbatim}
@@
 
-(define (f b s) s)
-(define (g fs) (fs #false))
-(define (h bs ns) 2)
-(define (i f) (f 2))
-(define (j f xs) (map f xs))
-(define (k f g x) (g (f x)))
-\end{SaveVerbatim}
-
-%\censor{\vspace{-2.5in}}
-\censor{%
-\noindent\BUseVerbatim{VerbEnv}}
-
-\newpage
-
-\noindent
-{\bf Problem 2 (12 points).}
-%
-For each of the following functions, provide the most general signature
-that correctly describes the function:
-
-\begin{alltt}
-;; method : \censor{Number -> Number}
-
-(define (method x)
-  (+ (sqr x) 2))
-
-;; raekwon : \censor{[Listof Number] -> Boolean}
-
-(define (raekwon x)
-  (ormap positive? x))
-
-;; rza : \censor{[Listof String] [Listof Number] -> [Listof Number]}
-
-(define (rza x y)
-  (cond [(empty? x) y]
-        [(cons? x)
-         (cons (string-length (first x))
-               (rza (rest x) y))]))
-
-;; gza : \censor{(String Number -> Number) -> Number}
-
-(define (gza x)
-  (foldr x 0 (list "a" "b" "c")))
-
-;; odb : \censor{(Number -> Number) -> (Number -> Number)}
-
-(define (odb x)
-  (lambda (y)   
-    (/ (- (x (+ y 0.001))
-          (x (- y 0.001)))
-       (* 2 0.001))))
-
-;; ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]}
-
-(define (ghostface x y)
-  (cond [(empty? x) '()]
-        [(cons? x)
-         (cons (y (first x)) 
-               (ghostface (rest x) y))]))
-\end{alltt}
-
-\newpage
-
-\noindent
-{\bf Problem 3 (20 points).}
-%
-Design a program called {\tt w-avg} that computes the weighted average
-of a list of numbers and a list of weights.
-
-For example, let's say a class grade is based on two midterms and a
-project where the project is worth twice as much as the midterms.  A
-student who gets an 80 and a 70 on the midterms and a 90 on the
-project would have a weighted average of 82.5: $((1 \times 80) + (1
-\times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
-  (w-avg (list 80 70 90) (list 1 1 2))}.  You should assume the two
-lists have the same length.
-
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
-;; avg : [Listof Number] -> Number
-;; Compute the average of a list of numbers
-;; Assume: the list is non-empty
-(check-expect (avg (list 3 4 5)) 4)
-(check-expect (avg (list 10 20)) 15)
-(define (avg lon)
-  (/ (sum lon) 
-     (length lon)))
-
-;; sum : [Listof Number] -> Number
-;; Sum the list of numbers
-(check-expect (sum '()) 0)
-(check-expect (sum (list 3 4 5)) 12)
-(define (sum lon)
-  (foldr + 0 lon))
-
-;; Alt: follow [Listof Number] template
-\end{SaveVerbatim}
-
-\censor{%
-\noindent
-\BUseVerbatim{VerbEnv}}
-
-\newpage 
-\noindent
-{\bf Problem 4 (20 points).}
-%
-Here is a parametric data definition for a tree of elements:
-\begin{verbatim}
-;; A [Tree X] is one of:
-;; - (make-leaf)
-;; - (make-node X [Tree X] [Tree X])
-;; Interp: a binary tree that is either empty (a leaf), or non-empty (a node)
-;; with an element and two sub-trees.
-(define-struct leaf ())
-(define-struct node (elem left right))
-\end{verbatim}
-{\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
-that operates on a binary tree of numbers, {\tt [Tree Number]}, and
-produces the sum of all the numbers in the tree.
-
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
-;; tree-sum : [Tree Number] -> Number
-;; Sum all the numbers in the tree
-(check-expect (tree-sum (make-leaf)) 0)
-(check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
-                                     (make-leaf)))
-              8)
-(define (tree-sum bt)
-  (cond [(leaf? bt) 0]
-        [(node? bt)
-         (+ (node-elem bt)
-            (tree-sum (node-left bt))
-            (tree-sum (node-right bt)))]))
+(check-expect (defn<? (make-defn "a" "first") (make-defn "b" "second")) #true)
+(check-expect (defn<? (make-defn "b" "second") (make-defn "a" "first")) #false)
+(define (defn<? d1 d2)
+  (string<? (defn-word d1) (defn-word d2)))
 
@@
 \noindent
-{\bf Problem 4(b) (10 points).}
-Here is a function that produces a list of all the elements in a tree:
+{\bf Problem 2 (15 points).}
+%
+A dictionary can be represented using a structure similar to a binary
+search tree as follows (where {\tt Defn} is defined in problem 1):
 \begin{verbatim}
-;; tree-elems : [X] . [Tree X] -> [Listof X]
-;; Produce a list of all elements in tree going top-down, left-to-right
-(check-expect (tree-elems (make-leaf)) '())
-(check-expect (tree-elems (make-node 7 
-                                     (make-node 9 (make-leaf) (make-leaf))
-                                     (make-node 2 (make-leaf) (make-leaf))))
-              (list 7 9 2))
-(define (tree-elems bt)
-  (cond [(leaf? bt) '()]
-        [(node? bt)
-         (cons (node-elem bt)
-               (append (tree-elems (node-left bt))
-                       (tree-elems (node-right bt))))]))
+;; A BSD (binary search dictionary) is one of:
+;; - (make-leaf)
+;; - (make-node d l r), where d is a Defn, l is a BSD, r is a BSD,
+;;      and every definition in l alphabetically comes before d
+;;      and d comes alphabetically before every definition in r
+;; ASSUME: words are defined at most once in a BSD
+(define-struct leaf ())
+(define-struct node (val left right))
 \end{verbatim}
-Here is an abstraction function for trees that is similar to
-{\tt foldr} for lists, but works on trees:
-\begin{verbatim}
-;; tree-fold : [X Y] . [X Y Y -> Y] Y [Tree X] -> Y
-;; The fundamental abstraction function for trees
-(define (tree-fold f b bt)
-  (cond [(leaf? bt) b]
-        [(node? bt)
-         (f (node-elem bt)
-            (tree-fold f b (node-left bt))
-            (tree-fold f b (node-right bt)))]))
-\end{verbatim}
-Give an equivalent definition of {\tt tree-elems} in terms of {\tt
-  tree-fold}.  (Just provide the code.)
+
+Design a function {\tt all-words} that consumes a {\tt BSD} and
+produces a list of all the words (just the words, not their meanings)
+defined in the dictionary, in alphabetic order.
 
@@
 
-(define (tree-elems bt)
-  (tree-fold (lambda (x l r) (cons x (append l r))) '() bt))
+(define a (make-defn "a" "first"))
+(define b (make-defn "b" "second"))
+(define c (make-defn "c" "third"))
+(define l (make-leaf))
+
+;; all-words : BSD -> [Listof String]
+;; List all words defined in given dictionary in alphabetic order
+(check-expect (all-words l) '())
+(check-expect (all-words (make-node b (make-node a l l) (make-node c l l)))
+              (list "a" "b" "c"))
+(define (all-words bsd)
+  (cond [(leaf? bsd) '()]
+        [(node? bsd)
+         (append (all-words (node-left bsd))
+                 (list (defn-word (node-val bsd)))
+                 (all-words (node-right bsd)))]))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
+\newpage
+
+\noindent
+{\bf Problem 3 (15 points).}
+%
+For each of the following functions, correctly reformulate the
+function in terms of an existing abstraction function.  You may use
+either lambda-expressions or local definitions if needed.
+
+\begin{alltt}
+;; any-zero? : [Listof Number] -> Boolean
+;; Are any of the numbers in given list equal to zero?
+(check-expect (any-zero? '(1 2 3)) #false)
+(check-expect (any-zero? '(1 0 3)) #true)
+(define (any-zero? xs)
+  (cond [(empty? xs) #false]
+        [(cons? xs)
+         (or (zero? (first xs))
+             (any-zero? (rest xs)))]))
+\end{alltt}
+
+\vspace{.8in}
+
+\begin{SaveVerbatim}{VerbEnv}
+SOLUTION:
+
+(define (any-zero? xs) (ormap zero? xs))
+
+\end{SaveVerbatim}
+
+\censor{%
+\vspace{-.8in}
+\noindent
+\BUseVerbatim{VerbEnv}}
+
+\begin{alltt}
+(define DOT (circle 10 "solid" "black"))
+(define MT (empty-scene 200 200))
+
+;; draw-dots : [Listof Posn] -> Image
+;; Draw dots for all given positions on an empty scene.
+(check-expect (draw-dots '()) MT)
+(check-expect (draw-dots (list (make-posn 10 20) (make-posn 20 30)))
+              (place-image DOT 10 20 (place-image DOT 20 30 MT)))
+(define (draw-dots ps)
+  (cond [(empty? ps) MT]
+        [(cons? ps)
+         (place-image DOT
+                      (posn-x (first ps))
+                      (posn-y (first ps))
+                      (draw-dots (rest ps)))]))
+\end{alltt}
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (draw-dots ps)
+  (foldr (lambda (p img) (place-image DOT (posn-x p) (posn-y p) img))
+         (empty-scene 200 200)
+         ps))
 
@@
 \noindent
-{\bf Problem 5 (15 points).}  
-%
-Design a program that takes a list of strings and produces the count
-of strings with length longer than 3 in the list.  For example, if the
-list contains {\tt "a"}, {\tt "dave"}, {\tt "abc"}, and {\tt "fred"},
-the count is 2.  You may assume the {\tt{[Listof String]}} data
-definition is defined.
-%
-For full credit, use list abstraction functions.  For partial credit,
-follow the template for {\tt{[Listof String]}}.
+{\bf Problem 3 (cont).}
+
+\begin{alltt}
+;; remove-short-words : Number [Listof Defn] -> [Listof Defn]
+;; Remove all definitions for words shorter than given length
+(check-expect (remove-short-words 2 (list (make-defn "a" "first") 
+                                          (make-defn "ab" "rest")))
+              (list (make-defn "ab" "rest")))
+(define (remove-short-words len ds)
+  (cond [(empty? ds) '()]
+        [(cons? ds)
+         (if (< (string-length (defn-word (first ds))) len)
+             (remove-short-words len (rest ds))
+             (cons (first ds)
+                   (remove-short-words len (rest ds))))]))
+\end{alltt}
 
@@
 
-;; count-3+ : [Listof String] -> Number
-;; Count the number of strings with length more than 3 in given list
-(check-expect (count-3+ '()) 0)
-(check-expect (count-3+ (list "a" "abcd" "fred" "c")) 2)
-(define (count-3+ los)
-  (foldr (lambda (s c) (if (> (string-length s) 3) (add1 c) c)) 0 los))
+(define (remove-short-words len ds)
+  (filter (lambda (d) (not (< (string-length (defn-word d)) len))) ds))
 
-;; Alts:
-(define (count-3+ los)
-  (local [;; count+ : String Number -> Number
-          ;; Bump count if string has length > 3
-          (define (count+ s c) 
-            (if (> (string-length s) 3) (add1 c) c))]
-    (foldr count+ 0 los))
-
-(define (count-3+ los)
-  (local [;; 3+? : String -> Boolean
-          ;; Does the string have length > 3?
-          (define (3+? s) 
-            (> (string-length s) 3))]
-    (length (filter 3+? los))))
-
-(define (count-3+ los)
-  (length (filter (lambda (s) (> (string-length s) 3)) los)))
 \end{SaveVerbatim}
@@
 \noindent
-{\bf Problem 6 (20 points).}
+{\bf Problem 4 (20 points).}
 %
-Here's a data definition for choose-your-own-adventure books:
+Design an abstraction of the following two functions and re-create
+the original functions in terms of your abstraction.
+
 \begin{verbatim}
-;; A CYOA is one:
-;; - String
-;; - (make-choice String CYOA CYOA)
-(define-struct choice (q yes no))
-;; Interp: a choose-your-own-adventure book where a string
-;; is a conclusion and a choice is a yes/no question and
-;; the two adventures that follow as a consequnce of the answer.
+;; sqr-pos : [Listof Number] -> [Listof Number]
+;; Square each number in the list that is positive, remove the rest
+(check-expect (sqr-pos '(2 -3 4)) '(4 16))
+(define (sqr-pos xs)
+  (cond [(empty? xs) '()]
+        [(cons? xs)
+         (if (positive? (first xs))
+             (cons (sqr (first xs))
+                   (sqr-pos (rest xs)))
+             (sqr-pos (rest xs)))]))
+
+;; lowercase-lengths : [Listof String] -> [Listof Number]
+;; Produce a list of lengths for all lowercase strings in given list
+(check-expect (lowercase-lengths '("Hey" "you" "guys")) '(3 4))
+(define (lowercase-lengths xs)
+  (cond [(empty? xs) '()]
+        [(cons? xs)
+         (if (string-lower-case? (first xs))
+             (cons (string-length (first xs))
+                   (lowercase-lengths (rest xs)))
+             (lowercase-lengths (rest xs)))]))
 \end{verbatim}
-
-\noindent
-Design a function called {\tt ans} that consumes a {\tt CYOA} and a
-{\tt [Listof Boolean]} that represents answers to the yes/no
-questions: {\tt \#true} means yes, {\tt \#false} means no.  The {\tt
-  ans} function produces the {\tt CYOA} adventure that results after
-giving all of the answers in the list.  If the list contains too many
-answers, {\tt ans} just produces the conclusion of the adventure.
 
@@
 SOLUTION:
+;; filter&map : [Listof X] [X -> Boolean] [X -> Y] -> [Listof Y]
+(define (filter&map xs p f)
+  (cond [(empty? xs) '()]
+        [(cons? xs)
+         (if (p (first xs))
+             (cons (f (first xs))
+                   (filter&map (rest xs) p f))
+             (filter&map (rest xs) p f))]))
 
-(define a0 "A0")
-(define a1 (make-choice "Q1" "A1" "A2"))
-(define a2 (make-choice "Q2" a0 a1))
+(define (sqr-pos xs)
+  (filter&map xs positive? sqr))
 
-;; ans : CYOA [Listof Boolean] -> CYOA
-;; ...
-(check-expect (adv a0 (list #true #false)) "A0")
-(check-expect (adv a1 (list #true #false)) "A1")
-(check-expect (adv a2 (list #false #true)) "A1")
-(check-expect (adv a2 (list #false)) a1)
-(define (ans adv loa)
-  (cond [(string? adv) adv]
-        [(empty? loa) adv]
-        [(first loa)
-         (ans (choice-yes adv) (rest adv))]
-        [else
-         (ans (choice-no adv) (rest adv))]))
+(define (lowercase-lengths xs)
+  (filter&map xs string-lower-case? string-length))
 \end{SaveVerbatim}
@@
 
+\newpage
+
+\noindent
+[Space for problem 4.]
+
+\newpage
+
+\noindent
+{\bf Problem 5 (20 points).}
+%
+Consider the following data definition (and examples) for representing sentences:
+\begin{verbatim}
+;; A Sentence is a [NEListof String]
+
+(define sent1 (list "a" "dog" "barked"))
+(define sent2 (list "a" "cat" "meowed"))
+\end{verbatim}
+The {\tt [NEListof X]} data definition is used to represent 
+non-empty lists of elements:
+\begin{verbatim}
+;; A [NEListof X] is one of:
+;; - (cons X '())
+;; - (cons X [NEListof X])
+\end{verbatim}
+Given a {\tt Sentence}, a \emph{bigram} is a list of two consecutive
+elements of the sentence. For example, {\tt sent1} has exactly two
+bigrams:
+\begin{verbatim}
+(list "a" "dog")
+(list "dog" "barked")
+\end{verbatim}
+Using the following data definition,
+\begin{verbatim}
+;; a Bigram is a (list String String)
+\end{verbatim}
+design a function {\tt bigrams} which takes a {\tt Sentence} and
+returns a list of {\tt Bigrams}.
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; bigrams : Sentence -> [Listof Bigram]
+;; List of all adjacent words in a sentence
+(check-expect (bigrams (list "one")) '())
+(check-expect (bigrams sent1)
+              (list (list "a" "dog") (list "dog" "barked")))
+(define (bigrams sent)
+  (cond [(empty? (rest sent)) '()]
+        [else
+         (cons (list (first sent) (second sent))
+               (bigrams (rest sent)))]))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
+\newpage
+\noindent
+[Space for problem 5.]
+
 \end{document}

### 2019-01-27 1/exams/f.tex
final.

@@
 (define S 1)
-(define (q z)
-  (cond [(< 4 z) (+ z S)]
-        [else 9]))
-(q 8)
+(define (w p) (< S p))
+
+(cond [(w 9) (* 8 2)]
+      [else 9])
 \end{verbatim}
@@
 
-(q 8)                                --[plug]-->
+(cond [(w 9) (* 8 2)] [else 9])      --[plug]-->
        ^^^^^
-(cond [(< 4 8) (+ 8 S)] [else 9])    --[arith]-->
+(cond [(< S 9) (* 8 2)] [else 9])    --[const]-->
+          ^
+(cond [(< 1 9) (* 8 2)] [else 9])    --[arith]-->
        ^^^^^^^
-(cond [#true (+ 8 S)] [else 9])      --[cond]-->
+(cond [#true (* 8 2)] [else 9])      --[cond]-->
 ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
-(+ 8 S)                              --[const]-->
-     ^
-(+ 8 1)                              --[arith]-->
+(* 8 2)                              --[arith]-->
 ^^^^^^^
-9
+16
 \end{SaveVerbatim}
@@
 \begin{verbatim}
-(define-struct qt (nw ne se sw))
+(define-struct roger (waters rabbit))
 \end{verbatim}
@@
 
-- make-qt  : Constructor
-- qt-nw    : Accessor
-- qt-ne    : Accessor
-- qt-se    : Accessor
-- qt-sw    : Accessor
-- qt?      : Predicate
+- make-roger   : Constructor
+- roger-waters : Accessor
+- roger-rabbit : Accessor
+- roger?       : Predicate
 \end{SaveVerbatim}
@@
 %
-Consider the following data definition (and examples) for representing sentences:
+Imagine designing the following two functions.  Now design an
+abstraction of both and define each original function in terms of that
+abstraction.  (It's not necessary to give the ``unabstracted''
+definitions unless it's helpful to you.)
 \begin{verbatim}
-;; A Sentence is a [NEListof String]
+;; find-even : [Listof Number] -> [Maybe Number]
+;; Find first even number, if there is one (or #false)
+(check-expect (find-even '(1 2 3 4)) 2)
+(check-expect (find-even '(1 3 5)) #false)
 
-(define sent1 (list "a" "dog" "barked"))
-(define sent2 (list "a" "cat" "meowed"))
+;; find-long : [Listof String] -> [Maybe String]
+;; Find first string longer than 5 letters, if there is one (or #false)
+(check-expect (find-long '("a" "really" "long")) "really")
+(check-expect (find-long '("a" "b" "c")) #false)
 \end{verbatim}
-Recall the following data definition of {\tt [NEListof X]}:
+Recall that {\tt [Maybe X]} is defined as:
 \begin{verbatim}
-;; A [NEListof X] is one of:
-;; - (cons X '())
-;; - (cons X [NEListof X])
+;; A [Maybe X] is one of:
+;; - X
+;; - #false
 \end{verbatim}
-Given a {\tt Sentence}, a \emph{bigram} is a list of two consecutive
-elements of the sentence. For example, {\tt sent1} has exactly two
-bigrams:
-\begin{verbatim}
-(list "a" "dog")
-(list "dog" "barked")
-\end{verbatim}
-Using the following data definition,
-\begin{verbatim}
-;; a Bigram is a (list String String)
-\end{verbatim}
-design a function {\tt bigrams} which takes a {\tt Sentence} and
-returns a list of {\tt Bigrams}.
 
@@
 
-;; bigrams : Sentence -> [Listof Bigram]
-;; List of all adjacent words in a sentence
-(check-expect (bigrams (list "one")) '())
-(check-expect (bigrams sent1)
-              (list (list "a" "dog") (list "dog" "barked")))
-(define (bigrams sent)
-  (cond [(empty? (rest sent)) '()]
-        [else
-         (cons (list (first sent) (second sent))
-               (bigrams (rest sent)))]))
+;; find : [Listof X] [X -> Boolean] -> [Maybe X]
+;; Find first element satisfying the predicate, if there is one (or #false)
+(define (find lon pred)
+  (cond [(empty? lon) #false]
+        [(cons? lon)
+         (cond [(pred (first lon)) (first lon)]
+               [else (find-even (rest lon))])]))
+
+(define (find-even lon)
+  (find lon even?))
+
+(define (find-long los)
+  (find los (lambda (s) (> (string-length s) 5))))
 \end{SaveVerbatim}
@@
 
+[Space for problem 3.]
+
+\newpage
+
 \noindent
@@
 %
-Here is a program that, given a sentence (see problem 3), computes a
-two element list consisting of the minimal and maximal words according
-to alphabetic ordering ({\tt string<?}).
-
+Here is a data definition for binary trees:
 \begin{verbatim}
-;; min-and-max : Sentence -> (list String String)
-;; Produces the min and max words in the sentence.
-(check-expect (min-and-max (list "the" "zoo" "has" "aardvarks"))
-              (list "aardvarks" "zoo"))
-(define (min-and-max s)
-  (cond [(empty? (rest s)) (list (first s) (first s))]
-        [else
-         (local [(define w (first s))
-                 (define m+m (min-and-max (rest s)))]
-            (cond [(string<? w (first m+m))
-                   (list w (second m+m))]
-                  [(string<? (second m+m) w)
-                   (list (first m+m) w)]
-                  [else m+m]))]))
+;; A [BT X] is one of:
+;; - (make-leaf)
+;; - (make-node X [BT X] [BT X])
+(define-struct node (val left right))
+(define-struct leaf ())
 \end{verbatim}
-Redesign this program to use an accumulator-based helper function with
-two accumulators: the mininmal and maximal elements of the sentence
-seen so far.
+Design a function that consumes a binary tree of strings and
+constructs a single string that results from concating the strings
+in left to val to right order.  For example:
+\begin{verbatim}
+(define l (make-leaf))
+(define t (make-node "b" (make-node "a" l l) (make-node "c" l l)))
+\end{verbatim}
+The function should produce {\tt "abc"} on {\tt t}.
 
@@
 
-(define (min-and-max s)
-  (min-and-max/a (rest s) (first s) (first s)))
-
-;; min-and-max/a : [Listof String] String String -> (list String String)
-;; Accumulator: min and max element seen so far
-(define (min-and-max/a s min max)
-  (cond [(empty? s) (list min max)]
-        [(cons? s)
-         (local [(define w (first s))]
-           (cond [(string<? w min) 
-                  (min-and-max/a (rest s) w max)]
-                 [(string<? max w)
-                  (min-and-max/a (rest s) min w)]
-                 [else
-                  (min-and-max/a (rest s) min max)]))]))
 \end{SaveVerbatim}
@@
 
-\newpage
+%% Using an accumulator-based design, design the program {\tt to10} that
+%% converts a list of digits into its correspdonding number.  The
+%% elements are listed in order of most to least significance,
+%% so for example, {\tt (to10 '(9 1 2 3))} should produce 9123.
+%% \begin{verbatim}
+%% ;; A Digit is a Natural in [0,9]
 
-[Space for problem 4.]
+%% ;; to10 : [Listof Digit] -> Number
+%% ;; Convert digits (in order of most to least significance) to a number
+%% (check-expect (to10 '(1 2 3)) 123)
+%% (check-expect (to10 '(7 2 8)) 728)
+%% \end{verbatim}
+%% Here is a hint, 728 can be computed as {\tt (+ 8 (* 10 (+ 2 (* 10 (+ 7 0)))))}.
+%% A design that does not use an accumulator will receive only partial credit.
+
+%% \begin{SaveVerbatim}{VerbEnv}
+
+%% SOLUTION:
+
+%% (define (to10 lod)
+%%   (local [(define (to10/a lod a)
+%%             (cond [(empty? lod) a]
+%%                   [(cons? lod)
+%%                    (to10/a (rest lod)
+%%                            (+ (first lod)
+%%                               (* 10 a)))]))]                              
+%%     (to10/a lod 0)))
+%% \end{SaveVerbatim}
+
+%% \censor{%
+%% \noindent
+%% \BUseVerbatim{VerbEnv}}
 
@@
 %
-Design the following program for determining if two sentences (see problem 3) are equal,
-meaning they are lists of the same length with the same elements (in the same order):
+Design the following variant of {\tt map} that takes a binary function
+and \emph{two} lists that are assumed to be of equal length and
+produces list of results formed by applying the function to the
+elements of each list.  Only tests are given, you have to come up with
+the rest:
 \begin{verbatim}
-;; sent=? : Sentence Sentence -> Boolean
-;; Are the two sentences the same?
+(check-expect (2map < '(1 4 3) '(3 1 2)) '(#true #false #false))
+(check-expect (2map string-ith '("fred" "wilma") '(1 0)) '("r" "w"))
 \end{verbatim}
-(You may not use {\tt equal?} in the design of this program.)
 
@@
 
-;; sent=? : Sentence Sentence -> Boolean
-;; Are the two sentences equal?
-(check-expect (sent=? (list "one") (list "one")) #true)
-(check-expect (sent=? (list "one") (list "two")) #false)
-(check-expect (sent=? (list "one") (list "one" "two")) #false)
-(check-expect (sent=? (list "one" "two") (list "one" "two")) #true)
-(define (sent=? s1 s2)
-  (cond [(empty? (rest s1))
-         (and (empty? (rest s2))
-              (string=? (first s1) (first s2)))]
-        [else
-         (and (not (empty? (rest s2)))
-              (string=? (first s1) (first s2))
-              (sent=? (rest s1) (rest s2)))]))
+;; 2map : [X Y -> Z] [Listof X] [Listof Y] -> [Listof Z]
+;; Binary variant of map
+;; ASSUME: lists are equal length
+(define (2map f xs ys)
+  (cond [(empty? xs) '()]
+        [(cons? xs)
+         (cons (f (first xs) (first ys))
+               (2map f (rest xs) (rest ys)))]))
 \end{SaveVerbatim}
@@
 Bioinformatics and Computational Biology (CBCB), you're asked to
-design a program for computing the \emph{reverse complement} of DNA strands.
+improve code written by last year's intern.  That intern designed a
+program for computing the \emph{reverse complement} of DNA strands.
 %
@@
 %
-Design a program that, given a DNA strand, computes its reverse complement.
-
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
+Last year's intern came up with the following:
+\begin{verbatim}
 ;; A DNA is a [Listof Base]
@@
         [(string=? b "G") "C"]))
+\end{verbatim}
+You notice that {\tt rev-comp} traverses the DNA sequence twice, once
+with {\tt map} and once with {\tt reverse}.  Improve this design by
+revising {\tt rev-comp} to use an accumulator to traverse the DNA
+sequence just once.
+
+\newpage
+
+\noindent
+[Space for problem 6.]
+
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+(define (rev-comp dna)
+  (local [;; DNA DNA -> DNA
+          ;; ACCUM: the reverse complement of the sequence so far
+          (define (rev-comp/a dna rcomp)
+            (cond [(empty? dna) rcomp]
+                  [(cons? dna)
+                   (rev-comp/a (rest dna)
+                               (cons (comp (first dna)) rcomp))]))]
+     (rev-comp/a dna '())))                      
 \end{SaveVerbatim}
@@
 strings and counts how many times the first string occurs within the
-second.
+second, {\bf including possibly overlapping occurences}.
 For example,
@@
 \item {\tt "ho"} occurs three times in {\tt "ho ho ho"}, while
-\item {\tt "ho ho"} occurs once in {\tt "ho ho ho"} (notice that occurences
-  cannot overlap).
+\item {\tt "ho ho"} occurs {\bf twice} in {\tt "ho ho ho"} (notice that occurences
+  overlap on the middle {\tt "ho"}).
 \end{itemize}
@@
 ;; occurs : String String -> Natural
-;; Count (non-overlapping) occurences of first string in second
+;; Count (possibly overlapping) occurences of first string in second
 ;; Assume: first string is non-empty
-;; Termination: because s1 is non-empty, each recursive call is on a strictly
+;; Termination: each recursive call is on a strictly
 ;; smaller s2, therefore eventually reaching first cond clause (trivial case).
 (check-expect (occurs "ho" "ho ho ho") 3)
-(check-expect (occurs "ho ho" "ho ho ho") 1)
+(check-expect (occurs "ho ho" "ho ho ho") 2)
 (define (occurs s1 s2)
@@
           [(string=? s1 (substring s2 0 s1-len))
-           (add1 (occurs s1 (substring s2 s1-len)))]
+           (add1 (occurs s1 (substring s2 1)))]
           [else

### 2019-07-10 0/exams/m1-practice.tex
Start of m1 practice.
Practice midterm 1.

--- 2018-10-18 1/exams/m1.tex
@@
 \begin{center}
-  {\Large {\bf CMSC 131A, Midterm 1}\\
+  {\Large {\bf CMSC 198Q, Midterm 1 (PRACTICE)}\\
     \censor{{\bf SOLUTION}} \ \\
     \ \\
-    Fall 2018 \ \\
+    Summer 2019 \ \\
   }
@@
 
-\noindent
-UID: \verb|__________________________________________|
-
-\vspace{2em}
-
 \begin{center}
@@
   3 & 15 \\ \hline
-  4 & 15 \\ \hline
-  5 & 40 \\ \hline
-  Total: & 90\\
+  4 & 10 \\ \hline
+  5 & 20 \\ \hline
+  Total: & 65\\
   \hline
@@
 device other than your brain and may not communicate with anyone.
-You have 50 minutes to complete the test.
+You have 60 minutes to complete the test.
 
@@
 
-(define (swap s i)
-  (string-append (substring s i)
-                 (substring s 0 i)))
+(define (dear p)
+  (string-append "Dear " 
+                 (title-abbrv (person-title p)) 
+                 " " 
+                 (person-last p) 
+                 ":"))
+
+;; title-string : Title -> String
+;; Render title as a string abbreviation
+(check-expect (title-abbrv "d") "Dr.")
+(check-expect (title-abbrv "r") "Mr.")
+(check-expect (title-abbrv "s") "Ms.")
+(define (title-abbrv t)
+  (cond [(string=? t "d") "Dr."]
+        [(string=? t "r") "Mr."]
+        [(string=? t "s") "Ms."]))
 \end{SaveVerbatim}
@@
 \noindent
-{\bf Problem 4 (15 points).}  
+{\bf Problem 4 (10 points).}
 %
-Design a program that takes a list of numbers and produces the sum of
-positive numbers in the list.  For example, if the list contains 3, 4,
--1, 0, -2, and 5, the sum of positive numbers (3+4+5) is 12.  You may
-use the following data definition:
+Do you ever mispell words like ``peice'' or ``acheive'' (which are
+correctly spelled ``piece'' and ``achieve'')?  Turns out a lot of
+people make \emph{transposition} errors like this when typing.  Based
+on this insight, you decide to make a next generation messaging app
+that helps users by giving them a \emph{transpose} operation.  The
+idea is that as a user types, they can place their cursor between two
+letters that need to be transposed and invoke {\tt transpose}.  To
+implement this feature, you define the following function for
+transposing letters at a given position in a string that has at
+least 2 letters in it:
+\begin{verbatim}
+;; transpose : String Index -> String
+;; Transpose characters at left and right of index i.
+;; Assumes string has length >= 2 and 1<=i, i+1<=length.
+(check-expect (transpose "ab" 1) "ba")
+(check-expect (transpose "peice" 2) "piece")
+(check-expect (transpose "student's" 8) "students'")
+(define (transpose s i) ...)
+\end{verbatim}
 
-\begin{verbatim}
-;; A LoN (list of numbers) is one of:
-;; - '()
-;; - (cons Number LoN)
-;; Interp: an arbitrarily long list of numbers
-\end{verbatim}
+\noindent
+Give a correct definition for \verb|transpose|.  (You only need to
+provide code, not the design steps.)
 
@@
 
-;; sum-pos : LoN -> Number
-;; Count the number of positive numbers in list
-(check-expect (sum-pos '()) 0)
-(check-expect (sum-pos (list 1 -2 8 0)) 9)
-(define (sum-pos lon)
-  (cond [(empty? lon) 0]
-        [(cons? lon)
-         (if (< (first lon) 0)
-             (sum-pos (rest lon))
-             (+ (first lon) (sum-pos (rest lon))))]))
+(define (transpose s i)
+  (string-append (substring s 0 (sub1 i))
+                 (substring s i (add1 i))
+                 (substring s (sub1 i) i)
+                 (substring s (add1 i))))
 \end{SaveVerbatim}
@@
 \newpage
+\noindent
+{\bf Problem 5 (20 points).}  Growing tired of Bug, you
+decide to build a new video game.  Part of the game consists of flying
+billiard balls that move in straight lines at varying velocities until
+hitting other balls or bouncing off walls and other obstacles.  After
+making some sketches, you decide on the following data representation
+for billiard balls:
+
+\begin{verbatim}
+;; A BB is a:
+;; (make-bb Color 
+;;          (make-posn Integer Integer) 
+;;          (make-vel Integer Integer))
+
+;; A Color is one of:
+;; - "red"
+;; - "yellow"
+;; - "green"
+
+(define-struct bb (color center vel))
+(define-struct vel (deltax deltay))
+\end{verbatim}
+The interpretation of a \verb|BB| is that the color is the color of
+the ball, the \verb|posn| is the location of the center of the ball,
+and the \verb|vel| structure describes the ball's velocity as a change
+along the $x$-axis (\verb|deltax|) and $y$-axis (\verb|deltay|) in one
+clock tick.
+
+\vskip 1em
 
 \noindent
-{\bf Problem 5 (40 points).} You've been hired by Nintendo to design
-their next blockbuster game: Dots.  It consists of, well, dots: lots
-and lots of dots, each of which occupies a coordinate in space.  You
-come up with the following data definitions for representing a single
-dot and a collection of dots:
-
-\begin{verbatim}
-;; A Dots is one of:
-;; - empty
-;; - (cons Dot Dots)
-;; Interp: collection of an arbitrary number of dots
-
-;; A Dot is a (make-posn Number Number)
-;; Interp: a dot's position in space
-\end{verbatim}
-
-\vspace{1em}
-
-\noindent
-{\bf Problem 5(a) [10 points].} Write templates for \verb|Dot| and
-\verb|Dots| functions.
+Design a function called \verb|tock : BB -> BB| that calculates where
+a given billiard ball will be, based on its velocity, after one tick
+of the clock and assuming it does not encounter any obstacle.
 
@@
 
-(define (dot-template d)
-  (... (posn-x d)
-       (posn-y d)
-       ...))
-
-(define (dots-template ds)
-  (cond [(empty? ds) ...]
-        [(cons? ds) (... (dot-template (first ds))
-                         (dots-template (rest ds)) ...)]))
-\end{SaveVerbatim}
-
-\vskip 18em
-
-\censor{\vspace{-15em}}
-\censor{%
-\noindent\BUseVerbatim{VerbEnv}}
-
-\noindent
-{\bf Problem 5(b) [10 points].} Write a stub for the following function:
-
-\begin{verbatim}
-;; nw : Dots -> Dots
-;; Move given dots in the northwest direction one unit
-(check-expect (nw empty) empty)
-(check-expect (nw (cons (make-posn 3 4) empty))
-              (cons (make-posn 4 5) empty))
-\end{verbatim}
-
-\begin{SaveVerbatim}{VerbEnv}
-
-SOLUTION:
-
-(define (nw ds) empty)
-
-\end{SaveVerbatim}
-
-\censor{%
-\noindent\BUseVerbatim{VerbEnv}}
-
-\newpage
-
-\noindent
-{\bf Problem 5(c) [20 points].}
-%
-Design two functions: {\tt dot-flip} reflects a given dot over the
-y-axis, for example, $(3,4)$ becomes $(-3,4)$, and {\tt dots-flip}
-reflects a collection of dots over the y-axis.
-\begin{SaveVerbatim}{VerbEnv}
-
-;; SOLUTION:
-
-;; dot-flip : Dot -> Dot
-;; Flip dot over y-axis
-(check-expect (dot-flip (make-posn 3 4)) (make-posn -3 4))
-(define (dot-flip d)
-  (make-posn (- (posn-x d)) (posn-y d)))
-
-;; dots-flip : Dots -> Dots
-;; Flip dots over y-axis
-(check-expect (dots-flip empty) empty)
-(check-expect (dots-flip (cons (make-posn 3 4) empty))
-              (cons (make-posn -3 4) empty))
-(define (dots-flip ds)
-  (cond [(empty? ds) empty]
-        [(cons? ds) 
-         (cons (dot-flip (first ds))
-               (dots-flip (rest ds)))]))                   
+;; tock : BB -> BB
+;; Move ball one tick based on velocity
+(check-expect (tock (make-bb "red" (make-posn 2 5) (make-vel 1 -2)))
+              (make-bb "red" (make-posn 3 3) (make-vel 1 -2)))
+(define (tock bb)
+  (make-bb (bb-color bb)
+           (make-posn (+ (posn-x (bb-center bb))
+                         (vel-deltax (bb-vel bb)))
+                      (+ (posn-y (bb-center bb))
+                         (vel-deltay (bb-vel bb))))
+           (bb-vel bb)))
 \end{SaveVerbatim}
@@
 
-\newpage
-
-\noindent
-[Extra space, should you need it.]
-
-%% Design a program that consumes an arbitrarily large collection of
-%% numbers and produces the count of non-negative numbers in the
-%% collection.
-
-%% \begin{SaveVerbatim}{VerbEnv}
-
-%% SOLUTION:
-
-%% ;; A LoN (list of numbers) is one of:
-%% ;; - '()
-%% ;; - (cons Number LoN)
-
-%% ;; count-non-neg : LoN -> Natural
-%% ;; Count the number of non-negative numbers in the given list.
-%% (check-expect (count-non-neg '()) 0)
-%% (check-expect (count-non-neg (cons -1 (cons 1 (cons -2 '())))) 1)
-%% (define (count-non-neg lon)
-%%   (cond [(empty? lon) 0]
-%%         [(cons? lon) 
-%%          (+ (if (<= 0 (first lon)) 1 0)
-%%             (count-non-neg (rest lon)))]))
-%% \end{SaveVerbatim}
-
-%% \censor{%
-%% \noindent
-%% \BUseVerbatim{VerbEnv}}
-
 \end{document}

### 2019-07-17 0/exams/m1.tex
Midterm.

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{times}
\usepackage{xcolor,censor}
\usepackage{fontspec}
\usepackage{fancyvrb}

\setmainfont{Times New Roman}

\censorruleheight=0ex
\StopCensoring

\begin{document}

\begin{center}
  {\Large {\bf CMSC 198Q, Midterm 1}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Summer 2019 \ \\
  }
\end{center}

\vspace{2em}

\noindent
NAME:\verb|________________________________________________________|

\vspace{2em}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 20 \\ \hline
  Total: & 70\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 60 minutes to complete the test.

\vskip 1em

\noindent
The phrase ``design a program'' or ``design a function'' means follow
the steps of the design recipe.  Unless specifically asked for, you do
not need to provide intermediate products like templates or stubs,
though they may be useful to help you construct correct solutions.

\vskip 1em

\noindent
You may use any of the data definitions given to you within this exam
and do not need to repeat their definitions.

\vskip 1em

\noindent
Unless specifically instructed otherwise, you may use any built-in BSL
functions or data types.

\vskip 1em

\noindent
When writing tests, you may use a shorthand for writing check-expects
by drawing an arrow between two expressions to mean you expect the
first to evaluate to same result as the second.  For example, you may
write \verb|(add1 3)| $\rightarrow$ \verb|4| instead of
\verb|(check-expect (add1 3) 4)|.

\newpage

\noindent
{\bf Problem 1 (10 points).}
%
For the following program, write out each step of computation. At each
step, underline the expression being simplified. Label each step as
being ``arithmetic'' (meaning any built-in operation),
``conditional'', ``plug'' (for plugging in an argument for a function
parameter), or ``constant'' for replacing a constant with its value.

\begin{verbatim}
(define R 2)
(define (q s) (< R (string-length s)))
(cond [(= 1 2) (q "fred")]
      [else (q "wilma")])
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(cond [(= 1 2) (q "fred")] [else (q "wilma")]) -->[arith]
       ^^^^^^^
(cond [#false (q "fred")] [else (q "wilma")]) -->[cond]
^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^   
(cond [else (q "wilma")])                     -->[cond]
^^^^^^^^^^^^^^^^^^^^^^^^^
(q "wilma")                                   -->[plug]
^^^^^^^^^^^
(< R (string-length "wilma"))                 -->[const]
   ^
(< 2 (string-length "wilma"))                 -->[arith]
     ^^^^^^^^^^^^^^^^^^^^^^^
(< 2 5)                                       -->[arith]
^^^^^^^
#true
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 2 (10 points).}
%
For the following structure definition, list the names of every
function it creates.  For each function, classify it as being either a
constructor, accessor, or predicate.

\begin{verbatim}
(define-struct up (side vote chuck))
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:

- make-up    : Constructor
- up-side    : Accessor
- up-vote    : Accessor
- up-chuck   : Accessor
- up?        : Predicate
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\vspace{3in}
\begin{verbatim}
(define U (make-up "exam" #false (make-up 2019 #true "red")))
\end{verbatim}
Using only accessor functions and {\tt U}, write an expression that
produces {\tt 2019}.

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:

(up-side (up-chuck U))
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 3 (15 points).}
%
You've been hired by Wheel of Fortune, a popular game show where
contestants try to reveal a hidden word or phrase by guessing letters.
So for example, if the hidden word is ``HELLO,'' they start knowing
there are five letters: \_ \_ \_ \_ \_.  Suppose they guess ``L,''
then the two occurrences are revealed: \_ \_ L L \_.  If they the
guess ``H,'' they'd see: H \_ L L \_, and so on.

You start by focusing on the representation of individual (potentially
hidden) letters and settle on the following data definition (here {\tt
  1String} just means a string of length 1):
\begin{verbatim}
;; A Letter is one of:
;; - (make-hidden 1String)
;; - 1String
;; Interp: (make-hidden s) means s has not yet been guessed and is 
;; being hidden, otherwise the letter has been guessed and revealed.
(define-struct hidden (letter))
\end{verbatim}
So {\tt (make-hidden "L")} represents the letter {\tt "L"} being 
hidden in a word or phrase.  Once guessed, it would become simply {\tt
  "L"}.

Desing a function {\tt guess} which takes a {\tt Letter} and a {\tt
  1String} and reveals the hidden letter if the guess is correct and
otherwise leaves it alone if it's incorrect or the letter is already
revealed.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; Letter 1String -> Letter
(check-expect (guess "a" "b") "a")
(check-expect (guess (make-hidden "a") "b") (make-hidden "a"))
(check-expect (guess (make-hidden "a") "a") "a")
(define (guess l s)
  (cond [(hidden? l) 
         (cond [(string=? s (hidden-letter l)) s]
               [else l])]
        [(string? l) l]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage
\noindent
{\bf Problem 4 (15 points).}
%
Points in three-dimensional space are a lot like their two-dimensional
counterparts, just with one more dimension.  So rather than pairs of
numbers $(x,y)$, 3D points are determined by triples $(x,y,z)$.
%
Similarly, the distance of a 3D point to the origin is a
straightforward extension of the formula for calculating the distance
of a 2D point to the origin:
\[
\sqrt (x^2 + y^2 + z^2)
\]
And the point $(2,3,6)$ has a distance of 7 to the origin.

Design a data representation for 3D points and a function for
computing the distance of a 3D point to the origin.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; A 3D is a (make-3d Number Number Number)
(define-struct 3d (x y z))

;; 3D -> Number
;; Compute distance to origin
(check-expect (dist (make-3d 2 3 6)) 7)
(define (dist p)
  (sqrt (+ (sqr (3d-x p)) (sqr (3d-y p)) (sqr (3d-z p)))))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage
\noindent
{\bf Problem 5 (20 points).} 
%
After completing the simple line text editor you developed in lab 4,
you decide to extend your design so that in addition to representing
text with a cursor somewhere within it, you can also represent text
where some portion of it has been ``selected'' (for example if the
user were to click and drag a cursor across some part of the text).

\begin{verbatim}
;; A Line is one of:
;; - (make-txt String String)
;; - (make-sel String String String)
;; Interpretation: (make-txt string1 string2) is a line of text consisting
;; of string1 and string2 with a cursor placed between them.
;; (make-sel string1 string2 string3) is a line of text consisting of
;; string1, string2, and string3, but string2 is currently selected.
(define-struct txt (left right))
(define-struct sel (left selected right))
\end{verbatim}
So for example, if you want to represent the text ``Hello'' with
``ll'' selected, it would be 
\begin{verbatim}
(make-sel "He" "ll" "o")
\end{verbatim}
Design a function \verb|select-all| that takes a line and selects all
available text in the line, regardless of where the cursor is or what
may currently be selected.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; Line -> Line
;; Select all of the text in the line
(check-expect (line-select (make-txt "a" "b")) (make-sel "" "ab" ""))
(check-expect (line-select (make-sel "a" "b" "c")) (make-sel "" "abc" ""))
(define (line-select-all l)
  (cond [(txt? l) (make-sel "" (string-append (txt-left l) (txt-right l)) "")]
        [(sel? l)
         (make-sel ""
                   (string-append (sel-left l)
                                  (sel-selected l)
                                  (sel-right l)) "")]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\end{document}

### 2019-07-25 0/exams/f-practice.tex
Practice exam.

--- 2017-10-28 1/exams/m2.tex
@@
 \usepackage{fullpage}
-\usepackage{times}
 \usepackage{xcolor,censor}
 \usepackage{fancyvrb}
+\usepackage{fontspec}
+\usepackage{alltt}
+
+\setmainfont{Times New Roman}
 
 \censorruleheight=0ex
-%\StopCensoring
+\StopCensoring
 
@@
 \begin{center}
-  {\Large {\bf CMSC 131A, Midterm 2 (Practice)}\\
+  {\Large {\bf CMSC 198Q, Final Exam (Practice)}\\
     \censor{{\bf SOLUTION}} \ \\
     \ \\
-    Fall 2017 \ \\
+    Summer 2019 \ \\
   }
@@
   1 & 12 \\ \hline
-  2 & 12 \\ \hline
+  2 & 15 \\ \hline
   3 & 15 \\ \hline
@@
   5 & 15 \\ \hline
-  6 & 10 \\ \hline
-  7 & 15 \\ \hline \hline
-  Total: & 94\\
+  6 & 15 \\ \hline \hline
+  Total: & 87\\
   \hline
@@
 device other than your brain and may not communicate with anyone.
-You have 50 minutes to complete the test.
+You have 120 minutes to complete the test.
 
@@
 [Boolean -> String] -> String
-[Listof Boolean] [Listof Number] -> Number
-[X] . [Number -> X] -> X
-[X Y] . [X -> Y] [Listof X] -> [Listof Y]
-[X Y Z] . [X -> Y] [Y -> Z] X -> Z
+[Lo Boolean] [Lo Number] -> Number
+[Number -> X] -> X
+[X -> Y] [Lo X] -> [Lo Y]
+[X -> Y] [Y -> Z] X -> Z
 \end{verbatim}
@@
 \noindent
-{\bf Problem 2 (12 points).}
-%
-For each of the following functions, provide the most general signature
-that correctly describes the function:
+{\bf Problem 2 (15 points).} Here is a data definition:
+\begin{verbatim}
+;; A Zoinks is one of:
+;; - Jinkies
+;; - Jeepers
 
-\begin{verbatim}
-;; f :
+;; A Jinkies is a (make-velma String Number)
+(define-struct velma (my glasses))
 
-(define (f x)
-  (+ (sqr x) 2))
+;; A Jeepers is a (make-daphne Zoinks Zoinks)
+(define-struct daphne (ruh roh))
+\end{verbatim}
 
-;; g :
+\noindent
+1) Define a template for functions that consume the following kind of data.
 
-(define (g x)
-  (map add1 x))
+\noindent
+2) Give three different examples of {\tt Zoinks}, including at least one {\tt Jeepers}.
 
-;; h :
+\begin{SaveVerbatim}{VerbEnv}
 
-(define (h x y)
-  (cond [(empty? x) y]
-        [(cons? x)
-         (cons (sqr (first x))
-               (h (rest x) y))]))
+SOLUTION:
 
-;; i :
+(define (zoinks-template z)
+  (cond [(velma? z) (... (velma-my z) (velma-glasses z) ...)]
+        [(daphne? z) 
+         (... (zoinks-template (daphne-ruh z))
+              (zoinks-template (daphne-roh z)) ...)]))
 
-(define (i x)
-  (sqr (x 5)))
+(make-velma "" 0)
+(make-velma "a" 1)
+(make-daphne (make-velma "" 0)
+             (make-velma "a" 1))
 
-;; j :
+\end{SaveVerbatim}
 
-(define (j x)
-  (lambda (y)
-    (+ x y)))
-
-;; k :
-
-(define (k x y)
-  (cond [(= y 0) (list (x 0))]
-        [else
-          (cons (x y) (k x (sub1 y)))]))
-\end{verbatim}
+%\censor{\vspace{-2.5in}}
+\censor{%
+\noindent\BUseVerbatim{VerbEnv}}
 
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; avg : [Lo Number] -> Number
+;; Compute the average of a list of numbers
+;; Assume: the list is non-empty
+(check-expect (avg (list 3 4 5)) 4)
+(check-expect (avg (list 10 20)) 15)
+(define (avg lon)
+  (/ (sum lon) 
+     (length lon)))
+
+;; sum : [Lo Number] -> Number
+;; Sum the list of numbers
+(check-expect (sum '()) 0)
+(check-expect (sum (list 3 4 5)) 12)
+(define (sum lon)
+  (foldr + 0 lon))
+
+;; Alt: follow [Lo Number] template
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \newpage 
@@
 \begin{verbatim}
-;; A [NEListof X] is one of:
+;; A [NELo X] is one of:
 ;; - (cons X '())
-;; - (cons X [NEListof X])
+;; - (cons X [NELo X])
 ;; Interp: an arbitrarily long but non-empty sequence of items.
@@
 
-;; smallest : [NEListof Number] -> Number
+;; smallest : [NELo Number] -> Number
 ;; Produce the smallest number in the list
@@
 
-;; longest : [NEListof String] -> String
+;; longest : [NELo String] -> String
 ;; Produce the first longest string in the list
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; mostest : [NELo X] [X X -> X] -> X
+;; Produce the first most-est element in the list
+(define (mostest xs most)
+  (cond [(empty? (rest xs)) (first xs)]
+        [else
+          (most (first xs)
+                (mostest (rest xs) most))]))
+
+(define (longest los)
+  (mostest los longer))
+
+(define (smallest lon)
+  (mostest lon min))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \newpage
@@
 For full credit, use list abstraction functions.  For partial credit,
-follow the template for {\tt{[Listof Number]}}.
+follow the template for {\tt{[Lo Number]}}.
 
@@
 
+;; count-pos : [Lo Number] -> Number
+;; Count the number of non-negative numbers in given list
+(check-expect (count-pos '()) 0)
+(check-expect (count-pos (list 3 -1 0 2)) 2)
+(define (count-pos lon)
+  (foldr (lambda (n c) (if (positive? n) (add1 c) c)) 0 lon))
+
+;; Alt:
+(define (count-pos lon)
+  (length (filter positive? lon)))
 \end{SaveVerbatim}
@@
 \noindent
-{\bf Problem 6 (10 points).}
-%
-It's been said that {\tt foldr} is \emph{the} fundamental list
-abstraction function because any function that follows the list
-template can be expressed in terms of {\tt foldr}.  Put that theory to
-the test by providing an equivalent definition for the following
-function in terms of {\tt foldr}:
-
-\begin{verbatim}
-;; map : [X Y] . [X -> Y] [Listof X] -> [Listof Y]
-;; Apply f to each element of list, collect a list of results
-(define (map f ls)
-  (cond [(empty? ls) '()]
-        [(cons? ls)
-         (cons (f (first ls))
-               (map f (rest ls)))]))
-\end{verbatim}
-
-\noindent
-(Just provide code, not a full design.)
-
-\newpage
-
-\noindent
-{\bf Problem 7 (15 points).}
+{\bf Problem 6 (15 points).}
 %
@@
 
+\begin{SaveVerbatim}{VerbEnv}
+
+SOLUTION:
+
+;; most-games : Tournament -> Number
+;; Compute most games any team has to win to win tournament
+(check-expect (most-games "NY") 0)
+(check-expect (most-games (make-bracket "LA" "HOU")) 1)
+(check-expect (most-games (make-bracket (make-bracket "LA" "DC") "HOU")) 2)
+(define (most-games t)
+  (cond [(string? t) 0]
+        [(bracket? t)
+         (add1 (max (most-games (bracket-t1 t))
+                    (most-games (bracket-t2 t))))]))
+\end{SaveVerbatim}
+
+\censor{%
+\noindent
+\BUseVerbatim{VerbEnv}}
+
 \end{document}
